### 2019-01-29 multi_body_interactions/papers/theory/permissions.tex
added document explaning permissions to reproduce published materials

\documentclass[nofootinbib,notitlepage,11pt]{revtex4-1}

%%% header / footer
\usepackage{fancyhdr} % easier header and footer management
\pagestyle{fancy} % page formatting style
\fancyhf{}
\renewcommand{\headrulewidth}{0pt} % horizontal line in header

\usepackage{hyperref,listings}
\setlength{\parindent}{0pt}
\parskip 6pt

\begin{document}

\title{Concerning permissions for reproducing material}

\maketitle
\thispagestyle{fancy}

The manuscript here submitted to {\it New Journal of Physics} contains
two figures which are reproduced from another manuscript currently in
press (accepted, but not yet published) with {\it Nature}, titled
``Emergence of multi-body interactions in a fermionic lattice clock''
(DOI: \verb|10.1038/s41586-018-0661-6|).  Both authors of this
manuscript are also authors of the one in {\it Nature}, and grant {\it
  New Journal of Physics} permission to use the reproduced material in
question.  Furthermore, it is editorial policy of Springer Nature, the
publishing company that owns {\it Nature}, that ownership of copyright
remains with the authors of its published material.  Verbatim from the
\href{https://www.nature.com/reprints/permission-requests.html}{Springer
  Nature permission request webpage}:
\begin{lstlisting}[breaklines]
  The author of articles published by SpringerNature do not usually
  need to seek permission for re-use of their material as long as the
  journal is credited with initial publication.

  Ownership of copyright in in original research articles remains with
  the Author
\end{lstlisting}
The publisher-to-be of the reproduced material thus also implicitly
grants permission to use it, conditional on proper citation practices.
As the reproduced material is currently in press and not yet
published, the manuscript submitted to {\it New Journal of Physics}
cites the arXiv version of the original material
(\href{https://arxiv.org/abs/1803.11282}{arXiv:1803.11282}).  Once the
original material is published, this citation will be updated
accordingly.

If we have in any way misunderstood editorial policy of either
Springer Nature or the Institute of Physics, please let us know and we
will take whatever actions necessary to sort out issues relating to
reproduced content and copyright.  In the meantime, we hope the
editors of {\it New Journal of Physics} can proceed to consider our
submitted manuscript without delay.

~\\

Sincerely,

Michael A. Perlin

Ana Maria Rey

\end{document}