### 2019-02-02 multi_body_interactions/papers/theory/paper.tex
updated multi-body paper with edits in response to referee comments

\documentclass[preprint,showkeys,nofootinbib]{revtex4-1}

% linking references
\usepackage{hyperref}
\hypersetup{
  breaklinks=true,
  colorlinks=true,
  linkcolor=blue,
  urlcolor=cyan,
}

% general physics / math packages and commands
\usepackage{physics,amsmath,amssymb,braket,dsfont,bm}
\renewcommand{\t}{\text} % text in math mode
\newcommand{\f}{\dfrac} % shorthand for fractions
\newcommand{\p}[1]{\left(#1\right)} % parenthesis
\renewcommand{\sp}[1]{\left[#1\right]} % square parenthesis
\renewcommand{\set}[1]{\left\{#1\right\}} % curly parenthesis
\newcommand{\bk}{\braket} % shorthand for braket

\newcommand{\g}{\text{g}}
\newcommand{\e}{\text{e}}
\newcommand{\x}{\text{x}}
\newcommand{\y}{\text{y}}
\newcommand{\z}{\text{z}}

\renewcommand{\c}{\hat{c}}
\newcommand{\n}{\hat{n}}

\newcommand{\A}{\mathcal{A}}
\newcommand{\B}{\mathcal{B}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\E}{\mathcal{E}}
\newcommand{\G}{\mathcal{G}}
\renewcommand{\H}{\mathcal{H}}
\newcommand{\I}{\mathcal{I}}
\newcommand{\K}{\mathcal{K}}
\renewcommand{\L}{\mathcal{L}}
\newcommand{\M}{\mathcal{M}}
\newcommand{\N}{\mathcal{N}}
\renewcommand{\O}{\mathcal{O}}
\renewcommand{\P}{\mathcal{P}}
\newcommand{\Q}{\mathcal{Q}}
\renewcommand{\S}{\mathcal{S}}
\newcommand{\U}{\mathcal{U}}
\newcommand{\1}{\mathds{1}}

\newcommand{\mA}{m_{\text{A}}} % symbol for the mass of an atom

% "left vector" arrow; requires tikz package
\newcommand{\lvec}[1]
{\reflectbox{\ensuremath{\vec{\reflectbox{\ensuremath{#1}}}}}}

% figures
\usepackage{graphicx} % for figures
\usepackage{grffile} % help latex properly identify figure extensions
\graphicspath{{./figures/}} % set path for all figures
\usepackage[caption=false]{subfig} % subfigures (via \subfloat[]{})

% inline lists
\usepackage[inline]{enumitem}
\setlist[enumerate,1]{label={(\roman*)}}

% for feynman diagrams
\usepackage{tikz,tikz-feynman}
\tikzset{
  baseline = (current bounding box.center)
}
\tikzfeynmanset{
  compat = 1.1.0,
  every feynman = {/tikzfeynman/small}
}
\newcommand{\shrink}[1]{\scalebox{0.8}{#1}} % for smaller diagrams

% color definitions (used in a figure)
\usepackage{xcolor}
\definecolor{lightblue}{RGB}{31,119,180}
\definecolor{orange}{RGB}{255,127,14}
\definecolor{green}{RGB}{44,160,44}
\definecolor{lightred}{RGB}{214,39,40}

% proper coloring inside math environment
\makeatletter
\def\mathcolor#1#{\@mathcolor{#1}}
\def\@mathcolor#1#2#3{
  \protect\leavevmode
  \begingroup
    \color#1{#2}#3
  \endgroup
}
\makeatother
\newcommand{\bmu}{\mathcolor{lightblue}{\mu}}
\newcommand{\onu}{\mathcolor{orange}{\nu}}
\newcommand{\grho}{\mathcolor{green}{\rho}}
\newcommand{\re}{\mathcolor{lightred}{\text{e}}}

\usepackage{setspace} % to modify spacing in table of contents

% leave a note in the text, visible in the compiled document
\newcommand{\note}[1]{\textcolor{red}{#1}}

% for strikeout text
% normalem included to prevent underlining titles in the bibliography
\usepackage[normalem]{ulem}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

\title{Effective multi-body SU($N$)-symmetric interactions of
  ultracold fermionic atoms on a 3-D lattice}

\author{M A Perlin}
\email{mika.perlin@gmail.com}
\author{A M Rey}
\affiliation{JILA, National Institute of Standards and Technology and
  University of Colorado, 440 UCB, Boulder, Colorado 80309, USA}
\affiliation{Center for Theory of Quantum Matter, 440 UCB, Boulder,
  Colorado 80309, USA}
\affiliation{Department of Physics, University of Colorado, 390 UCB,
  Boulder, Colorado 80309, USA}

\begin{abstract}
  Rapid advancements in the experimental capabilities with ultracold
  alkaline-earth-like atoms (AEAs) bring to a surprisingly near term
  the prospect of performing quantum simulations of spin models and
  lattice field theories exhibiting SU($N$) symmetry.  In order to
  analyze recent experiments preparing high density samples of
  strongly interacting ${}^{87}$Sr atoms in a three-dimensional
  optical lattice, we develop a low-energy effective theory of
  fermionic AEAs which exhibits emergent multi-body SU($N$)-symmetric
  interactions, where $N$ is the number of atomic nuclear spin levels.
  We fully characterize the low-lying excitations in our effective
  theory, and compare predictions of many-body interaction energies
  with direct measurements of many-body excitation spectra in an
  optical lattice clock.  Our work makes the first step in enabling a
  controlled, bottom-up experimental investigation of multi-body
  SU($N$) physics.
\end{abstract}

\keywords{\it ultracold atoms, optical lattices, SU(N) magnetism,
  multi-body interactions}

\maketitle

\begin{spacing}{1.3}
  \tableofcontents
\end{spacing}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:introduction}

Fermionic alkaline-earth atoms (AEAs), in addition to other atoms such
as ytterbium (Yb) sharing similar electronic structure, are currently
the building blocks of the most precise atomic clocks in the world
\cite{derevianko2011colloquium, katori2011optical, ludlow2015optical}.
These atoms have a unique, ultra-narrow optical transition between
metastable ${}^1S_0$ and ${}^3P_0$ electronic orbital states, i.e.~the
``clock states'', that allows for coherence times which can exceed 100
seconds \cite{porsev2004hyperfine, santra2004properties}.
Furthermore, AEAs can be trapped in fully controllable optical lattice
potentials and interrogated with ultra-stable lasers that can resolve
and probe their rich hyperfine spectra, consisting of $N$ different
nuclear spin levels with $N$ as large as 10 in strontium (${}^{87}$Sr)
and 6 in ytterbium (${}^{173}$Yb).

In 2015 the ${}^{87}$Sr optical lattice clock (OLC) at JILA, operated
in a one-dimensional (1-D) lattice at microkelvin temperatures,
achieved a total fractional uncertainty of $2\times10^{-18}$
\cite{bloom2014optical, nicholson2015systematic}.  More recently
(2017), a new generation of OLCs became operational at JILA,
interrogating a Fermi degenerate gas of ${}^{87}$Sr atoms in a 3-D
lattice at nanokelvin temperatures \cite{campbell2017fermidegenerate}.
All of these atoms' degrees of freedom, including the electronic
orbital, nuclear spin, and motional states, can be fully controlled
with high fidelity in a 3-D lattice \cite{daley2008quantum,
  gorshkov2009alkalineearthmetal, daley2011quantum,
  daley2011statedependent}.  With frequency measurements reaching the
$10^{-19}$ fractional uncertainty level, the new OLCs are thus
enabling an exciting opportunity to probe, for the first time, quantum
dynamics with sub-millihertz spectral resolution
\cite{campbell2017fermidegenerate}.

A wonderful consequence of the efforts to build better clocks is the
development of highly controllable quantum simulators of many-body
systems in the strongly-interacting regime, where inter-particle
interactions set the largest energy scale relevant for system dynamics
\cite{swallows2011suppression, lemke2011wave,
  campbell2017fermidegenerate}.  The marriage between precision clock
spectroscopy and quantum many-body physics \cite{taie2012su,
  martin2013quantum, scazza2014observation, cappellini2014direct,
  zhang2014spectroscopic, rey2014probing} has an enormous potential to
enable novel explorations of physics for the same reason that makes
AEAs such remarkable time-keepers.  Specifically, due to the lack of
electronic orbital angular momentum in the ${}^1S_0$ and ${}^3P_0$
states, AEAs exhibit decoupled orbital and nuclear spin degrees of
freedom.  For atoms with $N$ nuclear spin levels, this decoupling
leads to nuclear-spin-conserving SU($N$)-symmetric interactions
governed entirely by orbital-state parameters
\cite{cazalilla2009ultracold, taie2012su, zhang2014spectroscopic}.

The presence of this exotic SU($N$) symmetry in a highly controllable
experimental platform opens the door to experimental studies of
e.g.~the SU($N$) Heisenberg model, whose phase diagram is believed to
exhibit features such as a chiral spin liquid (CSL) phase with
topological order and fractional statistics \cite{hermele2009mott,
  hermele2011topological, chen2016syntheticgaugefield}.  In addition
to illuminating open questions in our understanding of the fractional
quantum Hall effect and unconventional superconductivity
\cite{lee2006doping, lee2008high, gong2014emergent}, the CSL can
support non-Abelian excitations which allow for universal topological
quantum computation \cite{freedman2004class, hermele2011topological}.
Harnessing the SU($N$)-symmetric interactions of AEAs might also
enable the simulation of various lattice gauge theories
\cite{wieseu.-j.2013ultracold, zohar2016quantum}, some of which share
important qualitative features with quantum chromodynamics such as
few-body bound states and confinement \cite{banerjee2013atomic,
  rico2018nuclear}.  These direct, quantum simulations have an
extraordinary potential to provide novel insights by circumventing
e.g.~severe sign problems which plague classical simulations of
strongly interacting fermionic systems \cite{wieseu.-j.2013ultracold,
  wu2003exact}.

In this work, we investigate the first experimental capabilities with
ultracold fermionic AEAs to prepare high-density samples in a 3-D
lattice with multiple occupation of individual lattice sites
\cite{goban2018emergence}.  Specifically, we consider ground-state
preparation of isolated few-body systems in the deep-lattice limit,
and carry out a bottom-up investigation of emergent multi-body
interactions on multiply-occupied lattice sites.  These multi-body
interactions appear in a low-energy effective theory of the atoms, and
inherit the SU($N$) symmetry of their bare, pair-wise interactions,
thereby enabling experimental studies of multi-body SU($N$) physics
through the exquisite capabilities with OLCs.  We briefly summarize
the experimental procedures relevant to our work in Appendix
\ref{sec:experiment_summary}.

Though effective multi-body interactions have previously been studied
in the context of harmonically \cite{johnson2012effective,
  yin2014universal} and lattice-confined \cite{johnson2009effective}
neutral bosons prepared in a single hyperfine state, our work deals
for the first time with fermions that have internal degrees of freedom
and multiple collisional parameters.  Nonetheless, due to the SU($N$)
symmetry of these collisions, we are able to find a simple way to
express effective multi-body Hamiltonians and fully characterize their
low-lying many-body energy eigenstates.  While some past work has
detected experimental signatures of multi-body interactions in the
form of quantum phase revivals \cite{will2010timeresolved}, we perform
a direct comparison of the many-body interaction energies predicted by
our low-energy effective theory with experimental measurements of
density-dependent atomic energy spectra performed in
ref.~\cite{goban2018emergence}, similarly to the measurements with
bosons performed in ref.~\cite{mark2011precision}.

The remainder of this paper is structured as follows.  In section
\ref{sec:overview} we provide an overview of the relevant one- and
two-body physics of ultracold atoms on a lattice, and summarize our
main result of effective multi-body interactions between these atoms.
In section \ref{sec:low_energy} we discuss our method for deriving a
low-energy effective theory, provide a perturbative expansion for the
net effective Hamiltonian, and compute all $M$-body Hamiltonians
through third order in the low-energy effective theory.  We then
analyze the low-lying excitations of the effective theory in section
\ref{sec:spectra}, comparing spectral predictions with experimental
measurements, and study the orbital-state dynamics of nuclear spin
mixtures interrogated via Rabi spectroscopy.  Finally, we summarize
and conclude our findings in section \ref{sec:summary}, and provide
some discussion of future outlooks.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Overview}
\label{sec:overview}

Consider a collection of fermionic atoms with total nuclear spin $I$
and two atomic orbital states loaded into a translationally invariant
3-D optical lattice.  While an external trapping potential will
generally break discrete translational symmetry of the lattice, any
background inhomogeneity can be made negligible by spectroscopically
addressing a sufficiently small region of the lattice
\cite{goban2018emergence}.  Throughout this paper, we work strictly in
the deep-lattice regime with negligible tunneling between lattice
sites.  We also assume a so-called ``magic-wavelength'' lattice for
which both atomic orbital states experience identical lattice
potentials \cite{ye2008quantum}.  The single-particle Hamiltonian of
the atoms can then be written in the form
\begin{align}
  H_0 = \sum_{i,n,\mu,s} E_n \c_{in\mu s}^\dag \c_{in\mu s},
  \label{eq:H_0}
\end{align}
where $\c_{in\mu s}$ is a fermionic operator which annihilates a
single atom on lattice site $i\in\mathbb{Z}^3$ in motional state
$n\in\mathbb{N}_0^3$ with nuclear spin $\mu\in\set{-I,-I+1,\cdots,I}$
(i.e.~projected onto a quantization axis) and orbital state
$s\in\set{\g,\e}$; and $E_n$ is the energy of a single atom in
motional state $n$.  In a harmonic trap approximation we would have
$E_n=\p{3/2+n_\x+n_\y+n_\z}\omega$ for an on-site angular trap
frequency $\omega$ (with the reduced Planck constant $\hbar=1$
throughout this paper), but in general the aharmonicity of the lattice
potential will cause a non-negligible shift in motional state
energies.

In the absence of hyperfine coupling, as when addressing the spinless
${}^1S_0$ ($\g$) and ${}^3P_0$ ($\e$) orbital states of AEAs,
interactions between any two atoms are governed by their orbital
states alone, and are therefore characterized by four scattering
lengths $a_X$ with $X\in\set{\g\g,\e\g^-,\e\g^+,\e\e}$, where the $+$
($-$) superscript denotes symmetrization (anti-symmetrization) of a
two-body orbital state under particle exchange.  In the low-energy
limit, we can write the bare two-body interaction Hamiltonian in the
form \cite{gorshkov2010twoorbital}
\begin{align}
  H_{\t{int}} = \sum_{\substack{\mu<\nu\\s}}
  G_s \int d^3x~ \hat\rho_{\mu s} \hat\rho_{\nu s}
  + G_+ \sum_{\mu,\nu}
  \int d^3x~ \hat\rho_{\mu,\e} \hat\rho_{\nu,\g}
  + G_- \sum_{\mu,\nu} \int d^3x~
  \hat\psi_{\mu,\e}^\dag \hat\psi_{\nu,\g}^\dag
  \hat\psi_{\nu,\e} \hat\psi_{\mu,\g},
  \label{eq:H_int_start}
\end{align}
where $\hat\psi_{\mu s}$ is a fermionic field operator for atoms with
nuclear spin $\mu$ and orbital state $s$;
$\hat\rho_{\mu s}\equiv\hat\psi_{\mu s}^\dag\hat\psi_{\mu s}$ is an
atomic density field operator; and the coupling constants $G_X$ are
defined in terms of the scattering lengths by
\begin{align}
  G_{s=\g,\e} \equiv \f{4\pi}{\mA}~ a_{ss}, &&
  G_\pm \equiv \f{2\pi}{\mA}~ \p{a_{\e\g+} \pm a_{\e\g-}},
  \label{eq:couplings}
\end{align}
where $\mA$ is the mass of a single atom.  Defining for brevity
\begin{align}
  G^{qr}_{st} \equiv \left\{
    \begin{array}{ll}
      G_q & ~ q = r = s = t \\
      G_+ & ~ q \ne r ~ \t{and} ~ (q,r) = (s,t) \\
      G_- & ~ q \ne r ~ \t{and} ~ (q,r) = (t,s) \\
      0 & ~ \t{otherwise}
    \end{array}\right.,
  \label{eq:coupling_tensor}
\end{align}
where $q,r,s,t\in\set{\g,\e}$ are orbital state indices, we can
alternately write the bare two-body interaction Hamiltonian in the
more compact form
\begin{align}
  H_{\t{int}}
  = \f12 \sum_{\substack{q,r,s,t\\\mu,\nu}} G^{qr}_{st} \int d^3x~
  \hat\psi_{\mu s}^\dag \hat\psi_{\nu t}^\dag
  \hat\psi_{\nu r} \hat\psi_{\mu q}.
  \label{eq:H_int_fields}
\end{align}
For nuclear spins $\mu,\nu$, the symbol $G^{qr}_{st}$ gives the
coupling constant between the two-atom states
$(\mu,q)+(\nu,r)\leftrightarrow(\mu,s)+(\nu,t)$.  Note that
\eqref{eq:H_int_start} and \eqref{eq:H_int_fields} neglect
momentum-dependent corrections to $s$-wave interactions, which are
generally captured (in a low-energy limit) by the use of an
energy-dependent interaction pseudo-potential \cite{blume2002fermi}.
These corrections will appear only at third order in our effective
theory, as corrections to two-body interaction energies, and do not
affect our discussions of multi-body interactions.  We therefore defer
further discussion of momentum-dependent corrections to the end of
section \ref{sec:two_body}, and their calculation to Appendix
\ref{sec:momentum_dependence}.

We now expand the field operators $\hat\psi_{\mu s}$ in the Wannier
basis for a 3-D lattice, such that
$\hat\psi_{\mu s}(x) = \sum_{i,n} \phi_{in}(x) \c_{in\mu s}$ with
spatial wavefunctions $\phi_{in}$ and fermionic annihilation operators
$\c_{in\mu s}$ indexed by lattice sites $i$ and motional states $n$.
Invoking the tight-binding approximation, we assume that the spatial
overlap integral in \eqref{eq:H_int_fields} is negligible unless all
wavefunctions are localized at the same lattice site; we discuss the
breakdown of this approximation and its consequences for our
low-energy effective theory in Appendix \ref{sec:error}.  The relevant
spatial overlap integral is then
\begin{align}
  K^{k\ell}_{mn}
  \equiv \int d^3x~ \phi_{im}^* \phi_{in}^* \phi_{i\ell} \phi_{ik},
  \label{eq:K_klmn}
\end{align}
which for a lattice with discrete translational invariance is
independent of the lattice site $i$.  The two-body interaction
Hamiltonian can be written in terms of this overlap integral as
\begin{align}
  H_{\t{int}}
  = \f12 \sum_{\substack{i,k,\ell,m,n\\q,r,s,t\\\mu,\nu}}
  K^{k\ell}_{mn} G^{qr}_{st}
  \c_{im\mu s}^\dag \c_{in\nu t}^\dag \c_{i\ell\nu r} \c_{ik\mu q}
  \equiv \f12 \sum K^{k\ell}_{mn} G^{qr}_{st}
  \c_{m\mu s}^\dag \c_{n\nu t}^\dag \c_{\ell\nu r} \c_{k\mu q},
  \label{eq:H_int}
\end{align}
where for brevity we will henceforth suppress the identical site index
($i$) on all operators, and implicitly sum over all free indices in a
summand (i.e.~indices which do not have a fixed value).  We may also
at times suppress motional state indices on the overlap integral
$K^{k\ell}_{mn}$, in which case the suppressed indices are implicitly
zero (corresponding to a motional ground state); i.e.
\begin{align}
  K^{\ell m}_n \equiv K^{\ell m}_{n,0},
  &&
  K^m_n \equiv K^{m,0}_{n,0},
  &&
  K_{mn} \equiv K^{0,0}_{mn},
  &&
  K_n \equiv K^{0,0}_{n,0},
  &&
  K \equiv K^{0,0}_{0,0}.
  \label{eq:K}
\end{align}
For simplicity, we will also generally work in a gauge for which all
two-body overlap integrals are real, such that
$K^{k\ell}_{mn}={K^{k\ell}_{mn}}^*=K^{mn}_{k\ell}$.  The existence of
such a gauge is guaranteed by the analytic properties of the Wannier
wavefunctions $\phi_{in}$ \cite{kohn1959analytic}.

Current experiments with ${}^{87}$Sr can prepare up to five atoms
atoms in the same (ground) orbital state on a single lattice site, and
coherently address states with a single orbital excitation per lattice
site (see Appendix \ref{sec:experiment_summary})
\cite{goban2018emergence}.  The temperature of the atoms in these
experiments is much smaller than single-atom motional excitation
energies $\Delta_n\equiv E_n-E_0$ for $n>0$, which implies a
negligible thermal occupation of excited motional states.  A naive
calculation of the multi-body orbital excitation spectrum, however,
assuming that atoms only occupy motional ground states with spatial
wavefunctions $\phi_{i,0}$, fails to reproduce the experimentally
observed orbital excitation spectrum of multiply-occupied lattice
sites.  The reason that such a calculation fails to reproduce the
observed excitation spectrum has to do with the fact that the two-body
Hamiltonian $H_{\t{int}}$ in \eqref{eq:H_int} is not diagonal in the
basis of single-particle motional eigenstates (i.e.~$\phi_{in}$).  For
interacting atoms, the true motional ground states are therefore
admixtures of the non-interacting motional eigenstates; when the
off-diagonal elements of $H_{\t{int}}$ are comparable in magnitude to
the excitation energies $\Delta_n$, this admixture can have
non-negligible weight outside the non-interacting motional
ground-state subspace.  In order to compute the correct many-body
orbital excitation spectrum, we develop a low-energy effective theory
for such atoms.  As our work is motivated by experiments that
initialize all atoms in identical (i.e.~ground) orbital and motional
states, to simplify our theory we assume that any $N$ atoms on a
single lattice site occupy $N$ nuclear spin states.  Ground-state
preparation implies that multiple occupation of a single nuclear spin
state on any given lattice site is initially forbidden by fermionic
statistics, and subsequently cannot occur in the absence of inter-site
effects or hyperfine coupling.

Our low-energy effective theory exhibits SU($N$)-symmetric multi-body
interactions, such that the effective interaction Hamiltonian can be
written in the form
\begin{align}
  H_{\t{int}}^{\t{eff}} = \sum_{M=2}^{2I+1} \sum_{p\ge1} H_M^{(p)},
\end{align}
where $H_M^{(p)}$ is an $M$-body Hamiltonian of order $p$ in the
coupling constants $G^{qr}_{st}$, $I$ is the total nuclear spin of
each atom (e.g.~$I=9/2$ for ${}^{87}$Sr), and the sum terminates at
$2I+1$ because this is the largest number of atoms which may initially
occupy a single lattice site.  We explicitly compute all $M$-body
Hamiltonians $H_M\equiv\sum_p H_M^{(p)}$ through order $p=3$, yielding
effective two-, three-, and four-body interactions.  To facilitate
understanding and comparison with experimental measurements of
density-dependent atomic energy spectra [34], in which states with
multiple orbital excitations on a single lattice site are
spectroscopically off-resonant, we also consider a restriction of our
theory to states with at most one orbital excitation per lattice site.
In this case, we find that SU($N$) symmetry allows us to express
multi-body Hamiltonians in the simple form
\begin{align}
  H_M = \sum_{\abs{\set{\mu_j}}=M}
  \p{U_{M,\g} \prod_{\alpha=1}^M \n_{\mu_\alpha,\g}
    + U_{M,+} \n_{\mu_1,\e} \prod_{\alpha=2}^M \n_{\mu_\alpha,\g}
    + U_{M,-} \c_{\mu_1,\g}^\dag \c_{\mu_2,\e}^\dag
    \c_{\mu_2,\g} \c_{\mu_1,\e} \prod_{\alpha=3}^M \n_{\mu_\alpha,\g}},
  \label{eq:H_M_preview}
\end{align}
where $\c_{\mu s}\equiv\c_{0,\mu s}$ is a fermionic annihilation
operator for an atom in the motional ground state with nuclear spin
$\mu$ and orbital state $s$;
$\n_{\mu s}\equiv \c_{\mu s}^\dag \c_{\mu s}$ is a number operator;
and the sum is performed over all choices of nuclear spins $\mu_j$
with $j=1,2,\cdots,M$ for which all $\mu_j$ are distinct, or
equivalently all choices of $\mu_j$ for which the set $\set{\mu_j}$
contains $M$ elements, for a total of ${2I+1\choose M}\times\p{M!}$
nuclear spin combinations.  The coefficients $U_X$ can be expressed in
terms of the coupling constants of the bare two-body interaction
Hamiltonian (see Appendix \ref{sec:U_X}).  The key feature of the
$M$-body interactions in \eqref{eq:H_M_preview} is that they take a
form identical to the two-body interactions between motional
ground-state atoms, but with the addition of $M-2$ spectator atoms.
This form is a direct consequence of the SU($N$) symmetry of
underlying two-body interactions.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Low-energy effective theory}
\label{sec:low_energy}

As the bare two-body Hamiltonian $H_{\t{int}}$ is not diagonal with
respect to single-particle motional states, the problem of determining
many-body spectra and eigenstates nominally involves all motional
degrees of freedom of the atoms involved.  If we are interested in
addressing internal atomic states at zero temperature, however, in
principle we need only to consider the motional ground state
associated with each many-body internal atomic state.  We can then
ignore all excited motional states, as they will be neither thermally
occupied nor externally interrogated.  This procedure makes the
many-body problem more tractable by greatly reducing the
dimensionality of the relevant Hilbert space.

We denote the space of all internal states of atoms in
single-particle, non-interacting (multi-particle, interacting)
motional ground states by $\H_{\t{ground}}^{\t{single}}$
($\H_{\t{ground}}^{\t{multi}}$).  For sufficiently weak interactions,
one can construct a unitary transformation $U$ acting on the full
many-body Hilbert space which rotates $\H_{\t{ground}}^{\t{multi}}$
into $\H_{\t{ground}}^{\t{single}}$ \cite{bravyi2011schrieffer}.  The
unitary $U$ can then be used to construct an {\it effective
  Hamiltonian} $H_{\t{int}}^{\t{eff}}$ on
$\H_{\t{ground}}^{\t{single}}$ that reproduces the spectrum of
$H_{\t{int}}$ on $\H_{\t{ground}}^{\t{multi}}$, namely
$H_{\t{int}}^{\t{eff}}=UH_{\t{int}}U^\dag$.  This method for
constructing an effective theory is commonly known as the
Schrieffer-Wolff transformation, named after the authors of its
celebrated application in relating the Anderson and Kondo models of
magnetic impurities in metals \cite{schrieffer1966relation}.

Using the machinery developed in ref.~\cite{bravyi2011schrieffer} for
performing a rotation between low-energy subspaces of a perturbed
(i.e.~interacting) and unperturbed (i.e.~non-interacting) Hamiltonian,
we derive an expansion for the effective interaction Hamiltonian
$H_{\t{int}}^{\t{eff}}$ in terms two-body interaction Hamiltonian
$H_{\t{int}}$ (see Appendix \ref{sec:eff_derivation}).  This expansion
takes the form
\begin{align}
  H_{\t{int}}^{\t{eff}} = \sum_{p\ge1} H_{\t{int}}^{(p)},
  \label{eq:H_int_eff}
\end{align}
where $H_{\t{int}}^{(p)}$ is order $p$ in $H_{\t{int}}$.  Letting
$\E_0$ denote the orthogonal complement of
$\H_{\t{ground}}^{\t{single}}$ (i.e.~$\E_0$ is the space of all states
with at least one atom in an excited motional state), $\B_0\p{\E_0}$
denote an eigenbasis of $\E_0$ with respect to the single-particle
Hamiltonian $H_0$, and $E_\alpha$ denote the motional energy (with
respect to $H_0$) of a state $\ket\alpha\in\E_0$ relative to the
corresponding motional ground-state energy, we define the operator
\begin{align}
  \I \equiv \sum_{\ket\alpha\in\B_0\p{\E_0}} \f{\op\alpha}{E_\alpha},
\end{align}
which sums over projections onto excited states with corresponding
energetic suppression factors.  The operator $\I$ together with the
projector $\P_0$ onto $\H_{\t{ground}}^{\t{single}}$ allows us
concisely write the first few terms in \eqref{eq:H_int_eff} as
\begin{align}
  H_{\t{int}}^{(1)} = \P_0 H_{\t{int}} \P_0,
  &&
  H_{\t{int}}^{(2)} = -\P_0 H_{\t{int}} \I H_{\t{int}} \P_0,
  \label{eq:H_int_1_2}
\end{align}
\begin{align}
  H_{\t{int}}^{(3)}
  = \P_0 H_{\t{int}} \I H_{\t{int}} \I H_{\t{int}} \P_0
  - \f12\sp{\P_0 H_{\t{int}} \P_0,
    \P_0 H_{\t{int}} \I^2 H_{\t{int}} \P_0}_+,
  \label{eq:H_int_3}
\end{align}
where $\sp{X,Y}_+\equiv XY+YX$.  The naive idea of neglecting all
excited states and using $H_{\t{int}}$ directly to describe the
spectrum of single-particle motional ground states is thus equivalent
to truncating our expansion for $H_{\t{int}}^{\t{eff}}$ at first
order.  In addition to this first order term, the expansion involves
{\it effective corrections} to the action of $H_{\t{int}}$ on the
single-particle motional ground states (i.e.~on
$\H_{\t{ground}}^{\t{single}}$) in the form of higher-order terms with
{\it intermediate} or {\it virtual occupation} of excited states,
manifest in $\I$.

Substituting the definition of $\I$ into
\eqref{eq:H_int_1_2}-\eqref{eq:H_int_3} yields expressions that are
highly reminiscent of standard non-degenerate perturbation theory in
quantum mechanics, but which nonetheless exhibit crucial differences.
The first, and most obvious difference is that these expressions are
operator equations, and that the sums over virtual states are
performed over a basis for the orthogonal complement of the subspace
$\H_{\t{ground}}^{\t{single}}$, rather than a basis for the orthogonal
complement of a single state, as in non-degenerate perturbation
theory.  Second, the non-degeneracy condition in standard perturbation
theory is here elevated to a restriction on the magnitude of the
perturbation $H_{\t{int}}$ relative to the spectral gap $\Delta$ of
the non-interacting Hamiltonian $H_0$ between
$\H_{\t{ground}}^{\t{single}}$ and the excited subspace $\E_0$.
Specifically, the validity of \eqref{eq:H_int_1_2}-\eqref{eq:H_int_3}
is conditional only on $\norm{H_{\t{int}}}\le\Delta/2$, where
$\norm{X}\equiv\max_{\ket\psi\in\H}\sqrt{\bk{\psi|X^\dag X|\psi}}$ is
the operator norm, with no restrictions on spectral gaps or
degeneracies within $\H_{\t{ground}}^{\t{single}}$ \cite{davis1969new,
  bravyi2011schrieffer}.  Finally, the effective theory involves no
corrections to the many-body energy eigenstates.  Including such
corrections would invalidate the effective theory, which preserves the
eigenstates of the non-interacting Hamiltonian $H_0$ on
$\H_{\t{ground}}^{\t{single}}$.

As a last comment, we note that our chosen method for constructing an
effective Hamiltonian is distinct from adiabatic elimination methods
which are commonly used in the atomic physics and quantum optics
communities to develop effective theories for e.g.~the low-lying
levels of a Lambda system \cite{brion2007adiabatic,
  james2007effective, reiter2012effective, sanz2016adiabatic}.  Unlike
the perturbative, but exact Schieffer-Wolff transformation, adiabatic
elimination methods use approximations which rely on the fast dynamics
of excited states.  While generally reasonable, these approximations
must be made carefully to avoid potential problems with
self-consistency (see section 3 of ref.~\cite{brion2007adiabatic}),
and yield no obvious or straightforward means to compute effective
corrections beyond second order in the couplings between low- and
high-energy sectors of a Hilbert space \cite{james2007effective,
  reiter2012effective, sanz2016adiabatic}.  While at least one attempt
at systematically computing higher-order corrections in the framework
of adiabatic elimination has recently been made
\cite{sanz2016adiabatic}, the resulting expressions do not lend
themselves as nicely to analytical treatment, and were in any case
found by the authors to be equivalent to a Schrieffer-Wolff
transformation.

\subsection{Diagrammatic representation of effective Hamiltonians}

The form of the bare two-body interaction Hamiltonian $H_{\t{int}}$ in
\eqref{eq:H_int} motivates a diagrammatic representation of terms in
the effective Hamiltonian $H_{\t{eff}}$ in \eqref{eq:H_int_eff},
similarly to the diagrams used to represent elements of the scattering
matrix in field theory.  Effective $M$-body interaction terms at order
$p$ in $H_{\t{int}}$ can be represented by directed graphs containing
$p$ internal vertices.  Each internal vertex and its associated edges
correspond respectively to a coupling constant and the associated
field operators in $H_{\t{int}}$.  An example 2-vertex diagram
representing an effective 3-body interaction term is provided in
figure \ref{fig:diagram}.  All diagrams are read from left to right to
construct a sequence of operators from right to left; the internal
vertices of a diagram are thus strictly ordered, with the $n$-th
internal vertex from the left corresponding to the $n$-th interaction
Hamiltonian $H_{\t{int}}$ from the right in \eqref{eq:H_int_1_2} or
\eqref{eq:H_int_3}.  Solid (dashed) lines represent field operators
acting on the lowest (arbitrary) motional states.  Spatial overlap
factors at each vertex are determined by the motional states of the
edges which connect to (i.e.~field operators associated with) that
vertex.  While it is possible to construct explicit rules for
determining the energetic suppression factors (i.e.~from $\I$) of the
term represented by a diagram, these factors are most easily
determined by examination of the effective Hamiltonians in
\eqref{eq:H_int_1_2} and \eqref{eq:H_int_3}.

\begin{figure}
  \centering
  \(
  \begin{tikzpicture}
    \begin{feynman}
      \vertex (v1);
      \vertex[above left = of v1] (f1) {$\bmu,\g$};
      \vertex[below left = of v1] (f2) {$\onu,\re$};
      \vertex[right = of v1] (vm);
      \vertex[right = of vm] (v2);
      \vertex[above = of vm] (f3) {$\bmu,\g$};
      \vertex[below = of vm] (f4) {$\grho,\g$};
      \vertex[below right = of v2] (f5) {$\grho,\re$};
      \vertex[above right = of v2] (f6) {$\onu,\g$};
      \vertex [left = 0.5em of v1] {$G^{\g\re}_{\g\re}$};
      \vertex [right = 0.5em of v2] {$G^{\re\g}_{\g\re}$};
      \diagram* {
        (f1) --[fermion, color = lightblue] (v1)
        --[fermion, color = lightblue] (f3),
        (f2) --[fermion, color = orange] (v1)
        --[charged scalar,
        edge label = {$\mathcolor{black}{n}
          \mathcolor{orange}{\nu}\mathcolor{black}{,\re}$},
        color = orange] (v2),
        (f4) --[fermion, color = green] (v2)
        --[fermion, color = green] (f5),
        (v2) --[fermion, color = orange] (f6), };
    \end{feynman}
  \end{tikzpicture}
  \begin{array}{ll}
    &= \P_0 \p{\f12 K_n G^{\re\g}_{\g\re}
      \c_{\onu,\g}^\dag \c_{\grho,\re}^\dag
      \c_{\grho,\g} \c_{n\onu,\re}}
      \f1{E_n}
      \p{\f12 K_n G^{\g\re}_{\g\re}
      \c_{\bmu,\g}^\dag \c_{n\onu,\re}^\dag
      \c_{\onu,\re} \c_{\bmu,\g}} \P_0 \\[1em]
    &= \f14 \f{K_n^2}{E_n}~ G^{\g\re}_{\g\re} G^{\re\g}_{\g\re}~
      \P_0 \c_{\bmu,\g}^\dag \c_{\onu,\g}^\dag \c_{\grho,\re}^\dag
      \c_{\grho,\g} \c_{\onu,\re} \c_{\bmu,\g} \P_0
  \end{array}
  \)
  \caption{\footnotesize An example second-order diagram and the
    corresponding three-body interaction term in $H_{\t{int}}^{(2)}$,
    with $n>0$ and $\c_{\mu s}\equiv\c_{0,\mu s}$.  For the sake of
    presentation, this diagram has colors associated with nuclear spin
    and orbital states, an arrow on each line to emphasize that they
    are directed left-to-right, and an explicit coupling constant
    written next to each vertex; we will generally not include these
    features, as they are not necessary to uniquely identify the term
    represented by a diagram.  We will also drop explicit appearances
    of the ground-state projector $\P_0$ in our expressions, with the
    understanding that the low-energy effective theory implicitly
    addresses only single-atom motional ground states.}
  \label{fig:diagram}
\end{figure}

The diagram in figure \ref{fig:diagram} explicitly labels all edges
with indices of the corresponding field operators, but in general we
may suppress these indices, in which case the diagram includes a sum
over the suppressed indices.  These sums are performed over all
allowed values of the suppressed indices, with the restriction that
virtual states (i.e.~vertical slices of the diagram between internal
vertices) represented with dashed lines must have at least one
motional excitation.  While we include factors of $1/2$ from
$H_{\t{int}}$ as expressed in \eqref{eq:H_int} in the definition of a
diagram, in all but the two-body case these factors of $1/2$ will be
cancelled out by corresponding symmetry factors, i.e.~the appearance
of duplicate diagrams which are equal up to a relabeling of indices
(see Appendix \ref{sec:diagrams}).  The explicit signs and factor of
$1/2$ which appear in the effective Hamiltonians in
\eqref{eq:H_int_1_2} and \eqref{eq:H_int_3} are not included in the
definition of a diagram, and must be kept track of manually.

\subsection{Effective two-body interactions and renormalization}
\label{sec:two_body}

The effective two-body Hamiltonian in \eqref{eq:H_int_eff} has
contributions at all orders in the coupling constants, and can be
expanded in the form
\begin{align}
  H_2 = \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v);
        \vertex[above left = of v] (f1);
        \vertex[below left = of v] (f2);
        \vertex[above right = of v] (f3);
        \vertex[below right = of v] (f4);
        \diagram* {
          (f1) -- (v),
          (f2) -- (v),
          (v) -- (f3),
          (v) -- (f4) };
      \end{feynman}
    \end{tikzpicture}}
  - \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[above left = of v1] (f1);
        \vertex[below left = of v1] (f2);
        \vertex[right = of v1] (v2);
        \vertex[above right = of v2] (f3);
        \vertex[below right = of v2] (f4);
        \diagram* {
          (f1) -- (v1),
          (f2) -- (v1),
          (v2) -- (f3),
          (v2) -- (f4),
          (v1) --[scalar, half left] (v2),
          (v1) --[scalar, half right] (v2) };
      \end{feynman}
    \end{tikzpicture}}
  + \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[above left = of v1] (f1);
        \vertex[below left = of v1] (f2);
        \vertex[right = of v1] (v2);
        \vertex[right = of v2] (v3);
        \vertex[above right = of v3] (f3);
        \vertex[below right = of v3] (f4);
        \diagram* {
          (f1) -- (v1),
          (f2) -- (v1),
          (v3) -- (f3),
          (v3) -- (f4),
          (v1)
          --[scalar, half left] (v2)
          --[scalar, half left] (v3),
          (v1)
          --[scalar, half right] (v2)
          --[scalar, half right] (v3) };
      \end{feynman}
    \end{tikzpicture}}
  + \cdots \equiv \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex[blob] (v) {};
        \vertex[above left = of v] (f1);
        \vertex[below left = of v] (f2);
        \vertex[above right = of v] (f3);
        \vertex[below right = of v] (f4);
        \diagram* {
          (f1) -- (v),
          (f2) -- (v),
          (v) -- (f3),
          (v) -- (f4) };
      \end{feynman}
    \end{tikzpicture}},
  \label{eq:H_2_expansion}
\end{align}
where the blob on the right schematically represents the net effective
two-body interaction.  On physical grounds, the net effective two-body
interaction must clearly be finite, but individual sums over excited
states in the loop diagrams of \eqref{eq:H_2_expansion} may generally
diverge \cite{johnson2012effective}.  These divergences ultimately
appear due to our use of effective field theory to describe
inter-atomic interactions in \eqref{eq:H_int_start}, rather than a
detailed microscopic description of two-atom scattering.  Divergences
of this sort are a generic feature of field theories, and can be dealt
with using standard techniques, in particular through renormalization.
We therefore renormalize our coupling constants by introducing
counter-terms $\tilde G_X$ into the effective interaction Hamiltonians
via $G_X\to G_X+\tilde G_X$.  The values of counter-terms can be fixed
by enforcing that $G_X$ is equal to the net two-body interaction
strength, which has the added benefit of allowing us to express the
effective two-body interactions in terms of net, physically observable
two-body coupling constants on the right-hand side of
\eqref{eq:H_2_expansion}, rather than in terms of long sums at all
orders of the bare coupling constants.  Diagrammatically, this
renormalization condition is
\begin{align}
  \substack{
    \shrink{
      \begin{tikzpicture}
        \begin{feynman}
          \vertex (v);
          \vertex[above left = of v] (f1);
          \vertex[below left = of v] (f2);
          \vertex[above right = of v] (f3);
          \vertex[below right = of v] (f4);
          \diagram* {
            (f1) -- (v),
            (f2) -- (v),
            (v) -- (f3),
            (v) -- (f4) };
        \end{feynman}
      \end{tikzpicture}}
    \\ \O(G)}
  + \substack{
    \shrink{
      \begin{tikzpicture}
        \begin{feynman}
          \vertex[crossed dot] (v) {};
          \vertex[above left = of v] (f1);
          \vertex[below left = of v] (f2);
          \vertex[above right = of v] (f3);
          \vertex[below right = of v] (f4);
          \diagram* {
            (f1) -- (v),
            (f2) -- (v),
            (v) -- (f3),
            (v) -- (f4) };
        \end{feynman}
      \end{tikzpicture}}
    \\ \O(\tilde G)}
  - \substack{
    \shrink{
      \begin{tikzpicture}
        \begin{feynman}
          \vertex (v1);
          \vertex[above left = of v1] (f1);
          \vertex[below left = of v1] (f2);
          \vertex[right = of v1] (v2);
          \vertex[above right = of v2] (f3);
          \vertex[below right = of v2] (f4);
          \diagram* {
            (f1) -- (v1),
            (f2) -- (v1),
            (v2) -- (f3),
            (v2) -- (f4),
            (v1) --[scalar, half left] (v2),
            (v1) --[scalar, half right] (v2) };
        \end{feynman}
      \end{tikzpicture}}
    \\ \O(G^2)}
  + \cdots = \substack{
    \shrink{
      \begin{tikzpicture}
        \begin{feynman}
          \vertex (v);
          \vertex[above left = of v] (f1);
          \vertex[below left = of v] (f2);
          \vertex[above right = of v] (f3);
          \vertex[below right = of v] (f4);
          \diagram* {
            (f1) -- (v),
            (f2) -- (v),
            (v) -- (f3),
            (v) -- (f4) };
        \end{feynman}
      \end{tikzpicture}}
    \\ \O(G)},
  \label{eq:counter_term}
\end{align}
where we represent counter-terms by a crossed dot (i.e.~$\otimes$) at
a vertex. To leading order, it follows that the counter-terms are
second order in the coupling constants, i.e.~$\tilde G\sim G^2$.  This
finding implies that counter-terms only come into play at
next-to-leading order (NLO) in the calculation of any effective
$M$-body interaction Hamiltonian.  By construction, these
counter-terms exactly cancel all terms beyond leading order in
\eqref{eq:H_2_expansion}, which implies that the effective two-body
interaction Hamiltonian is simply
\begin{align}
  H_2 = \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v);
        \vertex[above left = of v] (f1);
        \vertex[below left = of v] (f2);
        \vertex[above right = of v] (f3);
        \vertex[below right = of v] (f4);
        \diagram* {
          (f1) -- (v),
          (f2) -- (v),
          (v) -- (f3),
          (v) -- (f4), };
      \end{feynman}
    \end{tikzpicture}}
  = \f12 \alpha_2^{(1)} \sum_{\abs{\set{\mu,\nu}}=2} G^{qr}_{st}
  \c_{\mu s}^\dag \c_{\nu t}^\dag \c_{\nu r} \c_{\mu q},
  \label{eq:H_2}
\end{align}
where for consistency with existing literature
\cite{johnson2012effective} we define $\alpha_2^{(1)} \equiv K$ as the
overlap integral between two atoms occupying single-particle motional
ground states.

Before moving on to consider effective three-body interactions, there
are two comments we must make concerning the result in \eqref{eq:H_2}.
First, this effective two-body interaction takes the same form as the
bare two-body interaction in \eqref{eq:H_int}, but without excited
motional states, and with renormalized coupling constants in place of
bare ones.  Our choice of renormalization scheme, while convenient for
the analytical development of a low-energy effective theory, no longer
allows us to use the coupling constants $G_X$ as defined by the
free-space scattering lengths $a_X$ in \eqref{eq:couplings} to compute
interaction energies.  Instead, we must first compute the effective
coupling constants $G_X^{\t{lattice}}\p{\U}$ in a lattice, which now
depend on the lattice depth $\U$, and in turn use the effective
coupling constants to compute interaction energies.  We discuss the
calculation of effective coupling constants in Appendix
\ref{sec:renormalization}.

Second, the overview we provided in section \ref{sec:overview}, and
consequently our result in \eqref{eq:H_2}, neglects the effect of
momentum-dependent two-body scattering \cite{blume2002fermi}.  When
the effective range of momentum-dependent interactions is comparable
to the scattering lengths $a_X$, as is the case for ultracold
${}^{87}$Sr, these interactions are third order in the coupling
constants $G_X$.  Just as the $\O\p{G^2}$ counter-terms do not affect
$M$-body interactions until next-to-leading order (NLO), the
$\O\p{G^3}$ momentum-dependent interactions do not come into play
until next-next-leading order (NNLO).  Given that we develop our
low-energy effective theory through third order in the coupling
constants, these interactions will not appear in any of our three- and
four-body calculations, but they do need to be considered in the
calculation of pair-wise interaction energies.  The primary interest
of our work, however, concerns effective $M$-body interactions for
$M\ge3$; we therefore defer the calculation of momentum-dependent
two-body interactions to Appendix \ref{sec:momentum_dependence}.

\subsection{Effective three-body interactions at second order}

At second order in the coupling constants, the expansion for
$H_{\t{int}}^{(2)}$ in \eqref{eq:H_int_1_2} gives rise to effective
three-body interactions.  The virtual state of three-body terms in
$H_{\t{int}}^{(2)}$ cannot have two atoms in excited motional states,
as otherwise the second application of $H_{\t{int}}$ in
$H_{\t{int}}^{(2)}$ would have to address both of theses atoms to
bring them back down to the ground state, resulting in a two-body
process as in the second diagram of \eqref{eq:H_2_expansion}.  All
second-order three-body terms must therefore have only one excited
atom in the virtual state, and take the form
\begin{align}
  \begin{tikzpicture}
    \begin{feynman}
      \vertex (v1);
      \vertex[above left = of v1] (f1) {$\mu r$};
      \vertex[below left = of v1] (f2) {$\nu s$};
      \vertex[right = of v1] (vm);
      \vertex[right = of vm] (v2);
      \vertex[above = of vm] (f3) {$\mu r'$};
      \vertex[below = of vm] (f4) {$\rho t$};
      \vertex[below right = of v2] (f5) {$\rho t'$};
      \vertex[above right = of v2] (f6) {$\nu s''$};
      \diagram* {
        (f1) -- (v1) -- (f3),
        (f2) -- (v1) --  [scalar, edge label=$n\nu s'$] (v2),
        (f4) -- (v2) -- (f5),
        (v2) -- (f6), };
    \end{feynman}
  \end{tikzpicture}
  \propto K_n^2 G^{rs}_{r's'} G^{s't}_{s''t'}
  \c_{\mu r'}^\dag \c_{\nu s''}^\dag \c_{\rho t'}^\dag
  \c_{\rho t} \c_{\nu s} \c_{\mu r}
  \label{eq:H_3_2_diagram}.
\end{align}
Unlike for the two-body diagram in \eqref{eq:H_2}, the explicit
factors of $1/2$ which appear in the bare two-body Hamiltonian
$H_{\t{int}}$ in \eqref{eq:H_int} are now cancelled out by symmetry
factors which account for duplicate diagrams; this cancellation will
generally occur for all connected $M$-body diagrams with $M>2$ (see
Appendix \ref{sec:diagrams}).  The net effective three-body
interaction Hamiltonian at second order is then given by the sum over
all diagrams of the form in \eqref{eq:H_3_2_diagram}, i.e.
\begin{align}
  H_3^{(2)} = - \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[above left = of v1] (f1);
        \vertex[below left = of v1] (f2);
        \vertex[right = 4em of v1] (v2);
        \vertex[above right = of v1] (f3);
        \vertex[below left = of v2] (f4);
        \vertex[below right = of v2] (f5);
        \vertex[above right = of v2] (f6);
        \diagram* {
          (f1) -- (v1) -- (f3),
          (f2) -- (v1) --[scalar] (v2),
          (f4) -- (v2) -- (f5),
          (v2) -- (f6), };
      \end{feynman}
    \end{tikzpicture}}
  = -\alpha_3^{(2)} \sum_{\abs{\set{\mu,\nu,\rho}}=3}
  G^{rs}_{r's'} G^{s't}_{s''t'}
  \c_{\mu r'}^\dag \c_{\nu s''}^\dag \c_{\rho t'}^\dag
  \c_{\rho t} \c_{\nu s} \c_{\mu r},
  \label{eq:H_3_2}
\end{align}
where $\alpha_3^{(2)} \equiv \sum_{n>0} K_n^2/E_n$, and the preceding
minus sign is as prescribed by $H_{\t{int}}^{(2)}$ in
\eqref{eq:H_int_1_2}.

\subsection{Effective three-body interactions at third order}

The third-order effective interaction Hamiltonian $H_{\t{int}}^{(3)}$
in \eqref{eq:H_int_3} contains both three- and four-body terms.  To
compactly enumerate and evaluate all three-body diagrams at third
order, we introduce an expanded coupling symbol
\begin{align}
  G^{\mu q;\nu r}_{\rho s;\sigma t} \equiv \left\{
    \begin{array}{ll}
      G^{qr}_{st} & ~ \p{\mu,\nu} = \p{\rho,\sigma}
      \\
      - G^{qr}_{ts} & ~ \p{\mu,\nu} = \p{\sigma,\rho}
      \\
      0 & ~ \t{otherwise}
    \end{array}\right.
  \label{eq:G_expanded}
\end{align}
for more general
$\p{\mu,q}+\p{\nu,r}\leftrightarrow\p{\rho,s}+\p{\sigma,t}$ coupling
induced by terms proportional to
$\c_{\mu q}^\dag \c_{\nu r}^\dag \c_{\sigma t} \c_{\rho s}$.  The
minus sign in \eqref{eq:G_expanded} accounts for fermionic statistics:
if $\p{\mu,\nu}=\p{\sigma,\rho}$, then we are considering a term of
the form
\begin{align}
  G^{\mu q;\nu r}_{\nu s;\mu t}
  \c_{\nu s}^\dag \c_{\mu t}^\dag \c_{\nu r} \c_{\mu q}
  = -G^{qr}_{ts} \c_{\nu s}^\dag \c_{\mu t}^\dag \c_{\mu r} \c_{\nu q}
  = G^{qr}_{ts} \c_{\mu t}^\dag \c_{\nu s}^\dag \c_{\nu r} \c_{\mu q}.
\end{align}
At the cost of introducing an additional sum over new nuclear spin
indices, the expanded coupling symbol allows us to collect together
diagrams which have the same graph topology, but represent different
matrix elements of the effective Hamiltonian due to the exchange of
nuclear spins at a vertex.  The third order three-body diagrams in
$H_{\t{int}}^{(3)}$ are then
\begin{align}
  \begin{tikzpicture}
    \begin{feynman}
      \vertex (v1);
      \vertex[below right = of v1] (v2);
      \vertex[above right = of v2] (v3);
      \vertex[above left = of v1] (f1) {$\mu r$};
      \vertex[left = of v1] (f2) {$\nu s$};
      \vertex[below left = of v2] (f3) {$\rho t$};
      \vertex[above right = of v3] (f4) {$\mu r''$};
      \vertex[right = of v3] (f5) {$\nu's'''$};
      \vertex[below right = of v2] (f6) {$\rho't'$};
      \diagram* {
        (f1) -- (v1) --[scalar, edge label=$\ell\mu r'$] (v3) -- (f4),
        (f2) -- (v1)
        --[scalar, edge label'=$m\nu s'$] (v2)
        --[scalar, edge label'=$n\nu's''$] (v3)
        -- (f5),
        (f3) -- (v2) -- (f6), };
    \end{feynman}
  \end{tikzpicture}
  \propto K_{\ell m} K^m_n K_{\ell n}
  G^{rs}_{r's'} G^{\nu s';\rho t}_{\nu's'';\rho't'} G^{r's''}_{r''s'''}
  \c_{\mu r''}^\dag \c_{\nu's'''}^\dag \c_{\rho't'}^\dag
  \c_{\rho t} \c_{\nu s} \c_{\mu r},
  \label{eq:H_3_3_S}
\end{align}
\begin{align}
  \begin{tikzpicture}
    \begin{feynman}
      \vertex (v1);
      \vertex[right = of v1] (v2);
      \vertex[below right = of v2] (v3);
      \vertex[above left = of v1] (f1) {$\mu r$};
      \vertex[below left = of v1] (f2) {$\nu s$};
      \vertex[below left = of v3] (f3) {$\rho t$};
      \vertex[above right = of v2] (f4) {$\mu r''$};
      \vertex[above right = of v3] (f5) {$\nu s'''$};
      \vertex[below right = of v3] (f6) {$\rho t'$};
      \diagram* {
        (f1) -- (v1)
        --[half left, scalar, edge label=$\ell\mu r'$] (v2)
        -- (f4),
        (f2) -- (v1)
        --[half right, scalar, edge label'=$m\nu s'$] (v2)
        --[scalar, edge label=$n\nu s''$] (v3)
        -- (f5),
        (f3) -- (v3) -- (f6), };
    \end{feynman}
  \end{tikzpicture}
  \propto K_{\ell m} K^{\ell m}_n K_n
  G^{rs}_{r's'} G^{r's'}_{r''s''} G^{s''t}_{s'''t'}
  \c_{\mu r''}^\dag \c_{\nu s'''}^\dag \c_{\rho t'}^\dag
  \c_{\rho t} \c_{\nu s} \c_{\mu r},
  \label{eq:H_3_3_O}
\end{align}
and the mirror image of \eqref{eq:H_3_3_O}.  As prescribed by
$H_{\t{int}}^{(3)}$ in \eqref{eq:H_int_3}, these diagrams have an
associated minus sign if they contain only one excited virtual state,
and a factor of $1/2$ if they contain a virtual ground state.
Remembering that counter-terms are $\O\p{G^2}$, there are additionally
two third-order three-body diagrams in $H_{\t{int}}^{(2)}$, namely
\begin{align}
  \begin{tikzpicture}
    \begin{feynman}
      \vertex[crossed dot] (v1) {};
      \vertex[above left = 4em of v1] (f1) {$\mu r$};
      \vertex[below left = 4em of v1] (f2) {$\nu s$};
      \vertex[right = 4em of v1] (v2);
      \vertex[above right = 4em of v1] (f3) {$\mu r'$};
      \vertex[below left = of v2] (f4) {$\rho t$};
      \vertex[below right = of v2] (f5) {$\rho t'$};
      \vertex[above right = of v2] (f6) {$\nu s''$};
      \diagram* {
        (f1) -- (v1) -- (f3),
        (f2) -- (v1)
        --[scalar, edge label=$n\nu s'$] (v2),
        (f4) -- (v2) -- (f5),
        (v2) -- (f6), };
    \end{feynman}
  \end{tikzpicture}
  \propto K_n^2 \tilde G^{rs}_{r's'} G^{s't}_{s''t'}
  \c_{\mu r'}^\dag \c_{\nu s''}^\dag \c_{\rho t'}^\dag
  \c_{\rho t} \c_{\nu s} \c_{\mu r}
  \label{eq:H_3_3_counter_term}
\end{align}
and its mirror image, where $\tilde G^{qr}_{st}$ is equal to the
counter-term associated with $G^{qr}_{st}$.  To leading order in the
coupling constants, the renormalization condition in
\eqref{eq:counter_term} implies that
\begin{align}
  \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex[crossed dot] (v) {};
        \vertex[above left = of v] (f1);
        \vertex[below left = of v] (f2);
        \vertex[above right = of v] (f3);
        \vertex[below right = of v] (f4);
        \diagram* {
          (f1) -- (v),
          (f2) -- (v),
          (v) -- (f3),
          (v) -- (f4), };
      \end{feynman}
    \end{tikzpicture}}
  = \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[right = of v1] (v2);
        \vertex[above left = of v1] (f1);
        \vertex[below left = of v1] (f2);
        \vertex[above right = of v2] (f3);
        \vertex[below right = of v2] (f4);
        \diagram* {
          (f1) -- (v1) --[half left, scalar] (v2) -- (f3),
          (f2) -- (v1) --[half right, scalar] (v2) -- (f4), };
      \end{feynman}
    \end{tikzpicture}},
\end{align}
which means that the sum over all counter-term diagrams in
\eqref{eq:H_3_3_counter_term} is
\begin{align}
  \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex[crossed dot] (v1) {};
        \vertex[right = 4em of v1] (v2);
        \vertex[above left = of v1] (f1);
        \vertex[below left = of v1] (f2);
        \vertex[above right = of v1] (f3);
        \vertex[below left = of v2] (f4);
        \vertex[below right = of v2] (f5);
        \vertex[above right = of v2] (f6);
        \diagram* {
          (f1) -- (v1) -- (f3),
          (f2) -- (v1) --[scalar] (v2),
          (f4) -- (v2) -- (f5),
          (v2) -- (f6), };
      \end{feynman}
    \end{tikzpicture}}
  = \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[right = of v1] (v2);
        \vertex[below right = of v2] (v3);
        \vertex[above left = of v1] (f1);
        \vertex[below left = of v1] (f2);
        \vertex[below left = of v3] (f3);
        \vertex[above right = of v2] (f4);
        \vertex[above right = of v3] (f5);
        \vertex[below right = of v3] (f6);
        \diagram* {
          (f1) -- (v1) --[half left, scalar] (v2) -- (f4),
          (f2) -- (v1) --[half right, scalar] (v2)
          --[scalar] (v3) -- (f5),
          (f3) -- (v3) -- (f6), };
      \end{feynman}
    \end{tikzpicture}}.
  \label{eq:H_3_3_counter_term_sum}
\end{align}
The net contribution of the these counter-term diagrams together with
the two-particle-loop diagrams in \eqref{eq:H_3_3_O} is therefore
\begin{multline}
  - \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex[crossed dot] (v1) {};
        \vertex[right = 4em of v1] (v2);
        \vertex[above left = of v1] (f1);
        \vertex[below left = of v1] (f2);
        \vertex[above right = of v1] (f3);
        \vertex[below left = of v2] (f4);
        \vertex[below right = of v2] (f5);
        \vertex[above right = of v2] (f6);
        \diagram* {
          (f1) -- (v1) -- (f3),
          (f2) -- (v1) --[scalar] (v2),
          (f4) -- (v2) -- (f5),
          (v2) -- (f6), };
      \end{feynman}
    \end{tikzpicture}}
  + \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[right = of v1] (v2);
        \vertex[below right = of v2] (v3);
        \vertex[above left = of v1] (f1);
        \vertex[below left = of v1] (f2);
        \vertex[below left = of v3] (f3);
        \vertex[above right = of v2] (f4);
        \vertex[above right = of v3] (f5);
        \vertex[below right = of v3] (f6);
        \diagram* {
          (f1) -- (v1) --[half left, scalar] (v2) -- (f4),
          (f2) -- (v1) --[half right, scalar] (v2)
          --[scalar] (v3) -- (f5),
          (f3) -- (v3) -- (f6), };
      \end{feynman}
    \end{tikzpicture}}
  - \f12 \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[right = of v1] (v2);
        \vertex[below right = of v2] (v3);
        \vertex[above left = of v1] (f1);
        \vertex[below left = of v1] (f2);
        \vertex[below left = of v3] (f3);
        \vertex[above right = of v2] (f4);
        \vertex[above right = of v3] (f5);
        \vertex[below right = of v3] (f6);
        \diagram* {
          (f1) -- (v1) --[half left, scalar] (v2) -- (f4),
          (f2) -- (v1) --[half right, scalar] (v2)
          -- (v3) -- (f5),
          (f3) -- (v3) -- (f6), };
      \end{feynman}
    \end{tikzpicture}}
  - \f12 \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[right = of v1] (v2);
        \vertex[below right = of v2] (v3);
        \vertex[above left = of v1] (f1);
        \vertex[below left = of v1] (f2);
        \vertex[below left = of v3] (f3);
        \vertex[above right = of v2] (f4);
        \vertex[above right = of v3] (f5);
        \vertex[below right = of v3] (f6);
        \diagram* {
          (f1) -- (v1) --[half left] (v2) -- (f4),
          (f2) -- (v1) --[half right] (v2)
          --[scalar] (v3) -- (f5),
          (f3) -- (v3) -- (f6), };
      \end{feynman}
    \end{tikzpicture}} \\
  = - \f12 \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[right = of v1] (v2);
        \vertex[below right = of v2] (v3);
        \vertex[above left = of v1] (f1);
        \vertex[below left = of v1] (f2);
        \vertex[below left = of v3] (f3);
        \vertex[above right = of v2] (f4);
        \vertex[above right = of v3] (f5);
        \vertex[below right = of v3] (f6);
        \diagram* {
          (f1) -- (v1) --[half left, scalar] (v2) -- (f4),
          (f2) -- (v1) --[half right, scalar] (v2)
          -- (v3) -- (f5),
          (f3) -- (v3) -- (f6), };
      \end{feynman}
    \end{tikzpicture}}
  - \f12 \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[right = of v1] (v2);
        \vertex[below right = of v2] (v3);
        \vertex[above left = of v1] (f1);
        \vertex[below left = of v1] (f2);
        \vertex[below left = of v3] (f3);
        \vertex[above right = of v2] (f4);
        \vertex[above right = of v3] (f5);
        \vertex[below right = of v3] (f6);
        \diagram* {
          (f1) -- (v1) --[half left] (v2) -- (f4),
          (f2) -- (v1) --[half right] (v2)
          --[scalar] (v3) -- (f5),
          (f3) -- (v3) -- (f6), };
      \end{feynman}
    \end{tikzpicture}}.
  \label{eq:H_3_3_cancellation}
\end{multline}
An identical cancellation occurs for the mirror images of the diagrams
in \eqref{eq:H_3_3_O} and \eqref{eq:H_3_3_counter_term}, which results
in a contribution to the effective Hamiltonian that is equal to
\eqref{eq:H_3_3_cancellation}.  The contribution from the
three-particle-loop diagrams in \eqref{eq:H_3_3_S} is
\begin{multline}
  \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[below right = of v1] (v2);
        \vertex[above right = of v2] (v3);
        \vertex[above left = of v1] (f1);
        \vertex[left = of v1] (f2);
        \vertex[below left = of v2] (f3);
        \vertex[above right = of v3] (f4);
        \vertex[right = of v3] (f5);
        \vertex[below right = of v2] (f6);
        \diagram* {
          (f1) -- (v1) --[scalar] (v3) -- (f4),
          (f2) -- (v1) --[scalar] (v2) --[scalar] (v3) -- (f5),
          (f3) -- (v2) -- (f6), };
      \end{feynman}
    \end{tikzpicture}}
  - \f12 \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[below right = of v1] (v2);
        \vertex[above right = of v2] (v3);
        \vertex[above left = of v1] (f1);
        \vertex[left = of v1] (f2);
        \vertex[below left = of v2] (f3);
        \vertex[above right = of v3] (f4);
        \vertex[right = of v3] (f5);
        \vertex[below right = of v2] (f6);
        \diagram* {
          (f1) -- (v1) -- (v3) -- (f4),
          (f2) -- (v1) --[scalar] (v2) -- (v3) -- (f5),
          (f3) -- (v2) -- (f6), };
      \end{feynman}
    \end{tikzpicture}}
  - \f12 \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[below right = of v1] (v2);
        \vertex[above right = of v2] (v3);
        \vertex[above left = of v1] (f1);
        \vertex[left = of v1] (f2);
        \vertex[below left = of v2] (f3);
        \vertex[above right = of v3] (f4);
        \vertex[right = of v3] (f5);
        \vertex[below right = of v2] (f6);
        \diagram* {
          (f1) -- (v1) -- (v3) -- (f4),
          (f2) -- (v1) -- (v2) --[scalar] (v3) -- (f5),
          (f3) -- (v2) -- (f6), };
      \end{feynman}
    \end{tikzpicture}} \\
  = \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[below right = of v1] (v2);
        \vertex[above right = of v2] (v3);
        \vertex[above left = of v1] (f1);
        \vertex[left = of v1] (f2);
        \vertex[below left = of v2] (f3);
        \vertex[above right = of v3] (f4);
        \vertex[right = of v3] (f5);
        \vertex[below right = of v2] (f6);
        \diagram* {
          (f1) -- (v1) --[scalar] (v3) -- (f4),
          (f2) -- (v1) --[scalar] (v2) --[scalar] (v3) -- (f5),
          (f3) -- (v2) -- (f6), };
      \end{feynman}
    \end{tikzpicture}}
  - \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[below right = of v1] (v2);
        \vertex[above right = of v2] (v3);
        \vertex[above left = of v1] (f1);
        \vertex[left = of v1] (f2);
        \vertex[below left = of v2] (f3);
        \vertex[above right = of v3] (f4);
        \vertex[right = of v3] (f5);
        \vertex[below right = of v2] (f6);
        \diagram* {
          (f1) -- (v1) -- (v3) -- (f4),
          (f2) -- (v1) --[scalar] (v2) -- (v3) -- (f5),
          (f3) -- (v2) -- (f6), };
      \end{feynman}
    \end{tikzpicture}}.
\end{multline}
The third-order three-body interaction Hamiltonian is therefore
\begin{align}
  H_3^{(3)} &= \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[below right = of v1] (v2);
        \vertex[above right = of v2] (v3);
        \vertex[above left = of v1] (f1);
        \vertex[left = of v1] (f2);
        \vertex[below left = of v2] (f3);
        \vertex[above right = of v3] (f4);
        \vertex[right = of v3] (f5);
        \vertex[below right = of v2] (f6);
        \diagram* {
          (f1) -- (v1) --[scalar] (v3) -- (f4),
          (f2) -- (v1) --[scalar] (v2) --[scalar] (v3) -- (f5),
          (f3) -- (v2) -- (f6), };
      \end{feynman}
    \end{tikzpicture}}
  - \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[below right = of v1] (v2);
        \vertex[above right = of v2] (v3);
        \vertex[above left = of v1] (f1);
        \vertex[left = of v1] (f2);
        \vertex[below left = of v2] (f3);
        \vertex[above right = of v3] (f4);
        \vertex[right = of v3] (f5);
        \vertex[below right = of v2] (f6);
        \diagram* {
          (f1) -- (v1) -- (v3) -- (f4),
          (f2) -- (v1) --[scalar] (v2) -- (v3) -- (f5),
          (f3) -- (v2) -- (f6), };
      \end{feynman}
    \end{tikzpicture}}
   - \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[right = of v1] (v2);
        \vertex[below right = of v2] (v3);
        \vertex[above left = of v1] (f1);
        \vertex[below left = of v1] (f2);
        \vertex[below left = of v3] (f3);
        \vertex[above right = of v2] (f4);
        \vertex[above right = of v3] (f5);
        \vertex[below right = of v3] (f6);
        \diagram* {
          (f1) -- (v1) --[half left, scalar] (v2) -- (f4),
          (f2) -- (v1) --[half right, scalar] (v2) -- (v3) -- (f5),
          (f3) -- (v3) -- (f6), };
      \end{feynman}
    \end{tikzpicture}}
  - \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[right = of v1] (v2);
        \vertex[below right = of v2] (v3);
        \vertex[above left = of v1] (f1);
        \vertex[below left = of v1] (f2);
        \vertex[below left = of v3] (f3);
        \vertex[above right = of v2] (f4);
        \vertex[above right = of v3] (f5);
        \vertex[below right = of v3] (f6);
        \diagram* {
          (f1) -- (v1) --[half left] (v2) -- (f4),
          (f2) -- (v1) --[half right] (v2) --[scalar] (v3) -- (f5),
          (f3) -- (v3) -- (f6), };
      \end{feynman}
    \end{tikzpicture}} \nonumber \\[1em]
  &= \p{\alpha_3^{(3)} - \alpha_5^{(3)}} \H_{3,1}^{(3)}
  - \p{\alpha_{4,3}^{(3)} + \alpha_5^{(3)}} \H_{3,2}^{(3)},
  \label{eq:H_3_3}
\end{align}
where, letting $E_{mn}\equiv E_m+E_n$,
\begin{align}
  \alpha_3^{(3)} \equiv \sum_{\substack{\ell+m>0\\\ell+n>0}}
  \f{K_{\ell m} K^m_n K_{\ell n}}{E_{\ell m} E_{\ell n}},
  &&
  \alpha_{4,3}^{(3)}
  \equiv K \sum_{m+n>0} \f{K_{mn}^2}{E_{mn}^2},
  &&
  \alpha_5^{(3)}
  \equiv  K \sum_{n>0} \f{K_n^2}{E_n^2},
\end{align}
\begin{align}
  \H_{3,1}^{(3)} \equiv \sum_{\abs{\set{\mu,\nu,\rho}}=3}
  G^{rs}_{r's'} G^{\nu s'\rho t}_{\nu's''\rho't'} G^{r's''}_{r''s'''}
  \c_{\mu r''}^\dag \c_{\nu's'''}^\dag \c_{\rho't'}^\dag
  \c_{\rho t} \c_{\nu s} \c_{\mu r}
\end{align}
\begin{align}
  \H_{3,2}^{(3)} \equiv \sum_{\abs{\set{\mu,\nu,\rho}}=3}
  G^{rs}_{r's'} G^{r's'}_{r''s''} G^{s''t}_{s'''t'}
  \c_{\mu r''}^\dag \c_{\nu s'''}^\dag \c_{\rho t'}^\dag
  \c_{\rho t} \c_{\nu s} \c_{\mu r}.
\end{align}

\subsection{Effective four-body interactions at third order}

At third order in the coupling constants, we have four-body terms of
the form
\begin{align}
  \begin{tikzpicture}
    \begin{feynman}
      \vertex[label=0:$~m\mu q'$] (v1);
      \vertex[below right = of v1, label=0:$~n\mu'q''$] (v2);
      \vertex[below right = of v2] (v3);
      \vertex[above left = of v1] (f1) {$\mu q$};
      \vertex[below left = of v1] (f2) {$\nu r$};
      \vertex[below left = of v2] (f3) {$\rho s$};
      \vertex[below left = of v3] (f4) {$\sigma t$};
      \vertex[above right = of v1] (f5) {$\nu r'$};
      \vertex[above right = of v2] (f6) {$\rho's'$};
      \vertex[above right = of v3] (f7) {$\sigma t'$};
      \vertex[below right = of v3] (f8) {$\mu'q'''$};
      \diagram* {
        (f1) -- (v1) -- (f5),
        (f2) -- (v1) -- (f5),
        (v1) --[scalar] (v2) --[scalar] (v3) -- (f8),
        (f3) -- (v2) -- (f6),
        (f4) -- (v3) -- (f7),
      };
    \end{feynman}
  \end{tikzpicture}
  \propto K_m K^m_n K_n
  G^{qr}_{q'r'} G^{\mu q'\rho s}_{\mu'q''\rho's'} G^{q''t}_{q'''t'}
  \c_{\mu'q'''}^\dag \c_{\nu r'}^\dag \c_{\rho's'}^\dag \c_{\sigma t'}^\dag
  \c_{\sigma t} \c_{\rho s} \c_{\nu r} \c_{\mu q},
  \label{eq:H_4_3_C}
\end{align}
\begin{align}
  \begin{tikzpicture}
    \begin{feynman}
      \vertex (v1);
      \vertex[above right = 1.5em of v1] (v2);
      \vertex[below right = 2.5em of v1] (v3);
      \vertex[right = 0.5em of v1] (label)
      {$\let\scriptstyle\textstyle\substack{m\mu q'\\n\nu r'}$};
      \vertex[above left = of v1] (f1) {$\mu q$};
      \vertex[below left = of v1] (f2) {$\nu r$};
      \vertex[above left = of v2] (f3) {$\rho s$};
      \vertex[below left = of v3] (f4) {$\sigma t$};
      \vertex[above right = of v2] (f5) {$\rho s'$};
      \vertex[right = of v2] (f6) {$\mu q''$};
      \vertex[right = of v3] (f7) {$\nu r''$};
      \vertex[below right = of v3] (f8) {$\sigma t'$};
      \diagram* {
        (f1) -- (v1) --[scalar] (v2) -- (f6),
        (f2) -- (v1) --[scalar] (v3) -- (f7),
        (f3) -- (v2) -- (f5),
        (f4) -- (v3) -- (f8),
      };
    \end{feynman}
  \end{tikzpicture}
  \propto K_{mn} K_m K_n G^{qr}_{q'r'} G^{q's}_{q''s'} G^{r't}_{r''t'}
  \c_{\mu q''}^\dag \c_{\nu r''}^\dag \c_{\rho s'}^\dag \c_{\sigma t'}^\dag
  \c_{\sigma t} \c_{\rho s} \c_{\nu r} \c_{\mu q},
  \label{eq:H_4_3_B}
\end{align}
and the mirror image of \eqref{eq:H_4_3_B}.  As we are computing the
leading-order contribution to effective four-body interactions, there
are no counter-terms contributions.  In principle, there is now also
the possibility to make the disconnected diagrams of the form
\begin{align}
  \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[above left = of v1] (f1);
        \vertex[below left = of v1] (f2);
        \vertex[right = of v1] (v2);
        \vertex[above right = of v2] (f3);
        \vertex[below right = of v2] (f4);
        \vertex[below = 1em of f2] (f5);
        \vertex[below = of f5] (f6);
        \vertex[below = 1em of f4] (f7);
        \vertex[below = of f7] (f8);
        \diagram* {
          (f1) -- (v1),
          (f2) -- (v1),
          (v2) -- (f3),
          (v2) -- (f4),
          (v1) --[scalar, half left] (v2),
          (v1) --[scalar, half right] (v2),
          (f5) -- (f8),
          (f6) -- (f7) };
      \end{feynman}
    \end{tikzpicture}},
  &&
  \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[above left = of v1] (f1);
        \vertex[below left = of v1] (f2);
        \vertex[right = of v1] (v2);
        \vertex[above right = of v2] (f3);
        \vertex[below right = of v2] (f4);
        \vertex[left = of f2] (a1);
        \vertex[left = of f4] (a2);
        \vertex[below = 1em of a1] (f5);
        \vertex[below = of f5] (f6);
        \vertex[below = 1em of a2] (f7);
        \vertex[below = of f7] (f8);
        \diagram* {
          (f1) -- (v1),
          (f2) -- (v1),
          (v2) -- (f3),
          (v2) -- (f4),
          (v1) --[scalar, half left] (v2),
          (v1) --[scalar, half right] (v2),
          (f5) -- (f8),
          (f6) -- (f7) };
      \end{feynman}
    \end{tikzpicture}},
  &&
  \t{and}
  &&
  \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[above left = of v1] (f1);
        \vertex[below left = of v1] (f2);
        \vertex[right = of v1] (v2);
        \vertex[above right = of v2] (f3);
        \vertex[below right = of v2] (f4);
        \vertex[right = of f2] (a1);
        \vertex[right = of f4] (a2);
        \vertex[below = 1em of a1] (f5);
        \vertex[below = of f5] (f6);
        \vertex[below = 1em of a2] (f7);
        \vertex[below = of f7] (f8);
        \diagram* {
          (f1) -- (v1),
          (f2) -- (v1),
          (v2) -- (f3),
          (v2) -- (f4),
          (v1) --[scalar, half left] (v2),
          (v1) --[scalar, half right] (v2),
          (f5) -- (f8),
          (f6) -- (f7) };
      \end{feynman}
    \end{tikzpicture}}.
\end{align}
As prescribed by $H_{\t{int}}^{(3)}$ in \eqref{eq:H_int_3}, however,
the second and third of these diagrams pick up a factor of $-1/2$, so
the sum over disconnected diagrams vanishes.

Observing that the diagrams of the form in \eqref{eq:H_4_3_B} are
equal to their mirror image, the positive contributions to the
third-order four-body interaction Hamiltonian are
\begin{align}
  \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[above right = 1.5em of v1] (v2);
        \vertex[below right = 2.5em of v1] (v3);
        \vertex[above left = of v1] (f1);
        \vertex[below left = of v1] (f2);
        \vertex[above left = of v2] (f3);
        \vertex[below left = of v3] (f4);
        \vertex[above right = of v2] (f5);
        \vertex[right = of v2] (f6);
        \vertex[right = of v3] (f7);
        \vertex[below right = of v3] (f8);
        \diagram* {
          (f1) -- (v1) --[scalar] (v2) -- (f6),
          (f2) -- (v1) --[scalar] (v3) -- (f7),
          (f3) -- (v2) -- (f5),
          (f4) -- (v3) -- (f8),
        };
      \end{feynman}
    \end{tikzpicture}}
  + \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[below right = 2.5em of v1] (v3);
        \vertex[below left = 1.7em of v3] (v2);
        \vertex[above left = of v1] (f1);
        \vertex[left = of v1] (f2);
        \vertex[left = of v2] (f3);
        \vertex[below left = of v2] (f4);
        \vertex[above right = of v1] (f5);
        \vertex[above right = of v3] (f6);
        \vertex[below right = of v3] (f7);
        \vertex[below right = of v2] (f8);
        \diagram* {
          (f1) -- (v1) -- (f5),
          (f2) -- (v1) --[scalar] (v3) -- (f6),
          (f3) -- (v2) --[scalar] (v3) -- (f7),
          (f4) -- (v2) -- (f8), };
      \end{feynman}
    \end{tikzpicture}}
  + \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[below right = of v1] (v2);
        \vertex[below right = of v2] (v3);
        \vertex[above left = of v1] (f1);
        \vertex[below left = of v1] (f2);
        \vertex[below left = of v2] (f3);
        \vertex[below left = of v3] (f4);
        \vertex[above right = of v1] (f5);
        \vertex[above right = of v2] (f6);
        \vertex[above right = of v3] (f7);
        \vertex[below right = of v3] (f8);
        \diagram* {
          (f1) -- (v1) -- (f5),
          (f2) -- (v1) -- (f5),
          (v1) --[scalar] (v2) --[scalar] (v3) -- (f8),
          (f3) -- (v2) -- (f6),
          (f4) -- (v3) -- (f7),
        };
      \end{feynman}
    \end{tikzpicture}}
  = 2 \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[above right = 1.5em of v1] (v2);
        \vertex[below right = 2.5em of v1] (v3);
        \vertex[above left = of v1] (f1);
        \vertex[below left = of v1] (f2);
        \vertex[above left = of v2] (f3);
        \vertex[below left = of v3] (f4);
        \vertex[above right = of v2] (f5);
        \vertex[right = of v2] (f6);
        \vertex[right = of v3] (f7);
        \vertex[below right = of v3] (f8);
        \diagram* {
          (f1) -- (v1) --[scalar] (v2) -- (f6),
          (f2) -- (v1) --[scalar] (v3) -- (f7),
          (f3) -- (v2) -- (f5),
          (f4) -- (v3) -- (f8),
        };
      \end{feynman}
    \end{tikzpicture}}
  + \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[below right = of v1] (v2);
        \vertex[below right = of v2] (v3);
        \vertex[above left = of v1] (f1);
        \vertex[below left = of v1] (f2);
        \vertex[below left = of v2] (f3);
        \vertex[below left = of v3] (f4);
        \vertex[above right = of v1] (f5);
        \vertex[above right = of v2] (f6);
        \vertex[above right = of v3] (f7);
        \vertex[below right = of v3] (f8);
        \diagram* {
          (f1) -- (v1) -- (f5),
          (f2) -- (v1) -- (f5),
          (v1) --[scalar] (v2) --[scalar] (v3) -- (f8),
          (f3) -- (v2) -- (f6),
          (f4) -- (v3) -- (f7),
        };
      \end{feynman}
    \end{tikzpicture}},
\end{align}
while the negative contributions are
\begin{multline}
  \f12 \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[above right = 1.5em of v1] (v2);
        \vertex[below right = 2.5em of v1] (v3);
        \vertex[above left = of v1] (f1);
        \vertex[below left = of v1] (f2);
        \vertex[above left = of v2] (f3);
        \vertex[below left = of v3] (f4);
        \vertex[above right = of v2] (f5);
        \vertex[right = of v2] (f6);
        \vertex[right = of v3] (f7);
        \vertex[below right = of v3] (f8);
        \diagram* {
          (f1) -- (v1) --[scalar] (v2) -- (f6),
          (f2) -- (v1) -- (v3) -- (f7),
          (f3) -- (v2) -- (f5),
          (f4) -- (v3) -- (f8),
        };
      \end{feynman}
    \end{tikzpicture}}
  + \f12 \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[below right = 2.5em of v1] (v3);
        \vertex[below left = 1.7em of v3] (v2);
        \vertex[above left = of v1] (f1);
        \vertex[left = of v1] (f2);
        \vertex[left = of v2] (f3);
        \vertex[below left = of v2] (f4);
        \vertex[above right = of v1] (f5);
        \vertex[above right = of v3] (f6);
        \vertex[below right = of v3] (f7);
        \vertex[below right = of v2] (f8);
        \diagram* {
          (f1) -- (v1) -- (f5),
          (f2) -- (v1) -- (v3) -- (f6),
          (f3) -- (v2) --[scalar] (v3) -- (f7),
          (f4) -- (v2) -- (f8), };
      \end{feynman}
    \end{tikzpicture}}
  + \f12 \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[below right = of v1] (v2);
        \vertex[below right = of v2] (v3);
        \vertex[above left = of v1] (f1);
        \vertex[below left = of v1] (f2);
        \vertex[below left = of v2] (f3);
        \vertex[below left = of v3] (f4);
        \vertex[above right = of v1] (f5);
        \vertex[above right = of v2] (f6);
        \vertex[above right = of v3] (f7);
        \vertex[below right = of v3] (f8);
        \diagram* {
          (f1) -- (v1) -- (f5),
          (f2) -- (v1) -- (f5),
          (v1) --[scalar] (v2) -- (v3) -- (f8),
          (f3) -- (v2) -- (f6),
          (f4) -- (v3) -- (f7),
        };
      \end{feynman}
    \end{tikzpicture}}
  + \f12 \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[below right = of v1] (v2);
        \vertex[below right = of v2] (v3);
        \vertex[above left = of v1] (f1);
        \vertex[below left = of v1] (f2);
        \vertex[below left = of v2] (f3);
        \vertex[below left = of v3] (f4);
        \vertex[above right = of v1] (f5);
        \vertex[above right = of v2] (f6);
        \vertex[above right = of v3] (f7);
        \vertex[below right = of v3] (f8);
        \diagram* {
          (f1) -- (v1) -- (f5),
          (f2) -- (v1) -- (f5),
          (v1) -- (v2) --[scalar] (v3) -- (f8),
          (f3) -- (v2) -- (f6),
          (f4) -- (v3) -- (f7),
        };
      \end{feynman}
    \end{tikzpicture}} \\
  = \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[above right = 1.5em of v1] (v2);
        \vertex[below right = 2.5em of v1] (v3);
        \vertex[above left = of v1] (f1);
        \vertex[below left = of v1] (f2);
        \vertex[above left = of v2] (f3);
        \vertex[below left = of v3] (f4);
        \vertex[above right = of v2] (f5);
        \vertex[right = of v2] (f6);
        \vertex[right = of v3] (f7);
        \vertex[below right = of v3] (f8);
        \diagram* {
          (f1) -- (v1) --[scalar] (v2) -- (f6),
          (f2) -- (v1) -- (v3) -- (f7),
          (f3) -- (v2) -- (f5),
          (f4) -- (v3) -- (f8),
        };
      \end{feynman}
    \end{tikzpicture}}
  + \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[below right = of v1] (v2);
        \vertex[below right = of v2] (v3);
        \vertex[above left = of v1] (f1);
        \vertex[below left = of v1] (f2);
        \vertex[below left = of v2] (f3);
        \vertex[below left = of v3] (f4);
        \vertex[above right = of v1] (f5);
        \vertex[above right = of v2] (f6);
        \vertex[above right = of v3] (f7);
        \vertex[below right = of v3] (f8);
        \diagram* {
          (f1) -- (v1) -- (f5),
          (f2) -- (v1) -- (f5),
          (v1) --[scalar] (v2) -- (v3) -- (f8),
          (f3) -- (v2) -- (f6),
          (f4) -- (v3) -- (f7),
        };
      \end{feynman}
    \end{tikzpicture}},
\end{multline}
such that the net third-order four-body interaction Hamiltonian is
\begin{align}
  H_4^{(3)} &= 2 \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[above right = 1.5em of v1] (v2);
        \vertex[below right = 2.5em of v1] (v3);
        \vertex[above left = of v1] (f1);
        \vertex[below left = of v1] (f2);
        \vertex[above left = of v2] (f3);
        \vertex[below left = of v3] (f4);
        \vertex[above right = of v2] (f5);
        \vertex[right = of v2] (f6);
        \vertex[right = of v3] (f7);
        \vertex[below right = of v3] (f8);
        \diagram* {
          (f1) -- (v1) --[scalar] (v2) -- (f6),
          (f2) -- (v1) --[scalar] (v3) -- (f7),
          (f3) -- (v2) -- (f5),
          (f4) -- (v3) -- (f8),
        };
      \end{feynman}
    \end{tikzpicture}}
  - \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[above right = 1.5em of v1] (v2);
        \vertex[below right = 2.5em of v1] (v3);
        \vertex[above left = of v1] (f1);
        \vertex[below left = of v1] (f2);
        \vertex[above left = of v2] (f3);
        \vertex[below left = of v3] (f4);
        \vertex[above right = of v2] (f5);
        \vertex[right = of v2] (f6);
        \vertex[right = of v3] (f7);
        \vertex[below right = of v3] (f8);
        \diagram* {
          (f1) -- (v1) --[scalar] (v2) -- (f6),
          (f2) -- (v1) -- (v3) -- (f7),
          (f3) -- (v2) -- (f5),
          (f4) -- (v3) -- (f8),
        };
      \end{feynman}
    \end{tikzpicture}}
  + \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[below right = of v1] (v2);
        \vertex[below right = of v2] (v3);
        \vertex[above left = of v1] (f1);
        \vertex[below left = of v1] (f2);
        \vertex[below left = of v2] (f3);
        \vertex[below left = of v3] (f4);
        \vertex[above right = of v1] (f5);
        \vertex[above right = of v2] (f6);
        \vertex[above right = of v3] (f7);
        \vertex[below right = of v3] (f8);
        \diagram* {
          (f1) -- (v1) -- (f5),
          (f2) -- (v1) -- (f5),
          (v1) --[scalar] (v2) --[scalar] (v3) -- (f8),
          (f3) -- (v2) -- (f6),
          (f4) -- (v3) -- (f7),
        };
      \end{feynman}
    \end{tikzpicture}}
  - \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[below right = of v1] (v2);
        \vertex[below right = of v2] (v3);
        \vertex[above left = of v1] (f1);
        \vertex[below left = of v1] (f2);
        \vertex[below left = of v2] (f3);
        \vertex[below left = of v3] (f4);
        \vertex[above right = of v1] (f5);
        \vertex[above right = of v2] (f6);
        \vertex[above right = of v3] (f7);
        \vertex[below right = of v3] (f8);
        \diagram* {
          (f1) -- (v1) -- (f5),
          (f2) -- (v1) -- (f5),
          (v1) --[scalar] (v2) -- (v3) -- (f8),
          (f3) -- (v2) -- (f6),
          (f4) -- (v3) -- (f7),
        };
      \end{feynman}
    \end{tikzpicture}} \nonumber \\
  &= \p{2\alpha_{4,1}^{(3)} - \alpha_5^{(3)}} \H_{4,1}^{(3)}
  + \p{\alpha_{4,2}^{(3)} - \alpha_5^{(3)}} \H_{4,2}^{(3)},
  \label{eq:H_4_3}
\end{align}
where
\begin{align}
  \alpha_{4,1}^{(3)}
  \equiv \sum_{\substack{m\ge0\\n>0}} \f{K_{mn} K_m K_n}{E_{mn} E_n},
  &&
  \alpha_{4,2}^{(3)}
  \equiv \sum_{m,n>0} \f{K_m K^m_n K_n}{E_m E_n},
  &&
  \alpha_5^{(3)}
  \equiv  K \sum_{n>0} \f{K_n^2}{E_n^2},
\end{align}
\begin{align}
  \H_{4,1}^{(3)}
  \equiv \sum_{\abs{\set{\mu,\nu,\rho,\sigma}}=4}
  G^{qr}_{q'r'} G^{q's}_{q''s'} G^{r't}_{r''t'}
  \c_{\mu q''}^\dag \c_{\nu r''}^\dag \c_{\rho s'}^\dag \c_{\sigma t'}^\dag
  \c_{\sigma t} \c_{\rho s} \c_{\nu r} \c_{\mu q},
\end{align}
\begin{align}
  \H_{4,2}^{(3)}
  \equiv \sum_{\abs{\set{\mu,\nu,\rho,\sigma}}=4}
  G^{qr}_{q'r'} G^{\mu q'\rho s}_{\mu'q''\rho's'} G^{q''t}_{q'''t'}
  \c_{\mu'q'''}^\dag \c_{\nu r'}^\dag \c_{\rho's'}^\dag \c_{\sigma t'}^\dag
  \c_{\sigma t} \c_{\rho s} \c_{\nu r} \c_{\mu q}.
\end{align}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Low-excitation Hamiltonians, eigenstates, and spectra}
\label{sec:spectra}

Current experiments with ultracold ${}^{87}$Sr on a lattice can
coherently address ground states and their low-lying orbital
excitations for up to five atoms per lattice site
\cite{goban2018emergence} (see Appendix \ref{sec:experiment_summary}
for a summary of relevant experimental procedures).  Due to the
SU($N$) symmetry of inter-atomic interactions, manifest in the fact
that all coupling constants are independent of nuclear spin, we can
write all effective $M$-body Hamiltonians which address these states
in the form
\begin{align}
  H_M = \sum_{\abs{\set{\mu_j}}=M}
  \p{U_{M,\g} \prod_{\alpha=1}^M \n_{\mu_\alpha,\g}
    + U_{M,+} \n_{\mu_1,\e} \prod_{\alpha=2}^M \n_{\mu_\alpha,\g}
    + U_{M,-} \c_{\mu_1,\g}^\dag \c_{\mu_2,\e}^\dag
    \c_{\mu_2,\g} \c_{\mu_1,\e} \prod_{\alpha=3}^M \n_{\mu_\alpha,\g}},
  \label{eq:H_M}
\end{align}
where $\n_{\mu s}\equiv \c_{\mu s}^\dag\c_{\mu s}$ is a number
operator, and the coefficients $U_X$ can be determined from the
coupling constants $G_Y$ and prefactors $\alpha_Z^{(p)}$ of the
effective $M$-body Hamiltonians derived in section
\ref{sec:low_energy} (see Appendix \ref{sec:U_X}).  For a lattice with
$N\ge M$ atoms occupying nuclear spins $\N=\set{\mu_j}$ for
$j=1,2,\cdots,N$, the $M$-body Hamiltonian $H_M$ has a single ground
state $\ket{\N,0}$, and a singly-excited state $\ket{\N,+}$ which is
fully symmetric in the orbital degrees of freedom; these states are
\begin{align}
  \ket{\N,0}
  \equiv \p{\prod_{\mu\in\N} \c_{\mu,\g}^\dag} \ket{\t{vacuum}},
  &&
  \ket{\N,+} \equiv \f1{\sqrt{N}} \sum_{\mu\in\N}
  \c_{\mu,\e}^\dag \c_{\mu,\g} \ket{\N,0}.
  \label{eq:states_S}
\end{align}
As these states are fully symmetric in their orbital degrees of
freedom, they are anti-symmetric in their nuclear spin degrees of
freedom, forming an SU($N$) singlet.  Furthermore, the symmetric state
is particularly interesting as its orbital degrees of freedom form an
$N$-body entangled $W$ state, which belongs to a special class of
multi-partite entangled states that are robust against disposal or
loss of particles.  This state thus constitutes an important resource
for many quantum information processing and quantum communication
tasks \cite{zang2015generating}.

In addition to the states in \eqref{eq:states_S}, the multi-body
Hamiltonian $H_M$ in \eqref{eq:H_M} has an $\p{N-1}$-fold degenerate
excited-state eigenspace which is asymmetric in the orbital degrees of
freedom, spanned by the states
\begin{align}
  \ket{\N,-_j} \equiv \f1{\sqrt2}
  \p{\c_{\mu_1,\e}^\dag \c_{\mu_1,\g} - \c_{\mu_j,\e}^\dag \c_{\mu_j,\g}}
  \ket{\N,0}
  \label{eq:states_A}
\end{align}
for $j=2,\cdots,N$.  If $N>2$, the asymmetric states are not separable
in their orbital and nuclear spin degrees of freedom.  An important
feature of the excited states in \eqref{eq:states_S} and
\eqref{eq:states_A} is that they are entirely independent of $M$,
which implies that the effect of multi-body interactions is simply to
modify the many-body atomic energy spectra without affecting the
energy eigenstates.  The eigenvalues $E_{NX}^{(M)}=\bk{\N X|H_M|\N X}$
of $H_M$ associated with each of the eigenstates in
\eqref{eq:states_S} and \eqref{eq:states_A} are provided in table
\ref{tab:eigen}, both in terms of the coefficients $U_{MX}$ of $H_M$
as expressed in \eqref{eq:H_M} and the $M$-body eigenvalues
$E_{MX}^{(M)}$.  Due to the SU($N$) symmetry of the multi-body
Hamiltonian $H_M$, the eigenvalues $E_{NX}^{(M)}$ depend on the number
of nuclear spins on a lattice site, $N$, but not on the actual nuclear
spins $\mu\in\N$ which are occupied.  The total $N$-body interaction
energies $E_{NX}$ are given in terms of the $M$-body eigenvalues
$E_{NX}^{(M)}$ by $E_{NX}=\sum_ME_{NX}^{(M)}$.

\begin{table}
  \centering
  \caption{\footnotesize Low-excitation eigenvalues of $M$-body
    Hamiltonians $H_M$.  Many-body energy eigenstates are labeled by
    the nuclear spins they occupy (i.e.~$\N$ with $N\equiv\abs{\N}$)
    and whether they are in an orbital ground ($0$), singly-excited
    symmetric ($+$), or singly-excited asymmetric ($-$) state.  The
    corresponding $N$-body eigenvalues $E_{NX}^{(M)}$ of $H_M$ are
    given in terms of the coefficients $U_{MX}$ as appearing in
    \eqref{eq:H_M}, in addition to the $M$-body eigenvalues
    $E_{MX}^{(M)}$.}
  \label{tab:eigen}
  \begin{tabular}{c|c}
    Eigenstate
    & $H_M$ eigenvalue \\ \hline\hline
    $\ket{\N,0}$
    & $M! {N \choose M} U_{M,\g}$ \\ \hline
    $\ket{\N,+}$
    & $M! {N-1 \choose M} U_{M,\g}
      + \p{M-1}! {N-1 \choose M-1} \p{U_{M,+} + U_{M,-}}$ \\ \hline
    $\ket{\N,-}$
    & $M! {N-1 \choose M} U_{M,\g}
      + \p{M-1}! {N-1 \choose M-1} U_{M,+}
      - \p{M-2}! {N-2 \choose M-2} U_{M,-}$ \\ \hline\hline
    $\ket{\N,0}$
    & ${N \choose M} E_{M,0}^{(M)}$ \\ \hline
    $\ket{\N,+}$
    & ${N-1 \choose M} E_{M,0}^{(M)}
      + {N-1 \choose M-1} E_{M,+}^{(M)}$ \\ \hline
    $\ket{\N,-}$
    & ${N-1 \choose M} E_{M,0}^{(M)}
      + M^{-1} \sp{{N-1 \choose M-1} - {N-2 \choose M-2}} E_{M,+}^{(M)}
      + \sp{\p{1 - M^{-1}} {N-1 \choose M-1}
      + M^{-1} {N-2 \choose M-2}} E_{M,-}^{(M)}$
  \end{tabular}
\end{table}

\subsection{Many-body state spectroscopy}

Spectroscopic interrogation is a powerful means to probe the internal
structure and dynamics of a system under examination.  Consequently,
we consider Rabi spectroscopy of the low-lying energy eigenstates in
multiply-occupied lattice sites.  If we interrogate a lattice site by
a laser red-detuned by $\Delta$ from the single-atom orbital state
excitation energy, we realize the Hamiltonian
\begin{align}
  H_{\t{Rabi}}
  = \sum_X E_X \P_X
  + \sum_\mu\p{\Delta T_\mu^\z + \Omega_\mu T_\mu^\x},
  \label{eq:H_rabi}
\end{align}
where $E_X$ is an eigenvalue of the effective interaction Hamiltonian
$H_{\t{int}}^{\t{eff}}$, $\P_X$ is a projector onto the corresponding
eigenspace, and
\begin{align}
  T_\mu^\z \equiv \f12 \p{\c_{\mu,\e}^\dag \c_{\mu,\e}
    - \c_{\mu,\g}^\dag \c_{\mu,\g}},
  &&
  T_\mu^\x \equiv \f12 \p{\c_{\mu,\e}^\dag \c_{\mu,\g}
    + \c_{\mu,\g}^\dag \c_{\mu,\e}},
\end{align}
are single-atom pseudospin operators.  The Rabi frequency $\Omega_\mu$
is proportional to the Clebsch-Gordan coefficient
$\bk{I,\mu;1,0|I,\mu}\propto\mu$ for a photon-induced
nuclear-spin-conserving orbital state transition of an atom with
nuclear spin $\mu$.  We therefore define the ``bare'' Rabi frequency
$\Omega_0\equiv\Omega_\mu/\mu$ to explicitly factor out dependence on
nuclear spins $\mu$.

Consider now a single lattice site in the orbital ground state
$\ket{\N,0}$ with nuclear spins $\N\equiv\set{\mu_j}$ for
$j=1,2,\cdots,N$.  If we red-detune the interrogation laser by
$\delta$ from a many-body orbital state excitation energy, i.e.~set
$\Delta=\Delta_{NX}-\delta$ for $\Delta_{NX}\equiv E_{NX}-E_{N,0}$ and
$X\in\set{+,-}$, then in the subspace of the target states
$\set{\ket{\N,0},\ket{\N X}}$ the Hamiltonian in \eqref{eq:H_rabi}
becomes
\begin{align}
  H_{\N X} = \delta S_{\N X}^\z + \Omega_{\N X} S_{\N X}^\x,
  \label{eq:H_NX}
\end{align}
where
\begin{align}
  S_{\N X}^\z &\equiv \f12\p{\op{\N X}-\op{\N,0}},
  \label{eq:S_NX_z} \\
  S_{\N X}^\x &\equiv \f12\p{\op{\N X}{\N,0}+\op{\N,0}{\N X}},
  \label{eq:S_NX_x}
\end{align}
are many-body pseudospin operators, and the Rabi frequencies
$\Omega_{\N X}$ are determined by
\begin{align}
  H_{\t{Rabi}}\ket{\N,0}
  = \f12\Omega_0\sum_\mu\mu\c_{\mu,\e}^\dag\c_{\mu,\g}\ket{\N,0}
  = \f12\Omega_{\N,+}\ket{\N,+} + \f12\Omega_{\N,-}\ket{\N,-}.
  \label{eq:O_NX}
\end{align}
While the symmetric excited state $\ket{\N,+}$ is given in
\eqref{eq:states_S}, at this point we have not explicitly solved for
the asymmetric excited state $\ket{\N,-}$.  The asymmetric state is
implicitly defined by \eqref{eq:O_NX}, and lies somewhere in the span
of the $N-1$ asymmetric states given in \eqref{eq:states_A}.
Determining the symmetric-state Rabi frequency $\Omega_{\N,+}$ is
simply a matter of projecting the expression in \eqref{eq:O_NX} onto
$\ket{\N,+}$, which yields
\begin{align}
  \Omega_{\N,+}
  = \bk{\N,+|\Omega_0\sum_\mu\mu\c_{\mu,\e}^\dag\c_{\mu,\g}|\N,0}
  = \Omega_0 \sum_{\mu\in\N} \f{\mu}{\sqrt{N}}
  = \Omega_0 \sqrt{N} \bar\mu_\N,
  \label{eq:O_N+}
\end{align}
where $\bar\mu_\N\equiv\sum_{\mu\in\N}\mu/N$ is the average nuclear
spin $\mu\in\N$.  In order to determine the asymmetric-state Rabi
frequency $\Omega_{\N,-}$, we rearrange \eqref{eq:O_NX} to find
\begin{align}
  \Omega_{\N,-}\ket{\N,-}
  = \Omega_0\sum_\mu\mu\c_{\mu,\e}^\dag\c_{\mu,\g}\ket{\N,0}
  - \Omega_{\N,+}\ket{\N,+}
  = \Omega_0 \sum_{\mu\in\N} \p{\mu - \bar\mu_\N}
  \c_{\mu,\e}^\dag \c_{\mu,\g} \ket{\N,0}.
\end{align}
Denoting the standard deviation of nuclear spins $\mu\in\N$ by
$\sigma_\N$, normalization of $\ket{\N,-}$ thus determines the
asymmetric-state Rabi frequency
\begin{align}
  \Omega_{\N,-}
  = \Omega_0 \sp{\sum_{\mu\in\N}\p{\mu-\bar\mu_\N}^2}^{1/2}
  = \Omega_0 \sqrt{N} \sigma_\N,
  \label{eq:O_N-}
\end{align}
which in turn implies that the asymmetric excited state $\ket{\N,-}$
is
\begin{align}
  \ket{\N,-} = \f1{\sqrt{N}}\sum_{\mu\in\N}
  \p{\f{\mu-\bar\mu_\N}{\sigma_\N}}
  \c_{\mu,\e}^\dag \c_{\mu,\g} \ket{\N,0}.
\end{align}

\begin{figure}
  \centering
  \includegraphics{coefficients_table.pdf}
  \caption{\footnotesize Multiplicities of the magnitudes of reduced
    Rabi frequencies
    $\omega_{\N X}\equiv\Omega_{\N X}/\Omega_0\sqrt{N}$ in a lattice
    with a uniform mixture of nuclear spins with $I=9/2$ and
    single-site occupation numbers $N$ which are achievable in current
    ${}^{87}$Sr experiments.}
  \label{fig:coefficients}
\end{figure}

Figure \ref{fig:coefficients} shows multiplicities of the magnitudes
of reduced Rabi frequencies
$\omega_{\N X}\equiv\Omega_{\N X}/\Omega_0\sqrt{N}$ in a lattice with
a uniform mixture of nuclear spins with $I=9/2$ for single-site
occupation numbers $N$ which are achievable in current ${}^{87}$Sr
experiments \cite{goban2018emergence}.  On average, asymmetric-state
Rabi frequencies are greater in magnitude, which becomes more
pronounced for larger single-site occupation numbers.

\subsection{Experimental signatures and comparison}

\begin{figure}
  \centering
  \subfloat{\includegraphics{sweep_V54_O50.pdf}}
  \subfloat{\includegraphics{sweep_expt_V54.pdf}}
  \caption{\footnotesize Population (in arbitrary units) of the
    excited ${}^3P_0$ orbital state of ${}^{87}$Sr atoms in a uniform
    mixture of nuclear spins.  Atoms are prepared in the ground state
    of a lattice with depth $\U=54 E_{\t{R}}$, where
    $E_{\t{R}}\approx3.5\times2\pi~\t{kHz}$ is the lattice photon
    recoil energy of the atoms, and interrogated by a laser with Rabi
    frequency $\Omega_I=50\times2\pi~\t{Hz}$ for a time
    $t=\pi/\Omega_I$.  ({\bf Left}) Populations predicted by the
    low-energy effective theory (with $s$-wave scattering parameters
    retrieved from ref.~\cite{zhang2014spectroscopic}), averaged over
    all nuclear spin combinations of $N\in\set{1,\cdots,5}$ atoms per
    lattice site for a fixed total atom number.  ({\bf Right})
    Experimental measurements of ${}^3P_0$ populations retrieved from
    ref.~\cite{goban2018emergence}, with Lorentzian fits to each peak
    as a visual guide.  Resonance peaks are identified by the
    many-body orbital states which are excited at the peak.}
  \label{fig:sweep}
\end{figure}

We now consider samples of ${}^{87}$Sr atoms in a uniform mixture of
nuclear spins $\mu\in\set{-I,\cdots,I}$ prepared in motional ground
states of a rectangular lattice with depths
$\U=\p{\U_\x,\U_\y,\U_\z}=\p{41,55,69} E_{\t{R}}$, where
$E_{\t{R}}\approx3.5\times2\pi~\t{kHz}$ is the lattice photon recoil
energy of the atoms.  Such samples can be prepared in experiments
which can vary the single-site occupation number $N$, and which can
control for the total number of atoms that are addressed by an
external interrogation laser.  Figure \ref{fig:sweep} shows the
population of the excited ${}^3P_0$ orbital state when these atoms are
interrogated for a time $t=\pi/\Omega_I$ by a laser with Rabi
frequency $\Omega_I=50\times2\pi~\t{Hz}$ (i.e.~for individual atoms
with nuclear spin $\mu=I$) and detuning $\Delta$ from the single-atom
${}^1S_0\to{}^3P_0$ orbital excitation energy.  The ${}^3P_0$
population peaks when the laser detuning $\Delta$ is equal to the
many-body excitation energy $\Delta_{NX}\equiv E_{NX}-E_{N,0}$ for
$X\in\set{+,-}$, as this is precisely when the on-resonance condition
$\delta=0$ is satisfied in the many-body Rabi Hamiltonian $H_{\N X}$
in \eqref{eq:H_NX}.  Due to experimental uncertainties which vary with
single-site occupation number $N$, the heights of experimental peaks
in figure \ref{fig:sweep} are not well-calibrated between different
values of $N$.  Nonetheless, figure \ref{fig:sweep} exhibits
signatures of larger asymmetric-state Rabi frequencies than
symmetric-state ones in the form of higher asymmetric-state peaks for
fixed $N$.

\begin{figure}
  \centering
  \includegraphics{shifts_table.pdf}
  \caption{\footnotesize Multi-body excitation energies of ultracold
    ${}^{87}$Sr atoms at various lattice depths.  The top plot in each
    sub-figure with fixed $N,X$ shows the excitation energies
    $\Delta_{NX}\equiv E_{NX}-E_{N,0}$ measured experimentally in
    ref.~\cite{goban2018emergence} and those predicted by the
    low-energy effective theory at different orders in the coupling
    constants, when applicable both with and without four-body
    contributions.  The bottom plot in each sub-figure shows the
    relative error
    $\eta_{NX}\equiv
    \Delta_{NX}^{\t{theory}}/\Delta_{NX}^{\t{experiment}}-1$.  Error
    bars represent experimental error or conservatively estimated
    theoretical uncertainties from nearest-neighbor hopping of virtual
    states in the low-energy effective theory (see Appendix
    \ref{sec:error}).}
  \label{fig:shifts}
\end{figure}

\begin{figure}
  \centering
  \includegraphics{shifts_summary.pdf}
  \caption{\footnotesize Summary of the many-body excitation spectra
    in figure \ref{fig:shifts}, retrieved from
    ref.~\cite{goban2018emergence}.}
  \label{fig:shifts_summary}
\end{figure}

Identifying peaks in excitation spectra such as in figure
\ref{fig:sweep} constitutes a measurement of many-body excitation
energies, which was performed in ref.~\cite{goban2018emergence} to
detect signatures of effective multi-body interactions.  Figure
\ref{fig:shifts} shows a comparison between experimental measurements
of the many-body excitation energies $\Delta_{NX}$ for
$N\in\set{3,4,5}$ at various lattice depths $\U$, and the
corresponding values of $\Delta_{NX}$ predicted by the low-energy
effective theory at different orders in the coupling constants.  A
summary of the data in figure \ref{fig:shifts} is provided in figure
\ref{fig:shifts_summary}.  These figures show a clear improvement of
the low-energy effective theory we have developed with increasing
order in the coupling constants, in particular demonstrating the need
to consider multi-body interactions in a single-band model of the
atoms without motional excitations.

We expect that the dominant source of error in the third-order
effective theory comes from the fact that we neglect the inter-site
matrix elements of all Hamiltonians.  This source of error is
discussed in Appendix \ref{sec:error}, and leads to theoretical
uncertainties represented by error bars in figure \ref{fig:shifts}.
As these uncertainties are sometimes unable to account for the
residual disagreement between experimental measurements and the
effective theory, better agreement requires either a better
understanding of experimental procedures and error, a better
understanding of theoretical uncertainties, or consideration of yet
higher-order terms in the effective theory.  Due to a combinatorial
explosion in the number of diagrams which appear at increasing
perturbative orders of the effective theory, we need more systematic
methods to compute effective multi-body Hamiltonians at fourth order;
we therefore leave this calculation for future work.

\subsection{Orbital-state dynamics of a nuclear spin mixture}

In addition to spectral measurements of many-body interaction
energies, we consider the dynamics of multiply-occupied lattice sites
during spectroscopic interrogation.  While these dynamics do not
provide information about the nature or origin of effective multi-body
interactions, they provide tools and intuition for addressing the
low-lying orbital excitations which are readily accessible in an
experimental setting.  If we initialize all atoms in the $N$-body
ground state with an incoherent mixture of all nuclear spins, then we
prepare the mixed state $\rho_{N,0}=\P_{N,0}/\Tr\P_{N,0}$, where
\begin{align}
  \P_{NX} \equiv \sum_{\abs{\N}=N}\op{\N X}
  \label{eq:P_NX}
\end{align}
is a projector onto the space of the $N$-body orbital states
$\ket{\N X}$.  Interrogating the atoms for a time $t$ by a laser
resonant with the excitation energy $\Delta_{N,\pm}$ then gives us the
state
\begin{align}
  \rho_N^{(\pm)}\p{t} = \f1{\Tr\P_{N,0}}
  \sum_{\abs{\N}=N} \exp\p{-it\Omega_{\N,\pm}S_{\N,\pm}^\x}
  \op{\N,0} \exp\p{it\Omega_{\N,\pm}S_{\N,\pm}^\x},
\end{align}
where the Rabi frequencies $\Omega_{\N,+},\Omega_{\N,-}$ and
pseudo-spin operators $S_{\N,\pm}^\x$ are respectively given in
\eqref{eq:O_N+}, \eqref{eq:O_N-}, and \eqref{eq:S_NX_x}.  Denoting the
eigenstates of $S_{\N,\pm}^\x$ by
\begin{align}
  \ket{\N,\S_\pm} \equiv \f1{\sqrt2} \p{\ket{\N,0} + \ket{\N,\pm}},
  &&
  \ket{\N,\A_\pm} \equiv \f1{\sqrt2} \p{\ket{\N,0} - \ket{\N,\pm}},
\end{align}
and defining the identity operator projected to the relevant subspace,
\begin{align}
  \1_{\N,\pm} \equiv \op{\N,0} + \op{\N,\pm}
  = \op{\N,\S_\pm} + \op{\N,\A_\pm},
\end{align}
we can write the state $\rho_N^{(\pm)}\p{t}$ and excited-state
projectors $\P_{N,\pm}$ in the form
\begin{align}
  \rho_N^{(\pm)}\p{t} = \f1{\Tr\P_{N,0}} \sum_{\abs{\N}=N} \f12
  \sp{\1_{\N,\pm} + e^{i2t\Omega_{\N,\pm}}\op{\N,\S_\pm}{\N,\A_\pm}
    + e^{-i2t\Omega_{\N,\pm}}\op{\N,\A_\pm}{\N,\S_\pm}},
  \label{eq:rho_N_pm_t}
\end{align}
and
\begin{align}
  \P_{N,\pm} = \sum_{\abs{\N}=N} \f12
  \sp{\1_{\N,\pm}-\op{\N,\S_\pm}{\N,\A_\pm}-\op{\N,\A_\pm}{\N,\S_\pm}},
\end{align}
from which it follows that the net excited-state population at time
$t$ is
\begin{align}
  \bk{\P_{N,\pm}\p{t}}
  \equiv \Tr\sp{\rho_N^{(\pm)}\p{t}\P_{N,\pm}}
  = \f12 - \f12 \bk{\cos\p{2t\Omega_{\N,\pm}}}_{\abs{\N}=N},
  \label{eq:P_N_pm_t}
\end{align}
where $\bk{X}_{\abs{\N}=N}\equiv\sum_{\abs{\N}=N}X/\Tr\P_{N,0}$ is an
average of $X$ over all choices of $N$ distinct nuclear spins.

\begin{figure}
  \centering
  \includegraphics{time_table.pdf}
  \caption{\footnotesize Net population of the $N$-body orbital
    excited states $\set{\ket{\N,\pm}}$ after interrogation of an
    initial mixed state $\rho_{N,0}=\P_{N,0}/\Tr\P_{N,0}$ for a
    reduced time $\tau_N\equiv t\Omega_0/\sqrt{N}$ (i.e.~with real
    time $t$) by a laser with bare Rabi frequency $\Omega_0$ which is
    resonant with the $N$-body excitation energy $\Delta_{NX}$.  Here
    $\P_{NX}$, defined in \eqref{eq:P_NX}, is a projector onto the
    space of the $N$-body orbital states $\ket{\N X}$.}
  \label{fig:time}
\end{figure}

Figure \ref{fig:time} shows the excited-state population
$\bk{\P_{N,\pm}\p{t}}$ for several occupation numbers $N$.  With the
exception of $N=2$, the asymmetric-state populations generally have an
initial short period of growth before falling back to
$\bk{\P_{N,-}}\approx1/2$.  This behavior can be understood by the
fact that for fixed $N>2$, any pair of Rabi frequencies
$\Omega_{\N_1,-},\Omega_{\N_2,-}$ with
$\Omega_{\N_1,-}\ne\Omega_{\N_2,-}$ are mutually incommensurate, which
implies that at times $t$ with
$\min_{\abs{\N}=N}\set{2t\Omega_{\N,-}}\gtrsim1$ the averaging in
\eqref{eq:P_N_pm_t} effectively becomes a pseudo-random sampling
average of $\cos x$ over values of $x$, so
$\bk{\cos\p{2t\Omega_{\N,-}}}_{\abs{\N}=N}\approx0$.  When $N=2$, the
asymmetric-state Rabi frequencies essentially take on the same values
as the symmetric-state ones (see figure \ref{fig:coefficients}); the
behavior of asymmetric-state population dynamics for $N=2$ can
therefore be understood by the following discussion of symmetric-state
population dynamics.

To understand the periodic collapse and revival of symmetric-state
populations in figure \ref{fig:time}, we observe from \eqref{eq:O_N+}
that the symmetric-state phases in \eqref{eq:rho_N_pm_t} and
\eqref{eq:P_N_pm_t} take the form
\begin{align}
  2t\Omega_{\N,+} = \tau_N \sum_{\mu\in\N} 2\mu
  &&
  \t{with}
  &&
  \tau_N \equiv \f{t\Omega_0}{\sqrt{N}},
  \label{eq:2tO_N+}
\end{align}
where for fermionic atoms with half-integer nuclear spin, $2\mu$ is
always an odd integer, which implies that the sum in \eqref{eq:2tO_N+}
is an integer with the same parity (i.e.~even/odd) as the occupation
number $N$ (i.e.~the number of elements in $\N$).  At reduced times
$\tau_N=n\pi$ with integer $n$, therefore, if the occupation number
$N$ is even then all phases $2t\Omega_{\N,+}$ are integer multiples of
$2\pi$, which leads to a collapse of the excited-state populations as
$\left.\rho_N^{(\pm)}\p{t}\right|_{\tau_N=n\pi} = \rho_N^{(\pm)}\p{0}
= \rho_{N,0}$.  If the occupation $N$ is odd, meanwhile, then the
phases $2t\Omega_{\N,+}$ are all odd (even) integer multiples of $\pi$
for odd (even) $n$.  This alignment of phases implies a complete
population transfer to the excited state
$\rho_{N,-}\equiv\P_{N,-}/\Tr\P_{N,-}$ for odd $n$, and a collapse
back to the orbital ground state $\rho_{N,0}$ for even $n$, precisely
as observed in figure \ref{fig:time}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Summary and outlook}
\label{sec:summary}

Current 3-D optical lattice experiments with fermionic AEAs are
capable of operating in the low-temperature, high-density,
strongly-interacting limit where inter-atomic interactions set the
dominant energy scale governing system dynamics.  For AEAs with total
nuclear spin $I$ and $N=2I+1$ nuclear spin states, these interactions
exhibit an exotic SU($N$) symmetry which is of great interest for
near-term quantum simulations of SU($N$) spin models and lattice field
theories.  Working in the deep-lattice limit, we have derived a
low-energy effective theory of such atoms.  This theory exhibits
emergent multi-body interactions that inherit the SU($N$) symmetry of
the atoms' bare, two-body interactions.  When considering the
low-lying orbital excitations of our effective theory, the SU($N$)
symmetry of $M$-body Hamiltonians allowed us to express them in a
simple form, and to fully characterize their eigenstates and spectra.
Capitalizing on the extreme precision of state-of-the-art clock
spectroscopy, we have tested spectral predictions of our theory
against direct experimental measurements of the many-body ${}^{87}$Sr
excitation spectrum.  This comparison shows good agreement between
theory and experiment, clearly demonstrating the need to consider
multi-body effects for understanding the low-energy physics of high
density AEA samples on a 3-D lattice.  Finally, we analyzed the
many-body orbital-state dynamics of multiply-occupied lattice sites
prepared in a nuclear spin mixture and interrogated via Rabi
spectroscopy.  This analysis is useful for future experimental probes
of many-body state structures, as well as for the preparation of
long-lived states with multi-partite entanglement
(i.e.~$\ket{\N,\pm}$) which may be used as a resource to perform
quantum information processing tasks.

Despite the nominal success of our low-energy effective theory in
reproducing experimental observations, there remains room for
improvement in the form of controlled, systematic treatment of
higher-order and tunneling processes.  Nonetheless, our work makes a
major step towards the experimental investigation of multi-body
SU($N$) physics, providing the necessary framework for future studies
going beyond the deep-lattice limit to realize multi-body
super-exchange dynamics and orbital SU($N$) quantum magnetism with
AEAs.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section*{Acknowledgements}

We acknowledge helpful discussions with R.~B.~Hutson, G.~E.~Marti,
S.~L.~Campbell, J.~Ye, P.~Julienne, J.~P.~D'Incao, C.~Kennedy, and
L.~Radzihovsky; and in particular close correspondence and technical
contributions from A.~Goban.  This work is supported by the Air Force
Office of Scientific Research (AFOSR) grant FA9550-18-1-0319; the
AFOSR Multidisciplinary University Research Initiative (MURI) grant;
the Defense Advanced Research Projects Agency (DARPA) and Army
Research Office (ARO) grant W911NF-16-1-0576; the National Science
Foundation (NSF) grant PHY-1820885; JILA-NSF grant PFC-173400; and the
National Institute of Standards and Technology (NIST).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section*{Appendices}
\appendix

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Summary of experimental procedures}
\label{sec:experiment_summary}

The work in this paper is closely tied to the experimental work in
ref.~\cite{goban2018emergence}; for quick reference, we here provide a
short summary of the experimental procedures used therein.  The
experiment begins by preparing a degenerate gas of $10^4$-$10^5$
(fermionic) ${}^{87}$Sr atoms, in a uniform mixture of their ten
nuclear spin states, at $\sim0.1$ of their fermi temperature (10 to 20
nanokelvin) \cite{campbell2017fermidegenerate, marti2018imaging}.
This gas is loaded into a primitive cubic optical lattice at the
``magic wavelength'' for which both ground (${}^1S_0$) and
first-excited (${}^3P_0$) electronic orbital states of the atoms
experience the same lattice potential.  Lattice depths along the
principal axes of the lattice are comparable in magnitude, with a
geometric mean that can be varied from 30 to 80 $E_{\t{R}}$, where
$E_{\t{R}}\approx3.5\times2\pi~\t{kHz}$ is the lattice photon recoil
energy of the atoms.  These lattice depths are sufficiently large as
to neglect tunneling on the time scales relevant to the experiment.
Atoms are then addressed by an external (``clock'') interrogation
laser with an ultranarrow (26 mHz) linewidth, detuned by $\Delta$ from
the single-atom ${}^1S_0-{}^3P_0$ transition frequency.  When
$\Delta=0$, non-interacting atoms in singly-occupied lattice sites
thus undergo Rabi oscillations.  After a chosen time, the experiment
removes all ground-state (${}^1S_0$) atoms from the lattice and uses
absorption imaging to identify the remaining excited-state (${}^3P_0$)
atoms.  As discussed in section \ref{sec:spectra}, a similar procedure
with $\Delta\ne0$ leads to the spectrum shown in figure
\ref{fig:sweep}, thereby constituting a measurement of the many-body
atomic excitation spectrum.  The many-body atomic excitation spectrum
is different from the single-body atomic excitation spectrum due to
the presence of inter-atomic interactions.  We note that this
spectroscopic protocol addresses only singly-excited orbital states of
any lattice site.  Doubly-excited states are manifestly off-resonant
for single-photon absorption processes, while multi-photon absorption
processes are always off-resonant due to
\begin{enumerate*}
\item the non-linearity (on the scale of kHz) of on-site electronic
  orbital excitation energies (resulting from non-uniform coupling
  constants $G_X$), combined with
\item the ultranarrow linewidth (on the scale of mHz) of the
  interrogation laser.
\end{enumerate*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Derivation of the effective Hamiltonian expansion}
\label{sec:eff_derivation}

Suppose we have a Hamiltonian $H_0$ on a Hilbert space
$\H=\G_0\oplus\E_0$ for a zero-energy manifold $\G_0$ decoupled from a
positive-energy manifold $\E_0$, and that we perturb $H_0$ by an
operator $V$ which weakly couples $\G_0$ and $\E_0$.  For all
$\ket{\psi_\g}\in\G_0$ and $\ket{\phi_\e},\ket{\chi_\e}\in\E_0$, we
have $\bk{\psi_\g|H_0|\psi_\g}=0$, $\bk{\phi_\e|H_0|\psi_\g}=0$, and
$\abs{\bk{\phi_\e|V|\psi_\g}}\ll\bk{\chi_\e|H_0|\chi_\e}$.  The net
Hamiltonian $H=H_0+V$ will naturally admit a decomposition of the
Hilbert space as $\H=\G\oplus\E$ for a subspace $\G$ which is spanned
by the low-energy eigenstates of $H$ and has the same dimension as
$\G_0$, i.e.~$\abs{\G}=\abs{\G_0}$.

We can perform a canonical transformation between $\G$ and $\G_0$
which yields an {\it effective Hamiltonian} $H_{\t{eff}}$ on $\G_0$
that reproduces the spectrum of $H$ on $\G$
\cite{bravyi2011schrieffer}.  Given an eigenbasis
$\set{\ket{\alpha_0}}$ for $H_0$ on $\G_0$ and $\set{\ket\alpha}$ for
$H$ on $\G$, this transformation is implemented by a unitary $U$ for
which $\ket{\alpha_0}=U\ket\alpha$ and $U\to\1$ as $\norm{V}\to0$.
The effective Hamiltonian is then simply
\begin{align}
  H_{\t{eff}} = U H U^\dag.
  \label{eq:H_eff}
\end{align}
The prescription in \eqref{eq:H_eff} for constructing an effective
Hamiltonian is commonly known as a Schieffer-Wolff transformation
\cite{schrieffer1966relation}.  Unitaries $U$ which follow this
prescription are not unique, and different choices of $U$ amount to
different realizations of the Schieffer-Wolff transformation.  In
ref.~\cite{bravyi2011schrieffer}, the authors construct the unique
operator $S$ which generates a {\it direct} or {\it minimal} rotation
$U_{\t{min}}=e^S$ between $\G$ and $\G_0$, and use this construction
to expand \eqref{eq:H_eff} as a perturbative series in $V$.  The
rotation $U_{\t{min}}$ is minimal in the sense that it minimizes the
distance of candiate unitaries $U$ from the identity $\1$ with respect
to the Euclidian operator norm\footnote{The Euclidean operator norm is
  also known as the $L_{2,2}$, Hilbert-Schmidt, or Frobenius norm.}
$\norm{X}_{\t{E}}\equiv\sqrt{\Tr\p{X^\dag X}}$.  This rotation is
determined uniquely by enforcing
\begin{enumerate*}
\item that the generator $S$ is strictly block-off-diagonal with
  respect to $\G_0$ and $\E_0$,
\item that the norm $\norm{S}_{\t{E}}<\pi/2$, and
\item that the block-off-diagonal parts of \eqref{eq:H_eff} are zero.
\end{enumerate*}

To summarize the solution in ref.~\cite{bravyi2011schrieffer}, the
effective Hamiltonian $H_{\t{eff}}$ induced by a direct rotation can
be expanded as
\begin{align}
  H_{\t{eff}} = \sum_{p\ge0} H_{\t{eff}}^{(p)},
  \label{eq:H_eff_expansion}
\end{align}
where $H_{\t{eff}}^{(p)}$ is order $p$ in $V$.  Letting
$\Q_0\equiv1-\P_0$ denote the projector onto $\E_0$ and $X$ denote any
operator on $\H$, we define the superoperators
\begin{align}
  \D X \equiv \P_0 X \P_0 + \Q_0 X \Q_0,
  &&
  \O X \equiv \P_0 X \Q_0 + \Q_0 X \P_0,
\end{align}
which select out the diagonal ($\D$) and off-diagonal ($\O$) parts of
$X$ with respect to $\G_0$ and $\E_0$, and
\begin{align}
  \L X \equiv \sum_{\alpha,\beta}
  \f{\op\alpha \O X \op\beta}{E_\alpha - E_\beta}
  &&
  \t{where}
  &&
  H_0 = \sum_\alpha E_\alpha \op\alpha.
\end{align}
The first few terms of the expansion in \eqref{eq:H_eff_expansion} are
then, as derived in ref.~\cite{bravyi2011schrieffer},
\begin{align}
  H_{\t{eff}}^{(0)} = \P_0 H_0 \P_0,
  &&
  H_{\t{eff}}^{(1)} = \P_0 V \P_0,
  \label{eq:H_eff_0_1}
\end{align}
\begin{align}
  H_{\t{eff}}^{(2)} = - \f12 \P_0 \sp{\O V,\L V} \P_0,
  &&
  H_{\t{eff}}^{(3)} = \f12 \P_0 \sp{\O V, \L \sp{\D V, \L V}} \P_0.
\end{align}
Exploiting the fact that in our case $\bk{\psi|H_0|\psi}=0$ for all
$\ket\psi\in\G_0$, we let $\B_0\p{\E_0}$ denote an eigenbasis of $H_0$
for $\E_0$ and define the operator
\begin{align}
  \I \equiv \sum_{\ket\alpha\in\B_0\p{\E_0}}\f{\op\alpha}{E_\alpha},
\end{align}
which sums over projections onto excited states with corresponding
energetic suppression factors.  We then expand
\begin{align}
  \L X = \O\p{\L X} = \Q_0 \L X \P_0 + \P_0 \L X \Q_0
  = \I X \P_0 - \P_0 X \I,
\end{align}
which simplifies the expression for $H_{\t{eff}}^{(2)}$ as
\begin{align}
  H_{\t{eff}}^{(2)}
  = - \f12 \P_0 \p{\sp{\O V, \I V \P_0}
    - \sp{\O V, \P_0 V \I}} \P_0
  = - \P_0 V \I V \P_0.
  \label{eq:H_eff_2}
\end{align}
Working toward a similar expansion for $H_{\t{eff}}^{(3)}$, we compute
\begin{align}
  \sp{\D V, \L V}
  = \sp{\D V, \I V \P_0} - \sp{\D V, \P_0 V \I}
  = \O\p{V \I V} - \I V \P_0 V \P_0 - \P_0 V \P_0 V \I,
\end{align}
and in turn
\begin{align}
  H_{\t{eff}}^{(3)}
  &= \f12 \P_0 \p{\sp{\O V, \I \sp{\D V, \L V} \P_0}
    - \sp{\O V, \P_0 \sp{\D V, \L V} \I}} \P_0 \\
  &= \f12 \P_0 \p{V \I \sp{\D V, \L V}
    + \sp{\D V, \L V} \I V} \P_0 \\
  &= \P_0 V \I V \I V \P_0
  - \f12 \P_0 V \I^2 V \P_0 V \P_0
  - \f12 \P_0 V \P_0 V \I^2 V \P_0 \\
  &= \P_0 V \I V \I V \P_0
  - \f12 \sp{\P_0 V \P_0, \P_0 V \I^2 V \P_0}_+,
  \label{eq:H_eff_3}
\end{align}
where $\sp{X,Y}_+\equiv XY+YX$.  The expressions in
\eqref{eq:H_eff_0_1}, \eqref{eq:H_eff_2}, and \eqref{eq:H_eff_3}
complete the derivation for our expansion of the effective interaction
Hamiltonian $H_{\t{int}}^{\t{eff}}$ in \eqref{eq:H_int_eff} through
third order.  In the case of ultracold atoms on a lattice, the
motional ground-state subspace $\G_0$ actually contains many internal
atomic states with different energies.  Nonetheless, the total Hilbert
space is completely separable into uncoupled subspaces associated with
each symmetrized many-body internal atomic state.  One can therefore
diagonalize the interaction Hamiltonian with respect to these internal
states and derive an effective theory within each of the corresponding
subspaces, in each case setting the appropriate ground-state energy to
zero.  This procedure is equivalent to simultaneously calculating the
effective Hamiltonian $H_{\t{int}}^{\t{eff}}$ for all internal states
via the prescriptions we have provided, but letting $E_\alpha$ denote
only the motional excitation energy of states
$\ket{\alpha_0}\in\B_0\p{\E_0}$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Diagram counting and symmetry factors}
\label{sec:diagrams}

The fact that we include factors of $1/2$ from the bare two-body
interaction Hamiltonian $H_{\t{int}}$ in the definition of diagrams
implies that $p$-vertex diagrams acquire a factor of $1/2^p$.  In
practice, however, these factors are exactly cancelled out by
corresponding symmetry factors in all $M$-body diagrams with $M>2$.
As an illustrative example, consider the second-order effective
Hamiltonian $H_{\t{int}}^{(2)}$ in \eqref{eq:H_int_1_2}, which
expanded in full reads
\begin{align}
  H_{\t{int}}^{(2)} = -\sum_{m+n>0} \P_0
  \p{\f12 K_{mn} G^{st}_{s't'}
    \c_{\rho s'}^\dag \c_{\sigma t'}^\dag \c_{n\sigma t} \c_{m\rho s}}
  \f1{E_{mn}}
  \p{\f12 K_{mn} G^{qr}_{q'r'}
    \c_{m\mu q'}^\dag \c_{n\nu r'}^\dag \c_{\nu r} \c_{\mu q}} \P_0.
  \label{eq:H_int_2_full}
\end{align}
The three-body terms in this Hamiltonian have
$\abs{\set{\mu,\nu,\rho,\sigma}}=3$ and only one virtually excited
atom.  The non-vanishing three-body terms must therefore either have
$\rho\in\set{\mu,\nu}$ and contain a factor of the form
$\c_X^\dag \c_{\sigma t'}^\dag \c_{\sigma t} \c_Y$, or have
$\sigma\in\set{\mu,\nu}$ with a factor of the form
$\c_{\rho s'}^\dag \c_X^\dag \c_Y \c_{\rho s}$, where the labels $X,Y$
both address whichever nuclear spin (i.e.~$\mu$ or $\nu$) was excited
in the corresponding term.  Diagrammatically, we have terms of the
form
\begin{align}
  \begin{tikzpicture}
    \begin{feynman}
      \vertex (v1);
      \vertex[above left = of v1] (f1) {$\mu q$};
      \vertex[below left = of v1] (f2) {$\nu r$};
      \vertex[right = of v1] (vm);
      \vertex[right = of vm] (v2);
      \vertex[above = of vm] (f3) {$Z$};
      \vertex[below = of vm] (f4) {$\sigma t$};
      \vertex[below right = of v2] (f5) {$\sigma t'$};
      \vertex[above right = of v2] (f6) {$X$};
      \diagram* {
        (f1) -- (v1) -- (f3),
        (f2) -- (v1) --  [scalar, edge label=$Y$] (v2),
        (f4) -- (v2) -- (f5),
        (v2) -- (f6), };
    \end{feynman}
  \end{tikzpicture}
  &&
  \t{and}
  &&
  \begin{tikzpicture}
    \begin{feynman}
      \vertex (v1);
      \vertex[above left = of v1] (f1) {$\mu q$};
      \vertex[below left = of v1] (f2) {$\nu r$};
      \vertex[right = of v1] (vm1);
      \vertex[right = of vm1] (vm2);
      \vertex[right = of vm2] (v2);
      \vertex[above = of vm1] (f3) {$Z$};
      \vertex[above = of vm2] (f4) {$\rho s$};
      \vertex[above right = of v2] (f5) {$\rho s'$};
      \vertex[below right = of v2] (f6) {$X$};
      \diagram* {
        (f1) -- (v1) -- (f3),
        (f2) -- (v1) --  [scalar, edge label=$Y$] (v2),
        (f4) -- (v2) -- (f5),
        (v2) -- (f6), };
    \end{feynman}
  \end{tikzpicture}.
  \label{eq:rho_sigma}
\end{align}
Observing that
$\c_{\rho s'}^\dag \c_X^\dag \c_Y \c_{\rho s} = \c_X^\dag \c_{\rho
  s'}^\dag \c_{\rho s} \c_Y$, however, it is clear that both of the
terms represented in \eqref{eq:rho_sigma} are equal up to the
re-indexing $\p{\sigma,t,t'}\leftrightarrow\p{\rho,s,s'}$.  There is
therefore a symmetry factor of $2$ associated with the second vertex
of the diagrams in \eqref{eq:rho_sigma}, which cancels out with the
explicit factor of $1/2$ at that vertex, i.e.~the first factor of
$1/2$ in \eqref{eq:H_int_2_full}.  A symmetry factor of essentially
identical origin appears at every vertex with an ``incoming'' virtual
state, as in e.g.~the second and third vertices of
\begin{align}
  \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[below right = of v1] (v2);
        \vertex[above right = of v2] (v3);
        \vertex[above left = of v1] (f1);
        \vertex[left = of v1] (f2);
        \vertex[below left = of v2] (f3);
        \vertex[above right = of v3] (f4);
        \vertex[right = of v3] (f5);
        \vertex[below right = of v2] (f6);
        \diagram* {
          (f1) -- (v1) --[scalar] (v3) -- (f4),
          (f2) -- (v1) --[scalar] (v2) --[scalar] (v3) -- (f5),
          (f3) -- (v2) -- (f6), };
      \end{feynman}
    \end{tikzpicture}}
  &&
  \t{and}
  &&
  \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[below right = of v1] (v2);
        \vertex[below right = of v2] (v3);
        \vertex[above left = of v1] (f1);
        \vertex[below left = of v1] (f2);
        \vertex[below left = of v2] (f3);
        \vertex[below left = of v3] (f4);
        \vertex[above right = of v1] (f5);
        \vertex[above right = of v2] (f6);
        \vertex[above right = of v3] (f7);
        \vertex[below right = of v3] (f8);
        \diagram* {
          (f1) -- (v1) -- (f5),
          (f2) -- (v1) -- (f5),
          (v1) --[scalar] (v2) -- (v3) -- (f8),
          (f3) -- (v2) -- (f6),
          (f4) -- (v3) -- (f7),
        };
      \end{feynman}
    \end{tikzpicture}},
  \label{eq:diagram_example_1}
\end{align}
or the last vertex of
\begin{align}
  \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[below right = 2.5em of v1] (v3);
        \vertex[below left = 1.7em of v3] (v2);
        \vertex[above left = of v1] (f1);
        \vertex[left = of v1] (f2);
        \vertex[left = of v2] (f3);
        \vertex[below left = of v2] (f4);
        \vertex[above right = of v1] (f5);
        \vertex[above right = of v3] (f6);
        \vertex[below right = of v3] (f7);
        \vertex[below right = of v2] (f8);
        \diagram* {
          (f1) -- (v1) -- (f5),
          (f2) -- (v1) --[scalar] (v3) -- (f6),
          (f3) -- (v2) -- (v3) -- (f7),
          (f4) -- (v2) -- (f8), };
      \end{feynman}
    \end{tikzpicture}}.
  \label{eq:diagram_example_2}
\end{align}
We can thus account for cancellations of $1/2$ at all vertices except
those which address two ``initial'' ground-state atoms, as in the
first vertex of the diagrams in \eqref{eq:rho_sigma} and
\eqref{eq:diagram_example_1}, or the first two vertices of the diagram
in \eqref{eq:diagram_example_2}.  For such vertices, there are two
possibilities: either
\begin{enumerate}
\item both edges leaving the vertex in question (i.e.~leaving to the
  right) terminate at different vertices, as in the examples above, or
  \label{enum:different_vertices}
\item both edges leaving the vertex in question terminate at the same
  vertex, as in for example the first vertex of
  \label{enum:same_vertex}
  \begin{align}
    \shrink{
      \begin{tikzpicture}
        \begin{feynman}
          \vertex (v1);
          \vertex[right = of v1] (v2);
          \vertex[below right = of v2] (v3);
          \vertex[above left = of v1] (f1);
          \vertex[below left = of v1] (f2);
          \vertex[below left = of v3] (f3);
          \vertex[above right = of v2] (f4);
          \vertex[above right = of v3] (f5);
          \vertex[below right = of v3] (f6);
          \diagram* {
            (f1) -- (v1) --[half left, scalar] (v2) -- (f4),
            (f2) -- (v1) --[half right, scalar] (v2)
            --[scalar] (v3) -- (f5),
            (f3) -- (v3) -- (f6), };
        \end{feynman}
      \end{tikzpicture}}.
    \label{eq:diagram_example_3}
  \end{align}
\end{enumerate}
In the former case, \ref{enum:different_vertices}, the vertex in
question has an associated symmetry factor of 2 to account for the
possibility of a nuclear spin exchange at that vertex.  Considering
again our example of the second-order effective Hamiltonian
$H_{\t{int}}^{(2)}$ in \eqref{eq:H_int_2_full}, the non-vanishing
three-body terms must either have $m=0$ and contain the factor
$\c_{\mu q'}^\dag \c_{n\nu r'}^\dag \c_{\nu r} \c_{\mu q}$, or have
$n=0$ with the factor
$\c_{m\mu q'}^\dag \c_{\nu r'}^\dag \c_{\nu r} \c_{\mu q}$, which
diagrammatically translates to
\begin{align}
  \begin{tikzpicture}
    \begin{feynman}
      \vertex (v1);
      \vertex[above left = of v1] (f1) {$\mu q$};
      \vertex[below left = of v1] (f2) {$\nu r$};
      \vertex[right = of v1] (vm);
      \vertex[right = of vm] (v2);
      \vertex[above = of vm] (f3) {$\mu q'$};
      \vertex[below = of vm] (f4) {$\sigma t$};
      \vertex[below right = of v2] (f5) {$\sigma t'$};
      \vertex[above right = of v2] (f6) {$\nu r''$};
      \diagram* {
        (f1) -- (v1) -- (f3),
        (f2) -- (v1) --  [scalar, edge label=$n\nu r'$] (v2),
        (f4) -- (v2) -- (f5),
        (v2) -- (f6), };
    \end{feynman}
  \end{tikzpicture}
  &&
  \t{or}
  &&
  \begin{tikzpicture}
    \begin{feynman}
      \vertex (v1);
      \vertex[above left = of v1] (f1) {$\mu q$};
      \vertex[below left = of v1] (f2) {$\nu r$};
      \vertex[right = of v1] (vm);
      \vertex[right = of vm] (v2);
      \vertex[above = of vm] (f3) {$\nu r'$};
      \vertex[below = of vm] (f4) {$\sigma t$};
      \vertex[below right = of v2] (f5) {$\sigma t'$};
      \vertex[above right = of v2] (f6) {$\mu q''$};
      \diagram* {
        (f1) -- (v1) -- (f3),
        (f2) -- (v1) --  [scalar, edge label=$m\mu q'$] (v2),
        (f4) -- (v2) -- (f5),
        (v2) -- (f6), };
    \end{feynman}
  \end{tikzpicture}.
  \label{eq:mu_nu}
\end{align}
These terms are equal up to the re-indexing
$\p{\nu,r,r',r'',n}\leftrightarrow\p{\mu,q,q',q'',m}$, which implies
that there is a symmetry factor of $2$ associated with the first
vertex of the diagrams in \eqref{eq:mu_nu}.  A symmetry factor of
identical origin is associated with the first vertex of the diagrams
in \eqref{eq:diagram_example_1}, and the first two vertices of the
diagram in \eqref{eq:diagram_example_2}.

The final case we must consider is \ref{enum:same_vertex}, which
occurs in the first vertex of \eqref{eq:diagram_example_3}.  In this
case, the symmetry factor of $2$ which appears in case
\ref{enum:different_vertices} to account for the possibility of a
nuclear spin exchange simply gets ``pushed forward'' to the vertex at
which the two nuclear spins in question part ways, e.g.~to account for
the two possibilities
\begin{align}
  \begin{tikzpicture}
    \begin{feynman}
      \vertex (v1);
      \vertex[right = of v1] (v2);
      \vertex[below right = of v2] (v3);
      \vertex[above left = of v1] (f1) {$\mu r$};
      \vertex[below left = of v1] (f2) {$\nu s$};
      \vertex[below left = of v3] (f3) {$\rho t$};
      \vertex[above right = of v2] (f4) {$\mu r''$};
      \vertex[above right = of v3] (f5) {$\nu s'''$};
      \vertex[below right = of v3] (f6) {$\rho t'$};
      \diagram* {
        (f1) -- (v1)
        --[half left, scalar, edge label=$\ell\mu r'$] (v2)
        -- (f4),
        (f2) -- (v1)
        --[half right, scalar, edge label'=$m\nu s'$] (v2)
        --[scalar, edge label=$n\nu s''$] (v3)
        -- (f5),
        (f3) -- (v3) -- (f6), };
    \end{feynman}
  \end{tikzpicture}
  &&
  \t{and}
  &&
  \begin{tikzpicture}
    \begin{feynman}
      \vertex (v1);
      \vertex[right = of v1] (v2);
      \vertex[below right = of v2] (v3);
      \vertex[above left = of v1] (f1) {$\mu r$};
      \vertex[below left = of v1] (f2) {$\nu s$};
      \vertex[below left = of v3] (f3) {$\rho t$};
      \vertex[above right = of v2] (f4) {$\nu s''$};
      \vertex[above right = of v3] (f5) {$\mu r'''$};
      \vertex[below right = of v3] (f6) {$\rho t'$};
      \diagram* {
        (f1) -- (v1)
        --[half left, scalar, edge label=$\ell\mu r'$] (v2)
        -- (f4),
        (f2) -- (v1)
        --[half right, scalar, edge label'=$m\nu s'$] (v2)
        --[scalar, edge label=$n\mu r''$] (v3)
        -- (f5),
        (f3) -- (v3) -- (f6), };
    \end{feynman}
  \end{tikzpicture},
  \label{eq:mu_nu_loop}
\end{align}
which are equal up to the re-indexing
$\p{\nu,s,s',s'',s''',m}\leftrightarrow\p{\mu,r,r',r'',r''',\ell}$.
The arguments for a symmetry factor in cases
\ref{enum:different_vertices} and \ref{enum:same_vertex} fail only if
the two nuclear spins in question take identical paths through the
internal vertices of a diagram, such that there is no meaningful sense
in which two diagrams can be said to differ by a nuclear spin
exchange, as in \eqref{eq:mu_nu} and \eqref{eq:mu_nu_loop}.  If two
atoms take identical paths through the internal vertices of a diagram,
however, then they have only participated in a two-body process, as in
\begin{align}
  \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v);
        \vertex[above left = of v] (f1);
        \vertex[below left = of v] (f2);
        \vertex[above right = of v] (f3);
        \vertex[below right = of v] (f4);
        \diagram* {
          (f1) -- (v),
          (f2) -- (v),
          (v) -- (f3),
          (v) -- (f4) };
      \end{feynman}
    \end{tikzpicture}},
  &&
  \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[above left = of v1] (f1);
        \vertex[below left = of v1] (f2);
        \vertex[right = of v1] (v2);
        \vertex[above right = of v2] (f3);
        \vertex[below right = of v2] (f4);
        \diagram* {
          (f1) -- (v1),
          (f2) -- (v1),
          (v2) -- (f3),
          (v2) -- (f4),
          (v1) --[scalar, half left] (v2),
          (v1) --[scalar, half right] (v2) };
      \end{feynman}
    \end{tikzpicture}},
  &&
  \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[above left = of v1] (f1);
        \vertex[below left = of v1] (f2);
        \vertex[right = of v1] (v2);
        \vertex[right = of v2] (v3);
        \vertex[above right = of v3] (f3);
        \vertex[below right = of v3] (f4);
        \diagram* {
          (f1) -- (v1),
          (f2) -- (v1),
          (v3) -- (f3),
          (v3) -- (f4),
          (v1)
          --[scalar, half left] (v2)
          --[scalar, half left] (v3),
          (v1)
          --[scalar, half right] (v2)
          --[scalar, half right] (v3) };
      \end{feynman}
    \end{tikzpicture}},
  &&
  \cdots.
\end{align}
After summing over all free indices, therefore, all two-body diagrams
have a remaining factor of $1/2$ from the first vertex.  In all
connected $M$-body diagrams with $M>2$, meanwhile, every factor of
$1/2$ can be identified one-to-one with a corresponding symmetry
factor of $2$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Effective coupling constants in a lattice}
\label{sec:renormalization}

Due to our choice of renormalization scheme in section
\ref{sec:two_body}, the interaction energies prescribed by our
low-energy effective theory for multiply-occupied lattice sites are
not given directly by the coupling constants $G_X$ defined by the
free-space scattering lengths $a_X$ in \eqref{eq:couplings}.  Instead,
we must first compute effective coupling constants
$G_X^{\t{lattice}}\p{\U}$ in a lattice with depth $\U$, and in turn
use the effective coupling constants to compute interaction energies.
As the renormalization procedure $G_X\to G_X^{\t{lattice}}$ is
identical for all coupling constants, we henceforth drop the subscript
$X\in\set{\g\g,\e\g^-,\e\g^+,\e\e}$ on coupling constants $G_X$ in the
remainder of this Appendix.  To further simplify notation, we will
also neglect the explicit dependence of parameters on the lattice
depth $\U$, which we generally keep fixed.

Proper calculations of the interaction energy of two ultracold
fermions in an optical lattice were performed in
refs.~\cite{buchler2010microscopic} and \cite{wall2013strongly} using
a two-channel model of a Feshbach resonance, yielding prescriptions
for computing effective coupling constants in a lattice from
free-space interaction parameters.  These calculations, however, are
both analytically and numerically involved.  We therefore instead opt
to use a modified version of the considerably simpler single-channel
calculation in ref.~\cite{busch1998two} of the interaction energy of
two ultracold atoms in a harmonic trap.  Our approach is equivalent to
the calculation of Hubbard parameters performed in
ref.~\cite{dickerscheid2005feshbach}, and has been demonstrated to
reproduce correct results in the limit of a deep lattice (compared to
the lattice photon recoil energy) and small positive scattering
lengths (compared to the effective harmonic oscillator length)
\cite{buchler2010microscopic}.

The exact result in Eq. 16 of ref.~\cite{busch1998two} for the
interaction energy of two ultracold atoms in a harmonic oscillator
with angular trap frequency $\omega$ can be written in the form
\begin{align}
  \p{G_{\t{free}} K_{\t{HO}}/\omega}^{-1}
  = \f{\sqrt{\pi}~ \Gamma\p{-G_{\t{HO}}K_{\t{HO}}/2\omega}}
  {\Gamma\p{-G_{\t{HO}}K_{\t{HO}}/2\omega-1/2}},
  &&
  K_{\t{HO}} \equiv \int d^3x~ \abs{\phi_0^{\t{HO}}}^4,
  \label{eq:harmonic_gamma}
\end{align}
where $G_{\t{HO}}$ is an effective coupling constant in the harmonic
trap, $\phi_0^{\t{HO}}\p{x}$ is the corresponding single-particle
ground-state wavefunction, and $\Gamma$ is the gamma function.  The
expression in \eqref{eq:harmonic_gamma} can be solved numerically as
is, or expanded about $G_{\t{HO}}K_{\t{HO}}/\omega=0$ to get
\begin{align}
  G_{\t{free}}^{-1} = G_{\t{HO}}^{-1}
  \sum_{n=0}^\infty c_n \p{G_{\t{HO}}K_{\t{HO}}/\omega}^n,
  \label{eq:harmonic_series}
\end{align}
where the first few coefficients are
\begin{align}
  c_0 = 1,
  &&
  c_1 = 1 - \ln 2,
  &&
  c_2 = -\f{\pi^2}{24} - \ln 2 + \f12 \p{\ln 2}^2.
\end{align}
The series in \eqref{eq:harmonic_series} can in turn be inverted to
solve for $G_{\t{HO}}$ with an expansion of the form
\begin{align}
  G_{\t{HO}} = G_{\t{free}}
  \sum_{n=0}^\infty \tilde c_n \p{G_{\t{free}}K_{\t{HO}}/\omega}^n,
  \label{eq:inverted_series}
\end{align}
where if we truncate the series in \eqref{eq:harmonic_series} at
$n=2$, the first few coefficients of \eqref{eq:inverted_series} are
\begin{align}
  \tilde c_0 = 1,
  &&
  \tilde c_1 = 1 - \ln 2,
  &&
  \tilde c_2 = -\f{\pi^2}{24} - \ln 2 + \f12 \p{\ln 2}^2
  + \p{1 - \ln 2}^2.
  \label{eq:inverted_coefficients}
\end{align}
The coefficients $\tilde c_n$ thus found are consistent with the
coefficients $c_2^{(n+1)}$ reported in table 1 of
ref.~\cite{johnson2012effective}, in which the authors compute the
first few terms of the two-body Hamiltonian $H_2$ directly as
expressed in \eqref{eq:H_2_expansion} by using a renormalization
scheme which subtracts off divergences term by term.

All of the above results are exact for two atoms in a harmonic
oscillator interacting via $s$-wave scattering.  In order to adapt
these results for a lattice, we expand the lattice potential about a
lattice site centered at $x=(0,0,0)$ as
\begin{align}
  \U\sin^2\p{k_{\t{L}} \cdot x}
  \approx \U \sp{\p{k_{\t{L}}^\x k_{\t{L}}^\y k_{\t{L}}^\z}^{1/3} x}^2
  = \f12 \mA \omega_{\t{eff}}^2 x^2,
  &&
  \omega_{\t{eff}} \equiv \sqrt{2~ \U k_{\t{L}}^2 / \mA},
\end{align}
where $k_{\t{L}}=\p{k_{\t{L}}^\x,k_{\t{L}}^\y,k_{\t{L}}^\z}$ is the
lattice wavenumber, $\mA$ is the atomic mass, and $\omega_{\t{eff}}$
is an effective angular harmonic trap frequency.  We then use
$\omega_{\t{eff}}$ in place of $\omega$ in \eqref{eq:harmonic_gamma},
and use an overlap integral $K$ computed with the ground-state
wavefunctions $\phi_0$ in a lattice rather than those in a harmonic
oscillator.  We retrieve free-space $s$-wave scattering lengths
$a_{\t{free}}$ for ${}^{87}$Sr from ref.~\cite{zhang2014spectroscopic}
to determine the free-space coupling constants
$G_{\t{free}} \equiv \p{4\pi/\mA} a_{\t{free}}$.  This procedure
yields an effective coupling constant $G_{\t{lattice}}$ given by
\begin{align}
  \p{G_{\t{free}} K/\omega_{\t{eff}}}^{-1}
  = \f{\sqrt{\pi}~ \Gamma\p{-G_{\t{lattice}}K/2\omega_{\t{eff}}}}
  {\Gamma\p{-G_{\t{lattice}}K/2\omega_{\t{eff}}-1/2}},
\end{align}
with a solution
\begin{align}
  G_{\t{lattice}} = G_{\t{free}}
  \sum_{n=0}^\infty \tilde c_n \p{G_{\t{free}}K/\omega_{\t{eff}}}^n,
\end{align}
where the first few coefficients are provided in
\eqref{eq:inverted_coefficients}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Momentum-dependent $s$-wave interactions}
\label{sec:momentum_dependence}

In addition to the renormalization of coupling constants discussed in
Appendix \ref{sec:renormalization}, computing two-body interaction
energies $E_{NX}^{(2)}$ at third order in the low-energy effective
theory requires accounting for the contribution of momentum-dependent
$s$-wave interactions.  At next-to-leading order in the relative
momentum $k$ between two atoms, the effective momentum-dependent
scattering length $a_{\t{eff}}$ is given in terms of the zero-momentum
scattering length $a$ by \cite{blume2002fermi, chin2010feshbach}
\begin{align}
  \f1{a_{\t{eff}}}
  = \f1a - \f12 r_{\t{eff}} k^2
  = \f1a\p{1 - \f12 r_{\t{eff}} a k^2},
\end{align}
which for $r_{\t{eff}} a k^2 \ll 1$, implies that
\begin{align}
  a_{\t{eff}}
  \approx a\p{1 + \f12 r_{\t{eff}} a k^2}
  = a + \f12 r_{\t{eff}} a^2 k^2.
  \label{eq:a_eff}
\end{align}
Here $r_{\t{eff}}$ is an effective range of $\O\p{k^2}$ interactions,
determined in atomic units by the scattering length $a$ and van der
Waals $C_6$ coefficient by \cite{flambaum1999analytical}
\begin{align}
  r_{\t{eff}} = \f13 \xi^{-2} \chi \p{1 - 2\chi + 2\chi^2} a,
  &&
  \t{where}
  &&
  \xi \equiv \f{\Gamma\p{3/4}}{\Gamma\p{1/4}},
  &&
  \chi \equiv \sqrt{2}~ \xi~ \f{\p{\mA C_6}^{1/4}}{a},
  \label{eq:effective_range}
\end{align}
and $\Gamma$ is the gamma function.  As $\chi\sim1$ for ${}^{87}$Sr,
the momentum-dependent correction to the effective scattering length
$a_{\t{eff}}$ is $\O\p{a^3}$ without an additional separation of
scales (i.e.~which could have occurred if we had $\chi\ll1$ or
$\chi\gg1$).  The momentum-independent $\O\p{k^0}$ contribution to
$a_{\t{eff}}$ in \eqref{eq:a_eff} gives rise to the bare two-body
interactions in \eqref{eq:H_int} by use of an unregularized contact
(i.e.~$\delta$-function) potential, while the momentum-dependent
$\O\p{k^2}$ term gives rise to the interaction Hamiltonian
\cite{blume2002fermi, johnson2012effective}
\begin{align}
  H_{\t{int}}'
  \equiv \f12 \sum {G'}^{qr}_{st} \int d^3 x~d^3y~\delta\p{z}
  \sp{\hat\psi_{\mu q}^\dag\p{x} \hat\psi_{\nu r}^\dag\p{y}}
  \hat k_z^2\sp{\hat\psi_{\nu t}\p{y} \hat\psi_{\mu s}\p{x}},
  \label{eq:H_int_primed}
\end{align}
where
\begin{align}
  z \equiv x - y,
  &&
  \hat k_z^2
  \equiv -\f12 \p{\vec\nabla_z^2 + \lvec\nabla_z^2},
\end{align}
and the primed couplings ${G'}^{qr}_{st}$ are defined similarly to
unprimed couplings $G^{qr}_{st}$ in \eqref{eq:couplings} and
\eqref{eq:coupling_tensor}, but with scattering lengths
$a\to r_{\t{eff}}a^2/2$ and the effective range $r_{\t{eff}}$ defined
by \eqref{eq:effective_range} for each scattering length with an
appropriate $C_6$ coefficient.  We retrieve $C_6$ coefficients for
${}^{87}$Sr from the supplementary material of
ref.~\cite{zhang2014spectroscopic}.  The squared relative momentum
operator $k_z^2$ is represented by symmetrized left- and right-acting
derivative operators in order to preserve manifest Hermiticity of
$H_{\t{int}}'$.  At third order in the low-energy effective theory
developed in section \ref{sec:low_energy}, the bare momentum-dependent
interactions in \eqref{eq:H_2} yield only the effective two-body
Hamiltonian
\begin{align}
  H_2' \equiv \f12 K' \sum {G'}^{qr}_{st}
  \c_{\mu s}^\dag \c_{\nu t}^\dag \c_{\nu r} \c_{\mu q},
  \label{eq:H_2_prime}
\end{align}
where, letting $\Re\sp{X}$ denote the real part of $X$,
\begin{align}
  K' \equiv \f12 \int d^3x~
  \Re\sp{\p{\phi_0^*}^2
    \p{\vec\nabla\phi_0\cdot\vec\nabla\phi_0
      - \phi_0\vec\nabla^2\phi_0}}.
\end{align}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Bounds on theoretical uncertainties from inter-site effects}
\label{sec:error}

In our overview of the relevant one- and two-particle physics of
ultracold atoms on a lattice (section \ref{sec:overview}), we made two
approximations which introduce error into the low-energy effective
theory.  Both approximations concern the on-site locality of the
single- and two-body Hamiltonians: we assumed that
\begin{enumerate*}
\item tunneling between lattice sites and
\item inter-site interactions are negligible.
\end{enumerate*}
These approximations are justified for single-particle motional ground
states of atoms in a deep lattice, but generally break down when
considering virtual states occupying highly excited motional levels,
whose spatial wavefunctions can span multiple lattice sites.
Nonetheless, we can place upper bounds on the magnitude of inter-site
corrections to the effective on-site interaction Hamiltonians by
treating tunneling and inter-site interactions of virtual excited
states perturbatively and assuming no energetic penalty for
nearest-neighbor hopping.  These bounds can be used to diagnose the
breakdown of the on-site effective theory, and signal when a more
careful consideration of inter-site effects is necessary to make
precise predictions about many-body spectra and dynamics.

If we still assume negligible overlap between single-particle
ground-state wavefunctions in different lattice sites but consider
nearest-neighbor wavefunction overlaps of states with motional
excitations, our one-body and bare two-body Hamiltonians become
\begin{align}
  H_0 = \sum E_n \c_{in\mu s}^\dag \c_{in\mu s}
  - \sum_{\substack{\bk{i,j}\\m,n>0}}
  \p{t_{mn} \c_{jn\mu s}^\dag \c_{im\mu s} + \t{h.c.}},
  \label{eq:H_0_neighbor}
\end{align}
and
\begin{align}
  H_{\t{int}} &= \f12 \sum K^{k\ell}_{mn} G^{qr}_{st}
  \c_{im\mu s}^\dag \c_{in\nu t}^\dag \c_{i\ell\nu r} \c_{ik\mu q}
  + \f12 \sum_{\substack{\bk{i,j}\\n>0}} G^{\mu q;\nu r}_{\rho s;\sigma t}
  \p{\K_n \c_{j,0,\rho s}^\dag \c_{j,0,\sigma t}^\dag
    \c_{j,0,\nu r} \c_{in\mu q} + \t{h.c.}} \nonumber \\
  &\quad + \f12 \sum_{\substack{\bk{i,j}\\m,n>0}}
  G^{\mu q;\nu r}_{\rho s;\sigma t}
  \p{\K_{mn} \c_{j,0,\rho s}^\dag \c_{j,0,\sigma t}^\dag
    \c_{in\nu r} \c_{im\mu q} +
    \tilde\K_{mn} \c_{in\rho s}^\dag \c_{j,0,\sigma t}^\dag
    \c_{j,0,\nu r} \c_{im\mu q} + \t{h.c.}},
  \label{eq:H_int_neighbor}
\end{align}
where $t_{mn}$ is a tunneling rate; $\K_n,\K_{mn},\tilde\K_{mn}$ are
inter-site spatial overlap integrals; $\t{h.c.}$ denotes a Hermitian
conjugate, i.e.~$\p{X+\t{h.c.}}\equiv \p{X+X^\dag}$; and $\bk{i,j}$
denotes the set of all lattice sites $i$ together with their adjacent
sites $j$.  Note that we have neglected terms in
\eqref{eq:H_int_neighbor} which involve more than two field operators
addressing states with motional excitations, as these terms will not
appear in the leading-order corrections to the effective on-site
interaction Hamiltonians.  We also still neglect terms which involve
products of atomic wavefunctions for motional ground states in
different lattice sites.

Diagrammatically representing matrix elements of $H_0$ and
$H_{\t{int}}$ which are off-diagonal in lattice site by a dot
(i.e.~$\bullet$) and marking lines which represent field operators
addressing neighboring lattice sites by a cross (i.e.~$+$ or $\times$,
depending on the line orientation), the dominant terms in the
effective theory which we previously neglected by assuming on-site
locality are
\begin{align}
  \begin{tikzpicture}
    \begin{feynman}
      \vertex (v1);
      \vertex[above left = of v1] (f1);
      \vertex[below left = of v1] (f2);
      \vertex[right = of v1] (v2);
      \vertex[above right = of v2] (f3);
      \vertex[below right = of v2] (f4);
      \diagram* {
        (f1) -- (v1),
        (f2) -- (v1),
        (v1) --[scalar, half left] (v2),
        (v1) --[scalar, half right] (v2),
        (v2) --[insertion = 0.5] (f3),
        (v2) --[insertion = 0.5] (f4) };
      \draw[fill=black] (v2) circle(0.8mm);
    \end{feynman}
  \end{tikzpicture}
  &\sim \gamma_2^{(2)} G^2,
  & \gamma_2^{(2)} &\equiv \sum_{n+m>0} \f{K_{mn}\K_{mn}}{E_{mn}},
  \label{eq:g_2_2} \\[1em]
  \begin{tikzpicture}
    \begin{feynman}
      \vertex (v1);
      \vertex[above left = of v1] (f1);
      \vertex[below left = of v1] (f2);
      \vertex[right = 4em of v1] (v2);
      \vertex[above right = of v1] (f3);
      \vertex[below left = of v2] (f4);
      \vertex[below right = of v2] (f5);
      \vertex[above right = of v2] (f6);
      \diagram* {
        (f1) -- (v1) -- (f3),
        (f2) -- (v1) --[scalar] (v2),
        (f4) --[insertion = 0.5] (v2) --[insertion = 0.5] (f5),
        (v2) --[insertion = 0.5] (f6), };
      \draw[fill=black] (v2) circle(0.8mm);
    \end{feynman}
  \end{tikzpicture}
  &\sim \gamma_{3,1}^{(2)} G^2,
  & \gamma_{3,1}^{(2)} &\equiv \sum_{n>0} \f{K_n\K_n}{E_n},
  \label{eq:g_3_1_2} \\[1em]
  \begin{tikzpicture}
    \begin{feynman}
      \vertex (v1);
      \vertex[above left = of v1] (f1);
      \vertex[below left = of v1] (f2);
      \vertex[right = 2em of v1] (v2);
      \vertex[right = 2em of v2] (v3);
      \vertex[above right = of v1] (f3);
      \vertex[below left = of v3] (f4);
      \vertex[below right = of v3] (f5);
      \vertex[above right = of v3] (f6);
      \diagram* {
        (f1) -- (v1) -- (f3),
        (f2) -- (v1) --[scalar] (v2) --[scalar, insertion = 0.5] (v3),
        (f4) --[insertion = 0.5] (v3) --[insertion = 0.5] (f5),
        (v3) --[insertion = 0.5] (f6), };
      \draw[fill=black] (v2) circle(0.8mm);
    \end{feynman}
  \end{tikzpicture}
  &\sim \gamma_{3,2}^{(2)} G^2,
  & \gamma_{3,2}^{(2)} &\equiv \sum_{n,m>0} \f{K_mt_{mn}K_n}{E_mE_n},
  \label{eq:g_3_2_2}
\end{align}
where we have identified, up to an assignment of coupling constants
$G$, the magnitude of all nonzero matrix elements of the diagrams with
respect to an eigenbasis of the on-site single-particle Hamiltonian
$H_0$ in \eqref{eq:H_0}.

The terms in \eqref{eq:g_2_2}-\eqref{eq:g_3_2_2} can be used to
estimate an upper bound on the magnitude of dominant corrections to
the spectrum of the low-energy theory from off-diagonal (i.e.~in
lattice site) matrix elements of the Hamiltonians in
\eqref{eq:H_0_neighbor} and \eqref{eq:H_int_neighbor}.  Conservatively
assuming no energetic penalty and no Pauli blocking for any inter-site
process, the dominant correction $\delta E_N$ to the interaction
energy of a lattice site with $N$ atoms and $b$ neighboring sites
(e.g.~$b=6$ in a primitive cubic lattice) is roughly bounded as
\begin{align}
  \abs{\delta E_N} \lesssim b {N\choose 2}
  \max\set{\abs{\gamma_2^{(2)}},
    \p{N-1} \abs{\gamma_{3,1}^{(2)} + \gamma_{3,2}^{(2)}}} G^2,
  \label{eq:delta_E_N}
\end{align}
where the factor of $b$ accounts for the multiplicity of neighboring
sites; the factor of ${N\choose2}$ accounts for the number of on-site
pairs of atoms which are addressed by the diagrams in
\eqref{eq:g_2_2}-\eqref{eq:g_3_2_2}; and the factor of $N-1$ on
$\gamma_{3,X}^{(2)}$ accounts for the number of atoms in a neighboring
site which are addressed by the corresponding processes.  These
factors count the number of matrix elements in the Hamiltonian with
magnitude $\sim\gamma_X^{(2)}G^2$.  The maximization in
\eqref{eq:delta_E_N} is performed because the relevant two- and
three-body processes are mutually exclusive, requiring a different
number of atoms on neighboring lattice sites.  For a conservative
bound of $\abs{\delta E_N}$, the coupling factor $G^2$ in
\eqref{eq:delta_E_N} can simply be maximized over its allowed values
for a given state of atoms on a lattice site, e.g.~$G_\g^2$ for a
state with no orbital excitations, or $\max\set{G_\g^2,G_+^2,G_-^2}$
for a state with one net orbital excitation (in both cases, assuming
no orbital excitations in neighboring sites).  In the latter case, the
bound in \eqref{eq:delta_E_N} can also be reduced by observing that to
conserve energy, it must be the excited atom which moves to a
neighboring site, which reduces the factor of ${N\choose2}$ in down to
$N-1$.  We emphasize that the bound in \eqref{eq:delta_E_N} is by no
means an exact measure of error, and is merely intended to provide a
conservative range of energies and corresponding time scales for which
inter-site effects could become relevant despite negligible
single-particle ground-state tunneling rates.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Low-excitation $M$-body Hamiltonian coefficients}
\label{sec:U_X}

The $M$-body Hamiltonians addressing low-lying orbital state
excitations can be written in the form
\begin{align}
  H_M = \sum_{\abs{\set{\mu_j}}=M}
  \p{U_{M,\g} \prod_{\alpha=1}^M \n_{\mu_\alpha,\g}
    + U_{M,+} \n_{\mu_1,\e} \prod_{\alpha=2}^M \n_{\mu_\alpha,\g}
    + U_{M,-} \c_{\mu_1,\g}^\dag \c_{\mu_2,\e}^\dag
    \c_{\mu_2,\g} \c_{\mu_1,\e} \prod_{\alpha=3}^M \n_{\mu_\alpha,\g}},
\end{align}
where the coefficients can be expanded as
$U_{MX} = \sum_p U_{MX}^{(p)}$ with terms $U_{MX}^{(p)}$ at order $p$
in the coupling constants $G_Y$.  The terms $U_{MX}^{(p)}$ can be
determined from the $M$-body $p$-order Hamiltonians $H_M^{(p)}$
derived in section \ref{sec:low_energy}, i.e.~in \eqref{eq:H_2},
\eqref{eq:H_3_2}, \eqref{eq:H_3_3}, and \eqref{eq:H_4_3}.  For the
effective 2-, 3-, and 4-body Hamiltonians through third order in the
coupling constants, the coefficients are
\begin{align}
  U_{2,\g} = \f12 \alpha_2^{(1)} G_\g,
  &&
  U_{2,+} = \alpha_2^{(1)} G_+,
  &&
  U_{2,-} = \alpha_2^{(1)} G_-,
\end{align}
\begin{align}
  U_{3,\g}^{(2)} = - \alpha_3^{(2)} G_\g^2,
  &&
  U_{3,+}^{(2)} = - \alpha_3^{(2)} G_+ \p{G_+ + 2 G_\g},
\end{align}
\begin{align}
  U_{3,-}^{(2)} = - \alpha_3^{(2)} G_- \p{2 G_+ + G_- + 2 G_\g},
\end{align}
\begin{align}
  U_{3,\g}^{(3)}
  = \p{\alpha_3^{(3)} - \alpha_5^{(3)}} 2 G_\g^3
  - \p{\alpha_{4,3}^{(3)} + \alpha_5^{(3)}} G_\g^3,
\end{align}
\begin{align}
  U_{3,+}^{(3)}
  &= \p{\alpha_3^{(3)} - \alpha_5^{(3)}} \p{G_+^3 + 4 G_+^2 G_\g
    + G_+ G_-^2 + G_+ G_\g^2 + G_-^3 + G_-^2 G_\g} \nonumber \\
  &\quad - \p{\alpha_{4,3}^{(3)} + \alpha_5^{(3)}}
  \p{G_+^3 + G_+^2 G_\g + G_+ G_-^2 + G_+ G_\g^2 + G_-^2 G_\g},
\end{align}
\begin{align}
  U_{3,-}^{(3)}
  &= \p{\alpha_3^{(3)} - \alpha_5^{(3)}} G_- \p{3 G_+^2
    + 2 G_+ G_- + 8 G_+ G_\g + 3 G_- G_\g + G_\g^2} \nonumber \\
  &\quad - \p{\alpha_{4,3}^{(3)} + \alpha_5^{(3)}}
  G_- \p{3 G_+^2 + 2 G_+ G_- + 2 G_+ G_\g + G_-^2 + G_\g^2},
\end{align}
\begin{align}
  U_{4,\g}^{(3)}
  = \p{\alpha_{4,1}^{(3)} - \alpha_5^{(3)}} G_\g^3
  + \p{\alpha_{4,2}^{(3)} - \alpha_5^{(3)}} 2 G_\g^3,
\end{align}
\begin{align}
  U_{4,+}^{(3)}
  &= \p{\alpha_{4,1}^{(3)} - \alpha_5^{(3)}}
  2 G_+ G_\g \p{G_+ + G_\g} \nonumber \\
  &\quad + \p{\alpha_{4,2}^{(3)} - \alpha_5^{(3)}}
  G_+ \p{G_+^2 + 2 G_+ G_\g + 5 G_\g^2},
\end{align}
\begin{align}
  U_{4,-}^{(3)}
  &= \p{\alpha_{4,1}^{(3)} - \alpha_5^{(3)}}
  2 G_- G_\g \p{2 G_+ + G_- + G_\g} \nonumber \\
  &\quad + \p{\alpha_{4,2}^{(3)} - \alpha_5^{(3)}}
  G_- \p{3 G_+^2 + 3 G_+ G_- + 4 G_+ G_\g + G_-^2
    + 2 G_- G_\g + 5 G_\g^2},
\end{align}
In terms of the spatial overlap integrals defined in \eqref{eq:K_klmn}
and \eqref{eq:K}, the prefactors $\alpha_X^{(p)}$ on the coefficients
$U_X^{(p)}$ are
\begin{align}
  \alpha_2^{(1)} \equiv K,
  &&
  \alpha_3^{(2)} \equiv \sum_{n>0} \f{K_n^2}{E_n},
  &&
  \alpha_3^{(3)} \equiv \sum_{\substack{\ell+m>0\\\ell+n>0}}
  \f{K_{\ell m} K^m_n K_{\ell n}}{E_{\ell m} E_{\ell n}},
\end{align}
\begin{align}
  \alpha_{4,1}^{(3)}
  \equiv \sum_{\substack{m\ge0\\n>0}} \f{K_{mn} K_m K_n}{E_{mn} E_n},
  &&
  \alpha_{4,2}^{(3)}
  \equiv \sum_{m,n>0} \f{K_m K^m_n K_n}{E_m E_n},
  &&
  \alpha_{4,3}^{(3)}
  \equiv K \sum_{m+n>0} \f{K_{mn}^2}{E_{mn}^2},
\end{align}
\begin{align}
  \alpha_5^{(3)}
  \equiv  K \sum_{n>0} \f{K_n^2}{E_n^2}.
\end{align}

\bibliography{\jobname}

\end{document}