### 2019-08-28 soc_squeezing/squeezing_paper/prx/reviews/revision_summary.tex

\documentclass[reprint,onecolumn,12pt]{revtex4-2}

%%% linking references
\usepackage{hyperref}
\hypersetup{
  colorlinks=true,
  linkcolor=blue,
  urlcolor=cyan,
}

%%% physics and math packages
\usepackage{physics} % general physics package
\usepackage{amssymb} % math fonts and symbols
\usepackage{bm} % for making math symbols bold
\usepackage{braket} % for nice brackets (e.g. |X>)
\usepackage{mathtools} % for \coloneqq
\usepackage{graphicx} % for figures
\graphicspath{{./figures/}} % set path for all figures

%%% shorthands to use inside math environments
\renewcommand{\t}{\text} % text in math mode
\newcommand{\f}[2]{\dfrac{#1}{#2}} % shorthand for fractions
\newcommand{\p}[1]{\left(#1\right)} % parenthesis
\renewcommand{\sp}[1]{\left[#1\right]} % square parenthesis
\renewcommand{\set}[1]{\left\{#1\right\}} % square parenthesis
\renewcommand{\c}{\cdot} % inner product
\renewcommand{\d}{\text{d}} % integration measure, e.g. d^3 x
\renewcommand{\v}{\bm} % bold symbols (e.g. for vectors)
\newcommand{\uv}[1]{\hat{\bm #1}} % unit vectors, or vector operators
\newcommand{\bk}{\Braket}
\renewcommand{\ket}{\Ket}
\renewcommand{\bra}{\Bra}

\newcommand{\B}{\mathcal{B}}
\newcommand{\E}{\mathcal{E}}
\newcommand{\G}{\mathcal{G}}
\newcommand{\I}{\mathcal{I}}
\newcommand{\J}{\mathcal{J}}
\renewcommand{\L}{\mathcal{L}}
\renewcommand{\O}{\mathcal{O}}
\renewcommand{\P}{\mathcal{P}}
\newcommand{\Q}{\mathcal{Q}}
\renewcommand{\S}{\mathcal{S}}

\newcommand{\z}{\text{z}}
\newcommand{\x}{\text{x}}
\newcommand{\y}{\text{y}}

\newcommand{\up}{\uparrow}
\newcommand{\dn}{\downarrow}

\usepackage{enumitem}
\setlist[enumerate]{leftmargin=*}
\usepackage[dvipsnames]{xcolor}
\newcommand{\blue}[1]{{\color{blue} #1}}
\newcommand{\red}[1]{{\color{red} #1}}
\newcommand{\green}[1]{{\color{ForestGreen} #1}}

% for numbering points in response to referees
\newcounter{point}
\newcommand{\step}{\stepcounter{point}\setcounter{enumi}{0}}

\begin{document}

Dear editors,

In this document, we summarize all revisions made to our manuscript
since its initial submission.  Revisions made in our responses to the
referee are numbered R.$X$.$Y$, corresponding to the $Y$-th revision
in point $X$ in our response.  In addition, we summarize additional
revisions made to refine wording or fix errors in the original
manuscript; these revisions are numbered M.$Z$.

Excerpts from our original manuscript are written in \red{red}, and
excerpts from the revised version of our manuscript is written in
\green{green}.  All page, equation, and reference numbers generally
refer to those in our original manuscript, unless explicitly stated
otherwise.

\section*{Revisions in response to Referee}

\begin{enumerate}[label=(R.\arabic{point}.\arabic*)]
  \step \step
\item

  \step
\item

  \step
\item

  \step
\item

  \step
\item

\end{enumerate}

\section*{Miscellaneous revisions}

\begin{enumerate}[label=(M.\arabic*)]
\item To standardize our use of language, we have changed all
  instances of the words ``\red{interrogation laser}'' and
  ``\red{external interrogation laser}'' to ``\green{clock laser}'',
  as well as all instances of ``\red{SOC phase}'' ($\phi$) with
  ``\green{SOC angle}''.  For consistency, we have also prepended
  ``\green{Eqn.}'' or ``\green{Eqns.}'' as appropriate to all equation
  references.

\item To clarify the assumptions of the spin model that become invalid
  when $U/J\lesssim2$, we have changed the text (bottom of the left
  column on page 4):

  \red{... at $U/J\gtrsim2$ mode-changing collisions start to
    become relevant and invalidate assumptions of the spin model.}

  to:

  \green{... at $U/J\gtrsim2$ mode-changing collisions start to become
    relevant and invalidate the frozen-mode assumption of the spin
    model.}

\item We fixed a typo at the top of the right column on page 4,
  replacing the text:

  \red{... our protocol should work even in this highly dopped case;
    ...}

  by:

  \green{... our protocol should work even in this highly hole-doped
    case; ...}

\item We found an error in our summaries of the results in Figure
  4(b); to fix this error, we have changed the following text (both in
  the abstract and paragraph 1 on page 7):

  \red{... $\sim12-14$ dB of spin squeezing in $\sim1$ second with
    $10^2$--$10^3$ atoms ...}

  by:

  \green{... $\sim10-14$ dB of spin squeezing in $\sim1$ second with
    $10^2$--$10^4$ atoms ...}

\item In Appendix D (page 10), we have fixed an error in identifying
  the set jump operators corresponding to single-spin decay, replacing:

  \red{... sets of jump operators $\J_-\equiv s_-^{(j)}$ and
    $\J_\z\equiv\set{s_\z^{(j)}}$...}

  by:

  \green{... sets of jump operators $\J_-\equiv\set{s_-^{(j)}}$ and
    $\J_\z\equiv\set{s_\z^{(j)}}$...}

\end{enumerate}

\end{document}

% popular summary -- beginning of second paragraph -- added word "we"