### 2019-09-11 soc_squeezing/collective_paper/linear_response.tex

\section{Short-time linear response and two-time correlators}
\label{sec:linear_response}

Here we discuss the appearance of two-time correlation functions in
the short-time linear response of correlators to perturbations of a
Hamiltonian.  Consider an initial Hamiltonian $H$ perturbed by an
operator $V$ with $\norm{V}\ll\norm{H}$, where $\norm{\O}$ denotes the
operator norm of $\O$, such that the net Hamiltonian is
$\tilde H=H+V$.  We denote the generator of Heisenberg time evolution
under the perturbed (unperturbed) Hamiltonian by $\tilde T$ ($T$).
These generators are related by
\begin{align}
  \tilde T = T + i\ul V
\end{align}
where $\ul V$ is a superoperator whose action on operators $\O$ is
defined by
\begin{align}
  \ul V \O \equiv \sp{V,\O}_-.
\end{align}
Through quadratic order in the time $t$ and linear order in the
perturbation $\ul V$, we can say that
\begin{align}
  e^{t\tilde T}
  \approx \f12\sp{e^{tT}, e^{it\ul V}}_+
  \approx e^{tT} + \f12 it \sp{e^{t T}, \ul V}_+.
\end{align}
Defining perturbed and unperturbed Heisenberg operators
$\tilde\O\p{t}$ and $\O\p{t}$ that respectively satisfy
$\bk{\tilde\O\p{t}}=\bk{e^{t\tilde T}\O}$ and
$\bk{\O\p{t}}=\bk{e^{tT}\O}$, we thus find that for sufficiently small
times $t$ and weak perturbations $V$,
\begin{align}
  \bk{\tilde\O\p{t} - \O\p{t}}
  = \Bk{\p{e^{t\tilde T} - e^{tT}} \O}
  \approx \f12 i t \p{\bk{\sp{V,\O}_-\p{t}}
    + \bk{\sp{V,\O\p{t}}_-}}.
  \label{eq:response}
\end{align}
Two-time correlators $\bk{V\O\p{t}}$ and $\bk{\O\p{t}V}$, in addition
to the expectation values $\bk{\p{V\O}\p{t}}$ and
$\bk{\p{\O V}\p{t}}$, thus determine the short-time linear response of
correlators $\bk{\O\p{t}}$ to perturbations $V$ of a Hamiltonian.