### 2019-09-11 soc_squeezing/collective_paper/calculations_OAT.tex
simplified analytical expressions for OAT correlators

\section{Analytical results for the one-axis twisting model}
\label{sec:OAT}

The one-axis twisting (OAT) Hamiltonian for $N$ spin-1/2 particles
takes the form
\begin{align}
  H_{\t{OAT}}
  = \chi S_\z^2
  = \f12 \chi \sum_{j<k} \sigma_\z^{(j)} \sigma_\z^{(k)} + \f14 N \chi,
\end{align}
where $\sigma_\z^{(j)}$ represents a Pauli-$z$ operator acting on spin
$j$.  This model is a special case of the zero-field Ising Hamiltonian
previously solved in Ref.~[\citenum{foss-feig2013nonequilibrium}] via
exact, analytical treatment of the quantum trajectory Monte Carlo
method for computing expectation values.  The solution therein
accounts for coherent evolution in addition to decoherence via
uncorrelated single-spin decay, excitation, and dephasing respectively
at rates $\gamma_-$, $\gamma_+$, and $\gamma_\z$ (denoted by
$\Gamma_{\t{ud}}$, $\Gamma_{\t{du}}$, and $\Gamma_{\t{el}}$ in
Ref.~[\citenum{foss-feig2013nonequilibrium}]).  Letting $S\equiv N/2$
and $\mu,\nu\in\set{+1,-1}$, we adapt expectation values computed in
Ref.~[\citenum{foss-feig2013nonequilibrium}] for the initial state
$\ket\X\propto\p{\ket\up+\ket\dn}^{\otimes N}$ with
$S_\x\ket\X=S\ket\X$ evolving under $H_{\t{OAT}}$, finding
\begin{align}
  \bk{S_+\p{t}}
  &= S e^{-\kappa t} \Phi\p{\chi,t}^{N-1},
  \label{eq:S+_OAT} \\
  \bk{\p{S_\mu S_\z}\p{t}}
  &= -\f{\mu}{2}\bk{S_\mu\p{t}}
  + S \p{S-\f12} e^{-\kappa t}
  \Psi\p{\mu\chi,t} \Phi\p{\chi,t}^{N-2}, \\
  \bk{\p{S_\mu S_\nu}\p{t}}
  &= \p{1-\delta_{\mu\nu}} \p{S + \mu\bk{S_\z\p{t}}}
  + S \p{S-\f12} e^{-2\kappa t} \Phi\p{\sp{\mu+\nu}\chi,t}^{N-2},
  \label{eq:SS+-_OAT}
\end{align}
where
\begin{align}
  \Phi\p{X,t}
  \equiv e^{-\lambda t}
  \sp{\cos\p{\omega_X t} + \lambda t~\t{sinc}\p{\omega_X t}},
  &&
  \Psi\p{X,t}
  \equiv e^{-\lambda t}\p{\Delta+iX}t~\t{sinc}\p{\omega_X t},
\end{align}
for
\begin{align}
  \lambda \equiv \f12 \p{\gamma_+ + \gamma_-},
  &&
  \Delta \equiv \gamma_+ - \gamma_-,
  &&
  \kappa \equiv \f12 \p{\gamma_+ + \gamma_- + \gamma_\z},
  &&
  \omega_X \equiv \sqrt{X^2-\lambda^2-iX\Delta}.
\end{align}
In order to compute spin squeezing as measured by the Ramsey squeezing
parameter $\xi^2$ defined in \eqref{eq:squeezing}, we additionally
need analytical expressions for $\bk{S_\z\p{t}}$ and
$\bk{S_\z^2\p{t}}$.  As these operators commute with both the OAT
Hamiltonian and the single-spin operators $\sigma_\z^{(j)}$, their
evolution is governed entirely by decay-type decoherence (see Appendix
\ref{sec:decay_single}), which means
\begin{align}
  \f{d}{dt} \bk{S_\z\p{t}}
  &= S\p{\gamma_+-\gamma_-} - \p{\gamma_++\gamma_-} \bk{S_\z\p{t}},
  \\
  \f{d}{dt} \bk{S_\z^2\p{t}}
  &= S\p{\gamma_++\gamma_-}
  + 2\p{S-\f12}\p{\gamma_+-\gamma_-} \bk{S_\z\p{t}}
  - 2 \p{\gamma_++\gamma_-} \bk{S_\z^2\p{t}}.
\end{align}
The initial conditions $\bk{S_\z\p{0}}=0$ and $\bk{S_\z^2\p{0}}=S/2$
then imply
\begin{align}
  \bk{S_\z\p{t}} = S\p{\f{\gamma_+-\gamma_-}{\gamma_++\gamma_-}}
  \p{1-e^{-\p{\gamma_+ + \gamma_-} t}},
  &&
  \bk{S_\z^2\p{t}}
  = \f12 S + S \p{S-\f12} \p{\f{\bk{S_\z\p{t}}}{S}}^2.
  \label{eq:Sz_OAT}
\end{align}
With appropriate assumptions about the relevant sources of
decoherence, the expectation values in
\eqref{eq:S+_OAT}--\eqref{eq:SS+-_OAT} and \eqref{eq:Sz_OAT} are
sufficient to compute the spin squeezing parameter $\xi^2$ in
\eqref{eq:squeezing} at any time throughout evolution of the initial
state $\ket\X$ under $H_{\t{OAT}}$.