### 2019-09-25 sun_squeezing/sun_notes.tex
added a proof that the collective interaction Hamiltonian takes an identical form in any orthonormal basis of operators added a comment about Weyl operators

\documentclass[nofootinbib,notitlepage,11pt]{revtex4-2}

%%% linking references
\usepackage{hyperref}
\hypersetup{
  breaklinks=true,
  colorlinks=true,
  linkcolor=blue,
  filecolor=magenta,
  urlcolor=cyan,
}

%%% header / footer
\usepackage{fancyhdr} % easier header and footer management
\pagestyle{fancy} % page formatting style
\fancyhf{} % clear all header and footer text
\renewcommand{\headrulewidth}{0pt} % remove horizontal line in header
\usepackage{lastpage} % for referencing last page
\cfoot{\thepage~of \pageref{LastPage}} % "x of y" page labeling

%%% symbols, notations, etc.
\usepackage{physics,braket,bm,amssymb} % physics and math
\renewcommand{\t}{\text} % text in math mode
\newcommand{\f}[2]{\dfrac{#1}{#2}} % shorthand for fractions
\newcommand{\p}[1]{\left(#1\right)} % parenthesis
\renewcommand{\sp}[1]{\left[#1\right]} % square parenthesis
\renewcommand{\set}[1]{\left\{#1\right\}} % curly parenthesis
\newcommand{\bk}{\Braket} % shorthand for braket notation
\renewcommand{\v}{\bm} % bold vectors
\newcommand{\uv}[1]{\bm{\hat{#1}}} % unit vectors
\newcommand{\av}{\vec} % arrow vectors
\renewcommand{\d}{\text{d}} % for infinitesimals
\renewcommand{\c}{\cdot} % inner product

\usepackage{dsfont} % for identity operator
\newcommand{\1}{\mathds{1}}

\newcommand{\x}{\text{x}}
\newcommand{\y}{\text{y}}
\newcommand{\z}{\text{z}}

\newcommand{\B}{\mathcal{B}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\I}{\mathcal{I}}
\newcommand{\N}{\mathcal{N}}
\renewcommand{\O}{\mathcal{O}}
\newcommand{\Q}{\mathcal{Q}}
\newcommand{\T}{\mathcal{T}}
\renewcommand{\S}{\mathcal{S}}

\def\obra#1{\mathinner{({#1}|}}
\def\oket#1{\mathinner{|{#1})}}
\def\obk#1{\mathinner{({#1})}}
\def\oop#1#2{\oket{#1}\!\obra{#2}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\thispagestyle{fancy}

\title{Multi-level generalizations of collective spin dynamics and
  squeezing}%
\author{Michael A. Perlin}%
\date{\today}

\maketitle

\tableofcontents

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Collective SU($n$)-symmetric interactions}

Under the frozen-mode approximation, two-body collective
SU($n$)-symmetric interactions between fermions on a lattice take the
form\cite{perlin2019effective}
\begin{align}
  H_{\t{int}} = \f{u}{2} \sum_{p,q,\mu,\nu}
  \p{c_{p\mu}^\dag c_{q\nu}^\dag c_{q\nu} c_{p\mu}
    + c_{q\mu}^\dag c_{p\nu}^\dag c_{q\nu} c_{p\mu}},
\end{align}
where $u\equiv U/L$ is the two-body on-site interaction energy $U$
divided by the number of lattice sites $L$, and $c_{p\mu}$ is the
annihilation operator for a fermionic mode with quasi-momentum $p$ and
internal state (e.g.~nuclear spin) $\mu$.  Defining the spin operators
$s_{\mu\nu}^{(p)}\equiv c_{p\mu}^\dag c_{p\nu}$ for each
quasi-momentum $p$, we can alternately write
\begin{align}
  H_{\t{int}} = \f{u}{2} \sum_{p,q,\mu,\nu}
  \p{s_{\mu\mu}^{(p)} s_{\nu\nu}^{(q)} - s_{\mu\nu}^{(p)} s_{\nu\mu}^{(q)}}
  = \f{u}{2}\p{N^2 - \sum_{p,q,\mu,\nu} s_{\mu\nu}^{(p)} s_{\nu\mu}^{(q)}},
\end{align}
where $N$ is the total number of particles.  In the absence of
coherence between sectors of different particle number, we can neglect
the $\sim N^2$ term and simply write
\begin{align}
  H_{\t{int}}
  = - \f{u}{2} \sum_{p,q,\mu,\nu} s_{\mu\nu}^{(p)} s_{\nu\mu}^{(q)}.
\end{align}
In the case of SU(2), at this point we would expand the spin operators
$s_{\mu\nu}^{(p)}$ in terms of Pauli operators in order to convert
$H_{\t{int}}$ into an SU(2) spin Hamiltonian.  To generalize this
procedure to the case of SU($n$), we first define a vector
$\v s^{(p)}$ of all spin operators $s_{\mu\nu}^{(p)}$, and write
\begin{align}
  H_{\t{int}}
  = - \f{u}{2} \sum_{p,q} {\v s^{(p)}}^\dag \v s^{(q)}
  = - \f{u}{2} \v\S^\dag \c \v\S,
  &&
  \v\S\equiv \sum_p \v s^{(p)}.
  \label{eq:H_int_spin}
\end{align}
In the following sections, we identify various bases for the space of
operators on the Hilbert space of an $n$-level system.  For brevity,
we denote this space of operators by $\B_n$, and note that $\B_n$ is
itself an $n^2$-dimensional Hilbert space equipped with a trace
(Hilbert-Schmidt) inner product
\begin{align}
  \obk{\O|\Q} \equiv \tr\p{\O^\dag Q},
\end{align}
where $\oket{\Q}$ and $\obra{\O}$ respectively denote vectors in
$\B_n$ and its dual space $\B_n^*$.  Similarly to the set of spin
operators $\set{s_{\mu\nu}}$ on the Hilbert space at each
quasi-momentum, all of the bases we consider will be orthonormal with
respect to the trace inner product.  As we show in Appendix
\ref{sec:changing_bases}, the collective SU($n$)-symmetric interaction
Hamiltonian in \eqref{eq:H_int_spin} takes an identical form in any
orthonormal basis for $\B_n$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Generalized Gell-Mann (GGM) operators}

The most obvious generalization of Pauli operators to the case of
SU($n$) are the {\it generalized Gell-Mann (GGM)
  operators}\cite{hioe1981level, bertlmann2008bloch}, defined for
$j,k,\ell\in\set{0,1,\cdots,n-1}$ with $k<j$ and $\ell>0$ by
\begin{align}
  \lambda_0 \equiv \sqrt{\f{2}{n}}~ \1,
  &&
  \lambda_\ell \equiv \sqrt{\f{2}{\ell\p{\ell+1}}}
  \p{\sum_{n=0}^{\ell-1}\op{n} - \ell \op{\ell}},
  \label{eq:GGM_diag}
\end{align}
\begin{align}
  \lambda_{jk,+} \equiv \op{j}{k} + \op{k}{j},
  &&
  \lambda_{jk,-} \equiv i\p{\op{j}{k} - \op{k}{j}},
  \label{eq:GGM_off_diag}
\end{align}
where $\1$ is the identity operator.  The self-adjoint GGM operators
in \eqref{eq:GGM_diag} and \eqref{eq:GGM_off_diag} provide a complete
basis for the space of operators on the Hilbert space of an $n$-level
system, coincide exactly with the Pauli operators in the case of
$n=2$, and satisfy the orthonormality condition
\begin{align}
  \obk{\lambda_a|\lambda_b} = 2\delta_{ab}.
  \label{eq:GGM_inner}
\end{align}
Denoting a vector of all GGM operators by $\v\lambda$, we can
therefore write (see Appendix \ref{sec:changing_bases})
\begin{align}
  H_{\t{int}}
  = -\f{u}{4} \sum_{p,q} \v\lambda^{(p)} \c \v\lambda^{(q)}
  = -u \v\Lambda \c \v\Lambda,
  &&
  \v\Lambda \equiv \f12 \sum_p \v\lambda^{(p)}.
  \label{eq:H_int_GGM}
\end{align}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Transition operators}

The GGM operators in \eqref{eq:GGM_diag} and \eqref{eq:GGM_off_diag}
provide a convenient operator basis to describe dynamics obeying
SU($n$) symmetry.  An external driving field addressing $n$-level
nuclear spins, however, will generally violate SU($n$) symmetry, and
instead obeys the symmetries of the {\it polarization} or {\it
  transition operators} defined by\cite{kryszewski2006alternative,
  bertlmann2008bloch}
\begin{align}
  T_{JM} \equiv \sqrt{\f{2J+1}{2I+1}} \sum_{\mu,\nu}
  \bk{I\mu;I\nu,JM} \op{\mu}{\nu},
  \label{eq:trans_ops}
\end{align}
where $J\in\set{0,1,\cdots,n-1}$ and $M\in\set{-J,-J+1,J}$ index a
total spin and its projection onto a quantization axis;
$I\equiv\p{n-1}/2$ is the maximal angular momentum of a spin-$n$
system; $\mu,\nu\in\set{-I,-I+1,\cdots,I}$ index projections of an
$n$-level nuclear spin onto a quantization axis; and
$\bk{I\mu,JM;I\nu}$ is a Clebsch-Gordan coefficient.  Similarly to the
GGM operators in \eqref{eq:GGM_diag} and \eqref{eq:GGM_off_diag}, the
transition operators in \eqref{eq:trans_ops} provide a complete basis
for the space of operators on the Hilbert space of an $n$-level
system, and satisfy the orthonormality condition
\begin{align}
  \obk{T_{JM}|T_{J'M'}} = \delta_{JJ'} \delta_{MM'},
\end{align}
which implies
\begin{align}
  H_{\t{int}} = -\f{u}{2} \sum_{p,q} {\v T^{(p)}}^\dag \c \v T^{(q)}
  = -\f{u}{2} \v\T^\dag \c \v\T,
  &&
  \v\T \equiv \v\T\equiv\sum_p\v T^{(p)},
  \label{eq:H_int_trans}
\end{align}
where $\v T$ is a vector of all transition operators in
\eqref{eq:trans_ops}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Drive operators}

Unlike the GGM operators in \eqref{eq:GGM_diag} and
\eqref{eq:GGM_off_diag}, the transition operators in
\eqref{eq:trans_ops} are not self-adjoint.  In order to express
Hamiltonians in terms of self-adjoint operators, we define the {\it
  drive operators}
\begin{align}
  D_{J,0} \equiv \sqrt{2}~ T_{J,0},
  &&
  D_{J\Delta,+} \equiv -\p{T_{J\Delta} + T_{J\Delta}^\dag},
  &&
  D_{J\Delta,-} \equiv i\p{T_{J\Delta} - T_{J\Delta}^\dag},
  \label{eq:drive_ops}
\end{align}
for $\Delta\in\set{1,2,\cdots,J}$.  Similarly to the GGM operators in
\eqref{eq:GGM_diag} and \eqref{eq:GGM_off_diag}, the drive operators
satisfy the orthonormality condition
\begin{align}
  \obk{D_a|D_b} = 2\delta_{ab},
\end{align}
and are similarly equal to the Pauli operators in the case of $n=2$.
These operators have the convenient feature that $D_{1,1,\pm}$ is (up
to a gauge transformation) proportional to the Hamiltonian induced on
a spin-$n$ system by a classical driving laser.  In terms of the drive
operators,
\begin{align}
  H_{\t{int}} = -\f{u}{4} \sum_{p,q} \v D^{(p)}\c\v D^{(q)}
  = -u \v\D \c \v\D,
  &&
  \v\D \equiv \f12 \sum_p \v D^{(p)},
  \label{eq:H_int_drive}
\end{align}
where $\v D$ is a vector of all drive operators in
\eqref{eq:drive_ops}.

We note that Weyl operators\cite{bertlmann2008bloch} and self-adjoint
linear combinations thereof\cite{asadian2016heisenbergweyl} also
provide an interesting basis of operators that generalize the notion
of position and momentum to the case of a finite-dimensional space.
As our present work is motivated by multi-level systems that are
realized with nuclear spin (angular momentum) states, however, the
transition and drive operators are more natural to consider than the
Weyl operators, so we will not discuss the latter any further.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{External drive}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Spin-orbit coupling}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Classifying eigenstates}

\appendix

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Changing operator bases for collective SU($n$)-symmetric
  interactions}
\label{sec:changing_bases}

Here we show that the collective interaction Hamiltonian in
\eqref{eq:H_int_spin} takes an identical form in any orthonormal basis
for $\B_n$, i.e.~the space of operators on the Hilbert space of an
$n$-level system, equipped with the trace (Hilbert-Schmidt) inner
product
\begin{align}
  \obk{\O|\Q} \equiv \tr\p{\O^\dag \Q}.
\end{align}
To maintain generality, we choose an arbitrary basis $\set{X_j}$
satisfying the orthonormality condition
\begin{align}
  \obk{X_j|X_k} = \N_X \delta_{jk},
\end{align}
for $j\in\set{0,1,\cdots,n^2-1}$.  This orthonormality condition
implies that we can resolve the identity (super-)operator $\I$ with
$\I\oket{\O}=\oket{\O}$ for any $\oket\O\in\B_n$ as
\begin{align}
  \I = \f1{\N_X} \sum_j \oop{X_j}{X_j}
  = \f1{\N_X} \sum_j \oop{X_j^\dag}{X_j^\dag},
\end{align}
where we used the fact that $\set{X_j^\dag}$ is also an orthonormal
basis for $\B_n$, satisfying the same orthonormality condition as
$\set{X_j}$; these two bases are transformed into each other by
unitary $\sum_j\oop{X_j^\dag}{X_j}/\sqrt{\N_X}$.

There are two types of terms in the Hamiltonian in
\eqref{eq:H_int_spin}: two-body interactions and self-interactions,
respectively of the form
\begin{align}
  \sum_j X_j^\dag \otimes X_j
  = \sum_j \oket{X_j^\dag}\otimes\oket{X_j},
  &&
  \sum_j X_j^\dag X_j = \sum_j \oket{X_j^\dag X_j}.
\end{align}
Choosing a different basis $\set{Y_j}$ for $\B_n$ with norm $\N_Y$, by
resolution of the identity we find that
\begin{align}
  \sum_j \oket{X_j^\dag}\otimes\oket{X_j}
  = \f1{\N_Y^2} \sum_{j,k,\ell}
  \oket{Y_k^\dag} \obk{Y_k^\dag|X_j^\dag}
  \otimes \oket{Y_\ell} \obk{Y_\ell|X_j},
\end{align}
where
\begin{align}
  \obk{Y_k^\dag|X_j^\dag}
  = \tr\p{Y_k X_j^\dag}
  = \tr\p{X_j^\dag Y_k}
  = \obk{X_j|Y_k},
\end{align}
which allows us to simplify
\begin{align}
  \sum_j \oket{X_j^\dag}\otimes\oket{X_j}
  = \f1{\N_Y^2} \sum_{k,\ell}
  \oket{Y_k^\dag} \otimes \oket{Y_\ell}
  \sum_j \obk{Y_\ell|X_j} \obk{X_j|Y_k}
  = \f{\N_X}{\N_Y} \sum_k
  \oket{Y_k^\dag} \otimes \oket{Y_k}.
\end{align}
For the single-body self-interactions, we can similarly expand
\begin{align}
  X_j = \f1{\N_Y} \sum_k Y_k \obk{Y_k|X_j},
  &&
  X_j^\dag = \f1{\N_Y} \sum_k Y_k^\dag \obk{Y_k^\dag|X_j^\dag}
  = \f1{\N_Y} \sum_k Y_k^\dag \obk{X_j|Y_k},
\end{align}
which implies
\begin{align}
  \sum_j X_j^\dag X_j
  = \f1{\N_Y^2} \sum_{j,k,\ell} Y_k^\dag Y_\ell
  \obk{X_j|Y_k} \obk{Y_\ell|X_j}
  = \f{\N_X}{\N_Y} \sum_k Y_k^\dag Y_k.
\end{align}
The collective interaction Hamiltonian in \eqref{eq:H_int_spin} thus
takes an identical form in any orthonormal basis for the space $\B_n$
of operators on the Hilbert space of an $n$-level system, up to
appropriate re-scaling factors $\N_X/\N_Y$.

\bibliography{sun_notes.bib}

\end{document}