### 2019-10-10 collective_sun/sun_notes.tex
changed convensions in definition of transition operator computed product of drive operators

\documentclass[nofootinbib,notitlepage,11pt]{revtex4-2}

%%% linking references
\usepackage{hyperref}
\hypersetup{
  breaklinks=true,
  colorlinks=true,
  linkcolor=blue,
  filecolor=magenta,
  urlcolor=cyan,
}

%%% header / footer
\usepackage{fancyhdr} % easier header and footer management
\pagestyle{fancy} % page formatting style
\fancyhf{} % clear all header and footer text
\renewcommand{\headrulewidth}{0pt} % remove horizontal line in header
\usepackage{lastpage} % for referencing last page
\cfoot{\thepage~of \pageref{LastPage}} % "x of y" page labeling

%%% symbols, notations, etc.
\usepackage{physics,braket,bm,amssymb} % physics and math
\renewcommand{\t}{\text} % text in math mode
\newcommand{\f}[2]{\dfrac{#1}{#2}} % shorthand for fractions
\newcommand{\p}[1]{\left(#1\right)} % parenthesis
\renewcommand{\sp}[1]{\left[#1\right]} % square parenthesis
\renewcommand{\set}[1]{\left\{#1\right\}} % curly parenthesis
\newcommand{\bk}{\Braket} % shorthand for braket notation
\renewcommand{\v}{\bm} % bold vectors
\newcommand{\uv}[1]{\bm{\hat{#1}}} % unit vectors
\newcommand{\av}{\vec} % arrow vectors
\renewcommand{\d}{\text{d}} % for infinitesimals
\renewcommand{\c}{\cdot} % inner product

\usepackage{dsfont} % for identity operator
\newcommand{\1}{\mathds{1}}

\newcommand{\up}{\uparrow}
\newcommand{\dn}{\downarrow}

\newcommand{\x}{\text{x}}
\newcommand{\y}{\text{y}}
\newcommand{\z}{\text{z}}

\newcommand{\B}{\mathcal{B}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\I}{\mathcal{I}}
\newcommand{\N}{\mathcal{N}}
\renewcommand{\O}{\mathcal{O}}
\newcommand{\Q}{\mathcal{Q}}
\newcommand{\T}{\mathcal{T}}
\renewcommand{\S}{\mathcal{S}}

\DeclareMathOperator{\sign}{sign}
\DeclareMathOperator{\signs}{signs}

\def\obra#1{\mathinner{({#1}|}}
\def\oket#1{\mathinner{|{#1})}}
\def\obk#1{\mathinner{({#1})}}
\def\oop#1#2{\oket{#1}\!\obra{#2}}

\usepackage[inline]{enumitem} % in-line lists and \setlist{} (below)
\setlist[enumerate,1]{label={(\roman*)}} % default in-line numbering
\setlist{nolistsep} % more compact spacing between environments

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\thispagestyle{fancy}

\title{Multi-level generalizations of collective spin dynamics and
  squeezing}%
\author{Michael A. Perlin}%
\date{\today}

\maketitle

\tableofcontents

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Collective SU($n$)-symmetric interactions}
\label{sec:int}

We consider a collection of $n$-level fermions on a 1-D lattice, with
a single-particle Hamiltonian
\begin{align}
  H_{\t{lat}} = -J \sum_{q,\mu} \cos\p{qa} c_{q\mu}^\dag c_{q\mu},
  \label{eq:H_lat}
\end{align}
where $J$ is the nearest-neighbor tunneling rate, $a$ is the spacing
between neighboring lattice sites, and $c_{q\mu}$ is the fermionic
annihilation operator for a particle in quasi-momentum mode $q$ with
internal state (e.g.~nuclear spin) $\mu$.  Under the frozen-mode
approximation, two-body collective SU($n$)-symmetric interactions
between these fermions take form\cite{perlin2019effective}
\begin{align}
  H_{\t{int}} = \f{u}{2} \sum_{p,q,\mu,\nu}
  \p{c_{p\mu}^\dag c_{q\nu}^\dag c_{q\nu} c_{p\mu}
    + c_{q\mu}^\dag c_{p\nu}^\dag c_{q\nu} c_{p\mu}},
\end{align}
where $u\equiv U/L$ is the two-body on-site interaction energy $U$
divided by the number of lattice sites $L$.  Defining the spin
operators $s_{\mu\nu}^{(p)}\equiv c_{p\mu}^\dag c_{p\nu}$ for each
quasi-momentum $p$, we can alternately write
\begin{align}
  H_{\t{int}} = \f{u}{2} \sum_{p,q,\mu,\nu}
  \p{s_{\mu\mu}^{(p)} s_{\nu\nu}^{(q)} - s_{\mu\nu}^{(p)} s_{\nu\mu}^{(q)}}
  = \f{u}{2}\p{N^2 - \sum_{p,q,\mu,\nu} s_{\mu\nu}^{(p)} s_{\nu\mu}^{(q)}},
\end{align}
where $N$ is the total number of particles.  In the absence of
coherence between sectors of different particle number, we can neglect
the $\sim N^2$ term and simply write
\begin{align}
  H_{\t{int}}
  = - \f{u}{2} \sum_{p,q,\mu,\nu} s_{\mu\nu}^{(p)} s_{\nu\mu}^{(q)}.
\end{align}
In the case of SU(2), at this point we would expand the spin operators
$s_{\mu\nu}^{(p)}$ in terms of Pauli operators in order to convert
$H_{\t{int}}$ into an SU(2) spin Hamiltonian.  To generalize this
procedure to the case of SU($n$), we first define a vector
$\v s^{(p)}$ of all spin operators $s_{\mu\nu}^{(p)}$, and write
\begin{align}
  H_{\t{int}}
  = - \f{u}{2} \sum_{p,q} {\v s^{(p)}}^\dag \v s^{(q)}
  = - \f{u}{2} \v\S^\dag \c \v\S,
  &&
  \v\S\equiv \sum_p \v s^{(p)}.
  \label{eq:H_int_spin}
\end{align}
In the remainder of this section, we identify various bases for the
space of operators on the Hilbert space of an $n$-level system.  For
brevity, we denote this space of operators by $\B_n$, and note that
$\B_n$ is itself an $n^2$-dimensional Hilbert space equipped with a
trace (Hilbert-Schmidt) inner product
\begin{align}
  \obk{\O|\Q} \equiv \tr\p{\O^\dag Q},
\end{align}
where $\oket{\Q}$ and $\obra{\O}$ respectively denote vectors in
$\B_n$ and its dual space $\B_n^*$.  Similarly to the set of all spin
operators $s_{\mu\nu}\equiv\op{\mu}{\nu}$ on the single-particle
Hilbert space at each quasi-momentum, all of the bases we consider
will be orthonormal with respect to the trace inner product.  As we
show in Appendix \ref{sec:changing_bases}, the collective
SU($n$)-symmetric interaction Hamiltonian in \eqref{eq:H_int_spin}
takes an identical form in any orthonormal basis for $\B_n$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Generalized Gell-Mann (GGM) operators}
\label{sec:ggm_ops}

The most obvious generalization of Pauli operators to the case of
SU($n$) are the {\it generalized Gell-Mann (GGM)
  operators}\cite{hioe1981level, bertlmann2008bloch}, defined for
$j,k,\ell\in\set{0,1,\cdots,n-1}$ with $k<j$ and $\ell>0$ by
\begin{align}
  \lambda_0 \equiv \sqrt{\f{2}{n}}~ \1,
  &&
  \lambda_\ell \equiv \sqrt{\f{2}{\ell\p{\ell+1}}}
  \p{\sum_{n=0}^{\ell-1}\op{n} - \ell \op{\ell}},
  \label{eq:GGM_diag}
\end{align}
\begin{align}
  \lambda_{jk,+} \equiv \op{j}{k} + \op{k}{j},
  &&
  \lambda_{jk,-} \equiv i\p{\op{j}{k} - \op{k}{j}},
  \label{eq:GGM_off_diag}
\end{align}
where $\1$ is the identity operator.  The self-adjoint GGM operators
in \eqref{eq:GGM_diag} and \eqref{eq:GGM_off_diag} provide a complete
basis for the space of operators on the Hilbert space of an $n$-level
system, coincide exactly with the Pauli operators in the case of
$n=2$, and satisfy the orthonormality condition
\begin{align}
  \obk{\lambda_a|\lambda_b} = 2\delta_{ab},
  \label{eq:GGM_inner}
\end{align}
where $a,b$ index any GGM operator in \eqref{eq:GGM_diag} or
\eqref{eq:GGM_off_diag}.  Denoting a vector of all GGM operators by
$\v\lambda$, we can therefore write (see Appendix
\ref{sec:changing_bases})
\begin{align}
  H_{\t{int}}
  = -\f{u}{4} \sum_{p,q} \v\lambda^{(p)} \c \v\lambda^{(q)}
  = -\f{u}{4} \v\Lambda \c \v\Lambda,
  &&
  \v\Lambda \equiv \sum_p \v\lambda^{(p)}.
  \label{eq:H_int_GGM}
\end{align}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Transition operators}
\label{sec:trans_ops}

The GGM operators in \eqref{eq:GGM_diag} and \eqref{eq:GGM_off_diag}
provide a convenient operator basis to describe dynamics obeying
SU($n$) symmetry.  An external driving field addressing $n$-level
nuclear spins, however, will generally violate SU($n$) symmetry, and
instead obeys the symmetries of the {\it polarization} or {\it
  transition operators} defined by\cite{kryszewski2006alternative,
  bertlmann2008bloch}
\begin{align}
  T_{LM}
  \equiv \p{-1}^M \sqrt{\f{2L+1}{2I+1}} \sum_{\mu,\nu}
  \bk{I\mu;LM|I\nu} \op{\nu}{\mu},
  \label{eq:trans_ops}
\end{align}
where $L\in\set{0,1,\cdots,n-1}$ and $M\in\set{-L,-L+1,L}$ index a
total spin and its projection onto a quantization axis;
$I\equiv\p{n-1}/2$ is the maximal angular momentum of an $n$-level
spin; $\mu,\nu\in\set{-I,-I+1,\cdots,I}$ index projections of an
$n$-level nuclear spin onto a quantization axis; and
$\bk{I\mu;LM|I\nu}$ is a Clebsch-Gordan coefficient.  Note that our
definition of a transition operators differs from that in
Refs.~\cite{kryszewski2006alternative, bertlmann2008bloch} by the sign
$\p{-1}^M$; this choice of sign convention will simplify much of our
work below.

The transition operator $T_{LM}$ is proportional to the transition
induced on a nuclear spin by the absorption of a spin-$L$ particle
with spin projection $M$ onto a quantization axis.  Similarly to the
GGM operators in \eqref{eq:GGM_diag} and \eqref{eq:GGM_off_diag}, the
transition operators in \eqref{eq:trans_ops} provide a complete basis
for the space of operators on the Hilbert space of an $n$-level
system, and satisfy the orthonormality condition
\begin{align}
  \obk{T_{LM}|T_{L'M'}} = \delta_{LL'} \delta_{MM'},
\end{align}
which implies
\begin{align}
  H_{\t{int}} = -\f{u}{2} \sum_{p,q} {\v T^{(p)}}^\dag \c \v T^{(q)}
  = -\f{u}{2} \v\T^\dag \c \v\T,
  &&
  \v\T \equiv \sum_p\v T^{(p)},
  \label{eq:H_int_trans}
\end{align}
where $\v T$ is a vector of all transition operators in
\eqref{eq:trans_ops}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Drive opermators}
\label{sec:drive_ops}

Unlike the GGM operators in \eqref{eq:GGM_diag} and
\eqref{eq:GGM_off_diag}, the transition operators in
\eqref{eq:trans_ops} are not self-adjoint.  In order to express
Hamiltonians in terms of self-adjoint operators, we define the {\it
  drive operators}
\begin{align}
  D_{L,0} \equiv T_{L,0},
  &&
  D_{LM} \stackrel{M\ne0}{\equiv} \f{\sigma_M}{\sqrt{2}}
  \sp{T_{L\abs{M}} + \sign\p{M}T_{L\abs{M}}^\dag},
  &&
  \sigma_M \equiv \sign\p{M}^{3/2},
  \label{eq:drive_ops}
\end{align}
for integer $M$ with $\abs{M}\le L$.  The drive operator $D_{LM}$ is
proportional to the Hamiltonian induced on a nuclear spin by an
external classical field of spin-$L$ particles with spin projection
$M$ onto a quantization axis.  Similarly to the GGM operators in
\eqref{eq:GGM_diag} and \eqref{eq:GGM_off_diag}, the drive operators
are proportional to the Pauli operators in the case of $n=2$, and
satisfy the orthonormality condition
\begin{align}
  \obk{D_{LM}|D_{L'M'}} = \delta_{LL'}\delta_{MM'},
\end{align}
which implies
\begin{align}
  H_{\t{int}} = -\f{u}{2} \sum_{p,q} \v D^{(p)}\c\v D^{(q)}
  = -\f{u}{2} \v\D \c \v\D,
  &&
  \v\D \equiv \sum_p \v D^{(p)},
  \label{eq:H_int_drive}
\end{align}
where $\v D$ is a vector of all drive operators in
\eqref{eq:drive_ops}.

The drive operators $D_{1,0}$ and $D_{1,\pm1}$ are proportional to the
Hamiltonians induced on an $n$-level spin by an external magnetic
field and a classical spin-polarized driving laser; loosely speaking,
one can identify $\p{D_{1,0},D_{1,1},D_{1,-1}}\sim\p{S_\z,S_\x,S_\y}$,
where $S_\alpha$ is a collective SU(2) spin-$\alpha$
operator\cite{perlin2019shorttime}.  More concretely, we can embed the
Hilbert space of a single $n$-level spin into the
permutationally-symmetric (Dicke) manifold of states for $\p{n-1}$
2-level spins via
\begin{align}
  \ket{\mu}_{\t{single}} \to \ket{I+\mu}_{\t{collective}}
  \propto S_+^{I+\mu} \ket\dn^{\otimes\p{n-1}},
\end{align}
where $I\equiv\p{n-1}/2$ is a total spin,
$\mu\in\set{-I,-I+1,\cdots,I}$ indexes a state of the single $n$-level
spin, and $S_+$ is a collective spin-raising operator for $\p{n-1}$
2-level spins.  This embedding takes
\begin{align}
  \p{D_{1,0},D_{1,1},D_{1,-1}}
  \to \p{S_\z,S_\x,S_\y}
  \times \f{2\sqrt{3}}{\sqrt{\p{n-1}n\p{n+1}}}.
\end{align}

We note that Weyl operators\cite{bertlmann2008bloch} and self-adjoint
linear combinations thereof\cite{asadian2016heisenbergweyl} also
provide an interesting basis of operators, generalizing continuous
phase-space displacement operators to the case of a finite-dimensional
Hilbert space.  As our present work is motivated by multi-level
systems that are realized with nuclear spin (angular momentum) degrees
of freedom, however, the transition and drive operators are more
natural to consider than the Weyl operators, so we will not discuss
the latter any further.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{External drive and spin-orbit coupling}
\label{sec:drive_SOC}

An driving laser addressing spin-$1/2$ fermions on a 1-D lattice will
induce the Hamiltonian
\begin{align}
  \left. H_{\t{drive}}^{(\phi)} \right|_{n=2}
  = \Omega \sum_j
  \p{e^{-i\phi j} c_{j,\up}^\dag c_{j,\dn}
    + e^{i\phi j} c_{j,\dn}^\dag c_{j,\up}},
  \label{eq:H_drive_2}
\end{align}
where $\Omega$ is a (real) driving amplitude, $j\in\mathbb{Z}$ indexes
a lattice site, and $\phi$ is the relative phase of the drive between
adjacent lattice sites.  The multi-level analogue of
\eqref{eq:H_drive_2} is
\begin{align}
  H_{\t{drive}}^{(\phi)}
  = -\Omega \sum_j \p{e^{-i\phi j} T_{1,1}^{(j)}
    + e^{i\phi j} {T_{1,1}^{(j)}}^\dag}
  = \Omega \D_{1,1}^{(\phi)},
  \label{eq:H_drive}
\end{align}
where
\begin{align}
  \D_{LM}^{(\phi)}
  \equiv \sum_j \sp{\cos\p{M\phi j} D_{LM}^{(j)}
    + \sin\p{M\phi j} D_{L,-M}^{(j)}},
  \label{eq:drive_rot}
\end{align}
is an inhomogeneously rotated collective drive operator.  We can make
this drive spatially homogenous via the gauge transformation
$c_{j\mu}^\dag \to e^{i\mu\phi j} c_{j\mu}^\dag$, which takes
\begin{align}
  H_{\t{drive}}^{(\phi)} \to H_{\t{drive}} = \Omega \D_{1,1}.
\end{align}
The SU($n$) symmetry of collective interactions implies that the
interaction Hamiltonian $H_{\t{int}}$ in \eqref{eq:H_int_drive} is
invariant under this gauge transformation, but the single-particle
Hamiltonian in \eqref{eq:H_lat} transforms as (see Appendix
\ref{sec:lat_drive})
\begin{align}
  H_{\t{lat}}
  \to H_{\t{lat}}^{(\phi)}
  = -J \sum_{q,\mu} \cos\p{qa-\mu\phi} c_{q\mu}^\dag c_{q\mu}
  = \sum_{q,L} B_L^{(\phi,q)} D_{L,0}^{(q)},
  \label{eq:H_lat_SOC}
\end{align}
with effective inhomogeneous driving field amplitudes
\begin{align}
  B_L^{(\phi,q)} = -J w_L\p{qa} A_L^{(\phi)},
\end{align}
where
\begin{align}
  w_L\p{\theta} \equiv
  \begin{cases}
    \cos\theta & L~\t{even} \\
    \sin\theta & L~\t{odd}
  \end{cases},
  &&
  A_L^{(\phi)} \equiv \sqrt{\f{2L+1}{n}}
  \sum_\mu \bk{I\mu;L,0|I\mu} w_L\p{\mu\phi}.
\end{align}
Spin-orbit coupling with multi-level systems thus allows for the
simulation of synthetic gauge fields with large spin ($L>1$).  As we
will see in the following section, multi-laser drives can also
simulate synthetic large-spin fields with non-zero spin projection
($M\ne0$) onto a quantization axis.

In the regime of weak SOC with $\abs{\phi}\ll1$, by considering
\begin{enumerate*}
\item all $L\le12$ for arbitrary $n$, and
\item all $L<n$ for all $n\le25$
\end{enumerate*}
(and conjecturing the same pattern for all $L$ and $n$), we find that
% expand Clebsch-Gordan coefficients in a power series according to
% \url{functions.wolfram.com/07.38.06.0008.01}
\begin{align}
  A_L^{(\phi)} = c_L \sqrt{n_L}\, \phi^L + O\p{\phi^{L+2}},
  \label{eq:A_L_phi_small}
\end{align}
where
\begin{align}
  c_L \equiv \p{-1}^{\lfloor L/2\rfloor}
  \sqrt{2L+1}\, \f{L!}{\p{2L+1}!},
  &&
  n_L \equiv \prod_{\ell=-L}^L\p{n-\ell}.
\end{align}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Raman drive}
\label{sec:drive_raman}

Rather than coupling the states of a multi-level system directly with
a single driving laser, as in Section \ref{sec:drive_SOC}, we now
consider coupling through an off-resonant Raman transition to a highly
excited state.  Specifically, we consider a three-laser drive of the
form
\begin{align}
  H_{\t{Raman,bare}}^{(\phi)}
  = \sum_j \sum_{m\in\set{1,0,-1}} \p{\Omega_m e^{-i\phi_m j}
    T_{1,m}^{(j)} \otimes \op{\up}{\dn} + \t{h.c.}}
  + \f{\Delta}{2} \1 \otimes \p{\op{\up} - \op{\dn}},
  \label{eq:raman_bare}
\end{align}
where $j$ indexes a lattice site, $\Omega_m$ is a (real) driving
amplitude; $\phi_m$ is the relative phase of drive $m$ between
adjacent lattice sites; $\op{r}{s}$ for $r,s\in\set{\up,\dn}$ is a
pseudo-spin operator for the auxiliary degree of freedom that is
off-resonantly addressed by the lasers; $\Delta$ is a detuning of the
lasers from the energy of an excited state; and $\1$ is the identity
operator.  If the detuning is large, $\Delta\gg\Omega_m$, then at
second order in perturbation theory the effective action of
$H_{\t{Raman,bare}}^{(\phi)}$ within the $\ket{\dn}$-state manifold
takes the form
\begin{align}
  H_{\t{Raman}}^{(\phi)}
  = \sum_{\tau\in\set{0,1,2}} H_{\t{Raman},\tau}^{(\phi)}
\end{align}
where $H_{\t{Raman},\tau}^{(\phi)}$ generates a nuclear spin
transition $\mu\to\mu\pm\tau$:
\begin{align}
  H_{\t{Raman},0}^{(\phi)}
  &= -\sum_j\sum_m \f{\Omega_m^2}{\Delta}
  {T_{1,m}^{(j)}}^\dag T_{1,m}^{(j)}, & \label{eq:raman_0} \\
  H_{\t{Raman},1}^{(\phi)}
  &= -\sum_j\sum_{s\in\set{\pm1}}
  \f{\Omega_0\Omega_s}{\Delta} e^{-i\p{\phi_s-\phi_0} j}
  {T_{1,0}^{(j)}}^\dag T_{1,s}^{(j)} + \t{h.c.},
  \\
  H_{\t{Raman},2}^{(\phi)}
  &= -\sum_j\sum_{s\in\set{\pm1}}
  \f{\Omega_+\Omega_-}{\Delta} e^{-i\p{\phi_s-\phi_{-s}}j}
  {T_{1,-s}^{(j)}}^\dag T_{1,s}^{(j)}.
  \label{eq:raman_2}
\end{align}
In terms of the drive operators, we can expand
\begin{align}
  H_{\t{Raman}}^{(\phi)}
  = \sum_j \sum_{\substack{L\le2\\\abs{M}\le L}}
  \Omega_{LM}^{(\phi,j)} D_{LM}^{(j)},
  \label{eq:raman_drive}
\end{align}
where the coefficients $\Omega_{LM}^{(\phi,j)}$ are provided in
Appendix \ref{sec:drive_raman_coeff}.  If the driving lasers are
phase-matched with $\phi_m=m\phi$, such that a nuclear spin transition
$\mu\to\mu+m$ on site $j$ imprints the phase $-m\phi j$, then the
drive $H_{\t{Raman}}^{(\phi)}$ takes the form
\begin{align}
  \left. H_{\t{Raman}}^{(\phi)} \right|_{\phi_m=m\phi}
  = \sum_{\substack{L\le2\\0\le M\le L}} \Omega_{LM} \D_{LM}^{(\phi)},
  \label{eq:drive_raman_matched}
\end{align}
with rotated collective drive operators $\D_{LM}^{(\phi)}$ defined in
\eqref{eq:drive_rot} and uniform driving amplitudes $\Omega_{LM}$
provided in Appendix \ref{sec:drive_raman_coeff}.  Similarly to the
case of a direct drive considered in Section \ref{sec:drive_SOC}, the
Raman drive $H_{\t{Raman}}^{(\phi)}$ in \eqref{eq:drive_raman_matched}
can be made homogenous through the gauge transformation
$c_{j\mu}^\dag \to e^{i\mu\phi j} c_{j\mu}^\dag$, which takes
\begin{align}
  \left. H_{\t{Raman}}^{(\phi)} \right|_{\phi_m=m\phi}
  \to H_{\t{Raman}}
  = \sum_{\substack{L\le2\\0\le M\le L}} \Omega_{LM} \D_{LM}.
\end{align}
The effective driving amplitudes $\Omega_{LM}$ can be tuned through
the bare driving amplitudes $\Omega_m$, and satisfy
\begin{align}
  \Omega_{0,0} &\propto \Omega_+^2 + \Omega_-^2 + \Omega_0^2,
  &
  \Omega_{1,0} &\propto \Omega_+^2 - \Omega_-^2,
  &
  \Omega_{1,1} &\propto \Omega_0 \p{\Omega_+ + \Omega_-}, \\
  \Omega_{2,0} &\propto \Omega_+^2 + \Omega_-^2 - 2\Omega_0^2,
  &
  \Omega_{2,1} &\propto \Omega_0 \p{\Omega_+ - \Omega_-},
  &
  \Omega_{2,2} &\propto \Omega_+ \Omega_-.
\end{align}
In particular, the amplitudes $\Omega_{1,0}$, $\Omega_{2,0}$, and
$\Omega_{2,1}$ can be simultaneously tuned to zero by setting
$\Omega_0=\Omega_+=\Omega_-$, at which point the only remaining
effective drives in $H_{\t{Raman}}$ are $\sim\D_{1,1}$ and
$\sim\D_{2,2}$ (as $\D_{0,0}\propto\1$ is just constant shift in
energy).  If instead $\Omega_0=\pm\Omega_\pm$, then the amplitudes
$\Omega_{1,0}$, $\Omega_{2,0}$ and $\Omega_{1,1}$ vanish, leaving only
the drives $\D_{2,1}$ and $\D_{2,2}$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Future directions}

\begin{itemize}
\item Products and commutators of drive operators
\item Visualization with generalized Bloch spheres
  \begin{itemize}
  \item Single-particle eigenstates of drive operators
  \item Single-particle dynamics induced by drive operators
  \end{itemize}
\item Classifying collective states (e.g.~eigenstates of interactions,
  spin-wave-like states)
\item Implications of SOC
  \begin{itemize}
  \item Effective collective dynamics for $\abs{\phi}\ll1$
  \item Deriving \eqref{eq:A_L_phi_small}
  \item Special angles $\phi$
  \end{itemize}
\item Generalizations of squeezing beyond SU(2)?
\item Consider different interactions
  \begin{itemize}
  \item Include electronic states
  \item Super-exchange regime
  \item Non-uniform (``disordered'') interactions
  \end{itemize}
\item Consider different drives
\end{itemize}

\appendix

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Changing operator bases for SU($n$)-symmetric interactions}
\label{sec:changing_bases}

Here we show that the collective interaction Hamiltonian in
\eqref{eq:H_int_spin} takes an identical form in any orthonormal basis
for $\B_n$, i.e.~the space of operators on the Hilbert space of an
$n$-level system, equipped with the trace (Hilbert-Schmidt) inner
product
\begin{align}
  \obk{\O|\Q} \equiv \tr\p{\O^\dag \Q}.
\end{align}
To maintain generality, we choose an arbitrary basis
$X\equiv\set{X_j}$ satisfying the orthonormality condition
\begin{align}
  \obk{X_j|X_k} = \N_X \delta_{jk},
\end{align}
for $j\in\set{0,1,\cdots,n^2-1}$.  This orthonormality condition
implies that we can resolve the identity (super-)operator $\I$ with
$\I\oket{\O}=\oket{\O}$ for any $\oket\O\in\B_n$ as
\begin{align}
  \I = \f1{\N_X} \sum_j \oop{X_j}{X_j}
  = \f1{\N_X} \sum_j \oop{X_j^\dag}{X_j^\dag},
\end{align}
where we used the fact that $\set{X_j^\dag}$ is also an orthonormal
basis for $\B_n$, satisfying the same orthonormality condition as
$\set{X_j}$; these two bases are transformed into each other by
unitary $\sum_j\oop{X_j^\dag}{X_j}/\N_X$.

The collective interaction Hamiltonian in \eqref{eq:H_int_spin} takes
the form
\begin{align}
  H = \sum_{p,q,j} X_j^{(p)} {X_j^{(q)}}^\dag,
\end{align}
which essentially consists of two-body interactions ($p\ne q$) and
single-body terms ($p=q$), respectively of the form
\begin{align}
  \sum_j X_j \otimes X_j^\dag,
  &&
  \sum_j X_j X_j^\dag.
\end{align}
Choosing a different basis $Y\equiv\set{Y_j}$ for $\B_n$ with
corresponding norm $\N_Y$, we can expand
\begin{align}
  X_j = \f1{\N_Y} \sum_k Y_k \obk{Y_k|X_j},
  &&
  X_j^\dag = \f1{\N_Y} \sum_k Y_k^\dag \obk{Y_k^\dag|X_j^\dag}
  = \f1{\N_Y} \sum_k Y_k^\dag \obk{X_j|Y_k},
\end{align}
where we used the fact that
\begin{align}
  \obk{Y_k^\dag|X_j^\dag}
  = \tr\p{Y_k X_j^\dag}
  = \tr\p{X_j^\dag Y_k}
  = \obk{X_j|Y_k}.
\end{align}
By resolution of the identity, we can therefore find
\begin{align}
  \sum_j X_j \otimes X_j^\dag
  = \f1{\N_Y^2} \sum_{j,k,\ell} Y_k \obk{Y_k|X_j}
  \otimes Y_\ell^\dag \obk{X_j|Y_\ell}
  = \f{\N_X}{\N_Y^2} \sum_{k,\ell} Y_k \otimes Y_\ell^\dag
  \obk{Y_k|\I|Y_\ell}
  = \f{\N_X}{\N_Y} \sum_k Y_k \otimes Y_k^\dag,
\end{align}
and similarly
\begin{align}
  \sum_j X_j X_j^\dag
  = \f1{\N_Y^2} \sum_{j,k,\ell} Y_k Y_\ell^\dag
  \obk{Y_k|X_j} \obk{X_j|Y_\ell}
  = \f{\N_X}{\N_Y} \sum_k Y_k Y_k^\dag.
\end{align}
The collective interaction Hamiltonian in \eqref{eq:H_int_spin} thus
takes an identical form in any orthonormal basis for the space $\B_n$
of operators on the Hilbert space of an $n$-level system, up to
appropriate re-scaling factors $\N_X/\N_Y$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Spin-orbit coupling with drive operators}
\label{sec:lat_drive}

Here we derive the drive-operator ($D_{LM}$) expansion of the
spin-orbit coupled Hamiltonian $H_{\t{lat}}^{(\phi)}$ in
\eqref{eq:H_lat_SOC}.  In the basis of spin operators
$s_{\mu\mu}^{(q)} \equiv c_{q\mu}^\dag c_{q\mu}$, this Hamiltonian
takes the form
\begin{align}
  H_{\t{lat}}^{(\phi)}
  = -J \sum_{q,\mu} \cos\p{qa-\mu\phi} s_{\mu\mu}^{(q)}
  \label{eq:H_lat_spin}
\end{align}
Defining the diagonal ``spin-wave'' operators
\begin{align}
  \tilde d_{\phi+} \equiv \sum_\mu \cos\p{\mu\phi} s_{\mu\mu},
  &&
  \tilde d_{\phi-} \equiv \sum_\mu \sin\p{\mu\phi} s_{\mu\mu},
\end{align}
which are respectively even and odd under spin inversion
($\mu\to-\mu$), we can expand
\begin{align}
  \sum_\mu \cos\p{qa+\mu\phi} s_{\mu\mu}
  = \cos\p{qa} \tilde d_{\phi+} + \sin\p{qa} \tilde d_{\phi-},
\end{align}
and resolve the identity (super-)operator $\I$ with
$\I \tilde d_{\phi\pm} = \tilde d_{\phi\pm}$ in the basis of the drive operators
$D_{LM}$ (see Appendix \ref{sec:changing_bases}), finding
\begin{align}
  \sum_\mu \cos\p{qa+\mu\phi} s_{\mu\mu}
  = \cos\p{qa} \f12 \sum_{L,M} \obk{\tilde d_{\phi+}|D_{LM}} D_{LM}
  + \sin\p{qa} \f12 \sum_{L,M} \obk{\tilde d_{\phi-}|D_{LM}} D_{LM}.
\end{align}
We now note that the drive operators $D_{LM}$ are strictly
off-diagonal for $M\ne0$, and furthermore that $D_{L,0}$ with even
(odd) $L$ is even (odd) under spin inversion, which implies
\begin{align}
  \sum_{L,M} \obk{\tilde d_{\phi+}|D_{LM}} D_{LM}
  = \sum_{L\,\t{even}} \obk{\tilde d_{\phi+}|D_{L,0}} D_{L,0},
\end{align}
and likewise with $\tilde d_{\phi-}$ and odd $L$.  In total, we can
write
\begin{align}
  \sum_\mu \cos\p{qa+\mu\phi} s_{\mu\mu}
  = \cos\p{qa} \f12 \sum_{L\,\t{even}}
  \obk{\tilde d_{\phi+}|D_{L,0}} D_{L,0}
  + \sin\p{qa} \f12 \sum_{L\,\t{odd}}
  \obk{\tilde d_{\phi-}|D_{L,0}} D_{L,0},
\end{align}
or, more compactly,
\begin{align}
  \sum_\mu \cos\p{qa+\mu\phi} s_{\mu\mu}
  = \sum_L w_L\p{qa} A_L^{(\phi)} D_{L,0},
  &&
  w_L\p{\theta} \equiv
  \begin{cases}
    \cos\theta & L~\t{even} \\
    \sin\theta & L~\t{odd}
  \end{cases},
  \label{eq:compact_SOC_drive}
\end{align}
with effective driving field amplitudes
\begin{align}
  A_L^{(\phi)}
  \equiv \f12 \obk{\tilde d_{\phi,\p{-1}^L}|D_{L,0}}
  = \f12 \sum_\mu w_L\p{\mu\phi} \obk{s_{\mu\mu}|D_{L,0}}
  = \sqrt{\f{2L+1}{n}}
  \sum_\mu \bk{I\mu;L,0|I\mu} w_L\p{\mu\phi}.
\end{align}
Substituting the result in \eqref{eq:compact_SOC_drive} into the
single-particle Hamiltonian in \eqref{eq:H_lat_spin} yields
\begin{align}
  H_{\t{lat}}^{(\phi)} = \sum_{q,L} B_L^{(\phi,q)} D_{L,0}^{(q)},
  &&
  B_L^{(\phi,q)} \equiv -J w_L\p{qa} A_L^{(\phi)}.
\end{align}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Transition operator product expansion}
\label{sec:trans_prod}

Here we derive the structure constants
$f_{\ell_1 m_1;\ell_2 m_2}^{LM}$ of the transition operators on an
$n$-dimensional Hilbert space, which allows us to expand
\begin{align}
  T_{\ell_1 m_1} T_{\ell_2 m_2}
  = \sum_{L,M} f_{\ell_1 m_1;\ell_2 m_2}^{LM} T_{LM},
\end{align}
with
\begin{align}
  f_{\ell_1 m_1;\ell_2 m_2}^{LM}
  \equiv \obk{T_{LM} | T_{\ell_1 m_1} T_{\ell_2 m_2}}
  = \tr\p{T_{LM}^\dag T_{\ell_1 m_1} T_{\ell_2 m_2}}.
  \label{eq:trans_struct}
\end{align}
For reference, the transition operators $T_{LM}$ are defined for
integers $L,M$ with $0\le L<n$ and $\abs{M}\le L$ in terms of
Clebsch-Gordan coefficients $\bk{\ell_1 m_1;\ell_2 m_2|\ell_3 m_3}$ by
\begin{align}
  T_{LM} \equiv \p{-1}^M \sqrt{\f{2L+1}{2I+1}}
  \sum_{\mu,\nu} \bk{I\mu;LM|I\nu} \op{\nu}{\mu},
  &&
  I \equiv \f{n-1}{2}.
\end{align}
Using the symmetry properties of Clebsch-Gordan coefficients, namely
\begin{align}
  \bk{\ell_1 m_1; \ell_2 m_2| \ell_3 m_3}
  &= \p{-1}^{\ell_2+m_2} \sqrt{\f{2\ell_3+1}{2\ell_1+1}}
  \bk{\ell_3,-m_3; \ell_2 m_2| \ell_1,-m_1} \\
  \bk{\ell_1 m_1; \ell_2 m_2| \ell_3 m_3}
  &= \p{-1}^{\ell_1+\ell_2-\ell_3}
  \bk{\ell_1,-m_1; \ell_2,-m_2| \ell_3,-m_3},
\end{align}
we can find that
\begin{align}
  T_{LM}^\dag
  = \p{-1}^M \sqrt{\f{2L+1}{2I+1}}
  \sum_{\mu,\nu} \p{-1}^M \bk{I\nu;L,-M|I\mu} \op{\mu}{\nu}
  = \p{-1}^M T_{L,-M}.
\end{align}
Substituting this result into \eqref{eq:trans_struct} and expanding
Clebsch-Gordan coefficients in terms of Wigner 3-$j$ symbols as
\begin{align}
  \bk{\ell_1 m_1; \ell_2 m_2| \ell_3 m_3}
  = \p{-1}^{-\ell_1+\ell_2-m_3} \sqrt{\p{2\ell_3+1}}
  \begin{pmatrix}
    \ell_1 & \ell_2 & \ell_3 \\
    m_1 & m_2 & -m_3
  \end{pmatrix},
\end{align}
we find that
\begin{multline}
  f_{\ell_1 m_1;\ell_2 m_2}^{LM}
  = \delta_{M,m_1+m_2}
  \p{-1}^{m_1+m_2} \sqrt{\p{2L+1}\p{2\ell_1+1}\p{2\ell_2+1}} \\
  \times \sum_{\mu,\nu,\rho} \p{-1}^{-3I+L+\ell_1+\ell_2-\mu-\nu-\rho}
  \begin{pmatrix}
    I & L & I \\
    \mu & -M & -\nu
  \end{pmatrix}
  \begin{pmatrix}
    I & \ell_1 & I \\
    \nu & m_1 & -\rho
  \end{pmatrix}
  \begin{pmatrix}
    I & \ell_2 & I \\
    \rho & m_2 & -\mu
  \end{pmatrix},
\end{multline}
where we can evaluate the sum and introduce Wigner 6-$j$ symbols to
get
\begin{align}
  f_{\ell_1 m_1;\ell_2 m_2}^{LM}
  &= \delta_{M,m_1+m_2} \p{-1}^{2I+M}
  \sqrt{\p{2L+1}\p{2\ell_1+1}\p{2\ell_2+1}}
  \begin{pmatrix}
    L & \ell_1 & \ell_2 \\
    M & -m_1 & -m_2
  \end{pmatrix}
  \begin{Bmatrix}
    L & \ell_1 & \ell_2 \\
    I & I & I
  \end{Bmatrix} \\
  &= \delta_{M,m_1+m_2} \p{-1}^{2I+L}
  \sqrt{\p{2\ell_1+1}\p{2\ell_2+1}}
  \bk{\ell_1 m_1; \ell_2 m_2| LM}
  \begin{Bmatrix}
    \ell_1 & \ell_2 & L \\
    I & I & I
  \end{Bmatrix}.
  \label{eq:trans_prod}
\end{align}
The Wigner 6-$j$ symbol is symmetric under any permutation of its
columns, while Clebsch-Gordan coefficients satisfy
\begin{align}
  \bk{\ell_1 m_1; \ell_2 m_2| \ell_3 m_3}
  = \p{-1}^{\ell_1+\ell_2-\ell_3}
  \bk{\ell_2 m_2; \ell_1 m_1| \ell_3 m_3},
\end{align}
which implies that
\begin{align}
  f_{\ell_1 m_1;\ell_2 m_2}^{LM}
  = \p{-1}^{\ell_1+\ell_2-L} f_{\ell_2 m_2;\ell_1 m_1}^{LM}.
\end{align}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Drive operator product expansions}
\label{sec:drive_prod}

The drive operators are defined by
\begin{align}
  D_{L,0} \equiv T_{L,0},
  &&
  D_{LM} \stackrel{M\ne0}{\equiv} \f{\sigma_M}{\sqrt{2}}
  \sp{T_{L\abs{M}} + \sign\p{M}T_{L\abs{M}}^\dag},
  &&
  \sigma_M \equiv \sign\p{M}^{3/2},
\end{align}
and can alternately be expressed as
\begin{align}
  D_{LM} = \f{\tau_M}{\sqrt{2}} \p{T_{LM} + \epsilon_M T_{L,-M}},
\end{align}
where
\begin{align}
  \tau_M \equiv
  \begin{cases}
    \sqrt{2} & M = 0, \\
    \sign\p{M}^{M+1/2} & M \ne 0
  \end{cases},
  &&
  \epsilon_M \equiv \p{-1}^M \sign\p{M},
\end{align}
which also implies that
\begin{align}
  T_{LM} = \f{\tilde\tau_M}{\sqrt{2}} \p{D_{LM} + i D_{L,-M}},
  &&
  \tilde\tau_M \equiv
  \begin{cases}
    e^{-i\pi/4} & M = 0 \\
    \sign\p{M}^{M-1/2} & M \ne 0
  \end{cases}.
\end{align}
In order to evaluate the product of two drive operators, we can thus
expand them in terms of transition operators, evaluate the
corresponding products of transition operators (see Appendix
\ref{sec:trans_prod}), and finally expand the result in terms of drive
operators.  The general product of two drive operators takes the form
\begin{align}
  D_{\ell_1 m_1} D_{\ell_2 m_2}
  = \f{\tau_{m_1}\tau_{m_2}}{\sqrt{2}}
  \sum_L \p{f_{\ell_1 m_1;\ell_2 m_2}^{L,m_1+m_2}
    E_{\ell_1 m_1;\ell_2 m_2}^{(L,+)}
    + f_{\ell_1 m_1;\ell_2,-m_2}^{L,m_1-m_2}
    E_{\ell_1 m_1;\ell_2 m_2}^{(L,-)}},
\end{align}
with operators
\begin{align}
  E_{\ell_1 m_1;\ell_2 m_2}^{(L,+)}
  &\equiv \f12 \sum_{s\in\set{\pm1}} \sqrt{s} D_{L,s\p{m_1+m_2}}
  \sp{\tilde\tau_{m_1+m_2}
    + i s \p{-1}^{L+\ell_1+\ell_2}
    \epsilon_{m_1} \epsilon_{m_2} \tilde\tau_{-\p{m_1+m_2}}},
  \\
  E_{\ell_1 m_1;\ell_2 m_2}^{(L,-)}
  &\equiv \f12 \sum_{s\in\set{\pm1}} \sqrt{s} D_{L,s\p{m_1-m_2}}
  \sp{\epsilon_{m_2} \tilde\tau_{m_1-m_2}
    + i s \p{-1}^{L+\ell_1+\ell_2}
    \epsilon_{m_1} \tilde\tau_{-\p{m_1-m_2}}}.
\end{align}
The general product of drive operators simplifies somewhat in several
special cases.  If one of $m_1,m_2$ is zero, for example, then
\begin{align}
  D_{\ell_1 m_1} D_{\ell_2 m_2}
  \stackrel{0\in\set{m_1,m_2}}{=}
  \sum_L f_{\ell_1 m_1;\ell_2 m_2}^{L,m_1+m_2} \times
  \begin{cases}
    D_{L,m_1+m_2} & L + \ell_1 + \ell_2 ~\t{even} \\
    i D_{L,-\p{m_1+m_2}} & \t{otherwise}
  \end{cases}.
\end{align}
If $\abs{m_1}=\abs{m_2}$, then
\begin{multline}
  D_{\ell_1 m_1} D_{\ell_2 m_2}
  \stackrel{\abs{m_1}=\abs{m_2}}{=}
  \sqrt{s_m} \p{-1}^m \sum_L f_{\ell_1 m;\ell_2,-m}^{L,0} D_{L,0} \times
  \begin{cases}
    1 & \p{-1}^L = s_m \p{-1}^{\ell_1+\ell_2} \\
    0 & \t{otherwise}
  \end{cases} \\
  + \f{\sqrt{\sign\p{m}}}{\sqrt{s_m}} \f1{\sqrt{2}}
  \sum_L f_{\ell_1 m;\ell_2 m}^{L,2m} \times
  \begin{cases}
    D_{L,2m} & \p{-1}^L = s_m \p{-1}^{\ell_1+\ell_2} \sign\p{m} \\
    i D_{L,-2m} & \t{otherwise}
  \end{cases},
\end{multline}
where $m \equiv m_1$ and $s_m\equiv+1$ ($-1$) if $m_1=m_2$
($m_1\ne m_2$).  Finally, if $\abs{m_1}\ne\abs{m_2}$ with
$m_1,m_2\ne0$, then
\begin{multline}
  D_{\ell_1 m_1} D_{\ell_2 m_2}
  \stackrel{\substack{\abs{m_1}\ne\abs{m_2}\\m_1,m_2\ne0}}{=}
  \f1{\sqrt{2}} \sum_{s\in\set{\pm1}}\f1{\sqrt{s}}
  \tau_{m_1} \tau_{s m_2} \tau_{m_1+s m_2}
  \sign\p{m_1+sm_2} \\
  \times \sum_L f_{\ell_1 m_1;\ell_2 sm_2}^{L,m_1+sm_2} \times
  \begin{cases}
    D_{L,m_1+sm_2} & \p{-1}^L = \zeta_{\ell_1 m_1;\ell_2 m_2}^{(s)} \\
    i D_{L,-\p{m_1+sm_2}} & \t{otherwise}
  \end{cases},
\end{multline}
where
\begin{align}
  \zeta_{\ell_1 m_1;\ell_2 m_2}^{(s)}
  \equiv s \p{-1}^{\ell_1+\ell_2}
  \sign\p{m_1} \sign\p{sm_2} \sign\p{m_1+sm_2}.
\end{align}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Raman drive coefficients}
\label{sec:drive_raman_coeff}

Here we provide the coefficients $\Omega_{LM}^{(\phi,j)}$ of the
effective Raman driving Hamiltonian
\begin{align}
  H_{\t{Raman}}^{(\phi)}
  = \sum_j \sum_{\substack{L\in\set{0,1,2}\\\abs{M}\le L}}
  \Omega_{LM}^{(\phi,j)} D_{LM}^{(j)}
\end{align}
considered in Section \ref{sec:drive_raman}.  These coefficients are
computed using symmetry properties and structure constants of the
transition operator algebra (see Appendix \ref{sec:trans_prod}).  For
clarity, if any coefficient $\Omega_{LM}^{(\phi,j)}$ is independent of
$\phi$, we drop the explicit dependence on $\phi$, i.e.~taking
$\Omega_{LM}^{(\phi,j)}\to\Omega_{LM}$.
\begin{align}
  \Omega_{0,0} \equiv -\f1{\sqrt{n}}\,
  \sum_m\f{\Omega_m^2}{\Delta},
  &&
  \Omega_{1,0} \equiv \sqrt{\f{3}{n_1}}\,
  \f{\Omega_+^2 - \Omega_-^2}{\Delta},
  \label{eq:O_01_0}
\end{align}
\begin{align}
  \Omega_{2,0} \equiv \f1{\sqrt{5}}\, \f{\sqrt{n_2}}{n_1}\,
  \f{\Omega_+ + \Omega_- - 2\Omega_0}{\Delta},
  \label{eq:O_2_0}
\end{align}
\begin{align}
  \Omega_{L,1}^{(\phi,j)}
  &\equiv \sum_{s\in\set{\pm1}} \omega_{Ls} \cos\p{\phi_s j},
  &
  \Omega_{L,-1}^{(\phi,j)}
  &\equiv \sum_{s\in\set{\pm1}} s \omega_{Ls} \sin\p{\phi_s j}, \\
  \Omega_{2,2}^{(\phi,j)}
  &\equiv \omega_{2,2} \cos\p{\sp{\phi_+-\phi_-}j},
  &
  \Omega_{2,-2}^{(\phi,j)}
  &\equiv \omega_{2,2} \sin\p{\sp{\phi_+-\phi_-}j}, \\
  \omega_{1\pm}
  &\equiv -\sqrt{\f{6}{n_1}}\, \f{\Omega_0 \Omega_\pm}{\Delta},
  &
  \omega_{2\pm}
  &\equiv \mp \sqrt{\f65}\, \f{\sqrt{n_2}}{n_1}\,
  \f{\Omega_0\Omega_\pm}{\Delta}, \\
  \omega_{2,2}
  &\equiv 2\sqrt{\f35}\, \f{\sqrt{n_2}}{n_1}\,
  \f{\Omega_+\Omega_-}{\Delta},
  &
  n_L
  &\equiv \prod_{\ell=-L}^L \p{n+\ell}.
\end{align}
Here $\Omega_m$, $\phi_m$, and $\Delta$ are respectively the (real)
driving amplitudes, phases, and detuning that appear in the ``bare''
Raman drive in \eqref{eq:raman_bare}, and by choice of gauge we set
$\phi_-=0$, or equivalently re-define $\phi_m-\phi_0\to\phi_m$.  If
the driving lasers are phase-matched with $\phi_m=m\phi$, such that a
nuclear spin transition $\mu\to\mu+m$ on site $j$ imprints the phase
$-m\phi j$, then the drive $H_{\t{Raman}}^{(\phi)}$ takes the form
\begin{align}
  \left. H_{\t{Raman}}^{(\phi)} \right|_{\phi_m=m\phi}
  = \sum_{\substack{L\le2\\0\le M\le L}} \Omega_{LM} \D_{LM}^{(\phi)},
\end{align}
with inhomogeneously rotated collective drive operators
\begin{align}
  \D_{LM}^{(\phi)}
  \equiv \sum_j \sp{\cos\p{M\phi j} D_{LM}^{(j)}
    + \sin\p{M\phi j} D_{L,-M}^{(j)}},
\end{align}
and uniform coefficients given by \eqref{eq:O_01_0}, \eqref{eq:O_2_0},
and
\begin{align}
  \Omega_{1,-1} = \Omega_{2,-1} = \Omega_{2,-2} \equiv 0,
  &&
  \Omega_{L,1} \equiv \sum_{s\in\set{\pm1}} \omega_{Ls},
  &&
  \Omega_{2,2} \equiv \omega_{2,2}.
\end{align}

\bibliography{sun_notes.bib}

\end{document}