### 2019-10-23 collective_sun/multilevel_spin_notes.tex
added derivation of perturbative effect of SOC on the fully symmetric manifold

\documentclass[nofootinbib,notitlepage,11pt]{revtex4-2}

%%% linking references
\usepackage{hyperref}
\hypersetup{
  breaklinks=true,
  colorlinks=true,
  linkcolor=blue,
  filecolor=magenta,
  urlcolor=cyan,
}

%%% header / footer
\usepackage{fancyhdr} % easier header and footer management
\pagestyle{fancy} % page formatting style
\fancyhf{} % clear all header and footer text
\renewcommand{\headrulewidth}{0pt} % remove horizontal line in header
\usepackage{lastpage} % for referencing last page
\cfoot{\thepage~of \pageref{LastPage}} % "x of y" page labeling

%%% symbols, notations, etc.
\usepackage{physics,braket,bm,amssymb} % physics and math
\renewcommand{\t}{\text} % text in math mode
\newcommand{\f}[2]{\dfrac{#1}{#2}} % shorthand for fractions
\newcommand{\p}[1]{\left(#1\right)} % parenthesis
\renewcommand{\sp}[1]{\left[#1\right]} % square parenthesis
\renewcommand{\set}[1]{\left\{#1\right\}} % curly parenthesis
\newcommand{\bk}{\Braket} % shorthand for braket notation
\renewcommand{\v}{\bm} % bold vectors
\newcommand{\uv}[1]{\bm{\hat{#1}}} % unit vectors
\newcommand{\av}{\vec} % arrow vectors
\renewcommand{\d}{\text{d}} % for infinitesimals
\renewcommand{\c}{\cdot} % inner product

\usepackage{dsfont} % for identity operator
\newcommand{\1}{\mathds{1}}

\newcommand{\up}{\uparrow}
\newcommand{\dn}{\downarrow}

\newcommand{\x}{\text{x}}
\newcommand{\y}{\text{y}}
\newcommand{\z}{\text{z}}

\newcommand{\B}{\mathcal{B}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\E}{\mathcal{E}}
\renewcommand{\H}{\mathcal{H}}
\newcommand{\I}{\mathcal{I}}
\newcommand{\M}{\mathcal{M}}
\newcommand{\N}{\mathcal{N}}
\renewcommand{\O}{\mathcal{O}}
\renewcommand{\P}{\mathcal{P}}
\newcommand{\Q}{\mathcal{Q}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\T}{\mathcal{T}}
\renewcommand{\S}{\mathcal{S}}
\newcommand{\X}{\mathcal{X}}
\newcommand{\Z}{\mathcal{Z}}

\newcommand{\EE}{\mathbb{E}}
\renewcommand{\SS}{\mathbb{S}}
\newcommand{\ZZ}{\mathbb{Z}}

\newcommand{\FS}{\text{FS}}

\DeclareMathOperator{\sign}{sign}
\let\var\relax
\DeclareMathOperator{\var}{var}

\def\obra#1{\mathinner{({#1}|}}
\def\oket#1{\mathinner{|{#1})}}
\def\obk#1{\mathinner{({#1})}}
\def\oop#1#2{\oket{#1}\!\obra{#2}}

\usepackage[inline]{enumitem} % in-line lists and \setlist{} (below)
\setlist[enumerate,1]{label={(\roman*)}} % default in-line numbering
\setlist{nolistsep} % more compact spacing between environments

%%% text markup
\usepackage{color} % text color
\newcommand{\red}[1]{{\color{red} #1}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\thispagestyle{fancy}

\title{Multilevel generalizations of collective spin dynamics and
  squeezing}%
\author{Michael A. Perlin}%
\date{\today}

\maketitle

\tableofcontents

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Collective SU($n$)-symmetric interactions}
\label{sec:int}

We consider a collection of $n$-level fermions on a 1-D lattice, with
a single-particle Hamiltonian
\begin{align}
  H_{\t{lat}} = -t \sum_{q,\mu} \cos\p{qa} c_{q\mu}^\dag c_{q\mu},
  \label{eq:H_lat}
\end{align}
where $t$ is the nearest-neighbor tunneling rate, $a$ is the spacing
between neighboring lattice sites, and $c_{q\mu}$ is the fermionic
annihilation operator for a particle in quasi-momentum mode $q$ with
internal state (e.g.~nuclear spin) $\mu$.  Under the frozen-mode
approximation, collective two-body SU($n$)-symmetric interactions
between these fermions take form
\begin{align}
  H_{\t{int}} = \f{u}{2} \sum_{p,q,\mu,\nu}
  \p{c_{p\mu}^\dag c_{q\nu}^\dag c_{q\nu} c_{p\mu}
    + c_{q\mu}^\dag c_{p\nu}^\dag c_{q\nu} c_{p\mu}},
\end{align}
where $u\equiv U/L$ is the two-body on-site interaction energy $U$
divided by the number of lattice sites $L$.  Defining the spin
operators $S_{\mu\nu}^{(p)}\equiv c_{p\mu}^\dag c_{p\nu}$ for each
quasi-momentum $p$, we can alternately write
\begin{align}
  H_{\t{int}} = \f{u}{2} \sum_{p,q,\mu,\nu}
  \p{S_{\mu\mu}^{(p)} S_{\nu\nu}^{(q)} - S_{\mu\nu}^{(p)} S_{\nu\mu}^{(q)}}
  = \f{u}{2}\p{N^2 - \sum_{p,q,\mu,\nu}
    S_{\mu\nu}^{(p)} S_{\nu\mu}^{(q)}},
\end{align}
where $N$ is the total number of particles.  In the absence of
coherence between sectors of different particle number, we can neglect
the $\sim N^2$ term and simply write
\begin{align}
  H_{\t{int}}
  = - \f{u}{2} \sum_{p,q,\mu,\nu} S_{\mu\nu}^{(p)} S_{\nu\mu}^{(q)}.
\end{align}
In the case of SU(2), at this point we would expand the spin operators
$S_{\mu\nu}^{(p)}$ in terms of Pauli operators in order to convert
$H_{\t{int}}$ into an SU(2) spin Hamiltonian\cite{he2019engineering}.
To generalize this procedure to the case of SU($n$), we first define a
vector $\v S^{(p)}$ of all spin operators $S_{\mu\nu}^{(p)}$, and
write
\begin{align}
  H_{\t{int}}
  = - \f{u}{2} \sum_{p,q} {\v S^{(p)}}^\dag \c \v S^{(q)}
  = - \f{u}{2} \v\S^\dag \c \v\S,
  &&
  \v\S\equiv \sum_p \v S^{(p)}.
  \label{eq:H_int_spin}
\end{align}
In the bulk of this section, we identify various bases for the space
of linear operators on the Hilbert space $\H_n$ of an $n$-level
system.  For brevity, we denote this space of operators by
$\B\p{\H_n}$, and note that $\B\p{\H_n}$ is itself an
$n^2$-dimensional Hilbert space equipped with a trace
(Hilbert-Schmidt) inner product
\begin{align}
  \obk{\O|\Q} \equiv \tr\p{\O^\dag Q},
\end{align}
where $\oket{\Q}$ and $\obra{\O}$ respectively denote vectors in
$\B\p{\H_n}$ and its dual space $\B\p{\H_n}^*$.  Similarly to the set
of all spin operators $S_{\mu\nu}\equiv\op{\mu}{\nu}$ on the
single-particle Hilbert space at each quasi-momentum, all of the bases
we consider will be orthonormal with respect to the trace inner
product.  As we show in Appendix \ref{sec:changing_bases}, the
collective SU($n$)-symmetric interaction Hamiltonian in
\eqref{eq:H_int_spin} takes an identical form in any orthonormal basis
for $\B\p{\H_n}$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Generalized Gell-Mann (GGM) operators}
\label{sec:ggm_ops}

The most obvious generalization of Pauli operators to the case of
SU($n$) are the {\it generalized Gell-Mann (GGM)
  operators}\cite{hioe1981level, bertlmann2008bloch}, defined for
$j,k,\ell\in\set{0,1,\cdots,n-1}$ with $j<k$ and $\ell>0$ by
\begin{align}
  \lambda_0 \equiv \sqrt{\f{2}{n}}~ \1,
  &&
  \lambda_\ell \equiv \sqrt{\f{2}{\ell\p{\ell+1}}}
  \p{\sum_{n=0}^{\ell-1}\op{n} - \ell \op{\ell}},
  \label{eq:GGM_diag}
\end{align}
\begin{align}
  \lambda_{jk,\x} \equiv \op{j}{k} + \op{k}{j},
  &&
  \lambda_{jk,\y} \equiv i\p{\op{j}{k} - \op{k}{j}},
  \label{eq:GGM_off_diag}
\end{align}
where $\1$ is the identity operator.  The self-adjoint GGM operators
in \eqref{eq:GGM_diag} and \eqref{eq:GGM_off_diag} provide a complete
basis for the space of operators on the Hilbert space $\H_n$ of an
$n$-level system, coincide exactly with the Pauli operators in the
case of $n=2$, and satisfy the orthonormality condition
\begin{align}
  \obk{\lambda_a|\lambda_b} = 2\delta_{ab},
  \label{eq:GGM_inner}
\end{align}
where $a,b$ index any GGM operator in \eqref{eq:GGM_diag} or
\eqref{eq:GGM_off_diag}.  Denoting a vector of all GGM operators by
$\v\lambda$, we can therefore write (see Appendix
\ref{sec:changing_bases})
\begin{align}
  H_{\t{int}}
  = -\f{u}{4} \sum_{p,q} \v\lambda^{(p)} \c \v\lambda^{(q)}
  = -\f{u}{4} \v\Lambda \c \v\Lambda,
  &&
  \v\Lambda \equiv \sum_p \v\lambda^{(p)}.
  \label{eq:H_int_GGM}
\end{align}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Transition operators}
\label{sec:trans_ops}

The GGM operators in \eqref{eq:GGM_diag} and \eqref{eq:GGM_off_diag}
provide a convenient operator basis to describe dynamics obeying
SU($n$) symmetry.  An external driving field addressing $n$-level
nuclear spins, however, will generally violate SU($n$) symmetry, and
instead obeys the symmetries of the {\it polarization} or {\it
  transition operators} defined by\cite{kryszewski2006alternative,
  bertlmann2008bloch}
\begin{align}
  T_{LM}
  \equiv \sqrt{\f{2L+1}{2I+1}} \sum_{\mu,\nu}
  \bk{I\mu;LM|I\nu} \op{\nu}{\mu},
  \label{eq:trans_ops}
\end{align}
where $L\in\set{0,1,\cdots,n-1}$ and $M\in\set{-L,-L+1,L}$ index a
total spin and its projection onto a quantization axis;
$I\equiv\p{n-1}/2$ is the maximal angular momentum of an $n$-level
spin; $\mu,\nu\in\set{-I,-I+1,\cdots,I}$ index projections of an
$n$-level nuclear spin onto a quantization axis; and
$\bk{I\mu;LM|I\nu}$ is a Clebsch-Gordan coefficient.  The transition
operator $T_{LM}$ is closely related to the associated Legendre
polynomial $P_{LM}\p{x}$ for $x\in\sp{-1,1}$.  Operationally, $T_{LM}$
is proportional to the transition induced on a nuclear spin by the
absorption of a spin-$L$ boson with spin projection $M$ onto a
quantization axis.

Similarly to the GGM operators in \eqref{eq:GGM_diag} and
\eqref{eq:GGM_off_diag}, the transition operators in
\eqref{eq:trans_ops} provide a complete basis for the space of
operators on the Hilbert space $\H_n$ of an $n$-level system, and
satisfy the orthonormality condition
\begin{align}
  \obk{T_{LM}|T_{L'M'}} = \delta_{LL'} \delta_{MM'},
\end{align}
which implies
\begin{align}
  H_{\t{int}} = -\f{u}{2} \sum_{p,q} {\v T^{(p)}}^\dag \c \v T^{(q)}
  = -\f{u}{2} \v\T^\dag \c \v\T,
  &&
  \v\T \equiv \sum_p\v T^{(p)},
  \label{eq:H_int_trans}
\end{align}
where $\v T$ is a vector of all transition operators in
\eqref{eq:trans_ops}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Drive operators}
\label{sec:drive_ops}

Unlike the GGM operators in \eqref{eq:GGM_diag} and
\eqref{eq:GGM_off_diag}, the transition operators in
\eqref{eq:trans_ops} are not self-adjoint.  In order to express
Hamiltonians in terms of self-adjoint operators, we define the {\it
  drive operators}
\begin{align}
  D_{LM} \equiv \f{\eta_M}{\sqrt{2}}
  \sp{T_{L\abs{M}} + \sign\p{M} T_{L\abs{M}}^\dag},
  &&
  \eta_M \equiv
  \begin{cases}
    \sqrt{2} & M = 0 \\
    \p{-1}^M \sign\p{M}^{3/2} & M \ne 0
  \end{cases},
  \label{eq:drive_ops}
\end{align}
for integer $M$ with $\abs{M}\le L$.  The drive operator $D_{LM}$ is
proportional to the Hamiltonian induced on a nuclear spin by an
external classical field of spin-$L$ particles with spin projection
$M$ onto a quantization axis.  Similarly to the GGM operators in
\eqref{eq:GGM_diag} and \eqref{eq:GGM_off_diag}, the drive operators
are proportional to the Pauli operators in the case of $n=2$, and
satisfy the orthonormality condition
\begin{align}
  \obk{D_{LM}|D_{L'M'}} = \delta_{LL'}\delta_{MM'},
\end{align}
which implies
\begin{align}
  H_{\t{int}} = -\f{u}{2} \sum_{p,q} \v D^{(p)}\c\v D^{(q)}
  = -\f{u}{2} \v\D \c \v\D,
  &&
  \v\D \equiv \sum_p \v D^{(p)},
  \label{eq:H_int_drive}
\end{align}
where $\v D$ is a vector of all drive operators in
\eqref{eq:drive_ops}.

The drive operators $D_{1,0}$ and $D_{1,\pm1}$ are proportional to the
Hamiltonians induced on an $n$-level spin by an external magnetic
field and a classical spin-polarized driving laser; loosely speaking,
one can identify $\p{D_{1,0},D_{1,1},D_{1,-1}}\sim\p{S_\z,S_\x,S_\y}$,
where $S_\alpha$ is a collective SU(2) spin-$\alpha$ operator for
$\p{n-1}$ 2-level spins\cite{perlin2019shorttime}.  More concretely,
we can embed the Hilbert space of a single $n$-level spin into the
permutationally-symmetric (Dicke) manifold of states for $\p{n-1}$
2-level spins via
\begin{align}
  \ket{\mu}_{\t{single}} \to \ket{I+\mu}_{\t{collective}}
  \propto S_+^{I+\mu} \ket\dn^{\otimes\p{n-1}},
\end{align}
where $I\equiv\p{n-1}/2$ is a total spin,
$\mu\in\set{-I,-I+1,\cdots,I}$ indexes a state of the single $n$-level
spin, and $S_+$ is a collective spin-raising operator for $\p{n-1}$
2-level spins.  This embedding takes
\begin{align}
  \p{D_{1,0},D_{1,1},D_{1,-1}}
  \to \p{S_\z,S_\x,S_\y}
  \times \f{2\sqrt{3}}{\sqrt{\p{n-1}n\p{n+1}}}.
  \label{eq:spin_ops}
\end{align}
Furthermore, if we {\it define} the operators $S_\alpha$ for a single
$n$-level spin via the correspondence in \eqref{eq:spin_ops}, then the
familiar SU(2) transformation rule
\begin{align}
  e^{i\phi S_\z} S_\x e^{-i\phi S_\z} = S_\x^{(\phi)},
  &&
  S_\x^{(\phi)} \equiv \cos\phi S_\x - \sin\phi S_\y,
\end{align}
generalizes to
\begin{align}
  e^{i\phi S_\z} D_{LM} e^{-i\phi S_\z} = D_{LM}^{(\phi)},
  &&
  D_{LM}^{(\phi)} \equiv \cos\p{M\phi} D_{LM} - \sin\p{M\phi} D_{L,-M}.
\end{align}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Weyl operators}

Weyl operators\cite{bertlmann2008bloch} and self-adjoint linear
combinations thereof\cite{asadian2016heisenbergweyl} also provide an
interesting basis of operators for finite-dimensional Hilbert spaces.
These operators are related to the notions of position, momentum, and
phase-space displacement in a discrete space.  As our present work is
motivated by multilevel systems that are realized with nuclear spin
(angular momentum) degrees of freedom, however, the transition and
drive operators are more natural to use than the Weyl operators, so we
will not discuss Weyl operators any further.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Subsystem permutations and the fully symmetric manifold}
\label{sec:perm_ops}

For our final re-expression of collective SU($n$)-symmetric
interactions, we abandon the use of single-spin operators entirely to
express the interaction Hamiltonian $H_{\t{int}}$ in terms of
operators that permute tensor factors (subsystems) of a Hilbert space
with an $N$-fold tensor product structure (e.g.~an array of nuclear
spins).  Specifically, we denote the Hilbert space of an $n$-level
system by $\H_n$, denote its $N$-fold tensor product by
$\H_n^{\otimes N}$, and identify the set $\Pi\equiv\set{\Pi_\sigma}$
of operators that permute tensor factors of $\H_n^{\otimes N}$
according to elements $\sigma\in\SS_N$ of the symmetric group of order
$N$.  We then expand
\begin{align}
  H_{\t{int}}
  = -\f{u}{2} \sum_{p,q,\mu,\nu} S_{\mu\nu}^{(p)} S_{\nu\mu}^{(q)}
  = -u\sum_{\substack{p<q\\\mu,\nu}} S_{\mu\nu}^{(p)} S_{\nu\mu}^{(q)}
  - \f{u}{2} \sum_{p,\mu,\nu} S_{\mu\nu}^{(p)} S_{\nu\mu}^{(p)},
  \label{eq:H_int_perm_start}
\end{align}
where the first sum,
\begin{align}
  \sum_{\substack{p<q\\\mu,\nu}} S_{\mu\nu}^{(p)} S_{\nu\mu}^{(q)}
  = \sum_{\substack{p<q\\\mu,\nu}} \op{\mu}{\nu}^{(p)} \op{\nu}{\mu}^{(q)}
  = \sum_{\substack{p<q\\\mu,\nu}} \op{\mu\nu}{\nu\mu}^{(p,q)}
  = \sum_{p<q} \Pi_{pq}
  \label{eq:spin_perm}
\end{align}
is simply a sum over all permutations $\Pi_{pq}$ of subsystems $p$ and
$q$; and
\begin{align}
  \sum_{p,\mu,\nu} S_{\mu\nu}^{(p)} S_{\nu\mu}^{(p)}
  = \sum_{p,\mu,\nu} \op{\mu}{\nu}^{(p)} \op{\nu}{\mu}^{(p)}
  = n \sum_{p,\mu} \op{\mu}{\mu}^{(p)}
  = n \sum_p \1^{(p)}
  = n N
  \label{eq:spin_const}
\end{align}
is merely a constant.  Substituting \eqref{eq:spin_perm} and
\eqref{eq:spin_const} into \eqref{eq:H_int_perm_start} yields
\begin{align}
  H_{\t{int}} = -u \sum_{p<q} \Pi_{pq} - \f12 N n u
  \simeq -u\sum_{p<q} \Pi_{pq},
  \label{eq:H_int_perm}
\end{align}
where $\simeq$ denotes equality up to an overall constant with no
physical consequence.  The form of the interaction Hamiltonian
$H_{\t{int}}$ makes it evident that its ground-state manifold
$\M_{\FS}$ is the set of {\it fully symmetric} (FS) states that have
eigenvalue $1$ with respect to all pairwise permutation operators
$\Pi_{pq}$, and thereby also all $\Pi_\sigma\in\Pi$.  The fully
symmetric manifold $\M_\FS$ is spanned by a basis $\set{\ket{\v m}}$
of fully symmetric states that can be uniquely identified by a list of
integers $\v m\equiv\p{m_0,m_1,\cdots,m_{n-1}}\in\ZZ_{N+1}^n$, where
$m_s$ indicates the total occupation number of subsystem state $s$,
and $\sum_sm_s=N$.  In the case of $n=2$, the fully symmetric manifold
$\M_\FS$ is precisely the Dicke manifold\cite{dicke1954coherence} of
$N$ qubits, spanned by Dicke states $\set{\ket{m_0,m_1}}$ for which
$m_0+m_1=N$ and the integer $m_0$ ($m_1$) indicates the total
occupation number of qubit state $\ket{0}$ ($\ket{1}$).  The dimension
of fully symmetric manifold $\M_\FS$ is determined by the number of
ways to assign $N$ subsystems to $n$ states, which is
${N+n-1 \choose n-1}$, and the energy of a state
$\ket{\psi_\FS}\in\M_\FS$ with respect to the interaction Hamiltonian
$H_{\t{int}}$ in \eqref{eq:H_int_perm} is
$-u{N \choose 2}=-\p{1/4}N\p{N-1}u$.

In addition to the fact that the fully symmetric manifold $\M_\FS$ is
ground-state manifold of the interaction Hamltonian $H_{\t{int}}$, an
important feature of $\M_\FS$ is that it is closed under the action of
unitaries generated by collective operators of the form
$\Q=\sum_pQ^{(p)}$, as well as linear combinations and products
thereof.  Closure of $\M_\FS$ under collective dynamics is a
straightforward consequence of the permutational symmetry obeyed by
collective operators.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{External drive and spin-orbit coupling}
\label{sec:drive_SOC}

A driving laser addressing spin-$1/2$ fermions on a 1-D lattice will
induce the Hamiltonian
\begin{align}
  \left. H_{\t{drive}}^{(\phi)} \right|_{n=2}
  = \Omega \sum_j
  \p{e^{i\phi j} c_{j,\up}^\dag c_{j,\dn}
    + e^{-i\phi j} c_{j,\dn}^\dag c_{j,\up}},
  \label{eq:H_drive_2}
\end{align}
where $\Omega$ is a (real) driving amplitude, $j\in\mathbb{Z}$ indexes
a lattice site, and $\phi$ is the relative phase of the drive between
adjacent lattice sites.  The multilevel analogue of
\eqref{eq:H_drive_2} is
\begin{align}
  H_{\t{drive}}^{(\phi)}
  = -\Omega \sum_j \p{e^{i\phi j} T_{1,1}^{(j)}
    + e^{-i\phi j} {T_{1,1}^{(j)}}^\dag}
  = \Omega \D_{1,1}^{(\phi)},
  \label{eq:H_drive}
\end{align}
where
\begin{align}
  \D_{LM}^{(\phi)}
  \equiv \sum_j \sp{\cos\p{M\phi j} D_{LM}^{(j)}
    - \sin\p{M\phi j} D_{L,-M}^{(j)}},
  \label{eq:drive_rot}
\end{align}
is an inhomogeneously rotated collective drive operator.  We can make
this drive spatially homogenous via the gauge transformation
$c_{j\mu}^\dag \to e^{-i\mu\phi j} c_{j\mu}^\dag$, which takes
 \begin{align}
  H_{\t{drive}}^{(\phi)} \to H_{\t{drive}} = \Omega \D_{1,1},
  &&
  \D_{LM} \equiv \D_{LM}^{(0)} = \sum_j D_{LM}^{(j)}.
\end{align}
The SU($n$) symmetry of collective interactions considered in this
work implies that the interaction Hamiltonian $H_{\t{int}}$ in
\eqref{eq:H_int_perm} is unaffected by this gauge transformation.  The
single-particle Hamiltonian $H_{\t{lat}}$ in \eqref{eq:H_lat},
however, transforms as (see Appendix \ref{sec:lat_drive})
\begin{align}
  H_{\t{lat}}
  \to H_{\t{lat}}^{(\phi)}
  = -t \sum_{q,\mu} \cos\p{qa+\mu\phi} c_{q\mu}^\dag c_{q\mu}
  = \sum_{q,L} B_L^{(\phi,q)} D_{L,0}^{(q)},
  \label{eq:H_lat_SOC_B}
\end{align}
with effective inhomogeneous driving field amplitudes
\begin{align}
  B_L^{(\phi,q)} = -t w_L\p{qa} A_L^{(\phi)},
\end{align}
where
\begin{align}
  w_L\p{\theta} \equiv
  \begin{cases}
    \cos\theta & L~\t{even} \\
    \sin\theta & L~\t{odd}
  \end{cases},
  &&
  A_L^{(\phi)} \equiv \p{-1}^L \sqrt{\f{2L+1}{2I+1}}
  \sum_\mu \bk{I\mu;L,0|I\mu} w_L\p{\mu\phi}.
  \label{eq:A_L_phi}
\end{align}
The multilevel spin-orbit coupling (SOC) exhibited by the
singe-particle Hamiltonian $H_{\t{lat}}^{(\phi)}$ thus allows for the
simulation of synthetic gauge fields with large spin ($L>1$).  As we
will see in Section \ref{sec:drive_raman}, multi-laser drives can also
simulate synthetic large-spin fields with non-zero spin projection
($M\ne0$) onto a quantization axis.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Perturbative treatment of spin-orbit coupling}
\label{sec:pert_SOC}

The single-particle Hamiltonian $H_{\t{lat}}^{(\phi)}$ breaks
permutational symmetry, thereby coupling states within the fully
symmetric manifold $\M_\FS$ to states outsides it.  If SOC is
sufficiently weak, however, transitions outside the fully symmetric
manifold $\M_\FS$ are energetically suppressed by interactions, and we
can account for the effect of SOC perturbatively.  In the regime of
weak SOC with $\abs{\phi}\ll1$, by considering
\begin{enumerate*}
\item all $L\le12$ for arbitrary $n$, and
\item all $L<n$ for all $n\le25$
\end{enumerate*}
(and conjecturing the same functional form for all $L$ and $n$), we
find that
% expand Clebsch-Gordan coefficients in a power series according to
% \url{functions.wolfram.com/07.38.06.0008.01}
\begin{align}
  A_L^{(\phi)} = c_L \sqrt{n_L}\, \phi^L + O\p{\phi^{L+2}},
  \label{eq:A_L_phi_small}
\end{align}
where
\begin{align}
  c_L \equiv \p{-1}^{L+\lfloor L/2\rfloor}
  \sqrt{2L+1}\, \f{L!}{\p{2L+1}!},
  &&
  n_L \equiv \prod_{\ell=-L}^L\p{n+\ell}.
\end{align}

\vspace{3cm}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Raman drive}
\label{sec:drive_raman}

Rather than coupling the states of a multilevel system directly with a
single driving laser, as in Section \ref{sec:drive_SOC}, we now
consider coupling through an off-resonant Raman transition to a highly
excited state.  Specifically, we consider a three-laser drive of the
form
\begin{align}
  H_{\t{Raman,bare}}^{(\phi)}
  = \sum_j \sum_{m\in\set{1,0,-1}} \p{\Omega_m e^{i\phi_m j}
    T_{1,m}^{(j)} \otimes \op{\up}{\dn} + \t{h.c.}}
  + \f{\Delta}{2} \1 \otimes \p{\op{\up} - \op{\dn}},
  \label{eq:raman_bare}
\end{align}
where $j$ indexes a lattice site, $\Omega_m$ is a (real) driving
amplitude; $\phi_m$ is the relative phase of drive $m$ between
adjacent lattice sites; $\op{r}{s}$ for $r,s\in\set{\up,\dn}$ is a
pseudo-spin operator for the auxiliary degree of freedom that is
off-resonantly addressed by the lasers; $\Delta$ is a detuning of the
lasers from the energy of an excited state; and $\1$ is the identity
operator.  If the detuning is large, $\Delta\gg\Omega_m$, then at
second order in perturbation theory the effective action of
$H_{\t{Raman,bare}}^{(\phi)}$ within the $\ket{\dn}$-state manifold
takes the form
\begin{align}
  H_{\t{Raman}}^{(\phi)}
  = \sum_{\tau\in\set{0,1,2}} H_{\t{Raman},\tau}^{(\phi)}
\end{align}
where $H_{\t{Raman},\tau}^{(\phi)}$ generates a nuclear spin
transition $\mu\to\mu\pm\tau$:
\begin{align}
  H_{\t{Raman},0}^{(\phi)}
  &= -\sum_j\sum_m \f{\Omega_m^2}{\Delta}
  {T_{1,m}^{(j)}}^\dag T_{1,m}^{(j)}, & \label{eq:raman_0} \\
  H_{\t{Raman},1}^{(\phi)}
  &= -\sum_j\sum_{s\in\set{\pm1}}
  \f{\Omega_0\Omega_s}{\Delta} e^{i\p{\phi_s-\phi_0} j}
  {T_{1,0}^{(j)}}^\dag T_{1,s}^{(j)} + \t{h.c.},
  \\
  H_{\t{Raman},2}^{(\phi)}
  &= -\sum_j\sum_{s\in\set{\pm1}}
  \f{\Omega_+\Omega_-}{\Delta} e^{i\p{\phi_s-\phi_{-s}}j}
  {T_{1,-s}^{(j)}}^\dag T_{1,s}^{(j)}.
  \label{eq:raman_2}
\end{align}
In terms of the drive operators, we can expand
\begin{align}
  H_{\t{Raman}}^{(\phi)}
  = \sum_j \sum_{\substack{L\le2\\\abs{M}\le L}}
  \Omega_{LM}^{(\phi,j)} D_{LM}^{(j)},
  \label{eq:raman_drive}
\end{align}
where the coefficients $\Omega_{LM}^{(\phi,j)}$ are provided in
Appendix \ref{sec:drive_raman_coeff}.  If the driving lasers are
phase-matched with $\phi_m=m\phi$, such that a nuclear spin transition
$\mu\to\mu+m$ on site $j$ imprints the phase $m\phi j$, then the drive
$H_{\t{Raman}}^{(\phi)}$ takes the form
\begin{align}
  \left. H_{\t{Raman}}^{(\phi)} \right|_{\phi_m=m\phi}
  = \sum_{\substack{L\le2\\0\le M\le L}} \Omega_{LM} \D_{LM}^{(\phi)},
  \label{eq:drive_raman_matched}
\end{align}
with rotated collective drive operators $\D_{LM}^{(\phi)}$ defined in
\eqref{eq:drive_rot} and uniform driving amplitudes $\Omega_{LM}$
provided in Appendix \ref{sec:drive_raman_coeff}.  Similarly to the
case of a direct drive considered in Section \ref{sec:drive_SOC}, the
Raman drive $H_{\t{Raman}}^{(\phi)}$ in \eqref{eq:drive_raman_matched}
can be made homogenous through the gauge transformation
$c_{j\mu}^\dag \to e^{i\mu\phi j} c_{j\mu}^\dag$, which takes
\begin{align}
  \left. H_{\t{Raman}}^{(\phi)} \right|_{\phi_m=m\phi}
  \to H_{\t{Raman}}
  = \sum_{\substack{L\le2\\0\le M\le L}} \Omega_{LM} \D_{LM}.
\end{align}
The effective driving amplitudes $\Omega_{LM}$ can be tuned through
the bare driving amplitudes $\Omega_m$, and satisfy
\begin{align}
  \Omega_{0,0} &\propto \Omega_+^2 + \Omega_-^2 + \Omega_0^2,
  &
  \Omega_{1,0} &\propto \Omega_+^2 - \Omega_-^2,
  &
  \Omega_{1,1} &\propto \Omega_0 \p{\Omega_+ + \Omega_-}, \\
  \Omega_{2,0} &\propto \Omega_+^2 + \Omega_-^2 - 2\Omega_0^2,
  &
  \Omega_{2,1} &\propto \Omega_0 \p{\Omega_+ - \Omega_-},
  &
  \Omega_{2,2} &\propto \Omega_+ \Omega_-.
\end{align}
In particular, the amplitudes $\Omega_{1,0}$, $\Omega_{2,0}$, and
$\Omega_{2,1}$ can be simultaneously tuned to zero by setting
$\Omega_0=\Omega_+=\Omega_-$, at which point the only remaining
effective drives in $H_{\t{Raman}}$ are $\sim\D_{1,1}$ and
$\sim\D_{2,2}$ (as $\D_{0,0}\propto\1$ is just constant shift in
energy).  If instead $\Omega_0=\pm\Omega_\pm$, then the amplitudes
$\Omega_{1,0}$, $\Omega_{2,0}$ and $\Omega_{1,1}$ vanish, leaving only
the drives $\D_{2,1}$ and $\D_{2,2}$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Future directions}

\begin{itemize}
\item Visualization with generalized Bloch spheres
  \begin{itemize}
  \item Single-particle eigenstates of drive operators
  \item Single-particle dynamics induced by drive operators
  \end{itemize}
\item Classifying collective states (e.g.~eigenstates of interactions,
  spin-wave-like states)
\item Implications of SOC
  \begin{itemize}
  \item Special angles $\phi$
  \end{itemize}
\item Generalizations of squeezing beyond SU(2)?
\item Consider different interactions
  \begin{itemize}
  \item Include electronic states
  \item Super-exchange regime
  \item Non-uniform (``disordered'') interactions
  \end{itemize}
\item Consider different drives
\end{itemize}

\appendix

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Changing operator bases for SU($n$)-symmetric interactions}
\label{sec:changing_bases}

Here we show that the collective interaction Hamiltonian in
\eqref{eq:H_int_spin} takes an identical form in any orthonormal basis
for $\B\p{\H_n}$, i.e.~the space of linear operators on the Hilbert
space $\H_n$ of an $n$-level system, equipped with the trace
(Hilbert-Schmidt) inner product
\begin{align}
  \obk{\O|\Q} \equiv \tr\p{\O^\dag \Q}.
\end{align}
To maintain generality, we choose an arbitrary basis
$X\equiv\set{X_j}$ for $\B\p{\H_n}$ that satisfies the orthonormality
condition
\begin{align}
  \obk{X_j|X_k} = \N_X \delta_{jk},
\end{align}
for $j\in\set{0,1,\cdots,n^2-1}$.  This orthonormality condition
implies that we can resolve the identity (super-)operator $\I$ with
$\I\oket{\O}=\oket{\O}$ for any $\oket\O\in\B\p{\H_n}$ as
\begin{align}
  \I = \f1{\N_X} \sum_j \oop{X_j}{X_j}
  = \f1{\N_X} \sum_j \oop{X_j^\dag}{X_j^\dag},
\end{align}
where we used the fact that $X^\dag\equiv\set{X_j^\dag}$ is also an
orthonormal basis for $\B\p{\H_n}$, satisfying the same orthonormality
condition as $X$; these two bases are transformed into each other by
unitary $\sum_j\oop{X_j^\dag}{X_j}/\N_X$.

The collective interaction Hamiltonian in \eqref{eq:H_int_spin} takes
the form
\begin{align}
  H = \sum_{p,q,j} X_j^{(p)} {X_j^{(q)}}^\dag,
\end{align}
which essentially consists of single-body terms ($p=q$) and two-body
interactions ($p\ne q$) of the form
\begin{align}
  \sum_j X_j X_j^\dag,
  &&
  \sum_j X_j \otimes X_j^\dag.
\end{align}
Choosing an arbitrary orthonormal basis $Y\equiv\set{Y_j}$ for
$\B\p{\H_n}$ with corresponding norm $\N_Y$, we can expand
\begin{align}
  X_j = \f1{\N_Y} \sum_k Y_k \obk{Y_k|X_j},
  &&
  X_j^\dag = \f1{\N_Y} \sum_k Y_k^\dag \obk{Y_k^\dag|X_j^\dag}
  = \f1{\N_Y} \sum_k Y_k^\dag \obk{X_j|Y_k},
\end{align}
where we used the fact that
\begin{align}
  \obk{Y_k^\dag|X_j^\dag}
  = \tr\p{Y_k X_j^\dag}
  = \tr\p{X_j^\dag Y_k}
  = \obk{X_j|Y_k}.
\end{align}
Denoting an arbitrary bilinear operation by $\odot$, by resolution of
the identity we therefore find
\begin{align}
  \sum_j X_j \odot X_j^\dag
  = \f1{\N_Y^2} \sum_{j,k,\ell} Y_k \obk{Y_k|X_j}
  \odot Y_\ell^\dag \obk{X_j|Y_\ell}
  = \f{\N_X}{\N_Y^2} \sum_{k,\ell} Y_k \odot Y_\ell^\dag
  \obk{Y_k|\I|Y_\ell}
  = \f{\N_X}{\N_Y} \sum_k Y_k \odot Y_k^\dag,
\end{align}
which in particular implies that
\begin{align}
  \sum_j X_j X_j^\dag
  = \f{\N_X}{\N_Y} \sum_k Y_k Y_k^\dag,
  &&
  \sum_j X_j \otimes X_j^\dag
  = \f{\N_X}{\N_Y} \sum_k Y_k \otimes Y_k^\dag,
\end{align}
and in turn
\begin{align}
  H = \f{\N_X}{\N_Y} \sum_{p,q,j} Y_j^{(p)} {Y_j^{(q)}}^\dag.
\end{align}
The collective interaction Hamiltonian in \eqref{eq:H_int_spin} thus
takes an identical form in any orthonormal basis for the space
$\B\p{\H_n}$ of linear operators on the Hilbert space $\H_n$ of an
$n$-level system, up to appropriate re-scaling factors $\N_X/\N_Y$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Spin-orbit coupling with drive operators}
\label{sec:lat_drive}

Here we derive the drive-operator ($D_{LM}$) expansion of the
spin-orbit coupled Hamiltonian $H_{\t{lat}}^{(\phi)}$ in
\eqref{eq:H_lat_SOC_B}.  In the basis of spin operators
$S_{\mu\mu}^{(q)} \equiv c_{q\mu}^\dag c_{q\mu}$, this Hamiltonian
takes the form
\begin{align}
  H_{\t{lat}}^{(\phi)}
  = -t \sum_{q,\mu} \cos\p{qa+\mu\phi} S_{\mu\mu}^{(q)}
  \label{eq:H_lat_spin}
\end{align}
Defining the diagonal operators
\begin{align}
  \tilde d_{\phi+} \equiv \sum_\mu \cos\p{\mu\phi} S_{\mu\mu},
  &&
  \tilde d_{\phi-} \equiv \sum_\mu \sin\p{\mu\phi} S_{\mu\mu},
\end{align}
which are respectively even and odd under spin inversion
($\mu\to-\mu$), we can expand
\begin{align}
  \sum_\mu \cos\p{qa+\mu\phi} S_{\mu\mu}
  = \cos\p{qa} \tilde d_{\phi+} - \sin\p{qa} \tilde d_{\phi-},
\end{align}
and resolve the identity (super-)operator $\I$ with
$\I \tilde d_{\phi\pm} = \tilde d_{\phi\pm}$ in the basis of the drive operators
$D_{LM}$ (see Appendix \ref{sec:changing_bases}), finding
\begin{align}
  \sum_\mu \cos\p{qa+\mu\phi} S_{\mu\mu}
  = \cos\p{qa} \f12 \sum_{L,M} \obk{\tilde d_{\phi+}|D_{LM}} D_{LM}
  - \sin\p{qa} \f12 \sum_{L,M} \obk{\tilde d_{\phi-}|D_{LM}} D_{LM}.
\end{align}
We now note that the drive operators $D_{LM}$ are strictly
off-diagonal for $M\ne0$, and furthermore that $D_{L,0}$ with even
(odd) $L$ is even (odd) under spin inversion, which implies
\begin{align}
  \sum_{L,M} \obk{\tilde d_{\phi+}|D_{LM}} D_{LM}
  = \sum_{L\,\t{even}} \obk{\tilde d_{\phi+}|D_{L,0}} D_{L,0},
\end{align}
and likewise with $\tilde d_{\phi-}$ and odd $L$.  In total, we can
write
\begin{align}
  \sum_\mu \cos\p{qa+\mu\phi} S_{\mu\mu}
  = \cos\p{qa} \f12 \sum_{L\,\t{even}}
  \obk{\tilde d_{\phi+}|D_{L,0}} D_{L,0}
  - \sin\p{qa} \f12 \sum_{L\,\t{odd}}
  \obk{\tilde d_{\phi-}|D_{L,0}} D_{L,0},
\end{align}
or, more compactly,
\begin{align}
  \sum_\mu \cos\p{qa+\mu\phi} S_{\mu\mu}
  = \sum_L w_L\p{qa} A_L^{(\phi)} D_{L,0},
  &&
  w_L\p{\theta} \equiv
  \begin{cases}
    \cos\theta & L~\t{even} \\
    \sin\theta & L~\t{odd}
  \end{cases},
  \label{eq:compact_SOC_drive}
\end{align}
with effective driving field amplitudes
\begin{align}
  A_L^{(\phi)}
  \equiv \p{-1}^L \f12 \obk{\tilde d_{\phi,\p{-1}^L}|D_{L,0}}
  = \p{-1}^L \sqrt{\f{2L+1}{2I+1}}
  \sum_\mu \bk{I\mu;L,0|I\mu} w_L\p{\mu\phi}.
\end{align}
Substituting the result in \eqref{eq:compact_SOC_drive} into the
single-particle Hamiltonian in \eqref{eq:H_lat_spin} yields
\begin{align}
  H_{\t{lat}}^{(\phi)} = \sum_{q,L} B_L^{(\phi,q)} D_{L,0}^{(q)},
  &&
  B_L^{(\phi,q)} \equiv -J
  w_L\p{qa} A_L^{(\phi)}.
\end{align}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Transition operator product expansion}
\label{sec:trans_prod}

Here we derive multiplication rules for transition operators on an $n$-dimensional Hilbert space, which allows us to expand
\begin{align}
  T_{\ell_1 m_1} T_{\ell_2 m_2}
  = \sum_{L,M} f_{\ell_1 m_1;\ell_2 m_2}^{LM} T_{LM},
\end{align}
with structure constants
\begin{align}
  f_{\ell_1 m_1;\ell_2 m_2}^{LM}
  \equiv \obk{T_{LM} | T_{\ell_1 m_1} T_{\ell_2 m_2}}
  = \tr\p{T_{LM}^\dag T_{\ell_1 m_1} T_{\ell_2 m_2}}.
  \label{eq:trans_struct}
\end{align}
For reference, the transition operators $T_{LM}$ are defined for
integers $L,M$ with $0\le L<n$ and $\abs{M}\le L$ in terms of
Clebsch-Gordan coefficients $\bk{\ell_1 m_1;\ell_2 m_2|\ell_3 m_3}$ by
\begin{align}
  T_{LM} \equiv \sqrt{\f{2L+1}{2I+1}}
  \sum_{\mu,\nu} \bk{I\mu;LM|I\nu} \op{\nu}{\mu},
  &&
  I \equiv \f{n-1}{2}.
\end{align}
Using the symmetry properties of Clebsch-Gordan coefficients, namely
\begin{align}
  \bk{\ell_1 m_1; \ell_2 m_2| \ell_3 m_3}
  &= \p{-1}^{\ell_2+m_2} \sqrt{\f{2\ell_3+1}{2\ell_1+1}}
  \bk{\ell_3,-m_3; \ell_2 m_2| \ell_1,-m_1} \\
  \bk{\ell_1 m_1; \ell_2 m_2| \ell_3 m_3}
  &= \p{-1}^{\ell_1+\ell_2-\ell_3}
  \bk{\ell_1,-m_1; \ell_2,-m_2| \ell_3,-m_3},
\end{align}
we can find that
\begin{align}
  T_{LM}^\dag
  = \sqrt{\f{2L+1}{2I+1}}
  \sum_{\mu,\nu} \p{-1}^M \bk{I\nu;L,-M|I\mu} \op{\mu}{\nu}
  = \p{-1}^M T_{L,-M}.
\end{align}
Substituting this result into \eqref{eq:trans_struct} and expanding
Clebsch-Gordan coefficients in terms of Wigner 3-$j$ symbols as
\begin{align}
  \bk{\ell_1 m_1; \ell_2 m_2| \ell_3 m_3}
  = \p{-1}^{-\ell_1+\ell_2-m_3} \sqrt{\p{2\ell_3+1}}
  \begin{pmatrix}
    \ell_1 & \ell_2 & \ell_3 \\
    m_1 & m_2 & -m_3
  \end{pmatrix},
\end{align}
we find that
\begin{multline}
  f_{\ell_1 m_1;\ell_2 m_2}^{LM}
  = \delta_{M,m_1+m_2}
  \p{-1}^M \sqrt{\p{2L+1}\p{2\ell_1+1}\p{2\ell_2+1}} \\
  \times \sum_{\mu,\nu,\rho} \p{-1}^{-3I+L+\ell_1+\ell_2-\mu-\nu-\rho}
  \begin{pmatrix}
    I & L & I \\
    \mu & -M & -\nu
  \end{pmatrix}
  \begin{pmatrix}
    I & \ell_1 & I \\
    \nu & m_1 & -\rho
  \end{pmatrix}
  \begin{pmatrix}
    I & \ell_2 & I \\
    \rho & m_2 & -\mu
  \end{pmatrix},
\end{multline}
where we can evaluate the sum and introduce Wigner 6-$j$ symbols to
get
\begin{align}
  f_{\ell_1 m_1;\ell_2 m_2}^{LM}
  &= \delta_{M,m_1+m_2} \p{-1}^{2I+M}
  \sqrt{\p{2L+1}\p{2\ell_1+1}\p{2\ell_2+1}}
  \begin{pmatrix}
    L & \ell_1 & \ell_2 \\
    M & -m_1 & -m_2
  \end{pmatrix}
  \begin{Bmatrix}
    L & \ell_1 & \ell_2 \\
    I & I & I
  \end{Bmatrix} \\
  &= \delta_{M,m_1+m_2} \p{-1}^{2I+L}
  \sqrt{\p{2\ell_1+1}\p{2\ell_2+1}}
  \bk{\ell_1 m_1; \ell_2 m_2| LM}
  \begin{Bmatrix}
    \ell_1 & \ell_2 & L \\
    I & I & I
  \end{Bmatrix}.
  \label{eq:trans_prod}
\end{align}
The Wigner 6-$j$ symbol is symmetric under any permutation of its
columns, while Clebsch-Gordan coefficients satisfy
\begin{align}
  \bk{\ell_1 m_1; \ell_2 m_2| \ell_3 m_3}
  = \p{-1}^{\ell_1+\ell_2-\ell_3}
  \bk{\ell_2 m_2; \ell_1 m_1| \ell_3 m_3},
\end{align}
which implies that
\begin{align}
  f_{\ell_1 m_1;\ell_2 m_2}^{LM}
  = \p{-1}^{\ell_1+\ell_2-L} f_{\ell_2 m_2;\ell_1 m_1}^{LM}.
\end{align}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Drive operator product expansions}
\label{sec:drive_prod}

The drive operators are defined by
\begin{align}
  D_{LM} \equiv \f{\eta_M}{\sqrt{2}}
  \sp{T_{L\abs{M}} + \sign\p{M} T_{L\abs{M}}^\dag},
  &&
  \eta_M \equiv
  \begin{cases}
    \sqrt{2} & M = 0 \\
    \p{-1}^M \sign\p{M}^{3/2} & M \ne 0
  \end{cases},
\end{align}
or, equivalently,
\begin{align}
  D_{LM} \equiv \f{\tau_M}{\sqrt{2}}
  \sp{T_{LM} + \sign\p{M} T_{LM}^\dag},
  &&
  \tau_M \equiv
  \begin{cases}
    \sqrt{2} & M = 0 \\
    \p{-1}^M \sign\p{M}^{M+1/2} & M \ne 0
  \end{cases},
\end{align}
which implies that
\begin{align}
  T_{LM} = \f{\tilde\tau_M}{\sqrt{2}} \p{D_{LM} + i D_{L,-M}},
  &&
  \tilde\tau_M \equiv
  \begin{cases}
    \sqrt{2}/\p{1+i} & M = 0 \\
    \tau_M^* & M \ne 0
  \end{cases},
\end{align}
where $\tau_M^*$ is the complex conjugate of $\tau_M$.  In order to
evaluate the product of two drive operators, we can thus expand them
in terms of transition operators, evaluate the corresponding products
of transition operators (see Appendix \ref{sec:trans_prod}), and
finally expand the result in terms of drive operators.  The general
product of two drive operators takes the form
\begin{align}
  D_{\ell_1 m_1} D_{\ell_2 m_2}
  = \f{\tau_{m_1}\tau_{m_2}}{\sqrt{2}}
  \sum_L \p{f_{\ell_1 m_1;\ell_2 m_2}^{L,m_1+m_2}
    E_{\ell_1 m_1;\ell_2 m_2}^{(L,+)}
    + f_{\ell_1 m_1;\ell_2,-m_2}^{L,m_1-m_2}
    E_{\ell_1 m_1;\ell_2 m_2}^{(L,-)}},
\end{align}
where in terms of
\begin{align}
  \epsilon_M \equiv \p{-1}^M \sign\p{M},
\end{align}
we have
\begin{align}
  E_{\ell_1 m_1;\ell_2 m_2}^{(L,+)}
  &\equiv \sum_{s\in\set{\pm1}}
  \f12 \sp{\tilde\tau_{m_1+m_2}
    + i s \p{-1}^{L+\ell_1+\ell_2}
    \epsilon_{m_1} \epsilon_{m_2} \tilde\tau_{-\p{m_1+m_2}}}
  \sqrt{s} D_{L,s\p{m_1+m_2}},
  \\
  E_{\ell_1 m_1;\ell_2 m_2}^{(L,-)}
  &\equiv \sum_{s\in\set{\pm1}}
  \f12 \sp{\epsilon_{m_2} \tilde\tau_{m_1-m_2}
    + i s \p{-1}^{L+\ell_1+\ell_2}
    \epsilon_{m_1} \tilde\tau_{-\p{m_1-m_2}}}
  \sqrt{s} D_{L,s\p{m_1-m_2}}.
\end{align}
If $m_1\pm m_2=0$, these operators simplify to
\begin{align}
  E_{\ell_1 m;\ell_2,-m}^{(L,+)}
  &= \f1{\sqrt{2}} \sp{1 - \p{-1}^{L+\ell_1+\ell_2}
    \p{1-\delta_{m,0}}} D_{L,0}, \\
  E_{\ell_1 m;\ell_2 m}^{(L,-)}
  &= \f{\epsilon_m}{\sqrt{2}}
  \sp{1 + \p{-1}^{L+\ell_1+\ell_2}} D_{L,0},
\end{align}
and are otherwise
\begin{align}
  E_{\ell_1 m_1;\ell_2 m_2}^{(L,+)}
  &\stackrel{m_1+m_2\ne0}{=}
  \tau_{m_1+m_2}^* \sum_{s\in\set{\pm1}}
  \f12 \sp{1 + s \p{-1}^{L+\ell_1+\ell_2}
    \epsilon_{m_1} \epsilon_{m_2} \epsilon_{m_1+m_2}}
  \sqrt{s} D_{L,s\p{m_1+m_2}}, \\
  E_{\ell_1 m_1;\ell_2 m_2}^{(L,-)}
  &\stackrel{m_1-m_2\ne0}{=}
  \tau_{m_1-m_2}^* \sum_{s\in\set{\pm1}}
  \f12 \sp{\epsilon_{m_2} + s \p{-1}^{L+\ell_1+\ell_2}
    \epsilon_{m_1} \epsilon_{m_1-m_2}}
  \sqrt{s} D_{L,s\p{m_1-m_2}}.
\end{align}
If $\p{m_1,m_2}=\p{0,m}$, then we can use the fact that
$i\tau_m\tau_{-m}^*=\epsilon_m$ for $m\ne0$ to simplify
\begin{align}
  D_{\ell_1,0} D_{\ell_2 m}
  = \sum_L f_{\ell_1,0,\ell_2,m}^{Lm}
  \sum_{s\in\set{\pm1}} \f12 \sp{1 + s \p{-1}^{L+\ell_1+\ell_2}}
  \sqrt{s} D_{L,sm}.
\end{align}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Raman drive coefficients}
\label{sec:drive_raman_coeff}

Here we provide the coefficients $\Omega_{LM}^{(\phi,j)}$ of the
effective Raman driving Hamiltonian
\begin{align}
  H_{\t{Raman}}^{(\phi)}
  = \sum_j \sum_{\substack{L\in\set{0,1,2}\\\abs{M}\le L}}
  \Omega_{LM}^{(\phi,j)} D_{LM}^{(j)}
\end{align}
considered in Section \ref{sec:drive_raman}.  These coefficients are
computed using symmetry properties and structure constants of the
transition operator algebra (see Appendix \ref{sec:trans_prod}).  For
clarity, if any coefficient $\Omega_{LM}^{(\phi,j)}$ is independent of
$\phi$, we drop the explicit dependence on $\phi$, i.e.~taking
$\Omega_{LM}^{(\phi,j)}\to\Omega_{LM}$.
\begin{align}
  \Omega_{0,0} \equiv -\f1{\sqrt{n}}\,
  \sum_m\f{\Omega_m^2}{\Delta},
  &&
  \Omega_{1,0} \equiv \sqrt{\f{3}{n_1}}\,
  \f{\Omega_+^2 - \Omega_-^2}{\Delta},
  \label{eq:O_01_0}
\end{align}
\begin{align}
  \Omega_{2,0} \equiv \f1{\sqrt{5}}\, \f{\sqrt{n_2}}{n_1}\,
  \f{\Omega_+ + \Omega_- - 2\Omega_0}{\Delta},
  \label{eq:O_2_0}
\end{align}
\begin{align}
  \Omega_{L,1}^{(\phi,j)}
  &\equiv \sum_{s\in\set{\pm1}} \omega_{Ls} \cos\p{\phi_s j},
  &
  \Omega_{L,-1}^{(\phi,j)}
  &\equiv -\sum_{s\in\set{\pm1}} s \omega_{Ls} \sin\p{\phi_s j}, \\
  \Omega_{2,2}^{(\phi,j)}
  &\equiv \omega_{2,2} \cos\p{\sp{\phi_+-\phi_-}j},
  &
  \Omega_{2,-2}^{(\phi,j)}
  &\equiv -\omega_{2,2} \sin\p{\sp{\phi_+-\phi_-}j}, \\
  \omega_{1\pm}
  &\equiv \sqrt{\f{6}{n_1}}\, \f{\Omega_0 \Omega_\pm}{\Delta},
  &
  \omega_{2\pm}
  &\equiv \pm \sqrt{\f65}\, \f{\sqrt{n_2}}{n_1}\,
  \f{\Omega_0\Omega_\pm}{\Delta}, \\
  \omega_{2,2}
  &\equiv 2\sqrt{\f35}\, \f{\sqrt{n_2}}{n_1}\,
  \f{\Omega_+\Omega_-}{\Delta},
  &
  n_L
  &\equiv \prod_{\ell=-L}^L \p{n+\ell}.
\end{align}
Here $\Omega_m$, $\phi_m$, and $\Delta$ are respectively the (real)
driving amplitudes, phases, and detuning that appear in the ``bare''
Raman drive in \eqref{eq:raman_bare}, and by choice of gauge we set
$\phi_-=0$, or equivalently re-define $\phi_m-\phi_0\to\phi_m$.  If
the driving lasers are phase-matched with $\phi_m=m\phi$, such that a
nuclear spin transition $\mu\to\mu+m$ on site $j$ imprints the phase
$m\phi j$, then the drive $H_{\t{Raman}}^{(\phi)}$ takes the form
\begin{align}
  \left. H_{\t{Raman}}^{(\phi)} \right|_{\phi_m=m\phi}
  = \sum_{\substack{L\le2\\0\le M\le L}} \Omega_{LM} \D_{LM}^{(\phi)},
\end{align}
with inhomogeneously rotated collective drive operators
\begin{align}
  \D_{LM}^{(\phi)}
  \equiv \sum_j \sp{\cos\p{M\phi j} D_{LM}^{(j)}
    - \sin\p{M\phi j} D_{L,-M}^{(j)}},
\end{align}
and uniform coefficients given by \eqref{eq:O_01_0}, \eqref{eq:O_2_0},
and
\begin{align}
  \Omega_{1,-1} = \Omega_{2,-1} = \Omega_{2,-2} \equiv 0,
  &&
  \Omega_{L,1} \equiv \sum_{s\in\set{\pm1}} \omega_{Ls},
  &&
  \Omega_{2,2} \equiv \omega_{2,2}.
\end{align}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Weak SOC in the fully symmetric manifold}

Here we derive the effective Hamiltonian induced on the fully
symmetric manifold $\M_\FS$ by the single-particle SOC Hamiltonian
$H_{\t{lat}}^{(\phi)}$ in \eqref{eq:H_lat_SOC_B} in the weak SOC limit
$\abs{\phi}\ll1$.  For brevity, we work in units with the lattice
spacing $a=1$ throughout this section.  Perturbations $V$ to the
interaction Hamiltonian $H_{\t{int}}$ induce the following effective
Hamiltonians on the fully symmetric manifold
$\M_\FS$\cite{bravyi2011schrieffer, perlin2019effective}:
\begin{align}
  H_{\t{eff}}^{(1)} = \P_0 V \P_0,
  &&
  H_{\t{eff}}^{(2)} = -\P_0 V \E V \P_0,
  &&
  \E \equiv \sum_{E_\varepsilon>0} \f{\op{E_\varepsilon}}{E_\varepsilon},
  \label{eq:pert}
\end{align}
where $\P_0$ is a projector onto the fully symmetric manifold $\M_\FS$
and $\ket{E_\varepsilon}$ is an eigenstate of the interaction
Hamiltonian $H_{\t{int}}$ with energy $E_\varepsilon$ relative to the
interaction energy of fully symmetric states
$\ket{\psi_\FS}\in\M_\FS$.  The sum over all $E_\varepsilon>0$ is a
sum over a complete basis of states $\set{\ket{E_\varepsilon}}$ for
the orthogonal complement of $\M_\FS$.

\subsection{Effective SOC at first order}

The single-particle SOC Hamiltonian is
\begin{align}
  H_{\t{lat}}^{(\phi)}
  = -t \sum_{q,L} w_L\p{q} A_L^{(\phi)} D_{L,0}^{(q)},
  &&
  w_L\p{q} \equiv
  \begin{cases}
    \cos q & L~\t{even} \\
    \sin q & L~\t{odd}
  \end{cases},
  \label{eq:H_lat_SOC_A}
\end{align}
where $A_L^{(\phi)}$ is defined in \eqref{eq:A_L_phi}, and scales as
$A_L^{(\phi)} \sim \phi^L + O\p{\phi^{L+2}}$ when $\abs{\phi}\ll1$.
When the SOC angle $\phi$ is sufficiently small in magnitude, the
effective Hamiltonian induced on the fully symmetric manifold $\M_\FS$
at first order in the perturbation $H_{\t{lat}}^{(\phi)}$ is
\begin{align}
  H_{\t{eff}}^{(1)}
  = \P_0 H_{\t{lat}}^{(\phi)} \P_0
  = -t \sum_{q,L} w_L\p{q} A_L^{(\phi)} \P_0 D_{L,0}^{(q)} \P_0,
\end{align}
where we can freely change the index of each single-mode drive
operator $D_{L,0}^{(q)}$ within the fully symmetric manifold to
re-write
\begin{align}
  \P_0 D_{L,0}^{(q)} \P_0 = \f1N \sum_p \P_0 D_{L,0}^{(p)} \P_0
  = \f1N \P_0 \D_{L,0} \P_0,
\end{align}
which implies that
\begin{align}
  H_{\t{eff}}^{(1)}
  = -t \sum_L \EE_q\sp{w_L\p{q}} A_L^{(\phi)} \D_{L,0},
  &&
  \EE_q\sp{X\p{q}} \equiv \f1N \sum_q X\p{q}.
  \label{eq:H_eff_1}
\end{align}
Here $\EE_q\sp{X\p{q}}$ is an average of $X\p{q}$ over all occupied
quasi-momenta $q$.  Note that we have dropped the explicit projection
operators $\P_0$ in \eqref{eq:H_eff_1} with the understanding that we
are implicitly restricting all effective dynamics to the fully
symmetric manifold $\M_\FS$.

\subsection{Effective SOC at second order}

In order to expand the second-order effective Hamiltonian
$H_{\t{eff}}^{(2)}$ induced on the fully symmetric manifold $\M_\FS$
by the SOC Hamiltonian $H_{\t{lat}}^{(\phi)}$ when $\abs{\phi}\ll1$,
we expand $w_L\p{q}$ in \eqref{eq:H_lat_SOC_A} as
\begin{align}
  w_L\p{q} = \f12 \sum_{s\in\set{\pm1}} \chi_{Ls} e^{isq},
  &&
  \chi_{Ls} \equiv \p{-i}^{L\,\t{mod}\,2} s^L,
\end{align}
which implies that
\begin{align}
  H_{\t{lat}}^{(\phi)}
  = -\f{t}{2} \sum_{\substack{L\\s\in\set{\pm1}}}
  \chi_{Ls} A_L^{(\phi)} \Z_{Ls},
  &&
  \Z_{Ks} \equiv \sum_q e^{isq} D_{L,0}^{(q)},
\end{align}
and in turn
\begin{align}
  H_{\t{eff}}^{(2)}
  = -\f{t^2}{4} \sum_{\substack{J,K\\r,s\in\set{\pm1}}}
  \chi_{Jr} \chi_{Ks} A_J^{(\phi)} A_K^{(\phi)}
  \P_0 \Z_{Jr} \E \Z_{Ks} \P_0,
  \label{eq:H_eff_2_start}
\end{align}
where $\E$ is a sum over excited-state projectors with energetic
suppression factors, defined in \eqref{eq:pert}.

\subsubsection{Eliminating excited-state projectors in $\E$}

In order to simplify \eqref{eq:H_eff_2_start}, we first note that the
image of any fully symmetric state $\ket{\psi_{\FS}}\in\M_\FS$ under
the action of the ``spin-wave'' operator $\Z_{Ks}$ is an eigenvector
of the interaction Hamiltonian $H_{\t{int}}=-u\sum_{p<q}\Pi_{pq}$.  To
substantiate this claim, we expand
\begin{align}
  \sum_{p<q} \Pi_{pq} \Z_{Ks} \ket{\psi_\FS}
  = \sum_{\substack{p<q\\k\notin\set{p,q}}} e^{isk}
  \Pi_{pq} D_{K,0}^{(k)} \ket{\psi_\FS}
  + \sum_{\substack{p<q\\k\in\set{p,q}}} e^{isk}
  \Pi_{pq} D_{K,0}^{(k)} \ket{\psi_\FS}.
\end{align}
If $k\notin\set{p,q}$, then the permutation operator $\Pi_{pq}$
commutes with $D_{K,0}^{(k)}$, and in turn acts trivially on the
permutationally-symmetric state $\ket{\psi_\FS}$, giving us
\begin{align}
  \sum_{\substack{p<q\\k\notin\set{p,q}}} e^{isk}
  \Pi_{pq} D_{K,0}^{(k)} \ket{\psi_\FS}
  = \sum_k \sum_{\substack{p<q\\p,q\ne k}} e^{isk}
  D_{K,0}^{(k)} \ket{\psi_\FS}
  = { N-1 \choose 2 } \Z_{Ks} \ket{\psi_\FS}.
\end{align}
If $k\in\set{p,q}$, then we can resolve the identity operator
$\1=\Pi_{pq}^\dag \Pi_{pq}$ to expand
\begin{align}
  \sum_{\substack{p<q\\k\in\set{p,q}}} e^{isk}
  \Pi_{pq} D_{K,0}^{(k)} \ket{\psi_\FS}
  = \f12 \sum_{\substack{p\ne q\\k\in\set{p,q}}} e^{isk}
  \Pi_{pq} D_{K,0}^{(k)} \Pi_{pq}^\dag \Pi_{pq} \ket{\psi_\FS}
  = \sum_{q\ne k} e^{isk} D_{K,0}^{(q)} \ket{\psi_\FS},
\end{align}
where for $s\in\set{\pm1}$ we can simplify the sum
$\sum_{k\ne q} e^{isk} = -e^{isq}$,\footnote{We assume for now that
  $\sum_k e^{\pm ik}=0$, which is only valid when all quasi-momenta
  are occupied, such that the sum $\sum_k$ runs over all angles
  $k\in\ZZ_N\times2\pi/N$.  We will consider the slightly more
  complicated case with $\sum_k e^{\pm ik}\ne0$ later in this
  section.}  so
\begin{align}
  \sum_{\substack{p<q\\k\in\set{p,q}}} e^{isk}
  \Pi_{pq} D_{K,0}^{(k)} \ket{\psi_\FS}
  = - \sum_q e^{isq} D_{K,0}^{(q)} \ket{\psi_\FS}
  = - \Z_{Ks} \ket{\psi_\FS}.
\end{align}
In total, we have that
\begin{align}
  \sum_{p<q} \Pi_{pq} \Z_{Ks} \ket{\psi_\FS}
  = \sp{{ N-1 \choose 2 } - 1} \Z_{Ks} \ket{\psi_\FS},
\end{align}
which implies that $\Z_{Ks}\ket{\psi_\FS}$ is proportional to an
eigenstate of $H_{\t{int}}$ with interaction energy
$-u\sp{{ N-1 \choose 2 } - 1}$, which is $Nu$ above the interaction
energy of $-u{N\choose2}$ of states in the fully symmetric manifold,
so
\begin{align}
  \E \Z_{Ks} \P_0
  =  \f1{Nu} \Z_{Ks} \P_0
  = \f1{fU} \Z_{Ks} \P_0,
  \label{eq:proj_elim}
\end{align}
where $\E$ is defined in \eqref{eq:pert} and $f\equiv N/L$ is the
filling fraction of quasi-momentum modes.  We have thus far simplified
the effective Hamiltonian in \eqref{eq:H_eff_2_start} to
\begin{align}
  H_{\t{eff}}^{(2)}
  = -\f{t^2}{4fU} \sum_{\substack{J,K\\r,s\in\set{\pm1}}}
  \chi_{Jr} \chi_{Ks} A_J^{(\phi)} A_K^{(\phi)}
  \P_0 \Z_{Jr} \Z_{Ks} \P_0,
  \label{eq:H_eff_2_mid}
\end{align}
leaving us the task of evaluating the product
$\P_0 \Z_{Jr} \Z_{Ks} \P_0$, as well as the sum over $r,s$.

\subsubsection{The case of $r=s$}

To evaluate the product $\P_0 \Z_{Jr} \Z_{Ks} \P_0$ of spin-wave
operators projected onto the fully symmetric manifold $\M_\FS$, we
first consider the case of $r=s$, and use the permutational symmetry
enforced by the projectors $\P_0$ to expand
\begin{align}
  \P_0 \Z_{J\pm} \Z_{K\pm} \P_0
  &= \sum_{p,q} e^{\pm i\p{p+q}} \P_0 D_{J,0}^{(p)} D_{K,0}^{(q)} \P_0 \\
  &= \sum_p e^{\pm i2p} \P_0 D_{J,0}^{(0)} D_{K,0}^{(0)} \P_0
  + \sum_{p\ne q} e^{\pm i\p{p+q}} \P_0 D_{J,0}^{(0)} D_{K,0}^{(1)} \P_0,
\end{align}
where $\sum_p e^{\pm i2p}=0$, and
\begin{align}
  \sum_{p\ne q} e^{\pm i\p{p+q}}
  = \sum_p e^{\pm ip} \sum_{q\ne p} e^{\pm iq}
  = \sum_p e^{\pm ip} \p{-e^{\pm ip}}
  = -\sum_p e^{\pm i2p}
  = 0,
\end{align}
which implies that
\begin{align}
  \P_0 \Z_{J\pm} \Z_{K\pm} \P_0 = 0.
  \label{eq:ZZ_eq}
\end{align}

\subsubsection{The case of $r\ne s$}
\label{sec:H_eff_2_end}

For the case of $r\ne s$, we similarly expand
\begin{align}
  \P_0 \Z_{J\pm} \Z_{K\mp} \P_0
  &= \sum_{p,q} e^{\pm i\p{p-q}} \P_0 D_{J,0}^{(p)} D_{K,0}^{(q)} \P_0 \\
  &= N \P_0 D_{J,0}^{(0)} D_{K,0}^{(0)} \P_0
  + \sum_{p\ne q} e^{\pm i\p{p-q}} \P_0 D_{J,0}^{(0)} D_{K,0}^{(1)} \P_0,
\end{align}
where
\begin{align}
  \sum_{p\ne q} e^{\pm i\p{p-q}}
  = \sum_p e^{\pm ip} \sum_{q\ne p} e^{\mp iq}
  = \sum_p e^{\pm ip} \p{-e^{\mp ip}}
  = - \sum_p e^{\pm i\p{p-p}}
  = - N,
\end{align}
so
\begin{align}
  \P_0 \Z_{J\pm} \Z_{K\mp} \P_0
  = N \P_0 D_{J,0}^{(0)} D_{K,0}^{(0)} \P_0
  - N \P_0 D_{J,0}^{(0)} D_{K,0}^{(1)} \P_0.
  \label{eq:ZZ_neq_mid}
\end{align}
In order to express this product in terms of collective operators, we
expand
\begin{align}
  \P_0 \D_{J,0} \D_{K,0} \P_0
  = N \P_0 D_{J,0}^{(0)} D_{K,0}^{(0)} \P_0
  + N \p{N-1} \P_0 D_{J,0}^{(0)} D_{K,0}^{(1)} \P_0,
\end{align}
then solve for the last term and substitute the result into
\eqref{eq:ZZ_neq_mid}, giving us
\begin{align}
  \P_0 \Z_{J\pm} \Z_{K\mp} \P_0
  = \f{N^2}{N-1} \P_0 D_{J,0}^{(0)} D_{K,0}^{(0)} \P_0
  - \f1{N-1} \P_0 \D_{J,0} \D_{K,0} \P_0.
\end{align}
Using the structure constants $f_{\ell_1m_1;\ell_2m_2}^{LM}$ of the
transition operator algebra (see Appendix \ref{sec:trans_prod}), we
then expand
\begin{align}
  N \P_0 D_{J,0}^{(0)} D_{K,0}^{(0)} \P_0
  = N \sum_L f_{J,0;K,0}^{L,0} \P_0 D_{L,0}^{(0)} \P_0
  = \sum_L f_{J,0;K,0}^{L,0} \P_0 \D_{L,0} \P_0,
\end{align}
which implies that
\begin{align}
  \P_0 \Z_{J\pm} \Z_{K\mp} \P_0
  = \f{N}{N-1} \sum_L f_{J,0;K,0}^{L,0} \P_0 \D_{L,0} \P_0
  - \f1{N-1} \P_0 \D_{J,0} \D_{K,0} \P_0.
  \label{eq:ZZ_neq}
\end{align}

\subsubsection{Putting it all together}

Having evaluated the product $\P_0 \Z_{Jr} \Z_{Ks} \P_0$ of spin-wave
operators in \eqref{eq:ZZ_eq} and \eqref{eq:ZZ_neq}, the last
ingredient needed to simplify the effective Hamiltonian
$H_{\t{eff}}^{(2)}$ in \eqref{eq:H_eff_2_mid} is an evaluation of the
sum over $r,s\in\set{\pm1}$.  The product of spin-wave operators
enforces $r\ne s$, but is otherwise independent of $r,s$, so we need
only evaluate
\begin{align}
  \sum_{s\in\set{\pm1}} \chi_{J,-s} \chi_{Ks}
  =
  \begin{cases}
    2 & J = K \, \t{mod}\, 2 \\
    0 & \t{otherwise}
  \end{cases}.
\end{align}
Putting everything together, we thus find
\begin{align}
  H_{\t{eff}}^{(2)}
  = \f{t^2}{2\p{N-1}fU}
  \sum_{J=K\,\t{mod}\,2} A_J^{(\phi)} A_K^{(\phi)}
  \sp{\D_{J,0} \D_{K,0} - N\sum_L f_{J,0;K,0}^{L,0} \D_{L,0}},
  \label{eq:H_eff_2_unit}
\end{align}
where, as for $H_{\t{eff}}^{(1)}$ in \eqref{eq:H_eff_1}, we drop the
explicit projectors $\P_0$ in \eqref{eq:H_eff_2_unit} with the
understanding that we are implicitly restricting all effective
dynamics to the fully symmetric manifold $\M_\FS$.

Thus far, we have assumed that the sum over quasi-momenta
$\sum_ke^{\pm ik}=0$, which is only valid when all quasi-momenta are
occupied, such that the sum $\sum_k$ runs over all angles
$k\in\ZZ_N\times2\pi/N$.  If not all quasi-momenta are occupied, then
we can repeat all derivations above with mean-subtracted spin-wave
operators
\begin{align}
  \Z_{Ks} \equiv \sum_q e^{isq} D_{K,0}^{(q)}
  \to
  \tilde\Z_{Ks}
  \equiv \sum_q \p{e^{isq} - \EE_k\sp{e^{isk}}} D_{K,0}^{(q)},
\end{align}
where $\EE_k\sp{e^{isk}}$ is an average of $e^{isk}$ over all occupied
quasi-momenta $k$.  Note that the difference
$\tilde\Z_{Ks}-\Z_{Ks}=-\EE_k\sp{e^{isk}}\D_{K,0}$ is permutationally
symmetric, which implies that it only couples fully symmetric states
$\ket{\psi_\FS}\in\M_\FS$ to fully symmetric states.  In particular,
due to the fact that $\E$ is a linear combination of projectors onto
interaction-energy eigenstates strictly outside of $\M_\FS$,
\begin{align}
  \P_0 \tilde\Z_{Jr} \E \tilde\Z_{Ks} \P_0
  = \P_0 \Z_{Jr} \E \Z_{Ks} \P_0,
\end{align}
which implies that we can freely take
$\Z_{Jr},\Z_{Ks}\to\tilde\Z_{Jr},\tilde\Z_{Ks}$ with no consequence
for the effective Hamiltonian $H_{\t{eff}}^{(2)}$ in
\eqref{eq:H_eff_2_start}.

Repeating the previous calculations in this section with the
mean-subtracted spin-wave operators, we still find that, just as in
\eqref{eq:proj_elim},
\begin{align}
  \E \tilde\Z_{Ks} \P_0 = \f1{fU} \tilde\Z_{Ks} \P_0,
\end{align}
whereas, in contrast to \eqref{eq:ZZ_eq} and \eqref{eq:ZZ_neq},
\begin{align}
  \P_0 \Z_{J\pm} \Z_{K\pm} \P_0
  = -\f{\var_q\sp{e^{\pm iq}}}{N-1}
  \sp{\P_0 \D_{J,0} \D_{K,0} \P_0
    - N \sum_L f_{J,0;K,0}^{L,0} \P_0 \D_{L,0} \P_0},
\end{align}
where
$\var_q\sp{X_q} \equiv \EE_q\sp{\p{X_q-\EE_p\sp{X_p}}^2} =
\EE_q\sp{X_q^2}-\EE_q\sp{X_q}^2$ is the variance of $X_q$ over all
occupied quasi-momenta $q$, and
\begin{align}
  \P_0 \Z_{J\pm} \Z_{K\pm} \P_0
  = -\f{1-\EE_p\sp{e^{ip}}\EE_q\sp{e^{-iq}}}{N-1}
  \sp{\P_0 \D_{J,0} \D_{K,0} \P_0
    - N \sum_L f_{J,0;K,0}^{L,0} \P_0 \D_{L,0} \P_0}.
\end{align}

\vspace{3cm}

\red{To do: Work out SU($2$) explicitly as a special case in these
  notes.  Also, the current second-order derivation assumes unit
  filling, and needs to be adjusted for the possibility of $f<1$.  I
  think that in the case of $f<1$ the factor of $1/2$ in
  \eqref{eq:H_eff_2} turns into $\bk{w_J\p{q}w_K\p{q}}_q$, e.g.~when
  $J=K=1$, this factor becomes the variance of $\sin\p{q}$ over all
  occupied modes, as in the known case of SU($2$).  I should also work
  out explicitly the contributions through second order in $\phi$,
  which will simplify things somewhat as we will only need $J,K\le2$
  in $H_{\t{lat}}^{(1)}$, and $J=K=1$ in $H_{\t{lat}}^{(2)}$.  I can
  also do a third-order calculation of $H_{\t{eff}}^{(3)}$ to extract
  dependence on $\phi^3$, but that's probably unnecessary / not
  particularly interesting.}

\vspace{3cm}

\bibliography{multilevel_spin_notes.bib}

\end{document}