### 2020-09-17 sun_phases/notes.tex
removed content from notes.tex that was moved into sun_dynamics_notes.tex

\documentclass[nofootinbib,notitlepage,11pt]{revtex4-2}

%%% linking references
\usepackage{hyperref}
\hypersetup{
  breaklinks=true,
  colorlinks=true,
  linkcolor=blue,
  filecolor=magenta,
  urlcolor=cyan,
}

%%% header / footer
\usepackage{fancyhdr} % easier header and footer management
\pagestyle{fancy} % page formatting style
\fancyhf{} % clear all header and footer text
\renewcommand{\headrulewidth}{0pt} % remove horizontal line in header
\usepackage{lastpage} % for referencing last page
\cfoot{\thepage~of \pageref{LastPage}} % "x of y" page labeling

%%% symbols, notations, etc.
\usepackage{physics,braket,bm,amssymb} % physics and math
\renewcommand{\t}{\text} % text in math mode
\newcommand{\f}[2]{\dfrac{#1}{#2}} % shorthand for fractions
\newcommand{\p}[1]{\left(#1\right)} % parenthesis
\renewcommand{\sp}[1]{\left[#1\right]} % square parenthesis
\renewcommand{\set}[1]{\left\{#1\right\}} % curly parenthesis
\newcommand{\bk}{\Braket} % shorthand for braket notation
\renewcommand{\v}{\bm} % bold vectors
\newcommand{\uv}[1]{\bm{\hat{#1}}} % unit vectors
\newcommand{\av}{\vec} % arrow vectors
\renewcommand{\c}{\cdot} % inner product
\renewcommand{\d}{\partial} % partial derivative
\renewcommand{\dd}{\text{d}} % for infinitesimals
\renewcommand{\i}{\mathrm{i}\mkern1mu} % imaginary unit

\usepackage{dsfont} % for identity operator
\newcommand{\1}{\mathds{1}}

\usepackage{mathtools} % for \coloneqq

\newcommand{\up}{\uparrow}
\newcommand{\dn}{\downarrow}

\newcommand{\x}{\text{x}}
\newcommand{\y}{\text{y}}
\newcommand{\z}{\text{z}}
\newcommand{\X}{\text{X}}
\newcommand{\DCS}{\text{DCS}}

\newcommand{\B}{\mathcal{B}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\E}{\mathcal{E}}
\renewcommand{\H}{\mathcal{H}}
\newcommand{\I}{\mathcal{I}}
\newcommand{\J}{\mathcal{J}}
\newcommand{\M}{\mathcal{M}}
\newcommand{\N}{\mathcal{N}}
\renewcommand{\O}{\mathcal{O}}
\renewcommand{\P}{\mathcal{P}}
\newcommand{\Q}{\mathcal{Q}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\T}{\mathcal{T}}
\renewcommand{\S}{\mathcal{S}}
\newcommand{\V}{\mathcal{V}}
\newcommand{\Z}{\mathcal{Z}}

\newcommand{\EE}{\mathbb{E}}
\newcommand{\RR}{\mathbb{R}}
\renewcommand{\SS}{\mathbb{S}}
\newcommand{\ZZ}{\mathbb{Z}}

\newcommand{\PS}{\text{PS}}
\newcommand{\col}{\underline}

\let\var\relax
\DeclareMathOperator{\var}{var}

\usepackage[inline]{enumitem} % in-line lists and \setlist{} (below)
\setlist[enumerate,1]{label={(\roman*)}} % default in-line numbering
\setlist{nolistsep} % more compact spacing between environments

%%% text markup
\usepackage{color} % text color
\newcommand{\red}[1]{{\color{red} #1}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\thispagestyle{fancy}

% todo: find roots of lax vector

\title{Collective SU($n$) spin model dynamics}%
\author{Michael A. Perlin}%
\date{\today}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Initial states}

We can essentially prepare $N$-fold tensor products $\ket\Psi\equiv\ket\psi^{\otimes N}$ of three types of initial product states $\ket\psi$.
First and foremost, we can prepare states $\ket{\mu}$ of definite spin projection $\mu=\set{I,I-1,\cdots,-I}$ onto a quantization axis.
In addition, we can prepare arbitrary polarized states defined by the polar and azimuthal angles $\theta,\phi$:
\begin{align}
  \ket{\theta,\phi}
  \equiv e^{i\phi s_\z} e^{i\theta s_\y} \ket\up
  = \sum_\mu \zeta_\mu\p{\theta} e^{-\i\mu\phi} \ket\mu,
  \label{eq:polarized_state}
\end{align}
where $\ket\up\equiv\ket{I}$ is a state polarized along the $z$ axis, and\footnote{The expansion of $\ket{\theta,\phi}$ in \eqref{eq:polarized_state} and \eqref{eq:polarized_proj} is acquired by representing an $n$-level spin by $n-1$ qubits restricted to the (Dicke) manifold of permutationally symmetric states.}
\begin{align}
  \zeta_\mu\p{\theta} \equiv { 2I \choose I+\mu }^{1/2}
  \cos\p{\f{\theta}{2}}^{I+\mu} \sin\p{\f{\theta}{2}}^{I-\mu}.
  \label{eq:polarized_proj}
\end{align}
Finally, for even $n$ we can also prepare the following superposition of polarized states:
\begin{align}
  \ket{\DCS}
  \equiv \f1{\sqrt{2}}
  \p{e^{\i\gamma} \ket{\alpha,\beta}
    + e^{-\i\gamma} \ket{\alpha,-\beta}}
  = \sum_\mu \zeta_\mu\p{\alpha}
  \times \sqrt{2}\, \cos\p{\mu\beta-\gamma} \ket\mu
  \label{eq:double_state}
\end{align}
with
\begin{align}
  \alpha = \f{\pi}{2} + \arcsin\p{\f13},
  &&
  \beta = \f{\pi}{3},
  &&
  \gamma = -\f{2\pi}{3}\, I.
\end{align}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Non-interacting limit of a realistic 1D model}

We consider the single-particle Hamiltonian
\begin{align}
  H = -J \sum_{\mu,q} \cos\p{q + \mu\varphi} s_{\mu\mu q},
\end{align}
where $q$ is a quasi-momentum (in units with lattice spacing $a=1$); $\mu\in\set{I,I-1,\cdots,-I}$ is a nuclear spin projection onto a quantization axis, with $I\equiv\p{n-1}/2$ the total nuclear spin; $\varphi$ is a spin-orbit-coupling angle; and $J$ is essentially the single-particle bandwidth (proportional to the nearest-neighbor tunneling rate, $J/2$).
As dynamical order parameters, we consider the operators
\begin{align}
  s_{\mu\nu} \equiv \f1N \sum_q s_{\mu\nu q},
  &&
  s^2 \equiv \v s^\dag\cdot \v s
  = \f1{N^2} \sum_{\mu,\nu,p,q} s_{\mu\nu p} s_{\nu\mu q}.
\end{align}
In the Heisenberg picture, these operators at time $t$ are
\begin{align}
  s_{\mu\nu}\p{t} = \f1N \sum_q s_{\mu\nu q}
  \times e^{\i\eta_{\mu\nu q} J t},
  &&
  s^2\p{t} = \f1{N^2} \sum_{\mu,\nu,p,q} s_{\mu\nu p} s_{\nu\mu q}
  \times e^{\i \p{\eta_{\mu\nu p}-\eta_{\mu\nu q}} J t},
\end{align}
where
\begin{align}
  \eta_{\mu\nu q}
  \equiv -\sp{\cos\p{q+\mu\varphi} - \cos\p{q+\nu\varphi}}
  = 2 \sin\p{\varphi_{\mu\nu}^-} \sin\p{q+\varphi_{\mu\nu}^+},
  &&
  \varphi_{\mu\nu}^\pm \equiv \f{\mu\pm\nu}{2}\,\varphi.
\end{align}
In particular, we are interested in the expectation values
\begin{align}
  \bk{\Psi|s_{\mu\nu}\p{t}|\Psi}
  = \f1N \bk{\psi\op{\mu}{\nu}\psi} \sum_q e^{\i\eta_{\mu\nu q} J t}
  = \bk{\psi\op{\mu}{\nu}\psi} \chi_{\mu\nu}\p{\varphi},
\end{align}
and
\begin{align}
  \bk{\Psi|s^2\p{t}|\Psi}
  &= \f1{N^2} \sum_{\mu,\nu} \abs{\bk{\psi\op{\mu}{\nu}\psi}}^2
  \sum_{p\ne q} e^{\i\p{\eta_{\mu\nu p}-\eta_{\mu\nu q}} J t} + \f{n}{N} \\
  &= \sum_{\mu,\nu} \abs{\bk{\psi\op{\mu}{\nu}\psi}}^2
  \abs{\chi_{\mu\nu}\p{\varphi}}^2 + \f{n-1}{N},
\end{align}
where
\begin{align}
  \chi_{\mu\nu}\p{\varphi} \equiv \f1N \sum_q e^{\i\eta_{\mu\nu q} J t}
  = \f1N \sum_q e^{\i 2\sin\p{\varphi_{\mu\nu}^-} \sin\p{q+\varphi_{\mu\nu}^+} J t}.
\end{align}
As $N\to\infty$, this sum is well approximated by an integral:
\begin{align}
  \chi_{\mu\nu}\p{\varphi} \approx \f1{2\pi} \int_{-\pi}^\pi \dd\theta\,
  e^{\i 2\sin\p{\varphi_{\mu\nu}^-} \sin\p{\theta+\varphi_{\mu\nu}^+} J t}
  = \J_0\sp{2 \sin\p{\varphi_{\mu\nu}^-} J t},
\end{align}
where $\J_0$ is the zero-order Bessel function of the first kind.
To leading order in $N\gg1$, we thus find that
\begin{align}
  \bk{\Psi|s_{\mu\nu}\p{t}|\Psi}
  &\approx \bk{\psi\op{\mu}{\nu}\psi}
  \J_0\sp{2\sin\p{\varphi_{\mu\nu}^-}Jt},
  \\
  \bk{\Psi|s^2\p{t}|\Psi}
  &\approx \sum_{\mu,\nu} \abs{\bk{\psi\op{\mu}{\nu}\psi}}^2
  \J_0\sp{2\sin\p{\varphi_{\mu\nu}^-}Jt}^2.
  \label{eq:ops_time}
\end{align}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Limiting cases}

The constant (time independent) terms in \eqref{eq:ops_time} have $\mu=\nu$ (in which case all $\varphi_{\mu\nu}^-=0$), so up to $\O(1/N)$ corrections
\begin{align}
  \bk{\Psi|s_{\mu\nu}^{\t{const}}|\Psi}
  \approx \delta_{\mu\nu} \abs{\bk{\mu|\psi}}^2,
  &&
  \bk{\Psi|s^2_{\t{const}}|\Psi}
  \approx \sum_\mu \abs{\bk{\mu|\psi}}^4 \equiv r_\psi,
  \label{eq:ops_const}
\end{align}
where the quantity $r_\psi$ is sometimes called the inverse participation ratio of $\ket\psi$ with respect to the basis $\set{\ket\mu}$ of states with definite projection onto the $z$ axis.
Figure \ref{fig:limiting_vals} shows several values of $r_\psi$ for different $n$.

At ``short'' times $t$ for which $\sin\p{I\varphi}Jt\ll1$, we can Taylor expand the Bessel functions in \eqref{eq:ops_time} about $t=0$ to get
\begin{align}
  \bk{\Psi|s_{\mu\nu}\p{t}|\Psi}
  &\approx \bk{\psi\op{\mu}{\nu}\psi}
  \times \sp{1 - \sin\p{\varphi_{\mu\nu}^-}^2 \p{J t}^2}, \\
  \bk{\Psi|s^2\p{t}|\Psi}
  &\approx 1 - 2 \p{Jt}^2 \sum_{\mu,\nu}
  \sin\p{\varphi_{\mu\nu}^-}^2 \abs{\bk{\psi\op{\mu}{\nu}\psi}}^2.
\end{align}
In the limit of a small angle $\varphi$ for which $I\varphi\ll1$, we can further simplify
\begin{align}
  \bk{\Psi|s_{\mu\nu}\p{t}|\Psi}
  &\approx \bk{\psi\op{\mu}{\nu}\psi}
  \times \sp{1 - \p{\f{\mu-\nu}{2I}}^2
    \p{I\varphi J t}^2}, \\
  \bk{\Psi|s^2\p{t}|\Psi}
  &\approx 1 - \var_\psi\p{Z} \p{I\varphi J t}^2,
  \label{eq:small_time_angle}
\end{align}
where
\begin{align}
  \var_\psi\p{Z} \equiv \bk{\psi|Z^2|\psi} - \bk{\psi|Z|\psi}^2,
  &&
  Z \equiv \sum_\mu \f{\mu}{I} \op{\mu}.
\end{align}
Note that the result in \eqref{eq:small_time_angle} is valid for all $t\ll\p{I\varphi J}^{-1}$, which may be a long time given the small-angle assumption $I\varphi\ll1$.
Figure \ref{fig:limiting_vals} shows several values of $\var_\psi\p{Z}$ for different $n$.

\begin{figure}
  \centering
  \includegraphics{figures/oscillations/limiting_vals.pdf}
  \caption{Several values of $r_\psi$ and $\var_\psi\p{Z}$ for
    different $n$.  Computed with respect to the states
    $\ket\X\equiv\ket{\pi/2,0}$ and $\ket{\DCS}$, defined in
    \eqref{eq:polarized_state} and \eqref{eq:double_state}.}
  \label{fig:limiting_vals}
\end{figure}

\bibliography{notes.bib}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: