### 2020-11-23 sun_phases/spin_qudit_tomo.tex
added discussion of the connection between polarization operators and spherical harmonics (via phase-space representations) added references for ultracold molecules

\documentclass[notitlepage,twocolumn]{revtex4-2}

%%% linking references
\usepackage{hyperref}
\hypersetup{
  breaklinks=true,
  colorlinks=true,
  linkcolor=blue,
  filecolor=magenta,
  urlcolor=cyan,
}

%%% symbols, notations, etc.
\usepackage{physics,braket,bm} % physics and math
\renewcommand{\t}{\text} % text in math mode
\newcommand{\f}[2]{\dfrac{#1}{#2}} % shorthand for fractions
\newcommand{\p}[1]{\left(#1\right)} % parenthesis
\renewcommand{\sp}[1]{\left[#1\right]} % square parenthesis
\newcommand{\bk}{\Braket} % shorthand for braket notation
\renewcommand{\v}{\bm} % bold vectors
\newcommand{\uv}[1]{\bm{\hat{#1}}} % unit vectors
\renewcommand{\i}{\mathrm{i}\mkern1mu} % imaginary unit

% curly brackets for a set
\renewcommand{\set}[1]{\{#1\}} % curly parenthesis
\renewcommand{\Set}[1]{\left\{#1\right\}} % curly parenthesis

% double angle brackets
\newcommand{\bbk}[1]{\langle\!\langle #1 \rangle\!\rangle}
\newcommand{\Bbk}[1]
{\left\langle\!\!\left\langle #1 \right\rangle\!\!\right\rangle}

\newcommand{\x}{\text{x}}
\newcommand{\y}{\text{y}}
\newcommand{\z}{\text{z}}

\newcommand{\C}{\mathcal{C}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\E}{\mathcal{E}}
\newcommand{\I}{\mathcal{I}}
\newcommand{\N}{\mathcal{N}}
\renewcommand{\O}{\mathcal{O}}
\newcommand{\Q}{\mathcal{Q}}
\renewcommand{\S}{\mathcal{S}}

\def\obra#1{\mathinner{({#1}|}}
\def\oket#1{\mathinner{|{#1})}}
\def\obk#1{\mathinner{({#1})}}
\def\oop#1#2{\oket{#1}\!\obra{#2}}

\DeclareMathOperator{\cov}{cov}
\DeclareMathOperator{\diag}{diag}

%%% figures
\usepackage{graphicx} % for figures
\graphicspath{{./figures/qudit_tomo/}} % set path for all figures

\usepackage[inline]{enumitem} % in-line lists and \setlist{} (below)
\setlist[enumerate,1]{label={(\roman*)}} % default in-line numbering

%%% text markup
\usepackage{color} % text color
\newcommand{\red}[1]{{\color{red} #1}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

\newcommand{\JILA}{JILA, National Institute of Standards and Technology and
  University of Colorado, 440 UCB, Boulder, Colorado 80309, USA}
\newcommand{\CTQM}{Center for Theory of Quantum Matter, University of Colorado, Boulder, CO, 80309, USA}

\newcommand{\thetitle}{Spin qudit tomography}

\title{\thetitle}
\author{Michael A.~Perlin}
\email{mika.perlin@gmail.com}
\author{Diego Barberena}
\author{Ana Maria Rey}
\affiliation{\JILA}
\affiliation{\CTQM}
\date{\today}

\keywords{qudits; spin qudits; quantum state tomography}

\begin{abstract}
  We consider the task of performing quantum state tomography on a $d$-state spin qudit, using only measurements of spin projection onto different quantization axes.
  By an exact mapping onto the classical problem of signal recovery on the sphere, we prove that full reconstruction of arbitrary qudit states requires is possible with a minimal number of measurement axes, $r_d^{\t{min}}$, that is bounded by $2d-1\le r_d^{\t{min}}\le d^2$.
  We conjecture that $r_d^{\t{min}}=2d-1$, which we verify numerically for all $d\le200$.
  We then provide algorithms with $O\p{rd^3}$ serial runtime, parallelizable down to $O\p{rd^2}$, for (i) computing (a priori) upper bounds on the expected error with which spin projection measurements along $r$ given axes can reconstruct an unknown qudit state, and (ii) estimating (a posteriori) the statistical error in a reconstructed state.
  Our algorithms motivate a simple randomized tomography protocol, for which we numerically find that using more measurement axes can yield substantial benefits that plateau after $r\approx3d$.
\end{abstract}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

Quantum state tomography, or the task of reconstructing a quantum state by collecting and processing measurement data, is an essential primitive for quantum sensing, quantum simulation, and quantum information processing.
The central importance of quantum state tomography has led to the development of a variety of techniques based on compressed sensing \cite{gross2010quantum}, matrix product states \cite{cramer2010efficient}, maximum-likelihood estimation \cite{smolin2012efficient}, Bayesian inference \cite{huszar2012adaptive}, and neural networks \cite{torlai2018neuralnetwork}, among many others \cite{christandl2012reliable, qi2013quantum, opatrny1997leastsquares}
These techniques are typically developed in a general, information-theoretic setting, and make minimal assumptions about the physical medium of a quantum state.
As a consequence, even well-established techniques can nonetheless be ill-suited for physical platforms with unique or limited capabilities.

Due to advancements in experimental capabilities to address nuclear spins in ultracold atomic systems [\red{CITE}], as well as developments in the control of ultracold molecular systems \cite{zeppenfeld2012sisyphus, kozyryev2017sisyphus, puri2017synthesis, wu2017cryofuge, marco2019degenerate, liu2019molecular, anderegg2019optical, chou2020frequencycomb, lin2020quantum, covey2018approach, jamadagni2019quantum}, a particular setting of growing interest is the spin qudit, or a multilevel quantum angular momentum degree of freedom.
Spin qudits can provide advantages over their qubit counterparts for quantum sensing \cite{evrard2019enhanced}, enable quantum simulations of SU($d$) magnetism \cite{banerjee2013atomic, cazalilla2014ultracold, zhang2014spectroscopic, scazza2014observation, goban2018emergence}, and offer unique capabilities for quantum error correction \cite{albert2020robust, gross2020encoding}.
In all cases, quantum state tomography is necessary to take full advantage of a spin qudit.

The general problem of qudit tomography is not new, with an extensive literature on a variety of techniques \cite{manko1997spin, schmied2011tomographic, evrard2019enhanced, flammia2005minimal, thew2002qudit, salazar2012quantum, sosa-martinez2017quantum, ha2018minimal, stefano2019set, palici2020oam}.
However, existing protocols either rely infinite-dimensional representations of a quantum spin \cite{manko1997spin, schmied2011tomographic, evrard2019enhanced}, or require the capability to perform essentially arbitrary operations on a qudit \cite{thew2002qudit, flammia2005minimal, salazar2012quantum, sosa-martinez2017quantum, ha2018minimal, stefano2019set, palici2020oam}, generally resulting in tomographic protocols that can be highly inefficient or unachievable in practice.
The protocols based on infinite-dimensional representations of a quantum spin have the advantage of reconstructing its state from measurements of spin projection onto different spatial axes, which are generally accessible with any spin qudit.
Nonetheless, these protocols obfuscate the minimal requirements for performing full state tomography, provide no straightforward error bounds or guarantees of accuracy, and (with the exception of Ref.~\cite{schmied2011tomographic}) extract only a small faction of the information contained in measurement data.

In this work, we provide an efficient framework for performing spin qudit tomography using only measurements of spin projection onto different spatial axes.
We first introduce, in Section \ref{sec:polarization_ops}, a set of qudit operators that are a quantum analogue of the spherical harmonics, and which play a central role in our work.
We then map the quantum problem of spin qudit tomography onto the well-studied classical problem of signal recovery on the sphere in Section \ref{sec:signal_recovery}, thereby importing a host of existing literature and mathematical machinery that has been developed for this task \cite{mcewen2011novel, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality}.
In particular, this mapping allows us to bound the minimum number of measurement axes necessary to perform full tomography on a $d$-level spin qudit, $r_d^{\t{min}}$, by $2d-1\le r_d^{\t{min}}\le d^2$.
In Section \ref{sec:error} we provide (a priori) upper bounds and (a posteriori) estimates of the statistical error in a qudit state reconstructed from measurements of spin projection along a given set of $r$ measurement axes.
The capability to determine upper bounds on reconstruction error a priori motivates a simple randomized tomography protocol that we outline in Section \ref{sec:protocol}, and for which we numerically find that using more measurement axes yields substantial benefits that plateau after $r\approx3d$.

% todo: comment on code availability

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Polarization operators}
\label{sec:polarization_ops}

We begin by introducing a set of qudit operators that are analogous to the spherical harmonics $Y_{\ell m}$, and which play a central role in our work.
Consider a $d$-state spin qudit with total spin $s\equiv\frac{d-1}{2}$.
The defining property of a spin qudit is that it describes an angular momentum degree of freedom, which has specific implications for how a spin qudit should transform under the group SO(3) of rotations in 3D space.
Due to the central importance of these transformation rules for a spin qudit, we seek a basis of operators that transform nicely under 3D rotations \footnote{Technically speaking, we seek a basis of operators that transform as an irreducible representation of SO(3).}.
One such basis is that of the {\it polarization operators} \cite{kryszewski2006positivity, bertlmann2008bloch}, defined by
\begin{align}
  T_{\ell m} \equiv \sqrt{\f{2\ell+1}{2s+1}} \sum_{\mu,\nu=-s}^s
  \bk{s\mu;\ell m|s\nu} \op{\nu}{\mu},
  \label{eq:trans_op}
\end{align}
where $\bk{s\mu;\ell m|s\nu}$ is a Clebsh-Gordan coefficient that enforces $\ell\in\set{0,1,\cdots,d-1}$ and $m\in\set{-\ell,-\ell+1,\cdots,\ell}$, such that there are $d^2$ polarization operators in total.
For brevity, we will generally treat the value of $d$ (or $s$) as constant but arbitrary throughout this work, and suppress any explicit dependence of quantities or operators such as $T_{\ell m}$ on $d$.
The polarization operators are orthonormal with respect to the trace inner product, and transform nicely under conjugation:
\begin{align}
  \obk{T_{\ell m }|T_{\ell'm'}}
  = \delta_{\ell\ell'} \delta_{mm'},
  &&
  T_{\ell m}^\dag = \p{-1}^m T_{\ell,-m},
\end{align}
where for any $d\times d$ matrix $X=\sum_{\mu,\nu} X_{\mu\nu} \op{\mu}{\nu}$ we define the $d^2$-component vector $\oket{X} \equiv \sum_{\mu,\nu} X_{\mu\nu} \ket{\mu\nu}$; $\obra{X}$ is the conjugate transpose of $\oket{X}$, such that $\obk{X|Y}=\tr\p{X^\dag Y}$; and $\delta_{kk'}\equiv 1$ if $k=k'$ and $0$ otherwise.
These properties of the polarization operators allow us to expand any density operator $\rho$ in the polarization operator basis as
\begin{align}
  \rho = \sum_{\ell=0}^{d-1} \sum_{m=-\ell}^\ell
  \rho_{\ell m} T_{\ell m},
  &&
  \rho_{\ell m} \equiv \bk{T_{\ell m}^\dag}_\rho,
  \label{eq:trans_state}
\end{align}
where $\bk{X}_\rho\equiv\tr\p{\rho X}=\obk{\rho|X}$, and $\rho^\dag=\rho$ implies that $\rho_{\ell m}^*=\p{-1}^m\rho_{\ell,-m}$.
The polarization operators can be physically interpreted by the transition $\ket\psi\to T_{\ell m}\ket\psi$ induced on the state $\ket\psi$ of a spin-$s$ system by the absorption of a particle with total spin $\ell$ and spin projection $m$ onto a fixed quantization axis.
In analogy with the spherical harmonics $Y_{\ell m}$, we will refer to the index $\ell$ as the {\it degree} and $m$ as the {\it order} of $T_{\ell m}$.

The polarization operators are spherical tensor operators, whose degree is preserved under 3D rotations generated by the spin operators $S_\x,S_\y,S_\z$.
Moreover, the degree-$\ell$ polarization operators $T_{\ell m}$ transform identically to spin-$\ell$ particles under 3D rotations (see Appendix \ref{sec:rotations}).
For any triplet of Euler angles $\v\omega=\p{\alpha,\beta,\gamma}$, we can therefore define the rotation operator
\begin{align}
  R\p{\v\omega} \equiv e^{-\i\alpha S_\z} e^{-\i\beta S_\y} e^{-\i\gamma S_\z},
\end{align}
and expand polarization operators in a rotated frame as
\begin{align}
  T_{\v\omega\ell m} \equiv
  R\p{\v\omega}^\dag T_{\ell m} R\p{\v\omega}
  = \sum_{n=-\ell}^\ell D_{nm}^\ell\p{\v\omega} T_{\ell n},
  \label{eq:trans_rot}
\end{align}
where
\begin{align}
  D_{mn}^\ell\p{\v\omega} \equiv \bk{\ell m|R\p{\v\omega}|\ell n}
\end{align}
are (Wigner) rotation matrix elements.
Throughout this work, we will primarily consider rotations of the sphere that take the north pole to a point $\v v=\p{\alpha,\beta}$ at azimuthal angle $\alpha$ and polar angle $\beta$.
For ease of notation, we will therefore implicitly associate Euler angle doublets $\v v=\p{\alpha,\beta}$ with the triplet $\p{\alpha,\beta,0}$, such that $R\p{\v v} \equiv R\p{\alpha,\beta,0}$, $D^\ell_{mn}\p{\v v} \equiv D^\ell_{mn}\p{\alpha,\beta,0}$, and $T_{\v v\ell m} \equiv T_{\p{\alpha,\beta,0},\ell m}$.
We will also frequently consider polarization operators with order $m=0$.
Symbols with a suppressed order therefore implicitly set this order to zero, which is to say that $T_\ell\equiv T_{\ell,0}$, $T_{\v v\ell}\equiv T_{\v v\ell,0}$, and $D^\ell_m\equiv D^\ell_{m,0}$.

As a final point, we mention that the polarization operators $T_{\ell m}$ share a connection to the spherical harmonics $Y_{\ell m}$ that goes beyond the rules for their transformation under 3D rotations.
In fact, the phase-space (Wigner function) representation of $T_{\ell m}$ is essentially equal to $Y_{\ell m}$ (up to complex conjugation and a normalization factor).
The phase-space representation of an operator $X$ assigns, to each point $\v v$ on the sphere, the complex number
\begin{align}
  \tilde X\p{\v v} \equiv \bk{s_{\v v}|X|s_{\v v}},
\end{align}
where $\ket{s_{\v v}}\equiv R\p{\v v}\ket{s}$ is the state of a spin qudit polarized along $\v v$.
This representation is faithful in the sense that $X$ is uniquely determined by the phase-space values $\tilde X\p{\v v}$ at all points $\v v$ on the sphere.
The transformation rules for polarization operators in Eq.~\eqref{eq:trans_rot}, together with the fact that $\bk{s|T_{\ell m}|s}=0$ unless $m=0$, suffice to show that their phase-space distributions are
\begin{align}
  \tilde T_{\ell m}\p{\v v} = c_\ell Y_{\ell m}\p{\v v}^*,
\end{align}
where the proportionality factor $c_\ell$ simply enforces that $\obk{T_{\ell m}|T_{\ell m}}=1$ (see Appendix \ref{sec:rotations}).
In this way, the polarization operators $T_{\ell m}$ are a quantum analogue of spherical harmonics $Y_{\ell m}$, which play an important role in phase-space formalisms for spin qudits \cite{li2013weylwignermoyal}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Spin tomography as signal recovery on the sphere}
\label{sec:signal_recovery}

% todo: add figure for signal recovery problem?

Our goal is to reconstruct an arbitrary state $\rho$ of a spin qudit from measurements of spin projection onto different quantization axes.
We are thus nominally restricted to measuring projectors $\Pi_{\v v\mu} \equiv \op{\mu_{\v v}}$, where $\ket{\mu_{\v v}}\equiv R\p{\v v}\ket{\mu}$ is a state with spin projection $\mu$ onto the measurement axis $\v v$.
For any fixed axis $\v v$, the sets $\set{\Pi_{\v v\mu}}$ and $\set{T_{\v v\ell}}$ are both complete bases for the space of operators that are diagonal in the basis $\set{\ket{\mu_{\v v}}}$.
Measuring the projectors $\set{\Pi_{\v v\mu}}$ is therefore equivalent to measuring the polarization operators $\set{T_{\v v\ell}}$, and provides data on the expectation values $\bk{T_{\v v\ell}}_\rho$.

In order to reconstruct an arbitrary density operator $\rho$ from the expectation values $\bk{T_{\v v\ell}}_\rho$, we essentially need to to find a set of coefficients $C_{\ell mk}\p{\v v}$ that would allow us to recover any matrix element $\rho_{\ell m}$ of $\rho$ through
\begin{align}
  \rho_{\ell m} = \bk{T_{\ell m}^\dag}_\rho
  = \sum_{\v v,k} C_{\ell mk}\p{\v v} \bk{T_{\v vk}^\dag}_\rho.
  \label{eq:state_recon}
\end{align}
Expanding the rotated polarization operators $T_{\v vk}$ into a sum of un-rotated polarization operators, according to Eq.~\eqref{eq:trans_rot}, we find that the recovery condition in Eq.~\eqref{eq:state_recon} is satisfied when
\begin{align}
  T_{\ell m}^\dag
  = \sum_{\v v,k,n} C_{\ell mk}\p{\v v}
  D^k_n\p{\v v}^* T_{kn}^\dag.
\end{align}
Orthogonality of the polarization operators motivates the ansatz $C_{\ell mk}\p{\v v}=\delta_{k\ell}C_{\ell m}\p{\v v}$, which in turn implies that
\begin{align}
  \sum_{\v v} C_{\ell m}\p{\v v} D^\ell_n\p{\v v}^* = \delta_{mn}
  \label{eq:tomo_recovery}
\end{align}
for all $\ell$.

In fact, the problem of finding suitable axes $V$ and coefficients $C_{\ell m}\p{\v v}$ to satisfy Eq.~\eqref{eq:tomo_recovery}, solving which would enable the recovery of arbitrary spin qudit states, can be mapped onto the well-studied problem of signal recovery on the sphere \cite{mcewen2011novel, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality}.
The signal recovery problem can be stated as follows: given a square-integrable function $f$ on the sphere, with the spherical harmonic expansion
\begin{align}
  f\p{\v v} = \sum_{\ell,m} f_{\ell m} Y_{\ell m}\p{\v v},
\end{align}
where $f_{\ell m}$ are complex coefficients, find a set of points $V=\set{\v v}$ and associated coefficients $\tilde C_{\ell m}\p{\v v}$ with which we can reconstruct $f$, or equivalently its coefficients $f_{\ell m}$, from knowledge of the function's value $f\p{\v v}$ at all points $\v v\in V$; that is
\begin{align}
  f_{\ell m} = \sum_{\v v} \tilde C_{\ell m}\p{\v v} f\p{\v v}
  = \sum_{\v v,k,n} \tilde C_{\ell m}\p{\v v} Y_{kn}\p{\v v} f_{kn}.
\end{align}
Reconstruction of functions with arbitrary coefficients $f_{\ell m}$ implies that
\begin{align}
  \sum_{\v v} \tilde C_{\ell m}\p{\v v} Y_{kn}\p{\v v}
  = \delta_{k\ell} \delta_{mn},
  \label{eq:full_recovery}
\end{align}
which is a stronger version of the condition that we found for the spin qudit tomography problem in Eq.~\eqref{eq:tomo_recovery}.
We will refer to Eq.~\eqref{eq:full_recovery} as the {\it full recovery problem}, and Eq.~\eqref{eq:tomo_recovery} as the the {\it reduced recovery problem}.
Due to the fact that $D^k_n\p{\v v}^* = \sqrt{\frac{4\pi}{2\ell+1}}\, Y_{kn}\p{\v v}$, any solution to the full recovery problem automatically solves the reduced recovery problem by setting $C_{\ell m}\p{\v v} = \sqrt{\frac{2\ell+1}{4\pi}}\, \tilde C_{\ell m}\p{\v v}$.
In principle, this mapping allows us to import a host of existing signal recovery algorithms \cite{mcewen2011novel, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality} for the task of spin qudit tomography.
In practice, spin qudits typically have only a modest dimension $d$, which allows for simpler and optimized tomography protocols that are practical despite worse scaling with $d$ (see Section \ref{sec:protocol}).

If the function $f$ is {\it band-limited} at degree $L$, which is to say that $f_{\ell m}=0$ for all $\ell\ge L$, then the full recovery problem in Eq.~\eqref{eq:full_recovery} is provably solvable with a suitable choice of $\abs{V}=L^2$ points on the sphere \cite{freeden2008spherical, freeden2018spherical}.
The existence of these solutions to the full recovery problem in turn implies the existence of $d^2$ measurement axes that suffice to reconstruct arbitrary states of $d$-level spin qudit, whose possible states are essentially ``band-limited'' at degree $d$ \footnote{Spin qudit states can be faithfully represented by quasi-probability distributions on a sphere.
  For a $d$-level spin, these quasi-probability distributions are band-limited at degree $d$.}.
Moreover, for any fixed degree $\ell$, finding solutions to the reduced recovery problem in Eq.~\eqref{eq:tomo_recovery} is equivalent to the recovery of a degree-$\ell$ function $f_\ell = \sum_m f_{\ell m} Y_{\ell m}$, which is provably possible with $\abs{V}=2\ell+1$ samples \cite{freeden2008spherical}.
In the case of spin qudit tomography, the degree $\ell$ takes a maximal value of $\ell_{\t{max}}\equiv d-1$, so state recovery requires at least as many measurement axes as there are polarization operators with degree $\ell_{\t{max}}$, namely $2\ell_{\t{max}}+1=2d-1$.

Since
\begin{enumerate*}
\item there are only $2\ell+1\le2d-1$ polarization operators of degree $\ell\le\ell_{\t{max}}$, and
\item measuring spin projection along any single axis provides measurement data for polarization operators of {\it all} degrees $\ell\le\ell_{\t{max}}$,
\end{enumerate*}
one might expect that $2d-1$ suitably chosen measurement axes would always suffice to perform full tomography of a $d$-level spin qudit.
In the language of classical signal recovery, we would expect there to exist a set of $2\ell_{\t{max}}+1$ points on the sphere that enable the recovery of {\it any} function with fixed degree $\ell\le\ell_{\t{max}}$, given prior knowledge of $\ell$ \footnote{We could further conjecture that if {\it any} set of points $V\subset\Omega_2$ enables the recovery of degree-$\ell_{\t{max}}$ functions on the sphere, then $V$ enables the recovery degree-$\ell$ functions with $\ell\le\ell_{\t{max}}$.
  However, this condition is stronger than necessary to guarantee full recovery of a $d$-level spin qudit state with $2d-1$ measurement axes, which only requires the existence of {\it some} such set of points.}.
Unfortunately, we have no general proof of this conjecture, which requires a deeper analysis of fixed-degree signal recovery on the sphere.
Nonetheless, in the following sections we numerically verify that $2d-1$ points suffice for all $d\le200$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Tomographic reconstruction error}
\label{sec:error}

For the practically minded, the mere existence of solutions to a problem is less interesting than the exposition of a particular solution.
On a high level, a spin qudit tomography protocol consists of
\begin{enumerate*}
\item selecting a set of measurement axes,
\item collecting measurement data on spin projection onto these axes, and then
  \label{tomo:measure}
\item processing the collected data to reconstruct the state of the spin qudit.
\end{enumerate*}
Whereas step \ref{tomo:measure} can involve a host of platform-dependent technical challenges, in the following two sections we discuss the steps to take before and after collecting measurement data.

To this end, we begin by asking a question: what is a ``good'' choice of measurement axes?
Intuitively, a good choice of axes should minimize the error with which one can reconstruct an unknown quantum state from associated measurement data.
If we can quantify this intuition, then we can optimize over different choices of measurement axes to find a set that (approximately) minimizes the error in reconstructed states.

A set of measurement axes $V=\set{\v v}$ nominally induces a set of projectors $\set{\Pi_{\v v\mu}}$ that will be measured in an experiment.
By a simple change of basis, measuring these projectors is equivalent to measuring the polarization operators $\set{T_{\v v\ell}}$.
Flattening each $d\times d$ matrix $T_{\v v\ell}$ into the $d^2$-component column vector $\oket{T_{\v v\ell}}$, we construct the {\it measurement matrix}
\begin{align}
  M_V \equiv \sum_{\v v,\ell} \oket{T_{\v v\ell}} \bra{\v v\ell},
  \label{eq:meas_mat}
\end{align}
Here $\v v$ and $\ell$ label a column of $M_V$, or equivalently label a standard basis vector $\ket{\v v\ell}$ of a $\p{\abs{V}\times d}$-dimensional vector space.
A necessary and sufficient condition for $V$ to allow for full state tomography is that the measured polarization operators in $\set{T_{\v v\ell}}$, or equivalently the columns of $M_V$, span the entire ($d^2$-dimensional) space of operators on a $d$-level spin qudit.
In this case $M_V$ must be full rank, with $d^2$ nonzero singular values.
Indexing these singular values $M^V_k$ and the corresponding (normalized) right singular vectors $\v x^V_k\equiv(x^V_{k,1},x^V_{k,2},\cdots,x^V_{k,\abs{V}})$ by an integer $k\in\set{1,2,\cdots,d^2}$, we can construct the orthonormal qudit operators
\begin{align}
  Q^V_k \equiv \sum_j q^V_{kj} T_j,
  &&
  \v q^V_k \equiv \f{\v x^V_k}{M^V_k},
\end{align}
where for shorthand we use a combined index $j=\p{\v v,\ell}$ to specify both a measurement axis $\v v$ and a degree $\ell$, which identify the polarization operator $T_j\equiv T_{\v v\ell}$.
These operators allow us to expand any state $\rho$ of a $d$-level spin qudit in the form
\begin{align}
  \rho = \sum_{k=1}^{d^2} \rho_k^V Q_k^V,
  &&
  \rho_k^V \equiv \bk{Q^V_k}_\rho.
\end{align}
Given empirical estimates $\tilde T_j$ of the expectation values $\bk{T_j}_\rho$, an empirical estimate $\tilde\rho_V$ of $\rho$ is then
\begin{align}
  \tilde\rho_V \equiv \sum_k \tilde\rho^V_k Q^V_k,
  \label{eq:reconstructed_state}
\end{align}
where
\begin{align}
  \tilde\rho^V_k \equiv \sum_j q^V_{kj} \tilde T_j
  \approx \sum_j q^V_{kj} \bk{T_j}_\rho
  = \bk{Q^V_k}_\rho
  = \rho^V_k.
\end{align}
The measurement matrix $M_V$ allows us to make concrete statements about the statistical error between the empirical estimate $\tilde\rho_V$ and the true state $\rho$.
Assume, for example, that the estimates $\tilde T_j$ are equal to $\bk{T_j}_\rho$ up to uncorrelated noise with variance at most $\epsilon^2$:
\begin{align}
  \tilde T_j = \bk{T_j}_\rho + \epsilon_j,
  &&
  \bbk{\epsilon_j \epsilon_{j'}} \le \epsilon^2 \delta_{jj'},
  \label{eq:error_assumption}
\end{align}
where $\set{\epsilon_j}$ are independent random variables and we use the double brackets $\bbk{\cdot}$ to denote statistical averaging over experimental trials that estimate $\bk{T_j}_\rho$.
In this case, the mean squared error with which $\tilde\rho^V_k$ approximates $\rho^V_k$ is
\begin{align}
  \Bbk{\abs{\tilde\rho^V_k-\rho^V_k}^2}
  &= \Bbk{\p{\tilde\rho^V_k-\rho^V_k}^*\p{\tilde\rho^V_k-\rho^V_k}} \\
  &= \sum_{j,j'} \p{q^V_{kj}}^* q^V_{kj'}\,
  \bbk{\epsilon_j \epsilon_{j'}} \\
  &\le \sum_j \abs{q^V_{kj}}^2 \epsilon^2
  = \p{\f{\epsilon}{M^V_k}}^2.
\end{align}
Using the fact that the operators $Q^V_k$ are orthonormal, we can therefore bound the mean squared (Euclidean) distance between $\tilde\rho_V$ and $\rho$ as
\begin{align}
  \E_V\p{\rho}^2
  \equiv \Bbk{\norm{\tilde\rho_V-\rho}^2}
  \le \epsilon^2 \S_V^2,
  \label{eq:bound_eps}
\end{align}
where $\norm{X}^2 \equiv \obk{X|X} = \tr\p{X^\dag X}$ is the squared (Frobenius, or Hilbert-Schmidt) norm of $X$, and the {\it classical error scale} $\S_V$ is defined by
\begin{align}
  \S_V^2 \equiv \sum_k \p{M^V_k}^{-2} = \norm{M_V^{-1}}^2,
\end{align}
where $M_V^{-1}$ is the right inverse of $M_V$, satisfying $M_V M_V^{-1}=1$.
We refer to the error scale $\S_V$ as ``classical'' because the bound in Eq.~\eqref{eq:bound_eps} applies in the presence of classical sources of measurement error.
Note that the classical error scale $\S_V$ diverges if the measurement matrix $M_V$ is singular, which indicates that measuring spin projections along all axes in $V$ does not provide sufficient information to reconstruct arbitrary quantum states.

Computing the classical error scale $\S_V$ and estimates $\tilde\rho^V_k\approx\rho^V_k$ requires building the measurement matrix $M_V$ and computing its singular value decomposition.
The complexity of this task can be greatly reduced by the fact that the degree $\ell$ of a polarization operator $T_{\ell m}$ is preserved under rotations, which implies that the unitary
\begin{align}
  U \equiv \sum_{\ell=0}^{d-1} \sum_{m=-\ell}^\ell
  \oket{T_{\ell m}} \bra{\ell m},
  \label{eq:trans_basis}
\end{align}
with vectors $\oket{T_{\ell m}}$ in a column indexed by integers $\p{\ell,m}$, block-diagonalizes the measurement matrix into $d$ blocks indexed by the degree $\ell$:
\begin{align}
  U^\dag M_V = \sum_{\ell=0}^{d-1} \op{\ell} \otimes M_{V\ell},
\end{align}
where the $\p{2\ell+1}\times\abs{V}$-sized blocks are
\begin{align}
  M_{V\ell} \equiv \sum_{m,\v v} \ket{m} \obk{T_{\ell m}|T_{\v v\ell}} \bra{\v v}
  = \sum_{m,\v v} D^\ell_m\p{\v v} \op{m}{\v v}.
\end{align}
As the singular values of $M_V$ are invariant under unitary transformations, it follows that
\begin{align}
  \S_V^2 = \sum_\ell \S_{V\ell}^2,
  &&
  \S_{V\ell}^2 \equiv \norm{M_{V\ell}^{-1}}^2,
\end{align}
where $M_{V\ell}^{-1}$ is the right inverse of $M_{V\ell}$.
Constructing the block $M_{V\ell}$ and computing its singular value decomposition takes at most $O(\abs{V}d^2)$ time.
If we assume that $\abs{V}\sim d$, then computing the classical error scale $\S_V$ takes $O(d^4)$ serial or $O(d^3)$ parallel runtime (see Figure \ref{fig:times}).

\begin{figure}
  \centering
  \includegraphics{qudit_times.pdf}
  \caption{Serial runtime to compute $\S_V$, $\epsilon_V$, or $\E_V\p{\rho}$ with $\abs{V}=2d-1$ randomly chosen measurement axes and a randomly chosen qudit state $\rho$.
    Each point is an average over $10^3$ calculations or 5 minutes of runtime, whichever comes first.
    These results do not count fixed runtimes to pre-compute quantities that can be recycled for every new choice of $V$ or $\rho$.
    Dashed lines show fits to a runtime $t=c d^\alpha$ for all $d\ge100$, finding $\alpha\approx 3.7$.
    These simulations found axes $V$ with $\epsilon_V<d$ for all $d\le200$, thereby verifying for these cases that $2d-1$ measurement axes are sufficient to perform full state tomography.}
  \label{fig:times}
\end{figure}

The assumption that observables can be estimated up to uncorrelated noise with maximal variance $\epsilon^2$, summarized by Eq.~\eqref{eq:error_assumption}, is reasonable when measurement error is dominated by classical sources of experimental noise.
However, this assumption breaks down when measurement error is limited by fundamental quantum (shot) noise.
We relax the assumption of Eq.~\eqref{eq:error_assumption} in Appendix \ref{sec:bound}, where we instead assume that $\tilde\rho_V$ is built from $n$ independent measurements of spin projection along every axis $\v v\in V$, with shot noise the dominant source of error.
In this case, the constraints that $\tr\p\rho=1$ and $\bk{\Pi_{\v v\mu}}_\rho\ge0$ allow us to bound the mean squared distance between $\tilde\rho_V$ and $\rho$ as
\begin{align}
  \E_V\p{\rho}^2 < \f{\epsilon_V^2}{n},
  &&
  \epsilon_V^2 \equiv \sum_\ell \Gamma_\ell^2 \S_{V\ell}^2,
  \label{eq:bound}
\end{align}
where the {\it quantum error scale} $\epsilon_V$, so called because it is set by quantum shot noise, is defined in terms of (half of) the spectral range of $T_\ell$:
\begin{align}
  \Gamma_\ell \equiv \f{\max_\mu t_{\ell\mu} - \min_\mu t_{\ell\mu}}{2},
  &&
  t_{\ell\mu} \equiv \bk{\mu|T_\ell|\mu}.
\end{align}
If $d$ is even or $\ell$ is odd, then $\Gamma_\ell=\max_\mu t_{\ell\mu}$.
For comparison with the ``classical'' error bound in Eq.~\eqref{eq:bound_eps}, we note that $\epsilon_V^2<\S_V^2/2$, so the previous bound still holds with the replacement $\epsilon^2\to1/2n$.
The factors $\Gamma_\ell^2$ are quick to compute and can be recycled for every new choice of axes $V$, so the complexity of computing $\epsilon_V$ is the same as that of $\S_V$ (see Figure \ref{fig:times}).

Though straightforward to compute, the bound in Eq.~\eqref{eq:bound} is not tight, as it is acquired by bounding the statistical error in the empirical estimate $\tilde T_{\v v\ell}$ of $\bk{T_{\v v\ell}}_\rho$ by $\bbk{\epsilon_{\v v\ell}^2}\le\Gamma_\ell^2$.
Though the individual bounds on $\bbk{\epsilon_{\v v\ell}^2}$ for each axis $\v v$ and degree $\ell$ are tight, these bounds cannot all be achieved simultaneously.
There is therefore still room for improvement on the bound in Eq.~\eqref{eq:bound} by maximizing $\E_V$ over the set of all qudit states $\rho$.
We discuss this maximization problem in Appendix \ref{sec:exact}.
We also note that the (squared) reconstruction error bound in Eq.~\eqref{eq:bound} obeys the ``standard quantum limit'' of $\sim1/n$ scaling in the number of measurements.
In principle, this scaling can be improved to $\sim1/n^2$ by preparing and measuring entangled copies of many qudits \cite{giovannetti2006quantum}.

The quantum error scale $\epsilon_V$ provides a pessimistic upper bound on statistical error, which can be calculated without prior knowledge of the true qudit state $\rho$.
The actual error in the reconstruction $\tilde\rho_V$ of a particular state $\rho$ may be considerably smaller, and depends on $\rho$ itself.
The exact mean squared distance between $\tilde\rho_V$ and $\rho$, derived in Appendix \ref{sec:exact}, can be written in the form
\begin{align}
  \E_V\p{\rho}^2
  = \f1n \sum_\ell \sp{\bk{\chi_{V\ell}|\rho_\ell}
    - \bk{\rho_\ell|\N_{V\ell}|\rho_\ell}},
  \label{eq:error}
\end{align}
where $\ket{\rho_\ell} \equiv \sum_m \rho_{\ell m}\ket{m}$, and $\ket{\chi_{V\ell}}$ and $\N_{V\ell}$ are defined below.
While the true error in $\tilde\rho_V$ cannot be known exactly without knowing $\rho$, this error can be estimated by $\E_V\p{\tilde\rho_V} \approx \E_V\p{\rho}$.

The matrix $\N_{V\ell}$ that appears in Eq.~\eqref{eq:error} is defined in terms of the measurement matrix $M_{V\ell}$ and its right inverse $M_{V\ell}^{-1}$ by
\begin{align}
  \N_{V\ell} \equiv M_{V\ell} \diag\sp{M_{V\ell}^{-1}
    \p{M_{V\ell}^{-1}}^\dag} M_{V\ell}^\dag,
\end{align}
where $\diag\sp{X}$ sets all off-diagonal parts of $X$ to zero.
The vector $\ket{\chi_{V\ell}}\equiv\sum_m \chi^V_{\ell m} \ket{m}$ in Eq.~\eqref{eq:error} is defined by
\begin{align}
  \chi^V_{LM} &\equiv
  \sum_\ell \obk{\N_{V\ell}|\D_M|\tilde g_{L\ell}}, \\
  \tilde g_{L\ell} &\equiv \sum_{m,m'}
  \obk{T_{L,m+m'}|T_{\ell m}^\dag T_{\ell m'}} \op{m}{m'}, \\
  \D_M &\equiv \sum_{m,m'} \delta_{M,m'-m} \op{mm'}.
\end{align}
Here $\tilde g_{L\ell}$ is essentially a matrix of structure constants for the polarization operator algebra (see Appendix \ref{sec:trans_prod}), and $\D_M$ simply picks off the $M$-th diagonal of the matrix it acts on, such that $\obk{\N_{V\ell}|\D_M|\tilde g_{L\ell}}$ is an inner product of the $M$-th diagonal of $\tilde g_{L\ell}$ with the $\p{-M}$-th diagonal of $\N_{V\ell}$.
Overall, the complexity of computing $\E_V\p{\rho}$ is the same as that of $\S_V$ and $\epsilon_V$ (see Figure \ref{fig:times}).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Tomography protocol}
\label{sec:protocol}

The ability to certify a statistical error bound on the empirical estimate $\tilde\rho_V$ of an unknown quantum state $\rho$ motivates the following simple protocol for spin qudit tomography: select a random set of measurement axes $V$ by uniformly sampling points on the sphere \footnote{To sample a point $\p{\alpha,\beta}$ from the uniform distribution on the sphere (with azimuthal angle $\alpha$ and polar angle $\beta$), you can sample a point $\p{a,b}\in[0,1]\times[0,1]$ from the uniform distribution on the unit square, and then set $\alpha=2\pi a$ and $\beta=\arccos\p{1-2b}$.}, and optimize (using any standard minimization algorithm) the $2\abs{V}$ parameters in $V$ (two angles for each point $\v v\in V$) to minimize the quantum error scale $\epsilon_V$.
If $\abs{V}$ is too large for such optimization to be practical, one can instead generate many random sets of measurement axes, and then choose the set with the smallest quantum error scale $\epsilon_V$.
After choosing a set of measurement axes $V$, (experimentally) make $n$ measurements of spin projection along every axis $\v v\in V$, and set $\tilde\Pi_{\v v\mu}\approx\bk{\Pi_{\v v\mu}}_\rho$ to the fraction of times in which the measurement outcome was $\mu$.
The estimates $\tilde\Pi_{\v v\mu}$ of $\bk{\Pi_{\v v\mu}}_\rho$ determine estimates $\tilde T_{\v v\ell} \equiv \sum_\mu \bk{\mu|T_\ell|\mu}\tilde\Pi_{\v v\mu}$ of $\bk{T_{\v v\ell}}_\rho$, which can in turn be used to construct the estimate $\tilde\rho_V$ of $\rho$.
The total error in $\tilde\rho_V$, measured by its root-mean-square distance from $\rho$, is approximately $\E_V\p{\tilde\rho_V}\approx\E_V\p{\rho}$.
If $\tilde\rho_V$ has negative eigenvalues, its distance from $\rho$ can be reduced with maximum-likelihood corrections \cite{smolin2012efficient}.

Note that any information about $\rho$, obtained either from prior knowledge or preliminary measurement data, can be used to construct tailored or adaptive measurement protocols \cite{pereira2018adaptive} that are more efficient in terms of the number of measurements required to estimate $\rho$ to a fixed precision.
We leave the development of tailored and adaptive measurement protocols to future work.

The tomography protocol we outline above leaves open the question of {\it how many} measurement axes to use.
Though $2d-1$ measurement axes may be sufficient to perform full state tomography, this is not necessarily the best choice of $\abs{V}$.
Increasing the number of measurement axes generally decreases the quantum error scale $\epsilon_V$, but at the cost of having to estimate more observables.
At a fixed total number of measurements, increasing $\abs{V}$ reduces the number of measurements $n$ devoted to each axis $\v v\in V$.
This trade-off begs the question: how should one choose the number of measurement axes, $\abs{V}$?

The reconstruction error bound in Eq.~\eqref{eq:bound} nominally provides a straightforward answer: at a fixed total number of measurements, $N=n\abs{V}$, the number of measurement axes should be chosen to minimize the (squared) reconstruction error $\E_V\p{\rho}^2<\epsilon_V^2/n\propto\epsilon_V^2\abs{V}$.
We therefore consider the measurement-adjusted error scale
\begin{align}
  \beta(p)
  \equiv \min\set{\epsilon_V:\abs{V}=2d-1+p} \times \sqrt{2d-1+p}
\end{align}
as a function of $p$, the number of measurement axes exceeding $2d-1$.
Though we cannot minimize over all suitable choices of measurement axes $V$ to compute $\beta(p)$, we can compute an empirical upper bound $\tilde\beta(p)\ge\beta(p)$ by minimizing over a large number of randomly chosen $V$.
Figure \ref{fig:axes} shows the results of such empirical minimization, where we find that $\tilde\beta(p)$ drops substantially before plateauing at $p\approx d$, after which there are only marginal benefits to using more measurement axes.
In the interest of reducing experimental complexity as well as the runtime of our tomography protocol, which grow linearly in $\abs{V}$, we therefore conclude that spin qudit tomography should be performed with $\abs{V}\approx3d$ measurement axes.

\begin{figure}[h]
  \centering
  \includegraphics{qudit_axes.pdf}
  \caption{Empirical measurement-adjusted error scales $\tilde\beta(p)$ with $p$ excess measurement axes, determined by minimizing over $10^3$ choices of measurement axes $V$ or 5 minutes of runtime (for each $p$), whichever comes first.
    Color indicates the qudit dimension $d$ in the legend.
    The rapid initial drop in $\tilde\beta(p)$ implies that using more measurement axes can substantially lower the upper bound on reconstruction error provided in Eq.~\eqref{eq:bound}, and that these benefits plateau after $p\approx d$.}
  \label{fig:axes}
\end{figure}

\bibliography{\jobname.bib}

\onecolumngrid
\appendix

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Rotating polarization operators}
\label{sec:rotations}

Denoting the state of a spin-$s$ particle spin spin projection $\mu$ onto a quantization axis by $\ket{s\mu}$, we define
\begin{align}
  S_\z \equiv \sum_{\mu=-s}^s \mu \op{s\mu},
  &&
  S_\pm \equiv \sum_{\mu=-s}^s
  \sqrt{s\p{s+1}-\mu\p{\mu\pm1}} \op{s,\mu\pm1}{s\mu},
  \label{eq:spin_ops}
\end{align}
as well as
\begin{align}
  S_\x \equiv \f12\p{S_+ + S_-},
  &&
  S_\y \equiv -\f\i2\p{S_+-S_-},
  &&
  \v S \equiv \p{S_\x,S_\y,S_\z}.
\end{align}
The spin vector $\v S$ generates rotations of a spin-$s$ system in 3D space.
Specifically, the operator $e^{-\i\theta\v S\cdot\uv n}$ rotates a spin-$s$ system by an angle $\theta$ about the unit vector $\uv n$.

Observing that $S_\z=T_{1,0}$ and $S_\pm\propto T_{1,\pm1}$, we can use the product operator expansion of the polarization operators (see Appendix \ref{sec:trans_prod}), properties of Clebsch-Gordan coefficients, properties of Wigner $6$-$j$ symbols, and a computer algebra system to simplify the commutators
\begin{align}
  \sp{S_\z,T_{\ell m}} = m\, T_{\ell m},
  &&
  \sp{S_\pm,T_{\ell m}} = \sqrt{\ell\p{\ell+1}-m\p{m\pm 1}}\, T_{\ell,m\pm1},
  \label{eq:spin_trans}
\end{align}
which implies that $T_{\ell m}$ is a spherical tensor operator, whose degree degree $\ell$ is preserved under rotations generated by $\v S$.
Moreover, by comparing Eqs.~\eqref{eq:spin_ops} and \eqref{eq:spin_trans} we see that the polarization operators $T_{\ell m}$ transform identically to spin-$\ell$ particles under the adjoint action of the spin operators $S_\z$ and $S_\pm$.
For any triplet of Euler angles $\v\omega=\p{\alpha,\beta,\gamma}$, we can therefore define the rotation operator
\begin{align}
  R\p{\v\omega} \equiv
  e^{-\i\alpha S_\z} e^{-\i\beta S_\y} e^{-\i\gamma S_\z},
\end{align}
and expand polarization operators in a rotated frame as
\begin{align}
  T_{\v\omega\ell m}
  \equiv R\p{\v\omega}^\dag T_{\ell m} R\p{\v\omega}
  = \sum_{n=-\ell}^\ell D_{nm}^\ell\p{\v\omega} T_{\ell n},
  \label{eq:trans_rot_apndx}
\end{align}
where
\begin{align}
  D_{mn}^\ell\p{\v\omega} \equiv \bk{\ell m|R\p{\v\omega}|\ell n}
  = \obk{T_{\ell m} | R\p{\v\omega}^\dag
    \otimes R\p{\v\omega}^* |T_{\ell n}}
\end{align}
are matrix elements of the rotation operator $R\p{\v\omega}$ for spin-$\ell$ particles.

Implicitly associating the Euler angle doublet $\v v\equiv\p{\alpha,\beta}$ with the triplet $\p{\alpha,\beta,0}$, the transformation rules in Eq.~\eqref{eq:trans_rot_apndx} imply that we can expand the phase-space distribution of $T_{\ell m}$ as
\begin{align*}
  \tilde T_{\ell m}\p{\v v}
  \equiv \bk{s_{\v v} | T_{\ell m} | s_{\v v}}
  = \bk{s | R\p{\v v}^\dag T_{\ell m} R\p{\v v} | s}
  = D^\ell_{0,m}\p{\v v} \bk{s | T_{\ell,0} | s},
\end{align*}
where
\begin{align*}
  \bk{s | T_{\ell,0} | s}
  = \sqrt{\f{2\ell+1}{2s+1}} \, \bk{ss;\ell,0|ss}
  = \sqrt{\f{2\ell+1}{2s+\ell+1}
    \p{\f{\p{2s}!}{\p{2s+\ell}!}}
    \p{\f{\p{2s}!}{\p{2s-\ell}!}}},
\end{align*}
and the properties of the rotation matrix elements $D^\ell_{mn}$ imply that
\begin{align*}
  D^\ell_{0,m}\p{\v v} = \p{-1}^m D^\ell_{-m,0}\p{\v v}^*
  = \sqrt{\f{4\pi}{2\ell+1}} \p{-1}^m Y_{\ell,-m}\p{\v v}
  = \sqrt{\f{4\pi}{2\ell+1}} \, Y_{\ell m}\p{\v v}^*,
\end{align*}
so
\begin{align*}
  \tilde T_{\ell m}\p{\v v}
  = \sqrt{\f{4\pi}{2s+\ell+1}
    \p{\f{\p{2s}!}{\p{2s+\ell}!}} \p{\f{\p{2s}!}{\p{2s-\ell}!}}} \,
  Y_{\ell m}\p{\v v}^*
  \propto Y_{\ell m}\p{\v v}^*.
\end{align*}
In this way, the polarization operators are a quantum analogue of spherical harmonics.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{An improved reconstruction error bound}
\label{sec:bound}

In Section \ref{sec:error} of the main text, we provided a reconstruction error bound using the assumption of Eq.~\eqref{eq:error_assumption}, namely that expectation values derived from spin projection measurements can be estimated up to uncorrelated errors with maximal variance $\epsilon^2$.
This assumption is reasonable if measurement error is dominated by experimental sources of noise, and it yields a simple derivation of the reconstruction bound in Eq.~\eqref{eq:bound_eps}.
Nonetheless, there are two problems with the assumption of Eq.~\eqref{eq:error_assumption}:
\begin{enumerate*}
\item there is no a priori guarantee for the value of $\epsilon$, which must be inferred from experimental outcomes, and
\item the assumption that all errors are uncorrelated is unjustified (and generally false).
\end{enumerate*}
Here, we relax the assumption of Eq.~\eqref{eq:error_assumption} and derive an explicit error bound in terms of the qudit dimension $d$ and the number of spin projection measurements made along every measurement axis.

To this end, we fix a particular set of measurement axes $V$, and consider performing $n$ measurements of spin projection along every axis $\v v\in V$, for a total of $N=\abs{V}\times n$ measurements.
Such a procedure is equivalent to making $N$ local measurements of the $N$-fold product state $\rho^{\otimes N}$.
For convenience, we index the tensor factors of $\rho^{\otimes N}$ by the integers $\p{i,j}$, with $i\in\set{1,2,\cdots,\abs{V}}$ specifying a measurement axis $\v v_i\in V$, and $j\in\set{1,2,\cdots,n}$ specifying the copy of $\rho$ prepared for the $j$-th measurement spin projection along a particular axis.
We then define the projectors $\Pi_{i\mu}\equiv\op{\mu_{\v v_i}}$ onto states $\ket{\mu_{\v v_i}}$ with definite spin projection $\mu$ along axis $\v v_i\in V$, and define $\Pi_{i\mu}^j$ to be an $N$-qudit operator with $\Pi_{i\mu}$ on the $\p{i,j}$-th tensor factor and the identity elsewhere.
We denote the experimental outcome of measuring $\Pi_{i\mu}$ in the $\p{i,j}$-th copy of $\rho$ by $\tilde \Pi_{i\mu}^j\in\set{0,1}$.
In other words, $\tilde \Pi_{i\mu}^j$ is the ``single-shot estimate'' of $\Pi_{i\mu}$, with $\tilde \Pi_{i\mu}^j=1$ if outcome $\mu$ was observed on the $\p{i,j}$-th experimental trial, and $\tilde \Pi_{i\mu}^j=0$ otherwise.
An empirical estimate of the expectation value $\bk{\Pi_{i\mu}}_\rho$ is provided by the fraction of times that outcome $\mu$ was observed when measuring spin projection along axis $\v v_i$, that is
\begin{align}
  \tilde \Pi_{i\mu} \equiv \f1n \sum_{j=1}^n \tilde \Pi_{i\mu}^j
  \approx \f1n \sum_{j=1}^n \tr\p{\rho^{\otimes N} \Pi_{i\mu}^j}
  = \tr\p{\rho \Pi_{i\mu}}.
\end{align}
For reasons that will become clear shortly, it will be useful to think of $\tilde \Pi_{i\mu}$ as an empirical estimate of $\bk{\bar \Pi_{i\mu}}_{\rho^{\otimes N}}$, where
\begin{align}
  \bar \Pi_{i\mu} \equiv \f1n \sum_{j=1}^n \Pi_{i\mu}^j
  \label{eq:mean_proj}
\end{align}
is the average of $\Pi_{i\mu}$ applied to all copies of $\rho$ for which spin projection is measured along the axis $\v v_i$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Covariance of errors in the spin-projection basis}

Finite sampling error (i.e.~shot noise) generally induces statistical error $\epsilon_\O$ into the empirical estimate $\tilde\O$ of an observable $\O$:
\begin{align}
  \epsilon_\O \equiv \tilde\O - \bk{\O},
\end{align}
where the single brackets $\bk{\cdot}$ denote an expectation value with respect to the measured quantum state.
On average, this statistical error will be zero, which is to say that
\begin{align}
  \bbk{\epsilon_\O} = \Bbk{\tilde\O - \bk{\O}}
  = \bk{\O - \bk{\O}}
  = 0,
\end{align}
where the double brackets $\bbk{\cdot}$ to denote statistical averaging over experimental trials.
However, the covariance between statistical errors $\epsilon_\O$ and $\epsilon_\Q$ on the empirical estimates $\tilde\O$ and $\tilde\Q$ of observables $\O$ and $\Q$ is
\begin{align}
  \bbk{\epsilon_\O \epsilon_\Q}
  = \Bbk{\p{\tilde\O - \bk{\O}} \p{\tilde\Q - \bk{\Q}}}
  = \bk{\p{\O - \bk{\O}} \p{\Q - \bk{\Q}}}
  = \bk{\O\Q} - \bk{\O} \bk{\Q}.
\end{align}
In the context of spin qudit tomography, we can therefore define the statistical error
\begin{align}
  \epsilon_{i\mu}
  \equiv \tilde \Pi_{i\mu} - \bk{\Pi_{i\mu}}_{\rho}
  = \tilde \Pi_{i\mu} - \bk{\bar\Pi_{i\mu}}_{\rho^{\otimes N}}
\end{align}
in the empirical estimate of $\bk{\Pi_{i\mu}}_{\rho}$ or $\bk{\bar\Pi_{i\mu}}_{\rho^{\otimes N}}$, and use Eq.~\eqref{eq:mean_proj} to expand
\begin{align}
  \bbk{\epsilon_{i\mu} \epsilon_{i'\mu'}}
  = \bk{\bar\Pi_{i\mu} \bar\Pi_{i'\mu'}}_{\rho^{\otimes N}}
  - \bk{\bar\Pi_{i\mu}}_{\rho^{\otimes N}}
  \bk{\bar\Pi_{i'\mu'}}_{\rho^{\otimes N}}
  = \f1{n^2} \sum_{j,j'=1}^n
  \sp{\bk{\Pi_{i\mu}^j \Pi_{i'\mu'}^{j'}}_{\rho^{\otimes N}}
    - \bk{\Pi_{i\mu}^j}_{\rho^{\otimes N}}
    \bk{\Pi_{i'\mu'}^{j'}}_{\rho^{\otimes N}}}.
  \label{eq:proj_cov_start}
\end{align}
If $\p{i,j}\ne\p{i',j'}$, then $\Pi_{i\mu}^j$ and $\Pi_{i'\mu'}^{j'}$ address different tensor factors of the product state $\rho^{\otimes N}$, so the expectation value of their product factorizes due to the identity $\tr\sp{\p{A\otimes B}\p{A'\otimes B'}} = \tr\p{AA'}\times\tr\p{BB'}$.
This factorization can also be seen as a consequence of the fact that if $\p{i,j}\ne\p{i',j'}$, then $\Pi_{i\mu}^j$ and $\Pi_{i'\mu'}^{j'}$ are ``spatially separated'' on $\rho^{\otimes N}$, which means that their expectation values cannot have quantum correlations.
The terms in Eq.~\eqref{eq:proj_cov_start} with $\p{i,j}\ne\p{i',j'}$ therefore vanish, so
\begin{align}
  \bbk{\epsilon_{i\mu} \epsilon_{i'\mu'}}
  &= \delta_{ii'} \times \f1{n^2} \sum_{j=1}^n \sp{\bk{\Pi_{i\mu}^j \Pi_{i\mu'}^j}_{\rho^{\otimes N}}
    - \bk{\Pi_{i\mu}^j}_{\rho^{\otimes N}}
    \bk{\Pi_{i\mu'}^j}_{\rho^{\otimes N}}} \\
  &= \delta_{ii'} \times \f1n \sp{\bk{\Pi_{i\mu} \Pi_{i\mu'}}_\rho
    - \bk{\Pi_{i\mu}}_\rho \bk{\Pi_{i\mu'}}_\rho} \\
  &= \delta_{ii'} \times \f1n \cov_\rho\p{\Pi_{i\mu}, \Pi_{i\mu'}},
\end{align}
where $\cov_\rho\p{X,Y} \equiv \bk{XY}_\rho - \bk{X}_\rho \bk{Y}_\rho$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Variance of errors in the polarization operator basis}

Rather than the statistical errors $\epsilon_{i\mu} \equiv \tilde\Pi_{i\mu} - \bk{\Pi_{i\mu}}_\rho$ in the estimates $\tilde\Pi_{i\mu}$ of the projectors $\Pi_{i\mu}$, we now consider the statistical errors $\epsilon_{i\ell} \equiv \tilde T_{i\ell} - \bk{T_{i\ell}}_\rho$ in the estimates $\tilde T_{i\ell}$ of the polarization operators $T_{i\ell} \equiv T_{\v v_i\ell}$.
We can expand the polarization operators $T_{i\ell}$ as a sum over projectors $\Pi_{i\mu}$ as
\begin{align}
  T_{i\ell} = \sum_\mu t_{\ell\mu} \Pi_{i\mu},
  &&
  t_{\ell\mu} \equiv \bk{\mu|T_\ell|\mu}
  = \sqrt{\f{2\ell+1}{d}} \bk{s\mu;\ell|s\mu},
\end{align}
and likewise $\tilde T_{i\ell} = \sum_\mu t_{\ell\mu} \tilde\Pi_{i\mu}$.
The covariance between errors in the polarization operator basis is then
\begin{align}
  \bbk{\epsilon_{i\ell} \epsilon_{i'\ell'}}
  = \sum_{\mu,\mu'} t_{\ell\mu} t_{\ell'\mu'}
  \bbk{\epsilon_{i\mu} \epsilon_{i'\mu'}}
  = \delta_{ii'} \times \f1n
  \sum_{\mu,\mu'} t_{\ell\mu} t_{\ell'\mu'}
  \cov_\rho\p{\Pi_{i\mu}, \Pi_{i\mu'}}
  = \delta_{ii'} \times \f1n \cov_\rho\p{T_{i\ell}, T_{i\ell'}},
\end{align}
where we used the fact that the covariance $\cov_\rho\p{X,Y}$ is linear in both $X$ and $Y$.
Due to the appearance of $\delta_{ii'}$ above and the orthogonality of rotated polarization operators $T_{i\ell}$ and $T_{i'\ell'}$ with degrees $\ell\ne\ell'$, only the variance $\bbk{\epsilon_{i\ell}^2}$ ultimately contributes to reconstruction error.
We therefore seek to find an upper bound on $\bbk{\epsilon_{i\ell}^2}$.

To this end, we define the probability $p^i_\mu\equiv\bk{\Pi_{i\mu}}_\rho$, collect these probabilities into the classical probability distribution $p^i = \p{p^i_{-s}, p^i_{-s+1}, \cdots, p^i_s}$, and define the vector $t_\ell \equiv \p{t_{\ell,-s},t_{\ell,-s+1},\cdots,t_{\ell s}}$ of all $t_{\ell\mu}\equiv\bk{\mu|T_\ell|\mu}$.
We then observe that
\begin{align}
  \bbk{\epsilon_{i\ell}^2} = \f1n \times \sigma_{p^i}^2\p{t_\ell},
  &&
  \sigma_p^2\p{X}
  \equiv \sum_\mu p_\mu X_\mu^2 - \p{\sum_\mu p_\mu X_\mu}^2,
\end{align}
where $\sigma_p^2\p{X}$ is the weighted variance of $X$.
This variance is maximal when $p$ has equal weight on the largest and smallest values of $X$, which implies that
\begin{align}
  \sigma_p^2\p{t_\ell} \le \Gamma_\ell^2,
  &&
  \Gamma_\ell \equiv \f{\max_\mu t_{\ell\mu} - \min_\mu t_{\ell\mu}}{2},
  &&
  \t{so}
  &&
  \bbk{\epsilon_{i\ell}^2} \le \f1n \times \Gamma_\ell^2.
\end{align}
Note that this bound on $\bbk{\epsilon_{i\ell}^2}$ is tight, as equality is achieved by the state
\begin{align}
  \rho_i^\star = \f12\p{\Pi_{i\mu_{\t{max}}} + \Pi_{i\mu_{\t{min}}}}.
\end{align}
To find a nicer analytical bound on $\bbk{\epsilon_{i\ell}^2}$, we can use normalization of the polarization operators, $\obk{T_{i\ell}|T_{i\ell}}=\sum_\mu t_{\ell\mu}^2 = 1$, and the fact that all probabilities $p_\mu\le1$ to bound
\begin{align}
  \sigma_p^2\p{t_\ell} \le \sum_\mu p_\mu t_{\ell\mu}^2
  \le \sum_\mu t_{\ell\mu}^2 = 1,
  &&
  \t{so}
  &&
  \bbk{\epsilon_{i\ell}^2} < \f1n.
\end{align}
We can get a tighter analytical bound on $\bbk{\epsilon_{i\ell}^2}$ by considering the fact that $t_{\ell\mu}^2 = t_{\ell,-\mu}^2$ due to the symmetries of the Clebsch-Gordan coefficients.
Denoting the index that maximizes $t_{\ell\mu}^2$ by $\mu_{\t{max}}$, it follows that if $\mu_{\t{max}}\ne0$ then
\begin{align}
  \sigma_p^2\p{t_\ell} \le \sum_\mu p_\mu t_{\ell\mu}^2
  \le t_{\ell\mu_{\t{max}}}^2
  = \f12 \p{t_{\ell\mu_{\t{max}}}^2 + t_{\ell,-\mu_{\t{max}}}^2}
  \stackrel{\mu_{\t{max}}\ne0}{\le} \f12 \sum_\mu t_{\ell\mu}^2
  = \f12.
\end{align}
If $\mu_{\t{max}}=0$, then similarly
\begin{align}
  t_{\ell\mu_{\t{max}}}^2 + 2 t_{\ell\mu_{\t{min}}}^2
  = t_{\ell\mu_{\t{max}}}^2 + t_{\ell\mu_{\t{min}}}^2
  + t_{\ell,-\mu_{\t{min}}}^2
  \stackrel{\mu_{\t{max}}=0}{\le}
  \sum_\mu t_{\ell\mu}^2 = 1,
  &&
  \t{so}
  &&
  \abs{t_{\ell\mu_{\t{min}}}}
  \stackrel{\mu_{\t{max}}=0}{\le}
  \sqrt{\f{1 - t_{\ell\mu_{\t{max}}}^2}{2}},
\end{align}
where $\mu_{\t{min}}$ is the index that minimizes $t_{\ell\mu}^2$.
This inequality lets us bound
\begin{align}
  \Gamma_\ell
  = \f12 \p{t_{\ell\mu_{\t{max}}} - t_{\ell\mu_{\t{min}}}}
  \le \f12 \p{t_{\ell\mu_{\t{max}}} + \abs{t_{\ell\mu_{\t{min}}}}}
  \stackrel{\mu_{\t{max}}=0}{\le}
  \f12 t_{\ell\mu_{\t{max}}}
  + \f12\sqrt{\f{1 - t_{\ell\mu_{\t{max}}}^2}{2}}
  \equiv \lambda\p{t_{\ell\mu_{\t{max}}}}.
\end{align}
If is straightforward to show that $\lambda\p{x}$ is maximally $\lambda^\star\equiv \max_x \lambda\p{x} = \sqrt{3/8}$, so
\begin{align}
  \Gamma_\ell^2
  \stackrel{\mu_{\t{max}}=0}{\le} \p{\lambda^\star}^2
  = \f38 < \f12.
\end{align}
Altogether, we thus find that in all cases
\begin{align}
  \sigma_p^2\p{t_\ell} \le \f12,
  &&
  \t{so}
  &&
  \bbk{\epsilon_{i\ell}} \le \f1{2n}.
\end{align}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Revisiting the reconstruction error bound}
\label{sec:revisiting}

We now revisit the derivation of reconstruction error in Section \ref{sec:error} to make use of the bounds on variances $\bbk{\epsilon_{i\ell}^2}$.
To recap, for a set of measurement axes $V=\set{\v v}$ and degrees $\ell\in\set{0,1,\cdots,d-1}$ we construct the measurement matrix
\begin{align}
  M_V \equiv \sum_{\v v,\ell} \oket{T_{\v v\ell}} \bra{\v v\ell},
\end{align}
which can be block diagonalized as
\begin{align}
  U^\dag M_V = \sum_\ell \op{\ell} \otimes M_{V\ell},
  &&
  U \equiv \sum_{\ell,m} \oket{T_{\ell m}} \bra{\ell m},
  &&
  M_{V\ell} = \sum_{m,\v v} D^\ell_m\p{\v v} \op{m}{\v v},
\end{align}
where $D^\ell_{mn}\p{\v v}\equiv\bk{\ell m|R\p{\v v}|\ell n}$ is a (Wigner) rotation matrix element for a spin-$\ell$ particle.
The block-diagonal structure of $M_V$ allows us to index its singular values $M^V_{\ell m}$ and corresponding (normalized) right singular vectors $\v x^V_{\ell m} = (x^V_{\ell m,1}, x^V_{\ell m,2}, \cdots )$ by the indices $\p{\ell,m}$, where the integer $\abs{m}\le\ell$.
These singular vectors and values define the orthonormal operators
\begin{align}
  Q^V_{\ell m} \equiv \sum_i q^V_{\ell mi} T_{i\ell},
  &&
  \v q^V_{\ell m} \equiv \f{\v x^V_{\ell m}}{M^V_{\ell m}},
\end{align}
where $i\in\set{1,2,\cdots,\abs{V}}$ indexes an axis $\v v_i\in V$.
The state $\rho$ can be expanded in the basis of these operators as
\begin{align}
  \rho = \sum_{\ell,m} \bk{Q^V_{\ell m}}_\rho Q^V_{\ell m},
\end{align}
and the estimates $\tilde T_{i\ell}$ of $\bk{T_{i\ell}}_\rho$ can be used to construct the following estimate $\tilde\rho_V$ of $\rho$:
\begin{align}
  \tilde\rho_V \equiv \sum_{\ell,m}
  \sp{\sum_i q^V_{\ell mi} \tilde T_{i\ell}} Q^V_{\ell m}
  \approx \sum_{\ell,m}
  \sp{\sum_i q^V_{\ell mi} \bk{T_{i\ell}}_\rho} Q^V_{\ell m}
  = \sum_{\ell,m} \bk{Q^V_{\ell m}}_\rho Q^V_{\ell m}
  = \rho.
\end{align}
Recalling that $\epsilon_{i\ell} \equiv \tilde T_{i\ell} - \bk{T_{i\ell}}_\rho$, we can use orthonormality of all $Q^V_{\ell m}$ to expand the mean squared distance between $\tilde\rho_V$ and $\rho$ as
\begin{align}
  \E_V\p{\rho}^2 \equiv \Bbk{\norm{\tilde\rho_V-\rho}^2}
  = \sum_{\ell,m,i,i'} \p{q^V_{\ell mi}}^* q^V_{\ell mi'} \bbk{\epsilon_{i\ell} \epsilon_{i'\ell}}
  = \sum_{\ell,m,i} \abs{q^V_{\ell mi}}^2 \bbk{\epsilon_{i\ell}^2}
  < \f1n \sum_\ell \Gamma_\ell^2 \S_{V\ell}^2,
  \label{eq:bound_apndx}
\end{align}
where we used the fact that $\bbk{\epsilon_{i\ell}^2}\le\Gamma_\ell^2/n$, and
\begin{align}
  \sum_{m,i} \abs{q^V_{\ell mi}}^2
  = \sum_m \p{M^V_{\ell m}}^{-2}
  = \norm{M_{V\ell}^{-1}}
  = \S_{V\ell}^2.
\end{align}
The fact that $\bbk{\epsilon_{i\ell}^2}<1/2n$ also implies that
\begin{align}
  \E_V\p{\rho}^2 < \f1{2n} \sum_\ell \S_{V\ell}^2 = \f{\S_V^2}{2n}.
\end{align}
Note that the bound in Eq.~\eqref{eq:bound_apndx} is not tight, as the individual bounds on the variances $\bbk{\epsilon_{i\ell}^2}$ cannot all be achieved simultaneously.
There is therefore still room for improvement on the bound in Eq.~\eqref{eq:bound} by maximizing $\E_V$ over the set of all qudit states $\rho$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Exact reconstruction error}
\label{sec:exact}

Here we find exact expressions for reconstruction error, which can be used to estimate the error in a given reconstruction $\tilde\rho_V$ of an unknown state $\rho$ after performing tomography.
To this end, we start with Eq.~\eqref{eq:bound_apndx} from Appendix \ref{sec:revisiting} to write
\begin{align}
  \E_V\p{\rho}^2
  = \sum_{\ell,m,i,i'} \p{q^V_{\ell mi}}^* q^V_{\ell mi'} \bbk{\epsilon_{i\ell}\epsilon_{i'\ell}}
  = \f1n \sum_{\ell,m,i} \abs{\tilde{\v q}_{\ell i}}^2 \cov_\rho\p{T_{i\ell},T_{i\ell}},
  &&
  \abs{\tilde{\v q}_{\ell i}}^2 = \sum_m \abs{q_{\ell mi}}^2,
  \label{eq:error_SM_start}
\end{align}
where $\tilde{\v q}_{\ell i} = (q_{\ell,-\ell,i},q_{\ell,-\ell+1,i},\cdots q_{\ell,\ell,i})$, and we used the fact that $\bbk{\epsilon_{i\ell}\epsilon_{i'\ell}} = \delta_{ii'}\times\cov_\rho\p{T_{i\ell},T_{i\ell}}/n$.
Identifying the singular value decomposition $M_{V\ell} = U_{V\ell} \Sigma_{V\ell} W_{V\ell}^\dag$, we then we observe that vector $\tilde{\v q}_{\ell i}$ is essentially the $i$-th row of $\Sigma_{V\ell}^{-1} W_{V\ell}^\dag$, which is to say that $\tilde{\v q}_{\ell i} = \Sigma_{V\ell}^{-1} W_{V\ell}^\dag \ket{\v v_i}$.
We can therefore simplify
\begin{align}
  \abs{\tilde{\v q}_{\ell i}}^2
  = \bk{\v v_i | W_{V\ell} \Sigma_{V\ell}^{-2} W_{V\ell}^\dag | \v v_i}
  = \bk{\v v_i| M_{V\ell}^{-1} \p{M_{V\ell}^{-1}}^\dag | \v v_i}.
\end{align}
Using the fact that all $T_{i\ell}=T_{i\ell}^\dag$, we can also expand
\begin{align}
  \cov_\rho\p{T_{i\ell},T_{i\ell}}
  = \cov_\rho\p{T_{i\ell}^\dag,T_{i\ell}}
  = \sum_{m,m'} D^\ell_m\p{\v v_i}^* D^\ell_{m'}\p{\v v_i}
  \cov_\rho\p{T_{\ell m}^\dag, T_{\ell m'}},
\end{align}
which implies that
\begin{align}
  \E_V\p{\rho}^2
  = \f1n \sum_{i,\ell,m,m'}
  D^\ell_{m'}\p{\v v_i} \abs{\tilde{\v q}_{\ell i}}^2
  D^\ell_m\p{\v v_i}^*
  \cov_\rho\p{T_{\ell m}^\dag, T_{\ell m'}}.
\end{align}
Altogether, this reconstruction error can be expressed more compactly by defining the {\it covariance matrix}
\begin{align}
  \C_\ell\sp{\rho} \equiv
  \sum_{m,m'} \cov_\rho\p{T_{\ell m}^\dag, T_{\ell m'}} \op{m}{m'},
\end{align}
and the {\it noise matrix}
\begin{align}
  \N_{V\ell} \equiv M_{V\ell} \diag\sp{M_{V\ell}^{-1}
    \p{M_{V\ell}^{-1}}^\dag} M_{V\ell}^\dag,
\end{align}
where $\diag\sp{X}$ sets all off-diagonal entries of $X$ to zero, in terms of which
\begin{align}
  \E_V\p{\rho}^2 = \f1n \sum_\ell \obk{\N_{V\ell}|\C_\ell\sp{\rho}},
  \label{eq:error_SM_mid}
\end{align}
where $\obk{X|Y}=\tr\p{X^\dag Y}$ is a trace inner product.

The result in Eq.~\eqref{eq:error_SM_mid} essentially expresses reconstruction error as a weighted sum of the covariances $\cov_\rho\p{T_{\ell m},T_{\ell m'}}$, where the weights are given by the corresponding matrix elements of the noise matrix $\N_{V\ell}$.
This expression is perhaps the most physically meaningful form of the reconstruction error $\E_V\p{\rho}$ that we will consider in this work, but in practice it turns out that Eq.~\eqref{eq:error_SM_mid} is inconvenient and inefficient to evaluate for any given state $\rho$.
To find a more practical expression of reconstruction error, we use the fact that
\begin{align}
  \bk{T_{\ell m}^\dag}_\rho
  = \obk{\rho|T_{\ell m}^\dag}
  = \tr\p{\rho T_{\ell m}^\dag}
  = \tr\p{T_{\ell m}^\dag \rho}
  = \obk{T_{\ell m}|\rho},
\end{align}
to expand the covariance matrix as
\begin{align}
  \C_\ell\sp{\rho}
  &= \sum_{m,m'} \op{m}{m'} \sp{\obk{\rho|T_{\ell m}^\dag T_{\ell m'}}
    - \obk{\rho|T_{\ell m}^\dag} \obk{\rho|T_{\ell m'}}} \\
  &= \sum_{m,m'} \op{m}{m'} \sp{\obk{T_{\ell m'}^\dag T_{\ell m}|\rho}
    - \obk{T_{\ell m}|\rho} \obk{T_{\ell m'}^\dag|\rho}} \\
  &= \sum_{m,m'} \op{m}{m'}\I \sp{\obk{T_{\ell m'} T_{\ell m}|\rho}
    - \obk{T_{\ell m}|\rho} \obk{T_{\ell m'}|\rho}}
\end{align}
where we define the inversion operator $\I\equiv\sum_m \p{-1}^m\op{-m}{m}$.
We then expand the product $T_{\ell m'} T_{\ell m}$ as
\begin{align}
  \obk{T_{\ell m'} T_{\ell m}|\rho}
  = \sum_L g_{\ell m'm}^L \obk{T_{L,m'+m}|\rho},
  &&
  g_{\ell m'm}^L \equiv \obk{T_{L,m'+m} | T_{\ell m'} T_{\ell m}}
  = f_{\ell m';\ell m}^{L,m'+m},
\end{align}
where the (real) factors $f_{\ell m';\ell m}^{L,m'+m}$ are provided in Appendix \ref{sec:trans_prod}.
Substituting the covariance matrix back into Eq.~\eqref{eq:error_SM_mid} and replacing $\obk{T_{\ell m}|\rho}\to\rho_{\ell m}$, we get
\begin{align}
  \E_V\p{\rho}^2
  = \f1n \sp{\sum_{\ell,m} \p{\chi^V_{\ell m}}^* \rho_{\ell m}
  - \sum_{\ell,m,m'} \bk{m'|\I\N_{V\ell}|m}
  \rho_{\ell m} \rho_{\ell m'}},
  \label{eq:error_comps}
\end{align}
where
\begin{align}
  \chi^V_{LM}
  &\equiv \sum_{\ell,m,m'} \delta_{M,m'+m}
  \bk{m'|\I\N_{V\ell}|m}^* g^L_{\ell m'm} \\
  &= \sum_{\ell,m,m'} \delta_{M,-m'+m}
  \bk{m|\N_{V\ell}|m'} \p{-1}^{m'} g^L_{\ell,-m',m} \\
  &= \sum_\ell \obk{\N_{V\ell}|\D_M|\I g_{L\ell}}
\end{align}
can be written in terms of the matrices
\begin{align}
  g_{L\ell} \equiv \sum_{m,m'} g^L_{\ell m'm} \op{m'}{m},
  &&
  \D_M \equiv \sum_{m,m'} \delta_{M,-m'+m} \op{m'm}.
\end{align}
Here $\D_M$ simply picks off the $M$-th diagonal of the matrix it acts on, such that $\obk{\N_{V\ell}|\D_M|\I g_{L\ell}}$ is an inner product of the $M$-th diagonal of $\I g_{L\ell}$ with the $\p{-M}$-th diagonal of $\N_{V\ell}$.
Defining the $\p{2\ell+1}$-component vectors
\begin{align}
  \ket{\rho_\ell} \equiv \sum_m \rho_{\ell m} \ket{m},
  &&
  \ket{\chi_{V\ell}} \equiv \sum_{\ell,m} \chi^V_{\ell m} \ket{m},
\end{align}
we can write the expansion in Eq.~\eqref{eq:error_comps} in the vectorized form
\begin{align}
  \E_V\p{\rho}^2
  = \f1n \sum_\ell \sp{\bk{\chi_{V\ell}|\rho_\ell}
    - \bk{\rho_\ell|\N_{V\ell}|\rho_\ell}}.
  \label{eq:error_SM}
\end{align}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Comments on a tight reconstruction error bound}
\label{sec:tight}

In principle, maximizing the reconstruction error in Eq.~\eqref{eq:error_SM} over all qudit states $\rho$ would provide a tight upper bound on reconstruction error for any set of axes $V$.
To simplify this task somewhat, we first maximize Eq.~\eqref{eq:error_SM} over all $\rho$ with $\tr\p\rho=1$: this maximum occurs at a ``state'' $\sigma_V^\star$ whose components are defined by
\begin{align}
  \ket{\sigma_{V\ell}^\star} \stackrel{\ell\ne0}{\equiv}
  \f12 \N_{V\ell}^{-1} \ket{\chi_{V\ell}},
  &&
  \ket{\sigma_{V,0}^\star} \equiv \f1{\sqrt{d}} \ket{0}.
\end{align}
The corresponding maximum of $\E_V$ is given by
\begin{align}
  \E_V\p{\sigma_V^\star}^2 = \f1n \sum_{\ell>0}
  \sp{\f14 \bk{\chi_{V\ell}|\N_{V\ell}^{-1}|\chi_{V\ell}}
  - \f1d \tr\p{\N_{V\ell}}},
\end{align}
where the $\tr\p{\N_{V\ell}}$ terms above come from simplifying the $\ell=0$ terms of Eq.~\eqref{eq:error_SM} with $\rho\to\sigma_V^\star$.
While $\E_V\p{\sigma_V^\star}$ is a strict upper bound on $\E_V\p{\rho}$ over all $\rho$ with $\tr\p{\rho}=1$, this bound turns out to be useless in practice, because $\sigma_V^\star$ will generally be a non-physical ``state'' with negative eigenvalues.
To find tight bound on $\E_V\p{\rho}$ over the space of physical qudit states $\rho$, we also need to constrain $\rho$ to have no negative eigenvalues.
Equipped with $\sigma_V^\star$ and $\E_V\p{\sigma_V^\star}$, we can expand
\begin{align}
  \E_V\p{\rho}^2 = \E_V\p{\sigma_V^\star}^2
  - \f1n \norm{\rho-\sigma_V^\star}_V^2,
  &&
  \norm{X}_V^2 \equiv \sum_\ell \bk{X_\ell|\N_{V\ell}|X_\ell},
\end{align}
where $X_\ell\equiv\sum_m\obk{T_{\ell m}|X}\ket{m}$ is a vector of the degree-$\ell$ components of $X$ in the polarization operator basis, and $\norm{X}_V$ is a noise-weighted norm of $X$.
Maximizing $\E_V$ over all qudit states $\rho$ thus amounts to finding the closest physical qudit state $\rho$ to $\sigma_V^\star$, with distance measured by the metric $D_V\p{X,Y}\equiv\norm{X-Y}_V$.
We leave this minimization problem to future work, and note that solving it will likely require making use of the positivity conditions derived in Ref.~\cite{kryszewski2006positivity}.
A loose lower bound on $\norm{\rho-\sigma_V^\star}$ can be found by minimization under the constraint $\norm{\rho}\le1$ (or $\bk{\rho_\ell|\rho_\ell}\le1$), which may provide a tighter upper bound on $\E_V\p{\rho}$ than that in Eq.~\eqref{eq:bound} of the main text.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Polarization operator product expansion}
\label{sec:trans_prod}

The polarization operators on the $d$-dimensional Hilbert space of a spin-$s$ system (with $s\equiv\frac{d-1}{2}$) are defined by
\begin{align}
  T_{\ell m} \equiv \sqrt{\f{2\ell+1}{2s+1}} \sum_{\mu,\nu=-s}^s
  \bk{s\mu;\ell m|s\nu} \op{\nu}{\mu},
\end{align}
where $\bk{s\mu;\ell m|s\nu}$ is a Clebsh-Gordan coefficient that enforces $\ell\in\set{0,1,\cdots,2s}$ and $m\in\set{-\ell,-\ell+1,\cdots,\ell}$.
We wish to compute the coefficients of the operator product expansion
\begin{align}
  T_{\ell_1 m_1} T_{\ell_2 m_2}
  = \sum_{L,M} f_{\ell_1 m_1;\ell_2 m_2}^{LM} T_{LM},
  &&
  f_{\ell_1 m_1;\ell_2 m_2}^{LM}
  \equiv \obk{T_{LM} | T_{\ell_1 m_1} T_{\ell_2 m_2}}.
\end{align}
Using the symmetry properties of Clebsch-Gordan coefficients, namely
\begin{align}
  \bk{\ell_1 m_1; \ell_2 m_2| L M}
  &= \p{-1}^{\ell_2+m_2} \sqrt{\f{2L+1}{2\ell_1+1}}
  \bk{L,-M; \ell_2 m_2| \ell_1,-m_1} \\
  \bk{\ell_1 m_1; \ell_2 m_2| L M}
  &= \p{-1}^{\ell_1+\ell_2-L}
  \bk{\ell_1,-m_1; \ell_2,-m_2| L,-M},
\end{align}
we can find that the polarization operators transform under conjugation as
\begin{align}
  T_{\ell m}^\dag
  = \sqrt{\f{2\ell+1}{2s+1}}
  \sum_{\mu,\nu} \p{-1}^m \bk{s\nu;\ell,-m|s\mu} \op{\mu}{\nu}
  = \p{-1}^m T_{\ell,-m},
\end{align}
which implies that
\begin{align}
  f_{\ell_1 m_1;\ell_2 m_2}^{LM}
  = \p{-1}^M \sqrt{\f{\p{2L+1}\p{2\ell_1+1}\p{2\ell_2+1}}
    {\p{2s+1}\p{2s+1}\p{2s+1}}}
  \sum_{\mu,\nu,\rho} \bk{s\nu;L,-M|s\mu}
  \bk{s\rho;\ell_1m_1|s\nu} \bk{s\mu;\ell_2m_2|s\rho}.
\end{align}
Replacing Clebsch-Gordan coefficients by Wigner 3-$j$ symbols with the identity
\begin{align}
  \bk{\ell_1 m_1; \ell_2 m_2| L M}
  = \p{-1}^{2\ell_2} \p{-1}^{L-M} \sqrt{2L+1}
  \begin{pmatrix}
    L & \ell_2 & \ell_1 \\
    -M & m_2 & m_1
  \end{pmatrix},
\end{align}
we can use the fact that $2\ell_2$ is always even (because $\ell_2$ is always an integer) to expand
\begin{align}
  f_{\ell_1 m_1;\ell_2 m_2}^{LM}
  = \p{-1}^M \sqrt{\p{2L+1}\p{2\ell_1+1}\p{2\ell_2+1}}
  \sum_{\mu,\nu,\rho} \p{-1}^{3s-\mu-\nu-\rho}
  \begin{pmatrix}
    s & L & s \\
    -\mu & -M & \nu
  \end{pmatrix}
  \begin{pmatrix}
    s & \ell_1 & s \\
    -\nu & m_1 & \rho
  \end{pmatrix}
  \begin{pmatrix}
    s & \ell_2 & s \\
    -\rho & m_2 & \mu
  \end{pmatrix}.
\end{align}
This sum can be simplified by the introduction of Wigner 6-$j$ symbols, giving us
\begin{align}
  f_{\ell_1 m_1;\ell_2 m_2}^{LM}
  &= \p{-1}^{2s+M} \sqrt{\p{2L+1}\p{2\ell_1+1}\p{2\ell_2+1}}
  \begin{pmatrix}
    L & \ell_1 & \ell_2 \\
    M & -m_1 & -m_2
  \end{pmatrix}
  \begin{Bmatrix}
    L & \ell_1 & \ell_2 \\
    s & s & s
  \end{Bmatrix} \\
  &= \p{-1}^{2s+L} \sqrt{\p{2\ell_1+1}\p{2\ell_2+1}}
  \bk{\ell_1 m_1; \ell_2 m_2| LM}
  \begin{Bmatrix}
    \ell_1 & \ell_2 & L \\
    s & s & s
  \end{Bmatrix}.
\end{align}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

signal recovery literature:
-- a $\sim2d^2$ signaling theorem \cite{mcewen2011novel}
-- a nice review of these methods \cite{mcewen2011sampling}
-- an empirically accurate (albeit without rigorous proofs/guarantees) $d^2$ sampling method \cite{khalid2014optimaldimensionality}
-- current status of the theory behind sparse signal recovery with random sampling \cite{rauhut2011sparse} (note: existing bounds are not believed to be tight)
-- nice numerical method for sparse signal recovery \cite{alem2012sparse}
-- useful textbook references \cite{freeden2008spherical, freeden2018spherical}