### 2021-03-08 sun_phases/figures/sketches/many_int.tex
adding materials for SU(n) spin model paper

\documentclass[preview,varwidth]{standalone}
\input{preamble.tex}

\begin{document}
\begin{tikzpicture}
  % energy levels
  \coordinate (bl) at (0,0);
  \coordinate (tl) at ($ (bl) + (0,3em) $);
  \coordinate (br) at ($ (bl) + (5em,0) $);
  \coordinate (tr) at ($ (tl) + (5em,0) $);
  \coordinate (bc) at ($ (bl)!.5!(br) $);
  \coordinate (tc) at ($ (tl)!.5!(tr) $);
  \draw[thick] (bl) -- (br);
  \draw[thick] (tl) -- (tr);
  \coordinate (bg) at ($ (bl)!.5!(br) + (0,.2em) $);
  \coordinate (tg) at ($ (tl)!.5!(tr) - (0,.2em) $);
  \draw[<->] (bg) -- (tg) node[pos=.5,right] {$U$};
  % spin sketches
  \node[right, xshift=.3em] at (tr) {\includegraphics[scale=0.7]{many_diff.pdf}};
  \node[right, xshift=.3em] at (br) {\includegraphics[scale=0.7]{many_same.pdf}};
  % state text
  \node[above,align=center] at (tc) {asymmetric};
  \node[below,align=center] at (bc) {permutationally \\[-.3em] symmetric};
\end{tikzpicture}
\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: