### 2021-04-22 collective_sun/sun_perturbation.tex
changed some formatting and revised intro

\documentclass[nofootinbib,notitlepage,11pt]{revtex4-2}
\usepackage[T1]{fontenc}

%%% linking references
\usepackage[dvipsnames]{xcolor}
\usepackage{hyperref}
\hypersetup{
  breaklinks=true,
  colorlinks=true,
  allcolors=BlueViolet,
}

%%% header / footer
\usepackage{fancyhdr} % easier header and footer management
\pagestyle{fancy} % page formatting style
\fancyhf{} % clear all header and footer text
\renewcommand{\headrulewidth}{0pt} % remove horizontal line in header
\usepackage{lastpage} % for referencing last page
\cfoot{\thepage~of \pageref{LastPage}} % "x of y" page labeling

%%% symbols, notations, etc.
\usepackage{physics,braket,bm,amssymb} % physics and math
\renewcommand{\t}{\text} % text in math mode
\newcommand{\f}[2]{\dfrac{#1}{#2}} % shorthand for fractions
\newcommand{\p}[1]{\left(#1\right)} % parenthesis
\renewcommand{\sp}[1]{\left[#1\right]} % square parenthesis
\newcommand{\spp}[1]{\sp{\sp{#1}}} % double square parenthesis
\renewcommand{\set}[1]{\left\{#1\right\}} % curly parenthesis
\newcommand{\bk}{\Braket} % shorthand for braket notation

\renewcommand{\c}{\cdot} % inner product
\newcommand{\m}{\bm} % bold symbol
\renewcommand{\v}{\vec} % arrow vector

\usepackage{dsfont} % for identity operator
\newcommand{\1}{\mathds{1}}

\newcommand{\up}{\uparrow}
\newcommand{\dn}{\downarrow}

\renewcommand{\i}{\mathrm{i}\mkern1mu}
\renewcommand{\d}{\text{d}}
\newcommand{\x}{\text{x}}
\newcommand{\y}{\text{y}}
\newcommand{\z}{\text{z}}

\newcommand{\A}{\mathcal{A}}
\newcommand{\B}{\mathcal{B}}
\newcommand{\C}{\mathcal{C}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\E}{\mathcal{E}}
\newcommand{\G}{\mathcal{G}}
\renewcommand{\H}{\mathcal{H}}
\newcommand{\I}{\mathcal{I}}
\renewcommand{\L}{\mathcal{L}}
\newcommand{\M}{\mathcal{M}}
\newcommand{\N}{\mathcal{N}}
\renewcommand{\O}{\mathcal{O}}
\renewcommand{\P}{\mathcal{P}}
\newcommand{\Q}{\mathcal{Q}}
\newcommand{\R}{\mathcal{R}}
\renewcommand{\S}{\mathcal{S}}
\newcommand{\T}{\mathcal{T}}
\newcommand{\Z}{\mathcal{Z}}

\newcommand{\PP}{\mathbb{P}}
\renewcommand{\SS}{\mathbb{S}}
\newcommand{\TT}{\mathbb{T}}
\newcommand{\ZZ}{\mathbb{Z}}

\newcommand{\PS}{\text{PS}}
\newcommand{\EQPS}{=_{\text{PS}}}

\newcommand{\oper}{\operatorname}

\usepackage{accents} % for the undertilde
\newcommand{\ut}{\undertilde}
\newcommand{\ot}{\widetilde}
\newcommand{\col}{\underline}
\newcommand{\mean}{\overline}

\newcommand{\lra}{\leftrightarrow}
\newcommand{\floor}[1]{\lfloor{#1}\rfloor}

\usepackage[inline]{enumitem} % in-line lists and \setlist{} (below)
\setlist[enumerate,1]{label={(\roman*)}} % default in-line numbering
\setlist{nolistsep} % more compact spacing between environments

%%% figures
\usepackage{graphicx} % for figures
\graphicspath{{./figures/}} % set path for all figures
\usepackage[export]{adjustbox} % for vertical alignment in math
\newcommand{\diagram}[1]
{\,\includegraphics[valign=c]{diagrams/#1.pdf}\,}

% alphanumeric footnotes
\renewcommand*{\thefootnote}{\alph{footnote}}

% for energy diagrams, etc.
\usepackage{tikz}
\usetikzlibrary{arrows,math,calc}
\tikzstyle{every picture}+=[remember picture]
\tikzset{
    thickest/.style={line width=3pt}
}

%%% text markup
\usepackage{color} % text color
\newcommand{\red}[1]{{\color{red} #1}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

\title{SU($n$) ferromagnets near the Heisenberg point}
\author{Michael A. Perlin}
\date{\today}

\maketitle

\tableofcontents

\section{Introduction}

SU($n$) symmetries play an important role in physics.
Underpinning much of high energy physics, the SU($n$) gauge theory known as Yang-Mills theory is central to our understanding of the electroweak and strong forces.
Extensions of Yang-Mills and SU($n$) symmetry feature in the most well-studied examples of holographic duality \cite{maldacena1999largen} and the connection between entanglement and gravity \cite{ryu2006holographic} through the anti-de Sitter/conformal field theory (AdS/CFT) correspondence.
In a condensed matter setting, SU($2$) appears ubiquitously as a symmetry of the Hubbard model, with important consequences for the study of quantum magnetism and high temperature superconductivity \cite{lee2006doping}.
The extension of SU($2$) Hubbard and spin models to SU($n$) has led to predictions of exotic phases of matter such as valence bond solids \cite{read1989valencebond, rokhsar1990quadratic, kaul2012lattice, hermele2011topological} and chiral spin liquids \cite{hermele2009mott, hermele2011topological, chen2016syntheticgaugefield, nataf2016chiral}, as well as the potential to perform universal topological quantum computation \cite{freedman2004class, nayak2008nonabelian} and other phenomena \cite{nataf2014exact, nataf2016exact}.
Furthermore, disordered SU($n$) spin models have opened analytically tractable avenues for studying quantum chaos and information scrambling \cite{sachdev1993gapless, bentsen2019integrable}.

The tremendous theoretical significance of SU($n$) symmetries makes it all the more exciting that they appear naturally in experimental atomic, molecular, and optical (AMO) platforms \cite{wu2003exact, cazalilla2009ultracold, gorshkov2010twoorbital, beverland2016realizing, cazalilla2014ultracold, taie2012su, hofrichter2016direct, cappellini2014direct, scazza2014observation, zhang2014spectroscopic, goban2018emergence, perlin2019effective} with exquisite degrees of microscopic control.
As a simple example, this symmetry arises through the independence of atomic orbital and interaction parameters on the $n$ nuclear spin states of alkaline-earth(-like) atoms, with e.g.~$n=10$ for ${}^{87}$Sr \cite{wu2003exact, cazalilla2009ultracold, cazalilla2014ultracold}.
As a result, AMO experiments can directly probe the role of SU($n$) interactions in controllable settings.
Recent progress includes the experimental observations of bosonization and faster thermalization in multi-component Fermi gas with SU($n$) symmetry \cite{song2020evidence, sonderhouse2020thermodynamics}, SU($n$) Hubbard phases and phase transitions \cite{taie2012su, hofrichter2016direct, taie2020observation}, two-orbital SU($n$) magnetism \cite{cappellini2014direct, scazza2014observation, zhang2014spectroscopic}, and multi-body SU($n$)-symmetric interactions \cite{goban2018emergence, perlin2019effective}.
In the spirit of quantum simulation, further investigations in controlled settings will play an important role in understanding the consequence of SU($n$) symmetries for fundamental questions in physics, as well as their practical use in technological applications.
For example, SU($2$)-symmetric spin interactions can be harnessed to develop quantum sensors that surpass classical limits on measurement precision \cite{he2019engineering, perlin2020spin}.
The prospect of similarly exploiting more general SU($n$) symmetries to achieve a technological advantage is still an unexplored avenue of research with untapped potential.

In this work, we generalize some of the techniques used to harness SU(2)-symmetric spin interactions for metrological advantage to the case of SU($n$).
In doing so, we develop a low-energy effective theory of gapped SU($n$) ferromagnets with weakly broken SU($n$) symmetry (e.g.~due to external fields or additional interaction terms).
The restriction of our theory to ferromagnetic interactions can be relaxed to the assumption of
\begin{enumerate*}
\item an initial state in the manifold of permutationally symmetric states, a multilevel generalization of the Dicke manifold for two-level spins, together with the assumption that
\item this manifold is gapped away from states that break permutational symmetry.
\end{enumerate*}
In addition to working out the effective Hamiltonian induced by weak SU($n$) symmetry breaking through second order in perturbation theory, we develop a diagrammatic method to systematically compute matrix elements of arbitrary spin operators within manifolds of weakly broken permutational symmetry, thereby enabling the numerical simulation of dynamics induced by non-perturbative modifications to the SU($n$) Heisenberg model.
While these simulations generally accumulate uncontrolled errors from the truncation of Hilbert space, we identify criteria under which numerical results correctly capture the qualitative dynamical behavior of a system.

In the following section, \ref{sec:bare_sun}, we introduce the SU($n$) Heisenberg model, finding a convenient way to express its Hamiltonian in terms of permutation (or SWAP) operations that are agnostic to the value of $n$.
In Section \ref{sec:breaking_sun}, we consider modifying the Hamiltonian of the bare Heisenberg model by the addition of spin operators, corresponding to e.g.~external fields or additional interactions, that weakly break SU($n$) symmetry.
We sketch out a derivation of the effective Hamiltonian induced by these terms through second order in perturbation theory, deferring more detailed calculations to appendices.
Finally, we discuss a numerical method to simulate dynamics within manifolds of weak permutational symmetry breaking in Section \ref{sec:shell_model}, before concluding in Section \ref{sec:conclusions}.
Wherever possible, we will fall back to the broadly familiar case of SU(2) in order to facilitate an understanding of the generalizations to SU($n$).

\red{[re-visit / revise the introduction after finishing the remaining text]}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The SU($n$) Heisenberg model}
\label{sec:bare_sun}

Before getting to the full SU($n$) Heisenberg model, we first revisit
the more common case of SU(2).
The SU(2) Heisenberg model consists of
$N$ two-level spins on a lattice, with spin-aligning interactions
\begin{align}
  H_0 = \sum_{i\ne j} h_{ij} \m s_i \c\m s_j,
  \label{eq:H_SU2}
\end{align}
where $i,j$ index individual spins; $h_{ij}$ are scalar coefficients with $h_{ij}=h_{ji}$ and $h_{ii}=0$; and $\m s_i\equiv(s_\x^{(i)},s_\y^{(i)},s_\z^{(i)})$ is a vector of spin operators $s_\alpha^{(i)}\equiv\sigma_\alpha^{(i)}/2$ defined by Pauli operators $\sigma_\alpha^{(i)}$ acting on spin $i$.
In the standard language of spin physics, each term of this Hamiltonian essentially imposes an energy penalty of $-2h_{ij}$ to the formation of a spin 0 singlet state $\ket{\up\dn}-\ket{\dn\up}$ between spins $i$ and $j$, favoring the formation of spin triplet states with total spin 1.
Another interpretation of the Hamiltonian in \eqref{eq:H_SU2} is illuminated by the observation that the two-spin operator $\m s_i\c\m s_j$ is, up to an overall constant, essentially the two-qubit SWAP operation $\Pi_{ij}$.
We will refer to $\Pi_{ij}$ as {\it permutation operator}, because its action on the Hilbert space $\H\equiv\bigotimes_{k=1}^N\H_k$ of $N$ spins with individual Hilbert spaces $\H_k$ is to permute the tensor factors of $\H$ associated with spins $i$ and $j$ (see Figure \ref{fig:sketch_spins}).
Up to identity terms with no physical consequence, we then have
\begin{align}
  H_0 = \sum_{i<j} h_{ij} \Pi_{ij},
  &&
  \Pi_{ij} \equiv \sum_{\mu,\nu} s_{\mu\nu}^{(i)} s_{\nu\mu}^{(j)}
  = \sum_{\mu,\nu} \op{\mu\nu}{\nu\mu}_{ij}
  \label{eq:H_0}
\end{align}
where $s_{\mu\nu}^{(i)}\equiv\op{\mu}{\nu}_i$ is a spin-transition operator for spin $i$, and $\mu,\nu\in\set{\up,\dn}$ index the states of a single spin.
In addition to providing a nice interpretation of the Hamiltonian $H_0$ as essentially swapping the states of spins $i$ and $j$ at a rate $h_{ij}$, the definition of the SU(2)-symmetric Heisenberg model in in \eqref{eq:H_0} is straightforward to generalize to SU($n$) with arbitrary $n$.
That is, rather than having $\mu,\nu\in\set{\up,\dn}$ index two spin states, we simply let $\mu,\nu\in\set{1,2,\cdots,n}$ index the states (sometimes called {\it colors} or {\it flavors}) of an $n$-level spin.

\begin{figure}
  \centering \includegraphics[valign=m]{sketch_spins.pdf} $\stackrel{\displaystyle\Pi_{ij}} {\displaystyle\longleftrightarrow}$ \includegraphics[valign=m]{sketch_spins_swap.pdf}
  \caption{The permutation operator $\Pi_{ij}$ swaps the states of interacting spins $i$ and $j$, which in the case of SU(2) can be represented by magnetization vectors in 3-dimensional space.
    In the case of SU($n$), these vectors live in a higher-dimensional space without a clear geometric interpretation, but the interpretation of the permutation operator $\Pi_{ij}$ remains the same.
  }
\label{fig:sketch_spins}
\end{figure}

If the coefficients $h_{ij}$ of the interaction Hamiltonian $H_0$ are all negative, then the ground-state manifold $\M_0$ of $H_0$ consists of spin-polarized, {\it permutationally symmetric} (PS) states that are simultaneous $+1$ eigenstates of all permutation operators $\Pi_{ij}$.
We will assume that all $h_{ij}<0$ throughout these notes, but note that this restriction can be relaxed to the assumption that the initial state of all spins is a permutationally symmetric state (e.g.~a spin-polarized state).
We will also assume that the PS manifold $\M_0$ is gapped away from all orthogonal states by an interaction energy $\Delta_{\t{gap}}>0$.
Power-law couplings of the form $h_{ij}\sim-1/\abs{i-j}^\alpha$ on a $D$-dimensional lattice, for example, always yield a non-vanishing spectral gap $\Delta_{\t{gap}}$ for long-range interactions with $\alpha\le D$ (see Appendix \ref{sec:spectral_gap}).

In the case of SU(2), the PS manifold $\M_0$ is precisely the {\it Dicke manifold} spanned by the PS states $\ket{m_\up,m_\dn}$ with a definite (integer) number $m_\mu\ge0$ of spins occupying the single-spin state $\mu\in\set{\up,\dn}$.
If there are $N$ spins total, then Dicke states $\ket{m_\up,m_\dn}$ must have $m_\up+m_\dn=N$.
In the case of multilevel spins, the PS manifold $\M_0$ is similarly spanned by states $\ket{m}=\ket{m_1,m_2,\cdots,m_n}$ with a definite occupation number $m_\mu\ge0$ of the single-spin state $\mu$, and $\sum_\mu m_\mu=N$.

These states are essentially labeled by an assignment $m=\p{m_1,m_2,\cdots,m_n}$ of $N$ (identical) spins to $n$ (distinct) states.
We denote the set of all such assignments by $\A_n\p{N}$, such that our standard basis for $\M_0$ is $\set{\ket{m}:m\in\A_n\p{N}}$.
The dimension of $\M_0$, or equivalently the ground-state degeneracy of $H_0$, is straightforward to determine with a standard ``stars and bars'' calculation: $\dim\p{\M_0}=\abs{\A_n\p{N}} = {N+n-1 \choose n-1} \sim N^{n-1}$.
The energy of any PS state $\ket\psi\in\M_0$ with respect to the interaction Hamiltonian $H_0$ in \eqref{eq:H_0} is
\begin{align}
  E_0 = \sum_{i<j} h_{ij} = \f12 \sum_{i\ne j} h_{ij}.
\end{align}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Breaking SU($n$) symmetry}
\label{sec:breaking_sun}

A common task, for example in state preparation protocols or quenched dynamics experiments, is to first initialize in some trivial uniform product state, $\ket\phi^{\otimes N}$, and then to let it evolve under a (possibly time-dependent) Hamiltonian that induces non-trivial dynamics.
The permutational symmetry of states such as $\ket\phi^{\otimes N}\in\M_0$, however, makes them eigenstates of the SU($n$)-symmetric interaction Hamiltonian $H_0$ in \eqref{eq:H_0}.
In order to induce non-trivial dynamics, we therefore need to add new terms to the Hamiltonian that break its SU($n$) symmetry.

As a concrete example, this sort of scenario can be used in the SU(2) case to generate spin squeezed states that harness entanglement to surpass classical limits on measurement precision\cite{he2019engineering, perlin2020spin}.
Starting with a general (gapped) SU(2)-symmetric Hamiltonian of the form in \eqref{eq:H_SU2} we can consider the addition of Ising interactions
\begin{align}
  H_{\t{Ising}} = \sum_{i\ne j} J_{ij} s_\z^{(i)} s_\z^{(j)}.
  \label{eq:Ising}
\end{align}
If these interactions are sufficiently weak for perturbative treatment, as in the case of the XXZ model near the Heisenberg point, then the effective Hamiltonian induced by $H_{\t{Ising}}$ on initially PS states $\ket\psi\in\M_0$ is the one-axis twisting (OAT) Hamiltonian\cite{ma2011quantum}
\begin{align}
  H_{\t{OAT}} = \chi S_\z^2,
  \label{eq:OAT}
\end{align}
where the collective operator $S_\z\equiv\sum_{j=1}^Ns_\z^{(j)}$, and the squeezing strength $\chi$ is equal to the mean Ising coupling strength $J_{ij}$.
If Ising interactions are unavailable, the same effect can be achieved by applying an inhomogeneous magnetic field,
\begin{align}
  H_{\t{B}} = \sum_i B_i s_\z^{(i)},
  \label{eq:magnetic}
\end{align}
which gives rise to an OAT Hamiltonian at second order in perturbation theory, with a squeezing strength $\chi$ that is related to the variance of the magnetic field strength $B_i$.

Similarly to single- and two-body operators such as $H_{\t{B}}$ and $H_{\t{Ising}}$ in \eqref{eq:magnetic} and \eqref{eq:Ising}, it will be necessary for our purposes to additionally consider the introduction of $M$-body operators with $M>2$. Generalizing the above examples to the case of SU($n$) systems, we therefore consider a general $M$-body operator of the form
\begin{align}
  \O\p{\m w,X} \equiv \sum_{k\in\D_N\p{M}} w_k X_k,
  \label{eq:multi_body_op}
\end{align}
where $\m w$ is a dimension-$M$ (i.e.~$M$-index) tensor of scalar coefficients $w_k\equiv w_{k_1k_2\cdots k_M}$; $X$ is an $M$-spin operator, e.g.~$s_\z\otimes s_\z$ in the case of Ising interactions with $M=2$; $k\equiv\p{k_1,k_2,\cdots,k_M}$ is a list of the individual spins $k_i\in\ZZ_N$ that the operator $X_k\equiv X_{k_1k_2\cdots k_M}$ acts on; and
\begin{align}
  \D_N\p{M} \equiv
  \set{ k \in \ZZ_N^M : \t{all entries $k_i$ of $k$ are distinct} },
  \label{eq:off_diags}
\end{align}
is the strictly ``off-diagonal'' part of $\ZZ_N^M$, which is necessary to identify for a consistent definition of $X_k$ as an $M$-body operator.
In this notation, we can write the magnetic field and Ising Hamiltonians in \eqref{eq:magnetic} and \eqref{eq:Ising} as $H_{\t{B}}=\O\p{\m B,s_\z}$ and $H_{\t{Ising}}=\O\p{\m J,s_\z\otimes s_\z}$.

If the addition $\O\p{\m w,X}$ to the SU($n$)-symmetric Hamiltonian $H_0$ in \eqref{eq:H_0} is sufficiently small, namely with operator norm $\norm{\O\p{\m w,X}}$ less than half the spectral gap $\Delta_{\t{gap}}$ of $H_0$, $\norm{\O\p{\m w,X}}<\Delta_{\t{gap}}/2$, then we can treat the effect of $\O\p{\m w,X}$ on the ground-state PS manifold $\M_0$ perturbatively.
The effective Hamiltonians $H_{\t{eff}}^{(1)}$ and $H_{\t{eff}}^{(2)}$ induced by $\O\p{\m w,X}$ on $\M_0$ at first and second order in perturbation theory are\cite{bravyi2011schrieffer}
\begin{align}
  H_{\t{eff}}^{(1)} = \P_0 \O\p{\m w,X} \P_0,
  &&
  H_{\t{eff}}^{(2)} = - \sum_{\Delta\ne0}
  \f1\Delta \P_0 \O\p{\m w,X} \P_\Delta \O\p{\m w,X} \P_0,
  \label{eq:H_eff}
\end{align}
where $\P_\Delta$ is a projector onto the eigenspace $\E_\Delta$ of the interaction Hamiltonian $H_0$ with interaction energy $\Delta$ above that of PS manifold $\M_0$.
In particular, $\P_0$ is a projector onto $\M_0$ itself, and $\E_0=\M_0$.
The first order effective Hamiltonian $H_{\t{eff}}^{(1)}$ simply projects $\O\p{\m w,X}$ onto the PS manifold $\M_0$, and takes the form
\begin{align}
  H_{\t{eff}}^{(1)} \EQPS \mean w \, \col X,
  \label{eq:H_eff_1}
\end{align}
where $\EQPS$ denotes equality under a restriction to $\M_0$; the coefficient $\mean w$ is the average of all $w_k$; and $\col X$ is a collective version of $X$:
\begin{align}
  \mean w \equiv \f1{\abs{\D_N\p{M}}}
  \sum_{k\in\D_N\p{M}} w_k,
  &&
  \col X \equiv \sum_{k\in\D_N\p{M}} X_k,
  \label{eq:mean_col}
\end{align}
with $\abs{\D_N\p{M}}=\prod_{j=0}^{M-1}\p{N-j}$.
In the case of a magnetic field $s_\z$ or Ising interactions $s_\z\otimes s_\z$, for example,
\begin{align}
  \col{s_\z} = \sum_i s_\z^{(i)} = S_\z,
  &&
  \col{s_\z\otimes s_\z}
  = \sum_{i\ne j} s_\z^{(i)} s_\z^{(j)}
  = S_\z^2 - \f14 N.
\end{align}
The second order effective Hamiltonian $H_{\t{eff}}^{(2)}$ in \eqref{eq:H_eff} takes more work to simplify due to the presence of a projector $\P_\Delta$ onto the manifold $\E_\Delta$ of states with excitation energy $\Delta$.
This projector essentially picks off the part of $\O\p{\m w,X}$ that is strictly off-diagonal with respect to the ground- and excited-state manifolds $\E_0$ and $\E_\Delta$.
We therefore need to decompose $\O\p{\m w,X}$ into components that generate states of definite excitation energy when acting on PS states $\ket\psi\in\M_0$.
The SU($n$) symmetry of $H_0$ enables such a decomposition to take the form
\begin{align}
  H_0 \O\p{\m w,X} \ket\psi
  = \sum_\Delta \p{E_0+\Delta} \O\p{\m w_\Delta,X} \ket\psi,
  \label{eq:excitation}
\end{align}
where, thinking of the tensor $\m w$ as a $\abs{\D_N\p{M}}$-component vector, the tensor $\m w_\Delta$ can be found by diagonalizing a matrix $\hat{\m h}$ of dimensions $\abs{\D_N\p{M}}\times\abs{\D_N\p{M}}\sim N^M\times N^M$, and projecting $\m w$ onto the eigenspace of $\hat{\m h}$ with eigenvalue $\Delta$ (see Appendix \ref{sec:eigenstates}).
The matrix $\hat{\m h}$ is determined entirely by the couplings couplings $h_{ij}$ of $H_0$.
While the computational cost of diagonalizing $\hat{\m h}$ is nominally $O\p{N^{3M}}$, spatial symmetries of $H_0$ and $\O\p{\m w,X}$ such as translational invariance or isotropy can drastically reduce the complexity of this eigenvalue problem, as we discuss in Appendix \ref{sec:symmetries}.

Equipped with the decomposition of $\O\p{\m w,X}$ into terms $\O\p{\m w_\Delta,X}$ that generate states of definite excitation energy $\Delta$, we can expand
\begin{align}
  H_{\t{eff}}^{(2)} = -\sum_{\Delta\ne0} \f1\Delta
  \P_0 \O\p{\m w_\Delta,X}^2 \P_0.
\end{align}
Simplifying the product $\P_0 \O\p{\m w_\Delta,X}^2 \P_0$ generally yields $r$-body operators for all positive integers $r\le2M$.
We provide a general prescription for simplifying such products in Appendix \ref{sec:operator_product}, and explicitly work out the single-body ($M=1$) and two-body ($M=2$) cases in Appendices \ref{sec:single_pair_prod} and \ref{sec:two_pair_prod}.
If $X$ is a single-body operator, then
\begin{align}
  H_{\t{eff}}^{(2)}
  = \sum_{\Delta\ne0} \f{\m w_\Delta\c\m w_\Delta}{\Delta} \,
  \mean{X\otimes X}
  = \sum_{\Delta\ne0} \f{\m w_\Delta\c\m w_\Delta}{N\p{N-1}\Delta}
  \p{\col{X}^2 - N \col{X^2}},
  \label{eq:H_eff_2_1}
\end{align}
where $\mean{X\otimes X}$ is the average of $\p{X\otimes X}_{ij}=X_iX_j$ over all $i\ne j$, similarly to the definition of $\mean{w}$ in Eq.~\eqref{eq:mean_col}.
In the SU(2) case, this result implies that an inhomogeneous magnetic field ($X=s_\z$) yields the OAT Hamiltonian $H_{\t{OAT}}$ in \eqref{eq:OAT} at second order perturbation theory.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Beyond ground-state perturbation theory}
\label{sec:shell_model}

In order to find the effective Hamiltonian $H_{\t{eff}}^{(2)}$ induced by the $M$-body operator $\O\p{\m w,X}$ on the PS manifold $\M_0$ at second order in perturbation theory, we constructed states with definite excitation energy $\Delta>0$ with respect to $H_0$.
Rather than treating these excited states as ``intermediate'' or ``virtual'' for the derivation of an effective ground-state Hamiltonian, we can perform numerical simulations in the joint Hilbert space of both ground and excited states.
When the perturbation $\O\p{\m w,X}$ is small, namely with operator norm $\norm{\O\p{\m w,X}}<\Delta_{\t{gap}}/2$, such simulations are equivalent (up to perturbative corrections) to a higher-order perturbation theory.
Moreover, the first-class treatment of excited states in finite systems enables simulating the dynamics induced by {\it non-perturbative} additions $\O\p{\m w,X}$ to the base Hamiltonian $H_0$.

The basic idea will be to identify the set of states (manifolds) $\M_r$ with a definite number ($r$) of spin-wave excitations, or {\it magnons}.
The ground-state manifold of PS states is simply the zero-magnon manifold $\M_0$.
If we construct a basis for all $\M_r$ with $r\le r_{\t{max}}$, we can perform numerical simulations in the combined manifold of all states with at most $r_{\t{max}}$ magnons; we refer to such simulations as {\it order-}$r_{\t{max}}$ {\it truncated shell} (TS$_{r_{\t{max}}}$) simulations.
An $M$-spin operator $\O\p{\m w,X}$ can only couple $r$-magnon states (in $\M_r$) to $q$-magnon states (in $\M_q$) with $\abs{q-r}\le M$.
For initially PS states, TS$_{r_{\t{max}}}$ simulations thus capture all virtual processes through order $2\floor{r_{\t{max}}/M}+1$ in perturbation theory.
When $\norm{\O\p{\m w,X}}\ge\Delta_{\t{gap}}/2$, however, perturbation theory formally breaks down, and the error incurred by neglecting all $s$-magnon states with $s>r_{\t{max}}$ is not clear a priori.
Nonetheless, TS$_{r_{\t{max}}}$ simulations can be benchmarked by examining the populations of all $r$-magnon manifolds $\M_r$ with $r\le r_{\t{max}}$.
As long as the maximal populations within large-$r$ manifolds $\M_r$ are small, one should expect TS$_{r_{\t{max}}}$ simulations to faithfully capture (at least qualitatively) the dynamical behavior of the spin system under consideration.

The remainder of this section is devoted to the construction of a basis for the $r$-magnon manifolds $\M_r$, and to the expansion of general spin operators $\O\p{\m w,X}$ in this basis.
We begin with a classification of excited states in the case of SU(2).
A common starting point is to first classify states by a definite (non-negative) spin length $S=\set{N/2,N/2-1,\cdots}$, and a spin projection $m_\z\in\set{S,S-1,\cdots,-S}$ onto a fixed quantization axis.
In fact, the manifold of states with definite fixed spin length $S$ is precisely the $\p{N/2-S}$-magnon manifold $\M_{N/2-S}$.
The PS manifold of Dicke states with $S=N/2$ is then $\M_0$, and more generally the manifold of $r$-magnon states with $S=N/2-r$ is $\M_r$.
Rather than the spin projection $m_\z$, we then identify states by the occupation numbers $m=\p{m_\up,m_\dn}\in\A_2\p{N}$, where $m_\mu$ denotes the total number of spins in the state $\mu$, with $m_\up+m_\dn=N$ and $m_\up-m_\dn=m_\z$.

The label $m$ suffices to identify the PS (Dicke) manifold $\M_0=\oper{span}\set{\ket{m}:m\in\A_2\p{N}}$, but we need additional labels to uniquely identify states in the $r$-magnon manifold $\M_r$ with $r>0$.
Single-magnon ($r=1$) spin-wave states of SU(2) spin systems can be written in the form
\begin{align}
  \ket{m,\m v,\sigma_z}
  \equiv \O\p{\m v,\sigma_\z} \ket{m}
  = \sum_{j=1}^N v_j \sigma_\z^{(j)} \ket{m},
\end{align}
where each $v_j$ is a scalar and $\sigma_\z$ is a Pauli-$z$ operator.
If an SU(2) spin system is translationally invariant, then spin-wave eigenstates have $v_j=e^{\i k_{\m v}\c j}$ with wavenumber $k_{\m v}\ne0$.
Similarly, $r$-magnon states for SU(2) take the form $\ket{m,\m w,\sigma_\z^{\otimes r}} \equiv \O\p{\m w,\sigma_\z^{\otimes r}}\ket{m}$, where now $\m w$ is an $r$-index tensor that encodes the structure of correlations in the excited state $\ket{m,\m w,\sigma_\z^{\otimes r}}$.
The coefficient tensor $\m w$ determines the energy of the excited state $\ket{m,\m w,\sigma_\z^{\otimes r}}$ through the expansion in \eqref{eq:excitation}.
Considering states of this form, built using tensor products of Pauli-$z$ ($\sigma_\z$) operators, suffices to construct a basis for all excited states of an SU(2) spin system (see Appendix \ref{sec:operator_content}).
Note that $r$-body operators with arbitrary coefficients $\m w$ will generally couple the PS manifold $\M_0$ to all $q$-magnon manifolds $\M_q$ with $q\le r$ (see Figure \ref{fig:manifold_sketch}).
In order to generalize the notion of $r$-magnon manifolds $\M_r$ to the case of SU($n$) with $n>2$, where the analogous notion of a spin length $S$ is less intuitive, we can define $\M_r$ for $r>0$ as the manifold of states that (i) can be generated by the action of $r$-body operators on $\M_0$, and (ii) are orthogonal to all states in $\bigoplus_{q<r}\M_q$.

\begin{figure}
  \centering
  \begin{tikzpicture}
    \draw[thickest] (0,0) -- ++(10em,0);
    \foreach \hh in {0.15em,0.05em,-0.05em,-0.15em} {
      \draw (0,3em+\hh) -- ++(10em,0);
    }
    \draw[->] (7em,.25em) arc (220:0:.4em);
    \draw[->] (7em,.25em) to [bend left]
    node[left] {$\O\p{\m v,\sigma_\z}$} (7em,2.7em);
    \node () at (5em,4.7em) {$\vdots$};
    \foreach \hh in {0.25em,0.15em,0.05em,-0.05em,-0.15em,-0.25em} {
      \draw (0,6em+\hh) -- ++(10em,0);
    }
    \draw[->] (-.4em,0) arc (220:0:.4em);
    \draw[->] (-.4em,0) to [bend left, in=130] (-.2em,3em);
    \draw[->] (-.4em,0) to [bend left, in=130] (-.2em,4.3em);
    \draw[->] (-.4em,0) to [bend left, in=130]
    node[left] {$\O\p{\m w,\sigma_\z^{\otimes r}}$} (-.2em,6em);
    \node[right] () at (10em,0) {$\M_0,S=\frac{N}{2}$};
    \node[right] () at (10em,3em) {$\M_1,S=\frac{N}{2}-1$};
    \node[right] () at (10em,6em) {$\M_r,S=\frac{N}{2}-r$};
  \end{tikzpicture}
  \caption{Schematic of relationships between $q$-magnon manifolds $\M_q$ with total spin $S=N/2-q$ in an SU(2) spin system.
    Single-body operators $\O\p{\m v,\sigma_\z}$ generally couple the ground-state manifold $\M_0$ to the single-magnon manifold $\M_1$ as well as $\M_0$ itself, while $r$-body operators $\O\p{\m w,\sigma_\z^{\otimes r}}$ generally couple $\M_0$ to all $q$-magnon manifolds $\M_q$ with $q\le r$.}
  \label{fig:manifold_sketch}
\end{figure}

In the case of an SU($n$) spin system, we can similarly start with the ground-state PS manifold $\M_0=\oper{span}\set{\ket{m}:m\in\A_n\p{N}}$, and generate excited states by the action of multi-body operators, $\ket{m,\m w,X}\equiv\O\p{\m w,X}\ket{m}$.
The choice of tensors $\m w$ for generating states in the $r$-magnon manifold $\M_r$ is the same as before, but we now have the additional complication of having to choose $r$-spin operators, $X$.
As with the choice of $\sigma_\z^{\otimes r}$ for SU(2), it suffices to choose tensor products of {\it diagonal} single-spin operators (that is, diagonal in a fixed single-particle basis), but there will now be $\sim\p{n-1}^r$ choices: one of $n-1$ diagonal operators (that are linearly independent of the identity) for each of $r$ tensor factors.
We discuss the exact choice of operators in more detail in Appendix \ref{sec:operator_content}.
To summarize, we define $n$ single-spin operators $Z_\nu\equiv1-2\op\nu=\sum_{\mu\ne\nu}\op\mu-\op\nu$, for which $Z_\nu=\pm\sigma_\z$ in the case of SU(2), and we choose a list of non-negative integers $z\equiv\p{z_1,z_2,\cdots,z_n}$ with $\sum_\nu z_\nu=r$ to define the $r$-spin operator
\begin{align}
  X_z \equiv \bigotimes_{\nu=1}^n Z_\nu^{\otimes z_\nu}
  = \underbrace{Z_1 \otimes \cdots \otimes Z_1}_{z_1~\t{times}}
  \otimes
  \underbrace{Z_2 \otimes \cdots \otimes Z_2}_{z_2~\t{times}}
  \otimes \cdots \otimes
  \underbrace{Z_n \otimes \cdots \otimes Z_n}_{z_n~\t{times}}.
\end{align}
Our starting basis for $\M_r$ then consists of states $\ket{m,\m w,X_z}$, where we enforce that $z_\nu\le m_\nu$ for all $\nu$ (where $m_\nu$ is the occupation number of the single-spin state $\nu$), and that $z$ is zero in an entry where $m$ is maximal: $z_\nu=0$ for $\nu=\min\set{\mu:m_\mu=\max\p{m}}$.
This second restriction is equivalent to removing an operator from $\set{Z_\nu:\nu\in\ZZ_n}$, making the remaining set linearly independent of the identity operator.
The basis $\set{\ket{m,\m w,X_z}}$ for $\M_r$ is generally complete, but not orthonormal, so as a final step we use the states $\ket{m,\m w,X_z}$ to (numerically) construct an orthonormal basis $\B_r$ for $\M_r$ (see Appendix \ref{sec:operator_content}).

By repeating the above process incrementally for all $r\le r_{\t{max}}$ (starting with $r=1$), we can construct an orthonormal basis $\col\B_{r_{\t{max}}}\equiv\bigoplus_{r\le r_{\t{max}}}\B_r$ for the combined manifold of all states with at most $r_{\t{max}}$ magnons.
During the construction of $\col\B_s$, an appropriate choice of tensors $\m w$ ensures that all states $\ket{m,\m w,X}\in\col\B_s$ are eigenstates of the base Hamiltonian $H_0$ (see Appendix \ref{sec:eigenstates}), which is to say that $H_0$ is diagonal with respect to $\col\B_s$.
Matrix elements of general spin operators $\O\p{\m v,Y}$, meanwhile, take the form
\begin{align}
  \bk{\ell,\m u,X|\O\p{\m v,Y}|m,\m w,Z}
  = \bk{\ell|\O\p{\m u,X}\O\p{\m v,Y}\O\p{\m w,Z}|m}.
  \label{eq:matrix_element}
\end{align}
In order to compute these matrix elements, we simplify general products of multi-body operators in Appendix \ref{sec:operator_product}.

% dispersion relation? magnon interactions?
% lax formulation for single-particle terms
% solve collective model? Casmir operators?

\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}
\label{sec:conclusions}

\red{[TODO: write section]}

\begin{acknowledgments}
  We thank Asier Pi\~neiro Orioli for helpful discussions and technical contributions to this work.
\end{acknowledgments}

\bibliography{\jobname.bib}

\newpage
\appendix

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Existence of a spectral gap with long-range power-law
  interactions}
\label{sec:spectral_gap}

Here we show that in $D$ spatial dimensions, power-law SU($n$)-symmetric interactions of the form in \eqref{eq:H_0} with $h_{ij}=-1/\abs{i-j}^\alpha$ yield a non-vanishing spectral gap when $\alpha\le D$.
For definiteness, we will consider a periodic hypercubic lattice of $N=L^D$ spins.
The spectral gap of the interaction Hamiltonian is then determined by well-known spin-wave theory (see Ref.~[\citenum{ashcroft1978solid}] or Appendix \ref{sec:single_body_eigenstates})
\begin{align}
  \Delta
  = \sum_{d\in\ZZ_L^D} h_{0,d} \sp{\cos\p{d \c k_{\t{SE}}}-1}
  = \sum_{\substack{d\in\ZZ_L^D\\\abs{d}\ge1}}
  \f{1-\cos\p{d \c k_{\t{SE}}}}{\abs{d}^\alpha},
\end{align}
where $\ZZ_L$ denotes the set of integers modulo $L$; $k_{\t{SE}}\in\ZZ_L^D\times2\pi/L$ is a wavenumber for a single-magnon spin-wave state; and the norm $\abs{d}$ for a vector $d$ on a periodic lattice is implicitly understood to mean the smallest Euclidean distance of $d$ from a fixed origin.
In order to yield the smallest excitation energy $\Delta$, the wavenumber $k_{\t{SE}}$ should maximize the contribution of the cosine term above, which is achieved by a wavenumber that minimizes the oscillations of this term when integrated over the entire lattice.
A suitable candidate for a minimal wavenumber is $k_{\t{SE}}=\p{2\pi/L,0,0,\cdots}$, which corresponds to an excitation energy
\begin{align}
  \Delta = \sum_{\substack{d\in\ZZ_L^D\\\abs{d}\ge1}}
  \f{1-\cos\p{d_12\pi/L}}{\abs{d}^\alpha}.
\end{align}
Defining $\epsilon\equiv2/L$ and a rescaled domain symmetrized about $0$, $\SS_\epsilon\simeq\ZZ_L/\epsilon$ with $\SS_\epsilon\subset\sp{-1,1}$, we substitute $x\simeq\epsilon d$ to get
\begin{align}
  \Delta
  = \sum_{\substack{x\in\SS_\epsilon^D\\\abs{x}\ge\epsilon}}
  \f{1-\cos\p{\pi x_1}}{\abs{x/\epsilon}^\alpha}
  = \epsilon^{\alpha-D} \sum_{\substack{x\in\SS_L^D\\\abs{x}\ge\epsilon}}
  \epsilon^D \f{1-\cos\p{\pi x_1}}{\abs{x}^\alpha}.
\end{align}
As $\epsilon\to0$ the discrete sum over $x$ is well approximated by an integral that avoids an infinitesimal region about the origin, i.e.
\begin{align}
  \Delta = \epsilon^{\alpha-D} \I_D\p{\epsilon},
  &&
  \I_D\p{\epsilon}
  \equiv \int_{\TT_1^D\setminus\TT_\epsilon^D} \d^Dx\,
  \f{1-\cos\p{\pi x_1}}{\abs{x}^\alpha},
\end{align}
where the integral $\I_D\p{\epsilon}$ is defined using the symmetric interval $\TT_a\equiv\p{-a/2,a/2}$.
The integrand of $\I_D\p{\epsilon}$ is strictly positive and well-behaved on the entirety of its domain except for the origin, where depending on the value of $\alpha$ the integrand may vanish or diverge as $\abs{x}\to0$.
Together, these facts mean that
\begin{align}
  \I_D\p{\epsilon} \stackrel{\epsilon\to0}{\sim} \epsilon^{-\gamma},
  &&
  \Delta \stackrel{\epsilon\to0}{\sim} \epsilon^{\alpha-D-\gamma},
\end{align}
for some $\gamma\ge0$, which implies that gap $\Delta$ is non-vanishing when $\alpha\le D\le D+\gamma$.

For the sake of completion, we add that in fact the gap $\Delta$ always vanishes in the thermodynamic limit when $\alpha>D$.
To see this behavior, we note that the asymptotic dependence of the integral $\I_D\p{\epsilon}$ on $\epsilon$ is determined by the behavior of its integrand when $\abs{x}\sim\epsilon$, in which case $1-\cos\p{\pi x_1}\sim x_1^2$, so
\begin{align}
  \I_D\p{\epsilon}
  \sim \int_{\TT_1^D\setminus\TT_\epsilon^D} \d^Dx\,
  \f{x_1^2}{\abs{x}^\alpha}.
\end{align}
We can then use the fact that $x_1^2\le\abs{x}^2$ and change to spherical coordinates to find that
\begin{align}
  \I_D\p{\epsilon} \lesssim
  \int_{\TT_1^D\setminus\TT_\epsilon^D} \d^Dx\,
  \f{\abs{x}^2}{\abs{x}^\alpha}
  \sim \int_\epsilon^1 \d x\, x^{D+1-\alpha}
  \sim
  \begin{cases}
    \epsilon^0 & \alpha < D+2 \\
    \log\p{1/\epsilon} & \alpha = D+2 \\
    \epsilon^{D+2-\alpha} & \alpha > D+2
  \end{cases}.
\end{align}
It follows that the spectral gap
\begin{align}
  \Delta \stackrel{\epsilon\to0}{\lesssim}
  \begin{cases}
    \epsilon^{\alpha-D} & \alpha < D+2 \\
    \epsilon^2 \log\p{1/\epsilon} & \alpha = D + 2 \\
    \epsilon^2 & \alpha > D+2
  \end{cases},
\end{align}
which vanishes as $\epsilon\to0$ for all $\alpha>D$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Generating excitation energy eigenstates}
\label{sec:eigenstates}

Both perturbation theory and beyond-perturbative simulations of the dynamics induced by an operator $\O\p{\m w,X}$ require us to decompose $\O\p{\m w,X}$ into a sum of terms that generate eigenstates of the SU($n$)-symmetric interaction Hamiltonian $H_0$ in \eqref{eq:H_0} when acting on permutationally symmetric (PS) states $\ket\psi\in\M_0$.
To perform this decomposition, in this section we simplify products of the form $H_0\O\p{\m w,X}\ket\psi$ with $\ket\psi\in\M_0$.
For reference, the $M$-body operator $\O\p{\m w,X}$ is defined by
\begin{align}
  \O\p{\m w,X} \equiv \sum_{k\in\D_N\p{M}} w_k X_k,
\end{align}
where $\m w$ is a dimension-$M$ (i.e.~$M$-index) tensor of scalar coefficients $w_k\equiv w_{k_1k_2\cdots k_M}$; $X$ is an $M$-spin operator; $k\equiv\p{k_1,k_2,\cdots,k_M}$ is a list of the individual spins $k_i\in\ZZ_N$ that the operator $X_k\equiv X_{k_1k_2\cdots k_M}$ acts on; and
\begin{align}
  \D_N\p{M} \equiv
  \set{ k \in \ZZ_N^M: \t{all entries $k_i$ of $k$ are distinct} },
\end{align}
is the strictly ``off-diagonal'' part of $\ZZ_N^M$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Single-body excitations}
\label{sec:single_body_eigenstates}

To build familiarity with the problem at hand, we first consider the simple case of a single-body ($M=1$) case and simplify
\begin{align}
  H_0 \O\p{\m w,X} \ket\psi
  = \f12 \sum_{i\ne j} \sum_k h_{ij} w_k \Pi_{ij} X_k \ket\psi.
  \label{eq:single_body_diagnosis_start}
\end{align}
The sum in \eqref{eq:single_body_diagnosis_start} has terms with $k\in\set{i,j}$, and terms with $k\notin\set{i,j}$.
In the case of $k\notin\set{i,j}$, the permutation operator $\Pi_{ij}$ commutes with $X_k$ and annihilates on $\ket\psi$, and we can replace the sum
\begin{align}
  \sum_{k\notin\set{i,j}} \to \sum_k - \sum_{k\in\set{i,j}},
\end{align}
allowing us to simplify
\begin{align}
  \f12 \sum_{i\ne j} \sum_{k\notin\set{i,j}}
  h_{ij} w_k \Pi_{ij} X_k \ket\psi
  = E_0 \O\p{\m w,X} \ket\psi
  - \f12 \sum_{i\ne j} \sum_{k\in\set{i,j}} h_{ij} w_k X_k \ket\psi,
\end{align}
where $E_0=\frac12\sum_{i\ne j}h_{ij}$ is the interaction energy of PS states $\ket\psi\in\M_0$.
Making the replacement
\begin{align}
  \sum_{i\ne j} \sum_{k\in\set{i,j}}
  \to \sum_k \sum_{\substack{i\ne j\\\set{i,j}\ni k}},
\end{align}
we can simplify
\begin{align}
  \f12 \sum_{\substack{i\ne j\\\set{i,j}\ni k}} h_{ij}
  = \f12 \sum_i h_{ik} + \f12 \sum_j h_{kj}
  = h_k,
  &&
  h_k \equiv \sum_i h_{ik},
\end{align}
which implies that the terms in \eqref{eq:single_body_diagnosis_start} with $k\notin\set{i,j}$ are
\begin{align}
  \f12 \sum_{i\ne j} \sum_{k\notin\set{i,j}}
  h_{ij} w_k \Pi_{ij} X_k \ket\psi
  = E_0 \O\p{\m w,X} \ket\psi - \sum_k h_k w_k X_k \ket\psi.
\end{align}
The terms in \eqref{eq:single_body_diagnosis_start} with $k\in\set{i,j}$, meanwhile, are
\begin{align}
  \f12 \sum_{\substack{i\ne j\\k\in\set{i,j}}}
  h_{ij} w_k \Pi_{ij} X_k \ket\psi
  = \sum_{j,k} h_{kj} w_k X_j \ket\psi.
\end{align}
so in total
\begin{align}
  H_0 \O\p{\m w,X} \ket\psi
  = E_0 \O\p{\m w,X} \ket\psi
  + \sum_k \sp{\sum_j h_{kj} w_j - h_k w_k} X_k \ket\psi.
  \label{eq:single_body_diagnosis_end}
\end{align}
The action of the single-body perturbation $\O\p{\m w,X}$ on a permutationally symmetric state therefore generates an eigenstate of $H_0$ with energy $E_0+\Delta$ if the vector $\m w=\sum_k w_k \ket{k}$ satisfies the eigenvalue equation
\begin{align}
  \p{\m h - \oper{diag}\v h} \c \m w = \Delta \m w,
  \label{eq:single_body_eig}
\end{align}
where $\m h\equiv\sum_{i,j} h_{ij} \op{i}{j}$ is a matrix of all couplings $h_{ij}$; the vector $\v h \equiv \sum_{i,j} h_{ij} \ket{i} = \sum_i h_i \ket{i}$ is the sum of all columns of $\m h$; and $\oper{diag}\v h$ is a matrix with $\v h$ on the diagonal and zeroes everywhere else.
When the interaction Hamiltonian $H_0$ is translationally invariant, the single-body eigenvalue problem in \eqref{eq:single_body_eig} is solvable analytically.
In this case, the couplings $h_{ij}$ depend only on the separation $\abs{i-j}$, so eigenvectors of $\m h$ are plane waves of the form
\begin{align}
  \m w_k \equiv \sum_{d\in\ZZ_L^D} e^{\i d\c k} \ket{d},
\end{align}
where on a $D$-dimensional periodic lattice of $N=L^D$ spins, lattice sites are indexed by vectors $d\in\ZZ_L^D$, and wavenumbers take on values $k\in\ZZ_L^D\times2\pi/L$.
The corresponding eigenvalues of $\m h$ can be determined by expanding
\begin{align}
  \m h \c \m w_k
  = \sum_{c,d\in\ZZ_L^D} h_{cd} e^{\i d\c k} \ket{c}
  = \sum_{c,d\in\ZZ_L^D} h_{c,c+d} e^{\i\p{c+d}\c k} \ket{c}
  = \sum_{d\in\ZZ_L^D} h_{0,d} \cos\p{d\c k} \m w_k,
\end{align}
where the imaginary contributions vanish in the sum over $d$ because $h_{0,d}=h_{0,-d}$.
The remainder of \eqref{eq:single_body_eig} that we need to sort out is $\oper{diag}\v h$, where all $h_i = \sum_{i,j}h_{ij} = \sum_d h_{0,d}$ are equal, which implies that $\oper{diag}\v h = \sum_d h_{0,d}$ is a scalar.
We thus find that
\begin{align}
  \p{\m h - \oper{diag}\v h} \c \m w_k = \Delta_k \m w_k,
  &&
  \Delta_k \equiv \sum_{d\in\ZZ_L^D} h_{0,d} \sp{\cos\p{d\c k}-1}.
\end{align}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Multi-body excitations}

We now consider the full $M$-body case
\begin{align}
  H_0 \O\p{\m w,X} \ket\psi
  = \sum_{\substack{k\in\D_N\p{M}\\\set{i,j}\in\C_N\p{2}}}
  h_{ij} w_k \Pi_{ij} X_k \ket\psi,
  \label{eq:diagnosis_start}
\end{align}
where for convenience we define
\begin{align}
  \C_N\p{q} \equiv \set{ s \subset \ZZ_N : \abs{s} = q }
  \label{eq:choices}
\end{align}
as the set of all subsets (``choices'') of $q$ elements from $\ZZ_N$.
The sum in \eqref{eq:diagnosis_start} has terms with $i,j\notin k$, terms with $i,j\in k$, and terms with only one of $i$ or $j\in k$.
The permutation operator $\Pi_{ij}$ acts trivially on $X_k\ket\psi$ when $i,j\notin k$, and we can replace the sum
\begin{align}
  \sum_{\substack{\set{i,j}\in\C_N\p{2}\\i,j\notin k}}
  \to \sum_{\set{i,j}\in\C_N\p{2}}
  - \sum_{\substack{\set{i,j}\in\C_N\p{2}\\i\in k~\t{or}~j\in k}},
\end{align}
allowing us to simplify
\begin{align}
  \sum_{k\in\D_N\p{M}} \sum_{\substack{\set{i,j}\in\C_N\p{2}\\i,j\notin k}}
  h_{ij} w_k \Pi_{ij} X_k \ket\psi
  = E_0 \O\p{\m w,X} \ket\psi
  - \sum_{k\in\D_N\p{M}} h^{\t{D}}_k w_k X_k \ket\psi,
\end{align}
where $E_0 = \sum_{\set{i,j}\in\C_N\p{2}} h_{ij}$ is the interaction energy of PS states $\ket\psi\in\M_0$, and
\begin{align}
  h^{\t{D}}_k \equiv
  \sum_{\substack{\set{i,j}\in\C_N\p{2}\\i\in k~\t{or}~j\in k}} h_{ij}
  = \sum_{i\in k} h_i - \sum_{\set{a,b}\in\C_M\p{2}} h_{k_ak_b},
  &&
  h_i \equiv \sum_j h_{ij},
  \label{eq:multi_body_op_diag_element}
\end{align}
with $k_a,k_b\in k\equiv\p{k_1,k_2,\cdots,k_M}$.
Defining a linear operator $\hat{\m h}_{\t{D}}$ on the space of dimension-$M$ tensors, whose action on $\m w$ is a new dimension-$M$ tensor with entries
\begin{align}
  \p{\hat{\m h}_{\t{D}}\c\m w}_k \equiv h^{\t{D}}_k w_k,
  \label{eq:multi_body_op_diag}
\end{align}
we can write the terms in \eqref{eq:diagnosis_start} with $i,j\notin k$ as
\begin{align}
  \sum_{k\in\D_N\p{M}} \sum_{\substack{\set{i,j}\in\C_N\p{2}\\i,j\notin k}}
  h_{ij} w_k \Pi_{ij} X_k \ket\psi
  = E_0 \O\p{\m w,X} \ket\psi
  - \sum_{k\in\D_N\p{M}} \p{\hat{\m h}_{\t{D}}\c\m w}_k X_k \ket\psi.
\end{align}
The terms in \eqref{eq:diagnosis_start} with $i,j\in k$ are
\begin{align}
  \sum_{k\in\D_N\p{M}} \sum_{\substack{\set{i,j}\in\C_N\p{2}\\i,j\in k}}
  h_{ij} w_k \Pi_{ij} X_k \ket\psi
  = \sum_{k\in\D_N\p{M}} \sum_{\substack{\set{a,b}\in\C_M\p{2}}}
  h_{k_ak_b} w_{k_{a\lra b}} X_k \ket\psi,
\end{align}
where $k_{a\lra b}$ denotes a list that is equal to $k=\p{k_1,k_2,\cdots,k_M}$ but with $k_a$ and $k_b$ swapped, and we have relabeled $k_a$ and $k_b$ to change $w_k X_{k_{a\lra b}}$ to $w_{k_{a\lra b}} X_k$.
To write down these terms succinctly, we define a linear operator $\hat{\m h}_{\t{P}}$ on the space of dimension-$M$ tensors, whose action on $\m w$ is a new dimension-$M$ tensor with entries
\begin{align}
  \p{\hat{\m h}_{\t{P}} \c \m w}_k
  \equiv \sum_{\substack{\set{a,b}\in\C_M\p{2}}}
  h_{k_ak_b} w_{k_{a\lra b}},
  \label{eq:multi_body_op_perm}
\end{align}
such that
\begin{align}
  \sum_{k\in\D_N\p{M}} \sum_{\substack{\set{i,j}\in\C_N\p{2}\\i,j\in k}}
  h_{ij} w_k \Pi_{ij} X_k \ket\psi
  = \sum_{k\in\D_N\p{M}} \p{\hat{\m h}_{\t{P}} \c \m w}_k X_k \ket\psi.
\end{align}
Finally, the terms in \eqref{eq:diagnosis_start} with only one of $i\in k$ or $j\in k$ are
\begin{align}
  \sum_{k\in\D_N\p{M}} \sum_{i\notin k} \sum_{j\in k}
  h_{ij} w_k \Pi_{ij} X_k \ket\psi
  = \sum_{k\in\D_N\p{M}} \sum_{i\notin k}
  \sum_{a\in\ZZ_M} h_{ik_a} w_{k_{a:i}} X_k \ket\psi,
\end{align}
where $k_{a:i}$ denotes a list that is equal to $k$ but with $k_a$ replaced by $i$, and we have relabeled $k_a$ and $i$ to change $w_k X_{k_{a:i}}$ to $w_{k_{a:i}} X_k$.
For each fixed $a$, the sum over $i$ above is essentially a contraction of the matrix $\m h$ with the $a$-th index of the tensor $\m w$.
Defining a linear operator $\hat{\m h}_a$ on the space of dimension-$M$ tensors, whose action on $\m w$ is a new dimension-$M$ tensor with entries
\begin{align}
  \p{\hat{\m h}_a\c\m w}_k
  \equiv \sum_{i\notin k} h_{ik_a} w_{k_{a:i}}
  = \sum_{i\notin k}
  h_{ik_a} w_{k_1 k_2 \cdots k_{a-1} i k_{a+1} \cdots k_M},
  \label{eq:multi_body_op_contract}
\end{align}
we can expand
\begin{align}
  \sum_{k\in\D_N\p{M}} \sum_{i\notin k} \sum_{j\in k}
  h_{ij} w_k \Pi_{ij} X_k \ket\psi
  = \sum_{k\in\D_N\p{M}} \sum_{a\in\ZZ_M}
  \p{\hat{\m h}_a\c\m w}_k X_k \ket\psi.
\end{align}
Altogether, defining the combined operator
\begin{align}
  \hat{\m h}
  \equiv \sum_{a\in\ZZ_M} \m h_a + \m h_{\t{P}} - \m h_{\t{D}},
\end{align}
we find that
\begin{align}
  H_0 \O\p{\m w,X} \ket\psi
  = E_0 \O\p{\m w,X} \ket\psi +
  \O\p{\hat{\m h}\c\m w, X} \ket\psi
  = \sum_\Delta \p{E_0+\Delta} \O\p{\m w_\Delta, X} \ket\psi,
\end{align}
where $\m w_\Delta$ is the projection of $\m w$ onto the eigenspace of $\hat{\m h}$ with eigenvalue $\Delta$, i.e.
\begin{align}
  \hat{\m h} \c \m w_\Delta = \Delta \m w_\Delta,
  \label{eq:multi_body_problem}
\end{align}
with
\begin{align}
  \hat{\m h} \equiv \sum_{k\in\D_N\p{M}} \ket{k}
  \sp{\sum_{a\in\ZZ_M} \sum_{\substack{i\in\ZZ_N\\i\notin k}}
    h_{ik_a} \bra{k_{a:i}}
    + \sum_{\set{a,b}\in\C_M\p{2}} h_{k_ak_b} \bra{k_{a\lra b}}
    - h^{\t{D}}_k \bra{k}}.
  \label{eq:multi_body_mat}
\end{align}
For reference, here $k_{a:i}$ is equal to $k\equiv\p{k_1,k_2,\cdots,k_M}$ but with $k_a$ replaced by $i$; $\C_M\p{2}$ is defined in \eqref{eq:choices}; $k_{a\lra b}$ is equal to $k$ but with $k_a$ and $k_b$ swapped; and $h^{\t{D}}_k$ is defined in \eqref{eq:multi_body_op_diag_element}.

If the coefficient tensor $\m w$ is permutationally symmetric, meaning that $\m w$ is invariant under arbitrary permutations of its tensor factors, then this symmetry is preserved by $\hat{\m h}$.
In this case, we can replace sums over $k\in\D_N\p{M}$ in \eqref{eq:multi_body_problem} and \eqref{eq:multi_body_mat} by sums over $k\in\C_N\p{M}$, and replace vectors $\ket{k_1,k_2,\cdots,k_M}\to\ket{\set{k_1,k_2,\cdots,k_M}}$, such that e.g.~$\ket{k_{a\lra b}}=\ket{k}$.
These replacements reduce the size of $\hat{\m h}$ from $\abs{\D_N\p{M}}\times\abs{\D_N\p{M}}$ to $\abs{\C_N\p{M}}\times\abs{\C_N\p{M}}$, where $\abs{\D_N\p{M}}=\prod_{j=0}^{M-1}\p{N-j}=M!\times{N\choose M}$ and $\abs{\C_N\p{M}}={N\choose M}$.
We discuss further reduction of the eigenvalue problem in \eqref{eq:multi_body_problem} by use of additional symmetries in Section \ref{sec:symmetries}.

Some special cases of practical interest: if $M=1$, then $h^{\t{D}}_k=h_k$, so
\begin{align}
  \hat{\m h} \stackrel{M=1}{=} \sum_{k\in\ZZ_N} \ket{k}
  \sp{\sum_{i\in\ZZ_N} h_{ik} \bra{i} - h_k \bra{k}}
  = \m h - \oper{diag}\v h,
\end{align}
which is precisely what we found in \eqref{eq:single_body_eig}.
If $M=2$, then $h_{k\ell}^{\t{D}}=h_k+h_\ell-h_{k\ell}$, so
\begin{align}
  \hat{\m h} \stackrel{M=2}{=}
  \sum_{\p{k,\ell}\in\D_N\p{2}} \ket{k\ell}
  \sp{\sum_{\substack{i\in\ZZ_N\\i\notin\set{k,\ell}}}
    \p{ h_{ik} \bra{i\ell} + h_{i\ell} \bra{ki} }
    + h_{k\ell} \bra{\ell k}
    - \p{h_k + h_\ell - h_{k\ell}} \bra{k\ell}}.
\end{align}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Exploiting spatial symmetries}
\label{sec:symmetries}

Solving the eigenvalue problem in \eqref{eq:multi_body_problem} to find $M$-body operators $\O\p{\m w_\Delta,X}$ that generate states of definite interaction energy nominally requires diagonalizing the matrix $\hat{\m h}$ with dimensions $\sim N^M\times N^M$.
Spatial symmetries of the interaction Hamiltonian $H_0$, however, generally endow $\hat{\m h}$ with a block-diagonal structure that can be used to reduce the complexity of this eigenvalue problem.
Translational invariance, for example, can reduce this problem to that of diagonalizing a matrix with dimensions $\sim N^{M-1}\times N^{M-1}$, and isotropy can further reduce the problem size by factors that depend on the lattice geometry.
The complexity of diagonalizing a matrix with dimensions $d\times d$ is $O\p{d^3}$, so even a reduction of $d$ by a factor of 2 yields practically significant gains (a factor of $2^3=8$) in the runtime of diagonalization.
Here, we provide a general prescription for exploiting the spatial symmetries of $H_0$ to simplify the eigenvalue problem in \eqref{eq:multi_body_problem}.

Our task is essentially to express the multi-body eigenvalue problem in \eqref{eq:multi_body_problem} in a basis that respects the spatial symmetries of $H_0$.
These symmetries essentially break the set $\D_N\p{M}$ into equivalence classes $\S_M$\footnote{For brevity, we will suppress the explicit dependence of $\S_M$ on $N$, as well as other factors such as lattice geometry.}, where elements $s\in\S_M$ are subsets $s\subset\D_N\p{M}$ that are invariant under the appropriate symmetry group action.
In the case of $M=2$ in a translationally invariant and isotropic system, for example, each equivalence class $s\in\S_2$ would consist of all pairs $\p{i,j}\in\D_N\p{2}$ that are a fixed distance $\abs{i-j}=d_s$ apart, i.e.~$s=\set{\p{i,j}\in\D_N\p{2}:\abs{i-j}=d_s}$.
If the tensor $\m w$ is invariant under arbitrary permutations of its tensor factors, this symmetry is preserved by $\hat{\m h}$, and we can place all $k\in\D_N\p{M}$ that are equal up to permutation into the same equivalence class.

For simplicity, we first assume that the $M$-body operator $\O\p{\m w,X}$ obeys the same spatial symmetries as $H_0$.
For any equivalence class $s\in\S_M$, we then define a uniform superposition $\ket{s}$ over all $k\in s$, as well as a projector $\P_\S$ onto the space of uniform superpositions:
\begin{align}
  \ket{s} \equiv \f1{\sqrt{\abs{s}}} \sum_{k\in s} \ket{k},
  &&
  \P_{\S_M} \equiv \sum_{s\in\S_M} \op{s},
\end{align}
where $\abs{s}$ is the number of elements in $s$.
In turn, we expand a vector of $M$-body operator coefficients $\m w=\sum_{k\in\D_N\p{M}}w_k\ket{k}$ in this symmetrized basis:
\begin{align}
  \m w = \P_{\S_M} \m w
  = \sum_{s\in\S_M} \sum_{k\in s} w_k \ket{s} \bk{s|k}
  = \sum_{s\in\S_M} w_s \sqrt{\abs{s}} \ket{s},
\end{align}
where all $w_k$ with $k\in s$ are equal by assumption, so $w_s\equiv\sum_{k\in s}w_k/\abs{s}=w_\ell$ for any $\ell\in s$.
Finally, we expand the projection of $\hat{\m h}$ onto the symmetric subspace:
\begin{align}
  \P_{\S_M} \hat{\m h} \P_{\S_M}
  = \sum_{s\in\S_M} \f{\ket{s}}{\sqrt{\abs{s}}} \sp{
    \sum_{\substack{k\in s\\a\in\ZZ_M}}
    \sum_{\substack{p\in\ZZ_N\\p\notin k}}
    h_{p k_a} \f{\bra{\sp{k_{a:p}}}}{\sqrt{\abs{\sp{k_{a:p}}}}}
    + \sum_{\set{a,b}\in\C_M\p{2}} h_{k_ak_b}
    \f{\bra{\sp{k_{a\lra b}}}}{\sqrt{\abs{\sp{k_{a\lra b}}}}}
    - h^{\t{D}}_s \f{\bra{s}}{\sqrt{\abs{s}}}},
\end{align}
where $h^{\t{D}}_s\equiv h^{\t{D}}_k$ for any $k\in s$, with $h^{\t{D}}_k$ defined in \eqref{eq:multi_body_op_diag_element}; and $\sp{k}$ denotes the equivalence class of $k$.
The reduced multi-body eigenvalue problem is then
\begin{align}
  \P_{\S_M} \hat{\m h} \P_{\S_M} \c \m w = \Delta \m w.
\end{align}
More generally, even if the couplings $\m w$ of $\O\p{\m w,X}$ do not obey the spatial symmetries as $H_0$, these symmetries still endow $\hat{\m h}$ with a block-diagonal structure that reduces the complexity of the multi-body eigenvalue problem.
We leave this reduction to future work.
% todo: perform reduction

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Choosing the operator content}
\label{sec:operator_content}

Throughout our construction of excitation energy eigenstates of the form $\O\p{\m w,X}\ket\psi$, where $\ket\psi\in\M_0$ is a PS state, we have thus far primarily considered the choice of scalar coefficients, $\m w$.
Here, we discuss the choice of operators, $X$, which allows us to construct a complete and convenient basis for the set of all excited states.
Our construction of a basis for the set of excited states begins with the standard basis $\set{\ket{m}:m\in\A_n\p{N}}$ for the PS manifold $\M_0$, where $m\equiv\p{m_1,m_2,\cdots,m_n}$ and $m_\mu$ is the (integer) occupation number of the single-spin state $\mu$.
Written in the ``computational basis'' of $N$ multilevel spins, these states are
\begin{align}
  \ket{m} =
  \f1{\sqrt{\C\p{m}}}
  \sum_{\substack{\t{distinct}\\\t{permutations}\\\Pi~\t{of}~\tilde m}}
  \Pi \ket{\tilde m},
  &&
  \ket{\tilde m} \equiv \bigotimes_\mu \ket{\mu}^{\otimes m_\mu},
  \label{eq:pops_expansion}
\end{align}
where $\C\p{m}$ is a normalization factor equal to the number of terms in the sum.
While $\ket{m}$ is the only PS state with occupation numbers $m$, the set of {\it all} states with occupation numbers $m$ is spanned by states of the form
\begin{align}
  \ket{m,s} = \f1{\sqrt{\C\p{m}}}
  \sum_{\substack{\t{distinct}\\\t{permutations}\\\Pi~\t{of}~\tilde m}}
  s\p{\Pi} \Pi \ket{\tilde m},
  \label{eq:pops_states}
\end{align}
where $s\p{\Pi}\in\set{\pm1}$ for all permutations $\Pi$.
As a consequence, when generating excited states of the form $\O\p{\m w,X}\ket{m}$, we can freely restrict ourselves to operators $X$ that only change the signs of individual terms (associated with permutations $\Pi$) in \eqref{eq:pops_expansion}, which suffices to construct any state $\ket{m,s}$ in \eqref{eq:pops_states}.
We therefore consider single-operators $X$ that have $\pm1$ on the diagonal (in the computational basis), and zero everywhere else.

The set of all operators that are $\pm1$ on the diagonal is spanned by sums and tensor products of the single-spin operators $Z_\nu\equiv1-2\op\nu=\sum_{\mu\ne\nu}\op\mu-\op\nu$.
Given a list of non-negative integers $z\equiv\p{z_1,z_2,\cdots,z_n}$ with $\sum_\mu z_\mu=r$, we thus define the $r$-spin operator
\begin{align}
  X_z \equiv \bigotimes_\mu Z_\mu^{z_\mu},
\end{align}
and nominally use states of the form $\ket{m,\m w,z}=\O\p{\m w,X_z}\ket{m}$ as our basis for the $r$-magnon manifold $\M_r$.
There are three points to address concerning the states $\ket{m,\m w,z}$:
\begin{enumerate*}
\item \label{pt:trivial} depending on the value of $m$, some choices of $z$ result in trivial states, $\ket{m,\m w,z}=0$;
\item \label{pt:dependent} the set of nontrivial states is generally overcomplete, in that any given state can be written as a nontrivial linear combination of others, and finally
\item \label{pt:orthogonal} even after identifying a linearly independent set of nontrivial states, any given pair of states $\ket{m,\m w,y},\ket{m,\m w,z}$ with $y\ne z$ generally have a non-zero overlap, which is to say that they are not orthogonal.
\end{enumerate*}
To address the first two points, \ref{pt:trivial} and \ref{pt:dependent}, we will place restrictions on the values of $z$ that we allow for any given $m$.
The last point, \ref{pt:orthogonal}, will require us to construct a basis of orthonormal states numerically for any given values of $m,\m w$.

First, we identify criteria under which we are guaranteed that $\ket{m,\m w,z}=0$.
We begin by assuming that $\ket{m,\m w,z}\in\M_r$ with $\sum_\nu z_\nu=r$.
We then observe that if $z_\nu>m_\nu$, then $m_\nu-z_\nu$ factors of $Z_\nu$ in $X_z$ are ``wasted,'' in the sense that they never yield relative signs in different terms of $\ket{m}$ when expressed in the computational basis, as in \eqref{eq:pops_expansion}.
More concretely, consider the individual terms of the expansion
\begin{align}
  \ket{m,\m w,z}
  = \f1{\sqrt{\C\p{m}}}
  \sum_{\substack{\t{distinct}\\\t{permutations}\\\Pi~\t{of}~\tilde m}}
  \sum_{k\in\D_N\p{r}} w_k \sp{X_z}_k \Pi \ket{\tilde m},
  \label{eq:excited_state_expansion}
\end{align}
where $\sp{X_z}_k$ denotes the action of $X_z$ on spins $k=\p{k_1,k_2,\cdots,k_r}$.
All terms in \eqref{eq:excited_state_expansion} have exactly $m_\nu$ spins in the state $\nu$.
For any fixed $k$, at least $m_\nu-z_\nu$ factors of $Z_\nu$ in $\sp{X_z}_k$ must therefore evaluate to $+1$.
We can eliminate these ``extra'' factors of $Z_\nu$ by replacing $m_\nu-z_\nu$ of the factors that evaluate to $+1$ by the identity operator.
Performing this replacement for all $\nu\in\ZZ_n$ yields a state $\ket{m,\m w_{mz},{\floor{z}_m}}=\ket{m,\m w,z}$, where $\floor{z}_m\equiv\p{\min\set{z_\nu,m_\nu}:\nu\in\ZZ_n}$ indexes the $q$-spin operator $X_{\floor{z}_m}$, and $\m w_{mz}$ is some new tensor of coefficients.
If $q<r$ (because $z_\nu>m_\nu$ for some $\nu$), then our initial assumption that $\ket{m,\m w,z}\in\M_r$ implies that $\ket{m,\m w,z}=0$.
In other words, a nontrivial basis $\set{\ket{m,\m w,z}}$ for $\M_r$ must have $z_\nu\le m_\nu$ for all $\nu$.

Next, we address the fact that the basis $\set{\ket{m,\m w,z}}$ for $\M_r$ is overcomplete.
This fact is a simple consequence of our choice of single-spin operators for the construction of $\ket{m,\m w,z}$, namely $\set{Z_\nu:\nu\in\ZZ_n}$.
In essence, our problem is the fact that $Z_\nu=\p{n-2}-\sum_{\mu\ne\nu}Z_\mu$, which implies that any state $\ket{m,\m w,z}\in\M_r$ can be written as a linear combination of states in which $z_\nu\to z_\nu-1$, and possibly $z_\mu\to z_\mu+1$ for some $\mu\ne\nu$.
We can eliminate this redundancy by removing an operator from the $\set{Z_\nu}$, or equivalently enforcing that $z_\nu=0$ for some $\nu$.
As a matter of choice, we therefore consider only states $\ket{m,\m w,z}$ for which $z$ is zero at the first entry where $m$ is maximal: $z_\nu=0$ for $\nu=\min\set{\mu:m_\mu=\max\p{m}}$.
Given our previous restriction that $z_\nu\le m_\nu$ for all $\nu$, enforcing that $z$ is zero at an entry where $m$ is maximal maximizes the number of candidate values of $z$ that we eliminate from consideration.

While the above restrictions on $z$ eliminate some trivial and redundant states from the basis $\set{\ket{m,\m w,z}}$ for $\M_r$, they do not guarantee that there will be {\it no} trivial states or redundancies left over.
Furthermore, even after restrictions the remaining basis is not orthogonal, which is to say that generally $\bk{m,\m v,y|m,\m w,z}\ne0$ even when $y\ne z$.
In order to construct an orthonormal basis for $\M_r$, we therefore need to build a matrix $B_m$ with matrix elements
\begin{align}
  B_m\p{\m v,y;\m w,z} \equiv \bk{m,\m v,y|m,\m w,z}
  = \bk{m| \O\p{\m v,X_y}^\dag \O\p{\m w,X_z} |m},
  \label{eq:basis_matrix}
\end{align}
where $y,z$ satisfy appropriate restrictions, and $\m v,\m w$ are chosen from the set of tensors that generate eigenstates of $H_0$.
Eigenvectors of $B_m$ are then mutually orthogonal vectors of the form $\sum_{\m w,z}c_{\m w z}\ket{m,\m w,z}$, whose squared norms are equal to their corresponding eigenvalues.
Note that the matrix $B_m$ has a block-diagonal structure, as it cannot (for example) couple states $\ket{m,\m v,y},\ket{m,\m w,z}$ whose tensors $\m v,\m w$ generate states with different interaction energies.
We therefore only need to build and diagonalize this matrix independently within each block.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Products of multi-body operators}
\label{sec:operator_product}

The construction of a basis $\B_r$ for the $r$-magnon manifold $\M_r$ requires multiplying two multi-body operators $\O\p{\m v,X_y}^\dag, \O\p{\m w,X_z}$, and projecting the result onto the permutationally symmetric (PS) manifold $\M_0$, as in \eqref{eq:basis_matrix}.
Similarly, computing matrix elements of operators $\O\p{\m v,Y}$ with respect to the basis $\col\B_{r_{\t{max}}}\equiv\bigotimes_{r\le r_{\t{max}}}\B_r$ requires multiplying three multi-body operators and projecting the result onto $\M_0$, as in \eqref{eq:matrix_element}.
In this section, we develop a diagrammatic technique for the sorts of calculations, and work out some special cases of common interest.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{A diagrammatic expansion}

Given a system of $N$ spins, we wish to project a product of multi-body operators onto the PS manifold.
For $p$ multi-body operators, such a product takes the form
\begin{align}
  \Q\p{\m w,\m X}
  \equiv \prod_{\alpha\in\ZZ_p} \O\p{w_\alpha,X_\alpha}
  = \prod_{\alpha\in\ZZ_p} \sum_{k_\alpha\in\D_N\p{M_\alpha}}
  w_\alpha\p{k_\alpha} X_\alpha\p{k_\alpha},
  \label{eq:sym_prod_start}
\end{align}
where $\m w\equiv\p{w_1,w_2,\cdots,w_p}$ is a vector of dimension-$M_\alpha$ tensors $w_\alpha$; $\m X\equiv\p{X_1,X_2,\cdots,X_p}$ is a vector of $M_\alpha$-spin operators $X_\alpha$; $k_\alpha=\p{k_{\alpha,1},k_{\alpha,2},\cdots,k_{\alpha,M_\alpha}}$ is a choice of $M_\alpha$ spins that $X_\alpha\p{k_\alpha}$ acts on; $w_\alpha\p{k_\alpha}$ is a scalar entry of $w_\alpha$; and $\D_N\p{M_\alpha}$ is the strictly off-diagonal part of $\ZZ_N^M$, defined in \eqref{eq:off_diags}.
For later convenience, we extend the definitions of $w_\alpha\p{k_\alpha}=X_\alpha\p{k_\alpha}$ to all $k_\alpha\in\ZZ_N^M$ by defining $w_\alpha\p{k_\alpha}=X_\alpha\p{k_\alpha}=0$ for any $k_\alpha$ with repeated entries.

To simplify $\Q\p{\m w,\m X}$, we first collect terms to express \eqref{eq:sym_prod_start} as a sum of products:
\begin{align}
  \Q\p{\m w,\m X} = \sum_{k\in\D_N\p{\m M}} \prod_{\alpha\in\ZZ_p}
  w_\alpha\p{k_\alpha} X_\alpha\p{k_\alpha},
  \label{eq:sym_prod_sum}
\end{align}
where $\m M\equiv\p{M_1,M_2,\cdots,M_p}$ is a vector of the dimensions $M_\alpha$; and $\D_N\p{\m M}\equiv\bigotimes_{M_\alpha\in\m M}\D_N\p{M_\alpha}$, such that each $k\in\D_N\p{\m M}$ has the decomposition $k=\p{k_1,k_2,\cdots,k_p}$ with $k_\alpha\in\D_N\p{M_\alpha}$.
We then classify the terms in \eqref{eq:sym_prod_sum} by the numbers $g_S$ of indices shared by all tensors $w_\alpha$ with $\alpha\in S\subset\ZZ_p$.
For example, a term of the form $w_1\p{a,b,c} w_2\p{b,d,e} w_3\p{b,c,d,e}$ with distinct indices $a,b,c,d,e$ would have
\begin{align}
  g_{\set{1}} = 1,
  &&
  g_{\set{1,2,3}} = 1,
  &&
  g_{\set{1,3}} = 1,
  &&
  g_{\set{2,3}} = 2,
\end{align}
and $g_S=0$ for all other subsets $S\subset\ZZ_3$.
This index assignment can be associated with the Venn diagram
\begin{align}
  \diagram{example_123},
  \label{eq:venn_diagram}
\end{align}
where $g_S$ is determined the number of dots at the intersection of circles $\alpha\in S$.
We denote the set of all nonempty subsets of $\ZZ_p$ by $\PP^*\p{\ZZ_p}$, and denote a choice of $g_S$ for all $S\in\PP^*\p{\ZZ_p}$ by $g$.
With representations such as \eqref{eq:venn_diagram} in mind, we will loosely refer to $g$ as simply a ``diagram''.
We define $\abs{g}\equiv\sum_{S\in\PP^*\p{\ZZ_p}}g_S$ to be the number of indices considered in $g$ (or equivalently the number of dots in the diagram $g$), and we denote the set of all diagrams for $p=\dim\p{\m M}$ tensors with dimensions $\m M$ by $\G\p{\m M}$.
For consistency, all diagrams $g\in\G\p{\m M}$ must associate exactly $M_\alpha$ indices with the dimension-$M_\alpha$ tensor $w_\alpha$, i.e.
\begin{align}
  \sum_{S\in\PP^*\p{\ZZ_p}\,:\,\alpha\in S} g_S
  = \sum_{R\in\PP^*\p{\ZZ_p\setminus\set{\alpha}}} g_{\set{\alpha}\cup R}
  = M_\alpha
\end{align}
for all $\alpha\in\ZZ_p$.
A classification of the terms in \eqref{eq:sym_prod_sum} by diagrams $g\in\G\p{\m M}$ allows us to expand
\begin{align}
  \Q\p{\m w,\m X} \EQPS \sum_{g\in\G\p{\m M}} \Q_g\p{\m w,\m X},
  \label{eq:sym_prod_diagrams}
\end{align}
where $\EQPS$ denotes equality under a restriction to $\M_0$, and $\Q_g\p{\m w,\m X}$ is the sum of all terms in \eqref{eq:sym_prod_sum} that are associated with the diagram $g$.

To find an explicit form of $\Q_g\p{\m w,\m X}$, we first consider all distinct ways to assign $\abs{g}$ indices to specific tensor factors of $\m w$ and $\m X$ in a manner consistent with $g$.
For example, if $\m X=\p{X_1,X_2}$ with dimensions $\p{M_1,M_2}=\p{2,1}$ and $\abs{g}=2$ with $g_{\set{1}}=g_{\set{1,2}}=1$, then $X_1\p{a,b} X_2\p{b}$ and $X_1\p{b,a} X_2\p{b}$ are distinct assignments of $\abs{g}$ indices to $\m X$ in a manner consistent with $g$, provided that $a\ne b$.
After assigning $\abs{g}$ indices to $\m w$ and $\m X$ in a manner consistent with $g$, we need to sum over all values that these indices can take, which amounts to a sum over all $k\in\D_N\p{\abs{g}}$.

A specific assignment of indices can be thought of as a choice, for each of the $M_\alpha$ tensor factors of $w_\alpha$ and $X_\alpha$, of which dot in the diagram $g$ that tensor factor is associated with (i.e.~out of the $M_\alpha$ dots within circle $\alpha$).
There are nominally $M_\alpha!$ such choices for a given $\alpha$, and therefore $\prod_{\alpha\in\ZZ_p}M_\alpha!$ choices total.
The only catch is that these choices treat all dots in $g$ as distinct, whereas the $g_S$ dots within region $S$ of $g$ are actually indistinguishable.
The assignment $X_1\p{a,b,c} X_2\p{a,b}$, for example, is the same as $X_1\p{b,a,c} X_2\p{b,a}$, because $a$ and $b$ are dummy indices.
Treating $g_S$ dummy indices as as distinguishable results in over-counting distinct index assignments by a factor of $g_S!$.
Denoting the set of all distinct index assignments according to $g$ by $\L\p{g}$, the number of such assignments is therefore
\begin{align}
  \abs{\L\p{g}} = \sp{\prod_{\alpha\in\ZZ_p}M_\alpha!}
  \times \sp{\prod_{S\in\PP^*\p{\ZZ_p}} g_S!}^{-1}.
  \label{eq:symmetry}
\end{align}
Altogether, we have
\begin{align}
  \Q_g\p{\m w,\m X} \equiv
  \sum_{\ell\in\L\p{g}} \sum_{k\in\D_N\p{\abs{g}}}
  \m w\p{k,\ell} \m X\p{k,\ell},
\end{align}
where $\m w\p{k,\ell}$ and $\m X\p{k,\ell}$ denote the scalar and operator acquired by assigning the indices $k=\p{k_1,k_2,\cdots,k_{\abs{g}}}$ to $\m w$ and $\m X$ according to $\ell$.
Upon projection to the PS manifold $\M_0$, this operator becomes
\begin{align}
  \Q_g\p{\m w,\m X} \equiv
  \sum_{\ell\in\L\p{g}} \col{\m w}\p{\ell} \mean{\m X}\p{\ell},
  \label{eq:sum_assignments}
\end{align}
where
\begin{align}
  \col{\m w}\p{\ell} \equiv \sum_{k\in\D_N\p{\abs{g}}} \m w\p{k,\ell},
  &&
  \mean{\m X}\p{\ell}
  \equiv \f1{\abs{\D_N\p{\abs{g}}}}
  \sum_{k\in\D_N\p{\abs{g}}} \m X\p{k,\ell},
\end{align}
and permutational symmetry implies that the averaging for $\mean{\m X}\p{\ell}$ is not actually necessary:
\begin{align}
  \mean{\m X}\p{\ell}
  \EQPS \m X\p{k,\ell} ~\t{for any}~ k\in\D_N\p{\abs{g}}.
\end{align}
If, furthermore, all tensors $w_\alpha$ in $\m w$ (or all operators $X_\alpha$ in $\m X$) are permutationally symmetric, meaning that they are invariant under arbitrary permutations of their tensor factors, then
\begin{align}
  \Q_g\p{\m w,\m X}
  \stackrel{\t{sym}}{=}_{\t{PS}} \col{\m w}\p{g} \mean{\m X}\p{g},
  \label{eq:sum_diagrams}
\end{align}
where $\stackrel{\t{sym}}{=}_{\t{PS}}$ denotes equality up to
\begin{enumerate*}
\item permutational symmetry of all $\m w$ or $\m X$, and
\item restriction to the PS manifold $\M_0$; and
\end{enumerate*}
\begin{align}
  \col{\m w}\p{g} \equiv \sum_{\ell\in\L\p{g}} \col{\m w}\p{\ell},
  &&
  \mean{\m X}\p{g} \equiv \f1{\abs{\L\p{g}}}
  \sum_{\ell\in\L\p{g}} \mean{\m X}\p{\ell}.
\end{align}
In summary, the product of operators $\Q\p{\m w,\m X}$ in \eqref{eq:sym_prod_start} admits the diagrammatic expansion in \eqref{eq:sym_prod_diagrams}, as a sum of $\Q_g\p{\m w,\m X}$ over all diagrams $g\in\G\p{\m M}$.
Each term $\Q_g\p{\m w,\m X}$ is a sum of $\col{\m w}\p{\ell}\mean{\m X}\p{\ell}$ over all distinct index assignments $\ell\in\L\p{g}$.
The scalar $\col{\m w}\p{\ell}$ is obtained by assigning indices $k=\p{k_1,k_2,\cdots,k_{\abs{g}}}$ to the tensors $\m w$ as prescribed by $\ell$, and summing over all values of $k\in\D_N\p{\abs{g}}$.
The collective operator $\mean{\m X}\p{\ell}$ is similarly constructed with an average over all $k\in\D_N\p{\abs{g}}$, rather than a sum.
If the tensors $\m w$ or operators $\m X$ are all permutationally symmetric, then $\Q_g\p{\m w,\m X}$ factorizes into $\col{\m w}\p{g}\mean{\m X}\p{g}$, where $\col{\m w}\p{g}$ and $\mean{\m X}\p{g}$ are respectively the sum and average of $\col{\m w}\p{\ell}$ and $\mean{\m X}\p{\ell}$ over all index assignments $\ell\in\L\p{g}$.

As a final point, we note that the choice to sum (average) for the scalar (operator) content of $\Q_g\p{\m w,\m X}$ in \eqref{eq:sum_assignments} and \eqref{eq:sum_diagrams} was arbitrary, and we could just as well have chosen to interchange sums and averages.
To simplify the diagrammatic formalism we develop in the following section, we therefore define the {\it mixed} sums and averages
\begin{align}
  \ut{\m w}\p{g} \equiv \f1{\abs{\L\p{g}}}
  \sum_{\ell\in\L\p{g}} \col{\m w}\p{\ell}
  = \f1{\abs{\L\p{g}}} \sum_{\ell\in\L\p{g}}
  \sum_{k\in\D_N\p{\abs{g}}} \m w\p{k,\ell},
\end{align}
\begin{align}
  \ot{\m X}\p{g} \equiv \sum_{\ell\in\L\p{g}} \mean{\m X}\p{\ell}
  \EQPS \sum_{\ell\in\L\p{g}} \m X\p{k,\ell}
  ~\t{for any}~ k\in\D_N\p{\abs{g}},
\end{align}
where $\ut{\m w}\p{g}$ sums over index values $k\in\D_N\p{\abs{g}}$, but averages over index assignments $\ell\in\L\p{g}$; and vice versa for $\ot{\m X}\p{g}$.
If all tensors $w_\alpha$ in $\m w$ (or all operators $X_\alpha$ in $\m X$) are permutationally symmetric, then
\begin{align}
  \Q_g\p{\m w,\m X} \stackrel{\t{sym}}{=}_{\t{PS}}
  \ut{\m w}\p{g} \ot{\m X}\p{g}.
  \label{eq:sum_diagrams_alt}
\end{align}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Formalizing diagrams and minimizing computational costs}
\label{sec:diagrams}

Diagrams such as \eqref{eq:venn_diagram} play a central role in simplifying products of multi-body operators, $\Q\p{\m w,\m X}$ through the expansion of these products in \eqref{eq:sym_prod_diagrams} as sum over $\Q_g\p{\m w,\m X}$.
Here, we formalize a diagrammatic representation of the scalars $\col{\m w}\p{\ell}$ and $\ut{\m w}\p{g}$ that appear in $\Q_g\p{\m w,\m X}$, which enables us to perform otherwise intractable calculations.
We then derive rules to ``reduce'' these scalars so as to minimize the computational cost of their numerical evaluation.

A diagram $g\in\G\p{\m M}$ with $\m M=\p{M_1,M_2,\cdots,M_p}$ consists of $p$ circles, and $g_S$ dots at the intersection of circles $\alpha\in S\subset\ZZ_p$.
By slight abuse of notation, labeling each circle $\alpha$ by a tensor $w_\alpha$ defines the scalar $g\p{\m w}\equiv\ut{\m w}\p{g}$.
If, for example, all tensors $w_\alpha$ are all permutationally symmetric, then
\begin{align}
  \diagram{example_w123}
  = \sum_{\p{a,b,c,d,e}\in\D_N\p{5}}
  w_1\p{a,b,c} w_2\p{b,d,e} w_3\p{b,c,d,e}.
  \label{eq:example_123}
\end{align}
When the tensors $\m w$ are not all permutationally symmetric, one can additionally equip a diagram $g$ with a specific index assignment $\ell\in\L\p{g}$, in which case the diagram is equal to $\ell\p{\m w}\equiv\col{\m w}\p{\ell}$.
Diagrams that are not equipped with a specific index assignment $\ell$ are then an average over all $\ell\in\L\p{g}$, i.e.~$g\p{\m w}\equiv\sum_{\ell\in\L\p{g}} \ell\p{\m w} / \abs{\L\p{g}}$.
To keep mathematical expressions simple, we will use symmetric tensors in all examples throughout this section, but our discussions apply just as well to the asymmetric case.

A diagram such as \eqref{eq:example_123} with $\abs{g}$ dots nominally takes $O\p{N^{\abs{g}}}$ time to compute due to the interdependence of all indices.
We wish to ``reduce'' such diagrams and minimize their computational cost as much as possible.
Our general strategy will be to replace constrained sums that make all indices interdependent by unconstrained sums that can be carried out independently, in sequence.
In order to represent different constraints, we need to define diagrams in which filled dots (``$\bullet$'') may be replaced by empty dots (``$\circ$'') or crosses (``$\bm\times$'').

Filled dots in a diagram correspond to indices that are constrained to have different values.
The presence of $F$ filled dots thus corresponds to a sum over $\D_N\p{F}$.
Every empty dot, meanwhile, corresponds to an index that is summed over $\ZZ_N$ without additional constraints, e.g.
\begin{align}
  \diagram{example_o}
  = \sum_{\p{a,b,e}\in\D_N\p{3}} \sum_{c,d,f\in\ZZ_N}
  v\p{a,b,c,d,e} w\p{e,f}.
\end{align}
Crosses, meanwhile, correspond to indices that are {\it only} summed over the values of ``filled dot indices'', e.g.
\begin{align}
  \diagram{example_x}
  = \sum_{\p{a,e}\in\D_N\p{2}} \sum_{b,c\in\set{a,e}} \sum_{d\in\ZZ_N}
  u\p{a,b,c} v\p{b,d,e} w\p{b,c,e}.
\end{align}
Empty dots and crosses allow us to decompose diagrams with a high computational cost into different diagrams with a lower computational cost.
We will essentially use two tricks to reduce diagrams: eliminating filled dots (which will add empty dots and crosses), and eliminating crosses (which will move around filled dots).
These two tricks will be applied in sequence until only empty dots remain.

Given a diagram containing only dots (filled or empty), we can decompose any filled dot into an empty dot and a cross.
For example,
\begin{align}
  \diagram{example_elim}
  = \diagram{example_elim_o}
  - \diagram{example_elim_x}
  \label{eq:example_elim}
\end{align}
where
\begin{align}
  \diagram{example_elim}
  = \sum_{\p{a,b,c}\in\D_N\p{3}} v\p{a,b} w\p{b,c},
\end{align}
\begin{align}
  \diagram{example_elim_o}
  = \sum_{\p{b,c}\in\D_N\p{2}} v\p{\circ,b} w\p{b,c},
  &&
  v\p{\circ,b} \equiv \sum_{a\in\ZZ_N} v\p{a,b},
  \label{eq:example_elim_o}
\end{align}
\begin{align}
  \diagram{example_elim_x}
  = \sum_{\p{b,c}\in\D_N\p{2}} \sum_{a\in\set{b,c}}
  v\p{a,b} w\p{b,c}.
  \label{eq:example_elim_x}
\end{align}
The decomposition in \eqref{eq:example_elim} essentially breaks up a sum over $a\in\ZZ_N\setminus\set{b,c}$ into a difference of sums over $a\in\ZZ_N$ and $a\in\set{b,c}$.
Together, the sums over $a\in\ZZ_N$ to compute all $v\p{\circ,b}$ have an $O\p{N^2}$ cost, and can be performed prior to evaluating the diagram in \eqref{eq:example_elim_o}.
The decomposition in \eqref{eq:example_elim} therefore splits one $O\p{N^3}$ diagram into two $O\p{N^2}$ diagrams.

After decomposing a filled dot into an empty dot and a cross, the next step is to immediately eliminate the cross.
To eliminate a cross from a diagram, we first note that the corresponding index can ignore other indices on tensors that contain the index, which is to say that we can replace the sum over $a\in\set{b,c}$ in \eqref{eq:example_elim_x} by a sum over $a\in\set{c}$.
For each remaining index addressed by a cross, meanwhile, the cross simply forces the corresponding indices to be equal, which is equivalent to moving a filled dot into a different region:
\begin{align}
  \diagram{example_elim_x}
  = \sum_{\p{b,c}\in\D_N\p{2}} \sum_{a\in\set{c}} v\p{a,b} w\p{b,c}
  = \sum_{\p{b,c}\in\D_N\p{2}} v\p{c,b} w\p{b,c}
  = \diagram{example_elim_x_full},
  \label{eq:example_elim_final}
\end{align}
If there are multiple filled dots for the cross to eliminate, then the result is a sum over all choices of filled dots:
\begin{align}
  \diagram{example_elim_color_start}
  \sim
  \diagram{example_elim_color_x}
  = \diagram{example_elim_color_x_r}
  + \diagram{example_elim_color_x_m}
  + \diagram{example_elim_color_x_b}
  \sim
  3 \diagram{example_elim_color_end}.
\end{align}
We can generally repeat the above process of sequentially eliminating filled dots and crosses until only empty dots remain, at which point we have made all indices in a diagram independent.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Matrix elements of multi-spin operators}

The process of constructing and simplifying diagrams to compute scalar coefficients of the expansions in \eqref{eq:sym_prod_diagrams} and \eqref{eq:sum_diagrams_alt} is systematic enough to perform algorithmically.
The last step necessary to automate the evaluation of these expansions is the projection of multi-spin operators $\ot{\m X}\p{\ell}$ onto the PS manifold $\M_0$.
Without loss of generality, we consider a single $M$-spin operator $X$.
Our task is essentially to find coefficients of the expansion
\begin{align}
  \P_0 X \P_0
  = \sum_{a,b\in\A_n\p{N}} \bk{a|X|b} \op{a}{b},
\end{align}
where $\A_n\p{N}$ is the set of all ways to assign $N$ (identical) spins to $n$ (distinct) states, such that for any $a\in\A_n\p{N}$, the state $\ket{a}=\ket{a_1,a_2,\cdots,a_n}$ is labeled by the occupation number $a_\mu$ of state $\mu$, with $\sum_\mu a_\mu = N$.
Written out explicitly,
\begin{align}
  \ket{a} = \f1{\sqrt{\C\p{a}}}
  \sum_{\substack{\t{distinct}\\\t{permutations}\\\Pi~\t{of}~\tilde a}}
  \Pi \ket{\tilde a},
  &&
  \ket{\tilde a} \equiv \bigotimes_\mu \ket{\mu}^{\otimes a_\mu},
  &&
  \C\p{a} \equiv \f{\p{\sum_\mu a_\mu}!}{\prod_\nu a_\nu!},
\end{align}
where the multinomial coefficient $\C\p{a}$ counts the number of distinct ways to permute the tensor factors of $\ket{\tilde a}$, and enforces $\bk{a|a}=1$.
Using these states, we can expand
\begin{align}
  \bk{a|X|b}
  = \sum_{\substack{\alpha,\beta\in\A_n\p{M}\\\alpha\le a,\,\beta\le b}}
  \delta_{a-\alpha,b-\beta}
  \sqrt{\f{\C\p{\alpha}\C\p{a-\alpha}\C\p{\beta}\C\p{b-\beta}}
    {\C\p{a}\C\p{b}}}
  \bk{\alpha|X|\beta},
  \label{eq:multi_body_eval}
\end{align}
where the restriction $\alpha\le a$ and difference $a-\alpha$ are evaluated element-wise, i.e.~$\alpha\le a\implies \alpha_\mu\le a_\mu$ and $\p{a-\alpha}_\mu=a_\mu-\alpha_\mu$ for all $\mu$; and $\delta_{cd}=1$ if $c=d$ and zero otherwise.
We sum over both $\alpha$ and $\beta$ above merely to keep the expression symmetric with respect to transposition; in practice, one can simply sum over $\alpha\in\A_n\p{M}$ and set $\beta=b-a+\alpha$ (throwing out terms with any $\beta_\mu<0$).
Note that, by slight abuse of notation, the operator $X$ on the left of \eqref{eq:multi_body_eval} acts on an arbitrary choice of $M$ spins (out of $N$), whereas the operator $X$ on the right of \eqref{eq:multi_body_eval} is simply an $M$-spin operator, with matrix elements $\bk{\alpha|X|\beta}$ evaluated with respect to the $M$-spin states $\ket\alpha,\ket\beta$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Two single-body operators}
\label{sec:single_pair_prod}

Here we simplify the single-body product
\begin{align}
  \O\p{\m v,X} \O\p{\m w,Y}
  = \sum_{i,j} v_i w_j X_i Y_j
  \EQPS \diagram{prod_b11_c0} X_1 Y_2
  + \diagram{prod_b11_c1} X_1 Y_1,
\end{align}
where $i,j$ index individual spins that the single-spin operators $X,Y$ act on; and $\m m\equiv\sum_i m_i\ket{i}$ for $\m m\in\set{\m v,\m w}$ are vectors of scalar coefficients.
This product appears, for example, in the calculation of the second-order effective Hamiltonian $H_1^{(2)}$ induced on the permutationally symmetric manifold $\M_0$ by a single-body perturbation.
Using the rules that we worked out in Appendix \ref{sec:diagrams} for reducing diagrams, we can simplify
\begin{align}
  \diagram{prod_b11_c0}
  = \diagram{prod_b11_c0_o} - \diagram{prod_b11_c0_x}
  = \diagram{prod_b11_c0_oo} - \diagram{prod_b11_c1}
  = \col{v}\,\col{w} - \m v \c \m w,
\end{align}
where $\col{m} \equiv \sum_i m_i$ for both $\m m\in\set{\m v,\m w}$, so
\begin{align}
  \O\p{\m v,X} \O\p{\m w,Y}
  \EQPS \p{\col{v}\,\col{w} - \m v\c\m w} X_1 Y_2
  + \m v\c\m w X_1 Y_1.
\end{align}
In order to write this result in terms of collective operators $\col{Z} \equiv \sum_i Z_i$, we expand
\begin{align}
  \col{X Y} \EQPS N X_1 Y_1,
  &&
  \col{X}\,\col{Y} \EQPS \col{XY} + N\p{N-1} X_1 Y_2,
\end{align}
which implies that
\begin{align}
  \O\p{\m v,X} \O\p{\m w,Y}
  &\EQPS \f{\col{v}\,\col{w} - \m v\c\m w}{N\p{N-1}} \,
  \col{X}\, \col{Y}
  - \f{\col{v}\,\col{w} - N \m v\c\m w}{N\p{N-1}} \, \col{XY} \\
  &\EQPS \f{\col{v}\,\col{w}}{N\p{N-1}}
  \p{\col{X}\,\col{Y} - \col{XY}}
  - \f{\m v\c\m w}{N\p{N-1}}
  \p{\col{X}\,\col{Y} - N\col{XY}}.
\end{align}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Two two-body operators}
\label{sec:two_pair_prod}

Here we simplify the two-body product
\begin{align}
  \O\p{\m v,X} \O\p{\m w,Y}
  = \sum_{i\ne j} \sum_{k\ne\ell}
  v_{ij} w_{k\ell} X_{ij} Y_{k\ell},
\end{align}
where $i,j,k,\ell$ index individual spins that the two-spin operators $X,Y$ act on; and $\m m\equiv\sum_{i,j}m_{ij}\op{i}{j}$ for $\m m\in\set{\m v,\m w}$ are matrices of scalar coefficients, with $m_{ii}=0$.
This product appears, for example, in the calculation of the second-order effective Hamiltonian $H_2^{(2)}$ induced on the permutationally symmetric manifold $\M_0$ by a two-body perturbation.
The diagrammatic expansion of this operator is
\begin{align}
  \O\p{\m v,X} \O\p{\m w,Y}
  \EQPS \diagram{prod_b22_c0} \spp{X,Y}_4 + \diagram{prod_b22_c1}
  \spp{X,Y}_3 + \diagram{prod_b22_c2} \spp{X,Y}_2,
  \label{eq:two_prod_pair}
\end{align}
where $\spp{X,Y}_q$ denotes a sum over all $q$-spin operators that can be constructed from $X$ and $Y$, i.e.
\begin{align}
  \spp{X,Y}_4 \EQPS X_{1,2} Y_{3,4}, &&
  \spp{X,Y}_2 \EQPS X_{1,2} Y_{1,2} + X_{1,2} Y_{2,1},
\end{align}
\begin{align}
  \spp{X,Y}_3 \EQPS X_{1,2} Y_{1,3} + X_{1,2} Y_{3,1}
  + X_{1,2} Y_{2,3} + X_{1,2} Y_{3,2}.
\end{align}
To minimize the computational cost of the diagrams in \eqref{eq:two_prod_pair}, we reduce
\begin{align}
  \diagram{prod_b22_c1}
  &= \sp{\diagram{prod_b22_c1_o}} - \sp{\diagram{prod_b22_c1_x}} \\
  &= \sp{\diagram{prod_b22_c1_oo} - \diagram{prod_b22_c1_ox}}
  - \sp{\diagram{prod_b22_c2}} \label{eq:mid_zero} \\
  &= \diagram{prod_b22_c1_ooo} - \diagram{prod_b22_c2_oo}
\end{align}
where the middle diagram in \eqref{eq:mid_zero} is zero from the fact that $w_{ii}=0$, and
\begin{align}
  \diagram{prod_b22_c0}
  &= \sp{\diagram{prod_b22_c0_o}} - \sp{\diagram{prod_b22_c0_x}} \\
  &= \sp{\diagram{prod_b22_c0_oo} - \diagram{prod_b22_c0_ox}}
  - \sp{2 \diagram{prod_b22_c1}} \\
  &= \sp{\diagram{prod_b22_c0_oooo} - 2 \diagram{prod_b22_c1_o}}
  - \sp{2 \diagram{prod_b22_c1_ooo} - 2 \diagram{prod_b22_c2_oo}} \\
  &= \diagram{prod_b22_c0_oooo} - 4 \diagram{prod_b22_c1_ooo}
  + 2 \diagram{prod_b22_c2_oo}.
\end{align}
If the matrices $\m v,\m w$ are symmetric, then we can define
\begin{align}
  \v{\m m} \equiv \sum_{i\ne j} m_{ij} \ket{ij},
  &&
  \v m \equiv \sum_{i,j} m_{ij} \ket{i},
  &&
  \col{m} \equiv \sum_{i,j} m_{ij},
\end{align}
for both $\m m\in\set{\m v,\m w}$, and expand
\begin{align}
  \diagram{prod_b22_c2} = \v{\m v} \c\v{\m w},
  &&
  \diagram{prod_b22_c1} =
  \v v \c \v w - \v{\m v} \c \v{\m w},
  &&
  \diagram{prod_b22_c0}
  = \col{v}\,\col{w} - 4 \v v\c\v w + 2 \v{\m v}\c\v{\m w}.
\end{align}
In the special case that $X=Y=Z\otimes Z$ with a single-spin operator $Z$ for which $Z^2=1$, we can simplify
\begin{align}
  \spp{Z\otimes Z,Z\otimes Z}_4 \EQPS Z^{\otimes 4},
  &&
  \spp{Z\otimes Z,Z\otimes Z}_3 \EQPS 4 Z^{\otimes 2},
  &&
  \spp{Z\otimes Z,Z\otimes Z}_2 \EQPS 2,
\end{align}
which implies that
\begin{align}
  \O\p{\m v,Z\otimes Z} \O\p{\m w,Z\otimes Z}
  \EQPS A_4 Z^{\otimes 4} + A_2 Z^{\otimes 2} + A_0,
\end{align}
with
\begin{align}
  A_4 \equiv \diagram{prod_b22_c0}
  = \col{v}\,\col{w} - 4 \v v\c\v w + 2 \v{\m v}\c\v{\m w},
  &&
  A_2 \equiv 4 \diagram{prod_b22_c1}
  = 4 \v v \c \v w - 4 \v{\m v} \c \v{\m w}
\end{align}
\begin{align}
  A_0 \equiv 2 \diagram{prod_b22_c2} = 2 \v{\m v} \c\v{\m w}.
\end{align}
In terms of collective operators $\col{Z}\equiv\sum_iZ_i$, we have
\begin{align}
  \O\p{\m v,Z\otimes Z} \O\p{\m w,Z\otimes Z}
  \EQPS \tilde A_4 \col{Z}^4 + \tilde A_2 \col{Z}^2 + \tilde A_0,
\end{align}
where
\begin{align}
  \tilde A_4 \equiv \f{A_4}{N!_4},
  &&
  \tilde A_2 \equiv \f{A_2}{N!_2} - 2\p{3N-4} \f{A_4}{N!_4},
  &&
  \tilde A_0 \equiv A_0 - \f{A_2}{N-1} + \f{3A_4}{\p{N-1}\p{N-3}},
\end{align}
and
\begin{align}
  N!_q \equiv \sum_{j=0}^{q-1} \p{N-j}
\end{align}
is a falling factorial.

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \subsection{Three single-body operators}
% \label{sec:single_trip_prod}

% Here we simplify the single-body product
% \begin{align}
%   \O\p{\m u,X} \O\p{\m v,Y} \O\p{\m w,Z}
%   = \sum_{i,j,k} u_i v_j w_k X_i Y_j Z_k,
% \end{align}
% where $i,j,k$ index individual spins that the single-spin operators $X,Y,K$ act on; and $\m m\equiv\sum_i m_i\ket{i}$ for $\m m\in\set{\m u,\m v,\m w}$ are vectors of scalar coefficients.
% This product appears, for example, in the calculation of matrix elements of the operator $\O\p{\m v,Y}$ within the single-magnon manifold.
% The diagrammatic expansion of this operator is
% \begin{multline}
%   \O\p{\m u,X} \O\p{\m v,Y} \O\p{\m w,Z}
%   \EQPS \diagram{prod_b111_c0} X_1 Y_2 Z_3
%   + \diagram{prod_b111_c10} X_1 Y_1 Z_1 \\
%   + \diagram{prod_b111_c11_u} X_1 Y_2 Z_2
%   + \diagram{prod_b111_c11_v} X_2 Y_1 Z_2
%   + \diagram{prod_b111_c11_w} X_2 Y_2 Z_1,
% \end{multline}
% where we can reduce (by hand or with a computer)

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Three two-body Ising-like operators}

Here we simplify the product
\begin{align}
  \O\p{\m u,\m v,\m w,Z}
  &\equiv \O\p{\m u,Z\otimes Z} \O\p{\m v,Z\otimes Z}
  \O\p{\m w,Z\otimes Z} \\
  &= \sum_{i\ne j} \sum_{k\ne\ell} \sum_{p\ne q}
  u_{ij} v_{k\ell} w_{pq} Z_i Z_j Z_k Z_\ell Z_p Z_q,
\end{align}
projected onto the PS manifold $\M_0$, where $i,j,k,\ell,p,q$ index individual spins; $Z$ is a single-spin Ising-like operator, with $Z^2=1$; and $\m m\equiv\sum_{i,j} m_{ij}\op{i}{j}$ for $\m m\in\set{\m u,\m v,\m w}$ are matrices of scalar coefficients, with $m_{ii}=0$.
This product appears, for example, in the projection of Ising interactions onto the two-magnon manifold $\M_2$.
Collecting terms according to the number of $Z$ operators that remain after accounting for $Z^2=1$, we find that
\begin{align}
  \O\p{\m u,\m v,\m w,Z}
  \EQPS A_6 Z^{\otimes 6} + A_4 Z^{\otimes 4} + A_2 Z^{\otimes 2} + A_0,
  \label{eq:triple_multi}
\end{align}
with
\begin{align}
  A_6 \equiv \diagram{prod_b222_c0},
  &&
  A_4 \equiv 4 \diagram{prod_b222_c01} + 8 \diagram{prod_b222_c1},
  &&
  A_0 \equiv 8 \diagram{prod_b222_c0111},
  \label{eq:triple_64}
\end{align}
\begin{align}
  A_2 \equiv 8 \diagram{prod_b222_c011} + 2 \diagram{prod_b222_c02}
  + 8 \diagram{prod_b222_c11} + 4 \diagram{prod_b222_c2},
  \label{eq:triple_2}
\end{align}
where a diagram with unlabeled circles denotes a sum over all distinct label assignments, e.g.
\begin{align}
  \diagram{prod_b222_c1} \equiv \diagram{prod_b222_c1_uvw},
  \label{eq:triple_uvw}
\end{align}
\begin{align}
  \diagram{prod_b222_c011}
  \equiv \diagram{prod_b222_c011_uvw}
  + \diagram{prod_b222_c011_vwu} + \diagram{prod_b222_c011_wuv}.
  \label{eq:triple_uvw_3}
\end{align}
The last diagram in each of $A_6,A_4,A_2,A_0$ as written in \eqref{eq:triple_64}, \eqref{eq:triple_2} thus has only one assignment of labels, as in \eqref{eq:triple_uvw}, while all other unlabeled diagrams have three assignments, as in \eqref{eq:triple_uvw_3}.

Having identified diagrammatic representations of the coefficients $A_6,A_4,A_2,A_0$, we now need to simplify all diagrams to reduce their computational complexity, as discussed in Appendix \ref{sec:diagrams}.
The process of eliminating filled dots and crosses from a diagram is algorithmic enough to execute on a computer, doing which we find
\begin{align}
  A_6 = D_6 - D_4 + D_2 - D_0,
  &&
  A_4 = D_4 - 2 D_2 + 3 D_0,
  &&
  A_2 = D_2 - 3 D_0,
  &&
  A_0 \equiv D_0,
\end{align}
where
\begin{align}
  D_6 \equiv \diagram{prod_b222_c0_o},
  &&
  D_4 \equiv 4 \diagram{prod_b222_c01_o} - 16 \diagram{prod_b222_c1_o},
  &&
  D_0 \equiv 8 \diagram{prod_b222_c0111_o},
\end{align}
\begin{align}
  D_2 \equiv 8 \diagram{prod_b222_c011_o}
  + 2 \diagram{prod_b222_c02_o}
  - 16 \diagram{prod_b222_c11_o}
  + 16 \diagram{prod_b222_c2_o}.
\end{align}
If $m_{ij}=m_{ji}$ for all $\m m\in\set{\m u,\m v,\m w}$, then we can define
\begin{align}
  \v{\m m} \equiv \sum_{i\ne j} m_{ij} \ket{ij},
  &&
  \v m \equiv \sum_i m_i \ket{i},
  &&
  m_i \equiv \sum_j m_{ij},
  &&
  \col{m} \equiv \sum_{i,j} m_{ij},
\end{align}
and expand
\begin{align}
  D_6 = \col{u}\,\col{v}\,\col{w},
  &&
  D_4 = 4\p{\col{u}\,\v v\c\v w + \col{v}\,\v w\c\v u
  + \col{w}\,\v u\c\v v} - 16 \sum_i u_i v_i w_i,
  &&
  D_0 = 8 \tr\p{\m u \c \m v \c \m w},
\end{align}
\begin{multline}
  D_2 = 8\p{\v u \c\m v\c\v w + \v v \c\m w\c\v u + \v w \c\m u\c\v v}
  + 2\p{\col{u}\,\v{\m v}\c\v{\m w} + \col{v}\,\v{\m w}\c\v{\m u}
  + \col{w}\,\v{\m u}\c\v{\m v}} \\
  - 16 \sum_{i,j} \p{u_i v_{ij} w_{ij}
    + v_i w_{ij} u_{ij} + w_i u_{ij} v_{ij}}
  + 16 \sum_{i,j} u_{ij} v_{ij} w_{ij}.
\end{multline}
Finally, we can also write the product $\O\p{\m u,\m v,\m w,Z}$ in terms of the collective operator $\col{Z} \equiv \sum_i Z_i$:
\begin{align}
  \O\p{\m u,\m v,\m w,Z}
  \EQPS \tilde A_6 \col{Z}^6 + \tilde A_4 \col{Z}^4
  + \tilde A_2 \col{Z}^2 + \tilde A_0,
\end{align}
where
\begin{align}
  \tilde A_6 \equiv \f{A_6}{N!_6},
  &&
  \tilde A_4 \equiv \f{A_4}{N!_4} - 5\p{3N-8} \f{A_6}{N!_6},
\end{align}
\begin{align}
  \tilde A_2 \equiv \f{A_2}{N!_2} - 2\p{3N-4} \f{A_4}{N!_4}
  + \p{45N^2-210N+184} \f{A_6}{N!_6},
\end{align}
\begin{align}
  \tilde A_0 \equiv A_0 - \f{A_2}{N-1}
  + \f{3A_4}{\p{N-1}\p{N-3}}
  - \f{15A_6}{\p{N-1}\p{N-3}\p{N-5}}.
\end{align}
and
\begin{align}
  N!_q \equiv \sum_{j=0}^{q-1} \p{N-j}
\end{align}
is a falling factorial.

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: